### 2017-08-04 almo-langevin.tex
August 4th
Sections: Abstract, Introduction, Theory, Finding gammas, implimentation. Introduction still needs re-written linear scaling methods.

%\documentclass[journal=jacsat]{achemso}

\documentclass[10pt,aps,prb,twocolumn,amsmath,amssymb,superscriptaddress,longbibliography]{revtex4-1}

%\documentclass[preprint,showpacs,preprintnumbers,amsmath,amssymb]{revtex4-1}
%\documentclass[twocolumn,showpacs,preprintnumbers,amsmath,amssymb]{revtex4}
% Some other (several out of many) possibilities
%\documentclass[preprint,aps]{revtex4}
%\documentclass[preprint,aps,draft]{revtex4}
%\documentclass[prb,amsmath,amssymb]{revtex4}% Physical Review B

\usepackage{graphicx}% Include figure files
\usepackage{dcolumn}% Align table columns on decimal point
\usepackage{bm}% bold math
\usepackage{amsmath}% bold math
\usepackage{siunitx}% si units
\usepackage{color}
\usepackage{graphicx}
%\usepackage[normalem]{ulem}
%\nofiles

\bibliographystyle{achemso}
%\bibliographystyle{naturemag}
%\bibliographystyle{abbrv}

\begin{document}

\title{
Low-cost linear-scaling \emph{ab initio} molecular dynamics for weakly-interacting systems
}

\author{Hayden Scheiber}
\email{hayden.scheiber@mcgill.ca}
\author{Yifei Shi}
\author{Rustam Z. Khaliullin}
\email{rustam.khaliullin@mcgill.ca}
\affiliation{Department of Chemistry, McGill University, 801 Sherbrooke St. West, Montreal, QC H3A 0B8, Canada}

\date{\today}

\begin{abstract}
Linear scaling density functional theory based on absolutely localized molecular orbitals is used in conjunction with the Langevin equation of motion to create a low-cost molecular dynamics method for weakly coupled molecular systems which retains sufficient accuracy to generate correct dynamical properties in water.

The approach utilizes a systematically inaccurate linear-scaling potential energy surface (PES) calculation method. 
Inaccuracies in the calculation of force using this method are balanced by a weighted version of the Langevin Equation of motion. 
The stochastic ``white noise'' of the Langevin Equation acts to smooth out PES discontinuities while also thermostating the system to maintain a constant thermodynamic temperature. 
We show that Langevin dynamics can be successfully implemented to reduce computational cost by allowing systematic error in PES calculation to be absorbed into the stochastic and damping forces associated with the Langevin equation. 
We also introduce a simple heuristic approach to find the best Langevin Equation parameters to use for balancing system-specific imperfect forces and maintaining correct dynamical behavior.
\end{abstract}

\maketitle
\section{Introduction}
Since the unification of molecular dynamics (MD) and density functional theory (DFT) in 1985 by Car and Parrinello,~\cite{a:thecpmd} much research has been focused on reducing the computational overhead of \emph{ab initio} molecular dynamics (AIMD) potential energy calculations.~\cite{b:aimd} 
% Remove Following text for some journals
AIMD provides a great advantage over traditional predefined-potential molecular dynamics in that it calculates its PES on-the-fly, rather than needing to be laboriously optimized to the specific system of interest beforehand. 
This allows for exploration of much more chemically complex systems than are possible with traditional MD methods. 
AIMD also allows for much greater predictive capability than traditional MD, as a first principles approach is not confined to a predetermined range of system parameters. 
The tradeoff is the computational complexity involved in calculating an accurate set of forces for each atom at every time step of the simulation.~\cite{a:2gen-cpmd}
% End remove

The computational cost of the straightforward approach to Kohn-Sham DFT is known to grow (or \emph{scale}) cubically with system size because it involves either diagonalization of the effective Kohn-Sham Hamiltonian or, if the direct optimization of Kohn-Sham orbitals is performed, the equally costly inversion of a preconditioner.~\cite{a:ot} % Add other references here?
This cubic scaling severely constricts the possible length and time scales accessible by AIMD using such an approach. 
In an attempt to address this issue, there has been much research into the possibility of linear scaling (LS) DFT methods for the calculation of forces.

% Most methods are based on localization of electrons into small sub domains of the system, so that long range charge transfer is ignored.~\cite{a:linscale3,a:lee-yang-1996,a:ls-scuseria-1997,a:ls-manolopoulos-1998,a:ls-helgaker-2001,a:ls-niklasson-2003,a:curvy2,a:ls-dm-sign,a:ls-parinello} So far the available methods still suffer from a high computational cost prefactor, even though they scale linearly for large systems. This has limited the available time scales for LS AIMD severely.

% Re-write the following section and include any recent advances
A number of such alternative methods explore the natural sparsity of the one-electron density matrix (DM)~\cite{a:ls-kohn-1996,a:ls-rev-1999} and are capable of yielding linear scaling (LS) for large systems.~\cite{a:linscale3,a:lee-yang-1996,a:ls-scuseria-1997,a:ls-manolopoulos-1998,a:ls-helgaker-2001,a:ls-niklasson-2003,a:curvy2,a:ls-dm-sign} 
Unfortunately, the variational optimization of the DM is very inefficient for accurate DFT calculations~\cite{a:ls-rev-1999,a:ls-dm-sign}, which require many basis functions per atom. 
Therefore, the applications of DM-based LS methods have been limited to minimal-basis tight-binding problems. 
The optimal basis variants of the DM methods~\cite{a:ls-stechel-1994,a:ls-gillan-1995,a:ls-gillan-1996,a:ls-onetep-2003} designed to address this issue contract the large basis set into a small number of new localized basis functions and then optimize the DM in the contracted basis. 
Although such methods have been successfully used for the evaluation of accurate DFT energies of very large systems~\cite{a:ls-onetep-2009,a:ls-conquest-2010,a:ls-onetep-2010-app1,a:ls-rev-2012} their application for AIMD is hampered by the computationally costly optimization of both the contracted orbitals and the density matrix.
%
From this point of view, methods based on the direct optimization of localized KS orbitals~\cite{a:ls-galli-parrinello-1992,a:ls-mauri-galli-car-1993,a:ls-ordejon-1993,a:ls-mauri-galli-1994,a:ls-ordejon-1995,a:ls-kim-mauri-galli-1995,a:ls-fattebert-2004,a:ls-fattebert-2006,a:burger-yang-2008} are advantageous since they require only the occupied orbitals and, thus, deal with fewer variational degrees of freedom than the DM methods. 
Unfortunately, the progress in the development of orbital-based LS methods has been hindered by the inherently difficult convergence of the localized-orbital optimization.~\cite{a:ls-mauri-galli-car-1993,a:ls-ordejon-1995,a:ls-fattebert-2004,a:ls-rev-1999,a:weitao-yang-2013} 
%Furthermore, numerous proposed LS functionals exhibit multiple local minima, which limits their applicability in molecular dynamics simulations because of the resulting poor conservation of the total energy~\cite{a:ls-rev-1999}.

Hence, the computational prefactor of the existing LS DFT methods~\cite{a:ls-rev-1995,a:ls-rev-1999,a:ls-rev-2002,a:ls-rev-2012} remains very high and most AIMD simulations are still performed using conventional low-prefactor cubic scaling DFT methods.

Here, we introduce a new low-cost linear scaling AIMD method for weakly-interacting molecular systems based on the recently developed absolutely localized molecular orbitals (ALMO) DFT approach.~\cite{a:almo-ls}

This method exploits non-orthogonal molecular orbitals (MOs) in a two-step self-consistent filed (SCF) approach that is designed to avoid the usual convergence issues.~\cite{a:ls-mauri-galli-car-1993,a:ls-ordejon-1995,a:ls-fattebert-2004,a:ls-rev-1999} 
We show that due to its linear scaling nature and relatively low computational prefactor, ALMO-based DFT calculations allow for simulation of very large weakly-coupled molecular systems. 
To decrease the cost even further, the present paper builds on the ALMO SCF method as well as other recent work involving Brownian dynamics to balance force calculation errors.~\cite{a:ls-parinello,a:2ndcpmd,a:langevin-why} 
We will show that our approach lowers the computation cost pre-factor for the ALMO SCF method, while maintaining LS and accuracy for calculation of dynamical properties. 
This is done by balancing a new computationally economical but systemically inaccurate force calculation method with an optimized version of the Langevin Equation of motion. 
We also introduce an easy to follow algorithmic approach to optimize a system for generating correct dynamical behaviour.

\maketitle
\section{Theory}

Our approach is based on the ALMO SCF two-step variational method.~\cite{a:almo-ls} 
Initially, the electrons are partitioned into groups based on a logical approach, generally corresponding to molecules. 
In the first step, the basis atomic orbitals (AOs) are constrained to their parent molecules and a variational optimization is performed under the Born-Oppenheimer approximation using the DFT-SCF method within the non-overlapping domains. 
This is akin to calculating the electronic structure of each molecule (with frozen nuclei) within a coulombic field produced by the surrounding molecules. 
Charge transfer is completely constrained at this step. 
In the second step, the variationally optimized but constrained MOs are allowed to delocalize onto nearby groups up to a desired cutoff radius $R_{c}$. 
This step requires a second variational optimization to be performed. 
Because of the logical partitioning of the groups and ignoring of long range forces, both steps of the electronic structure calculation are linear scaling with respect to system size. 
For more details on this method, see ref.\ \citenum{a:almo-ls}.

A known source of error in force calculation within the ALMO SCF paradigm occurs due to the discontinuous jump in potential energy when a neighboring molecule passes beyond the cutoff radius $R_{c}$ and is suddenly ignored in potential energy calculations. 
Setting $R_{c}$ to a larger value reduces this discontinuity, but greatly increases the computational cost as longer distance (and less important) interactions must be calculated. 
As long as the discontinuity in force is small compared with the random error of the calculation, it can be safely ignored. Previously, this meant that $R_{c}$ had to be kept relatively large to maintain good accuracy of dynamical properties.

In the present article, we propose a new approach which allows for relaxation of the constraint on $R_{c}$ by introducing related damping and stochastic terms to the force calculation by utilizing the Langevin Equation of motion.~\cite{a:Kubo-1986}

\begin{align}
\label{eq:langevin}
m_i \ddot{r}_{i\alpha} = f^{\text{SCF}}_{i\alpha} - \gamma_L m_i \dot{r}_{i\alpha} + R^{\gamma_L}_{i\alpha} (t),
\end{align}

Where $m_i$ refers to the mass of the i-th particle, and $r_{i\alpha}$ its position along dimension $\alpha$. 
$f^{\text{SCF}}_{i\alpha}$ is the force calculated using the ALMO SCF method described above for the i-th particle along dimension $\alpha$, $\gamma_L$ is the Langevin scaling factor which sets the strength of both the damping term  $\gamma_L m_i \dot{r}_{i\alpha}$ and the stochastic Gaussian ``white noise'' term $R^{\gamma_L}_{i\alpha} (t)$. 
In order to correctly sample the canonical distribution, the random force must be zero mean. 
As well, the damping and white noise terms must be related by the second fluctuation dissipation theorem.~\cite{a:Kubo-1986,a:langevin-why,b:tuckerman-stat}
%
\begin{align}
\label{eq:stochastic}
\langle R^{\gamma_L}_{i\alpha} (t) \rangle &= 0, \\
\langle R^{\gamma_L}_{i\alpha} (t)  R^{\gamma_L}_{j\beta} (t') \rangle &= 2 k_B T \gamma_L m_i \delta_{ij} \delta_{\alpha\beta} \delta(t-t')
\end{align}
%
This means the damping and white noise terms will work to maintain a constant temperature while also generating the correct Maxwell-Boltzmann distribution of velocities. 
It has previously been shown that Langevin dynamics can be utilized to correct imperfect force calculations while maintaining accurate dynamical properties.~\cite{a:langevin-why,a:2ndcpmd,b:tuckerman-stat,a:ceriotti} 
By utilizing Eq.(\ref{eq:langevin}) in conjunction with the ALMO DFT method of force calculation, we will show that discontinuity in the PES can be ignored for sufficiently large $\gamma_L$. 
Note that increasing $\gamma_L$ increases the sample size required to generate accurate dynamical properties, as random Gaussian noise is introduced.

The concept of balancing an error in force calculation with the Langevin equation can be taken even further, thereby reducing computational overhead even more. 
During the second step of the ALMO SCF method described above, a self-consistent variational approach is taken to converge the non-orthogonal MOs. 
Generally, this step utilizes a iterative search function in order to reduce the absolute norm of the energy gradient to some max allowable value $\epsilon_{err}$. 
Previously, $\epsilon_{err}$ had to be set very low to ensure sufficient accuracy of the electron structure calculations. 
By increasing the allowable $\epsilon_{err}$, a great reduction in computational cost can be achieved. 
Obviously, this introduces a significant error into the calculation of forces during each time step. 
We assume that this error has the form
%
\begin{align}
\label{eq:assumption}
f^{\text{SCF}}_{i\alpha} = f^{\text{APP}}_{i\alpha} + R^{\Delta}_{i\alpha} (t)
\end{align}
%
Where $f^{\text{SCF}}_{i\alpha}$ is the fully converged ALMO DFT force calculation, $f^{\text{APP}}_{i\alpha}$ is the approximate force calculated by allowing a large $\epsilon_{err}$, and $R^{\Delta}_{i\alpha} (t)$ is a stochastic ``white noise'' term that obeys
%
\begin{align}
\label{eq:stochastic2}
\langle R^{\Delta}_{i\alpha} (t) \rangle &= 0, \\
\label{eq:stochastic3}
\langle R^{\Delta}_{i\alpha} (t)  R^{\Delta}_{j\beta} (t') \rangle &= 2 k_B T \Delta m_i \delta_{ij} \delta_{\alpha\beta} \delta(t-t')
\end{align}
%
We will see that this assumption is sufficiently justified later (see fig. \ref{fig:randomforce}). 
The Langevin equation of motion (\ref{eq:langevin}) can be re-written with the two stochastic terms combined into one $\gamma = \gamma_L+\Delta$.
%
\begin{align}
\label{eq:langevin2}
m_i \ddot{r}_{i\alpha} = f^{\text{APP}}_{i\alpha} - \gamma_L m_i \dot{r}_{i\alpha} + R^{\gamma}_{i\alpha} (t)
\end{align}
%
Note that $\Delta$ can be either positive or negative depending on the way in which errors are introduced in the system of interest. 
$\Delta$ was found to be negative for all systems considered here.

\section{Optimizing the values of $\gamma_L$ and $\Delta$}

In theory, both $\gamma_L$ and $\Delta$ should be system-size independent, but in practice only $\gamma_L$ is.
$\Delta$ is dependent on system size due to finite size effects, which decay for increasing system size.
Fortunately, $\gamma_L$ and $\Delta$ can be separately optimized for a given system of interest, assuming a method of accurate force calculation can be applied to the system for a short trajectory.

An optimized $\gamma_L$ is one that is: large enough to adequately smooth force discontinuities; large enough to correctly thermostat the system of interest; and small enough not to significantly affect dynamical properties. 
The relative weight of the three is determined by the specifics of the system and what is most important to the researcher. 
To optimize $\gamma_L$ given the relative importance of the above properties, we suggest running several short trajectories on the system of interest (or a scaled-down version of it) using a tightly converged PES calculation and varying the value of $\gamma_L$. 
We found that a value of $\gamma_L$ between \SI{1e-5}{\per\fs} and \SI{1e-3}{\per\fs} was optimal for the purposes of the research presented here.

An optimized $\Delta$ is one that, in the absence of $\gamma_L$, perfectly offsets any systematic inaccuracies in the force calculation. 
A trajectory with inaccurate forces, \mbox{$\gamma_L = \SI{0}{\per\fs}$}, and optimized $\Delta$ should have no long term drift in the total energy of the system. Short term energy fluctuations should be expected, and can later be ``smoothed out'' by setting a finite $\gamma_L$.

$\Delta$ can therefore be optimized simply by trying various values of $\Delta$ (with \mbox{$\gamma_L = 0$}) until the simulation described by Eq. (\ref{eq:langevin2}) has long-term stable total energy and satisfies the equipartition theorem. 
%
\begin{align}
\label{eq:eqipartition}
\langle \dfrac{1}{2} m_i \dot{\bm{r}}^{2}_{i} \rangle &= \dfrac{3}{2} k_{B} T
\end{align}
%
Another approach based on Eq. (\ref{eq:assumption}) can be used to find $\Delta$ within an order of magnitude if one can afford computing tightly converged SCF forces for a short MD trajectory. 
First, run the simulation with $\Delta = 0$ and converged SCF forces. 
Second, compute approximate forces for snapshots along the trajectory and calculate the random term $R^{\Delta}_{i\alpha} (t)$ using Eq. \ref{eq:assumption}. 
Third, calculate the autocovariance function (ACF) of the noise $\text{ACF}(t) = \langle R^{\Delta}_{i\alpha} (0)  R^{\Delta}_{i\alpha} (t) \rangle$, which should be sharply peaked around zero lag time if our assumption is correct. 
Finally, integrate Eq.(\ref{eq:stochastic3}) to obtain the following expression for $\Delta$.
%
\begin{align}
\label{eq:delta}
\Delta &= (2 k_B T m_i )^{-1} \int_{-\infty}^{\infty}\langle R^{\Delta}_{i\alpha} (0)  R^{\Delta}_{i\alpha} (t) \rangle dt
\end{align}
%
In other words, we use the integral of the ACF to estimate the strength of the random force. 
In practice, this method for estimating $\Delta$ is not particularly accurate due to rapidly oscillating correlations between sequential values of $R^{\Delta}_{i\alpha} (t)$, the ACF of which is not possible to accurately integrate without running a reference trajectory with extremely short time-steps (see figure~\ref{fig:randomforce}).

Once $\gamma_L$ and $\Delta$ are both optimized, approximate forces can be used to perform long MD simulations efficiently.
\bigskip

\begin{figure}
\includegraphics[trim={2.2cm 0.5cm 3cm 0.2cm},clip,width=8cm]{125mol_ALMO_Early_Random_force.eps}
\caption{\label{fig:randomforce}The average ``Random force'' along a water molecule dimension $\langle R^{\Delta}_{i\alpha} \rangle (t)$ calculated as the difference between the fully converged Born-Oppenheimer forces and unconverged ALMO DFT forces, where $\epsilon_{err} = \num{1e-2}$. 
(a) Plot of $\langle R^{\Delta}_{i\alpha} \rangle (t)$ as a function of time. 
Here it is clear that Eq.\ (\ref{eq:stochastic2}) is satisfied. 
(b) Autocovariance function of $\langle R^{\Delta}_{i\alpha} \rangle (t)$. 
The ACF is clearly not a delta-function, and is characterized by rapid oscillation between positive and negative correlation. 
However, the autocorrelation decays rapidly enough that these forces can be considered random and uncorrelated to a good approximation for sufficiently long time scales. 
Calculations were performed with the PBE $\mathrm{E_{XC}}$ functional and the TZV2P basis set on 128 water molecules.}
\end{figure}

\maketitle
\section{Implementation}

The methods presented here are built into CP2K, an open source MD simulation package.~\cite{www:cp2k} 
The computation scheme used is mostly identical to that described in the implementation section of ref.\ \citenum{a:almo-ls}. 
For the dynamical properties calculations using ALMO DFT calculation methods, a periodic cell containing 128 water molecules was simulated for $\SI{40}{\ns}$ at a constant temperature of $\SI{298}{\K}$ and a constant density of $\SI{0.9966}{\g\per\cm^{3}}$. 
The system was thermostated using the Langevin ensemble (eq.\ \ref{eq:langevin2}) where $\gamma_L$ was set to $\SI{1e-3}{\per\fs}$, and $\Delta$ was optimized to the order of $10^{-5}$ - $10^{-4}\ \mathrm{fs^{-1}}$. 
Note that for timing benchmarks, $\Delta$ was not optimized. 
This should not have a perceptible effect on calculation time.
Simulations for timing benchmarks were run for ten time steps, which is sufficient for convergence of wavefunctions from an initial guess.
The TZV2P basis set was used to represent MOs, and a cutoff of $\SI{320}{Ry}$ used to represent electron density. 
The exchange-correlation energy was approximated using the PBE functional, which expands upon the local density approximation by considering derivatives of the electron density at each point.~\cite{a:PBEfunctional}

The reference PES was calculated using the orbital transformation (OT) method~\cite{a:ot,a:ot2} and the canonical sampling through velocity rescaling (CSVR) thermostat~\cite{a:CSVR} with a time constant of $\SI{200}{fs}$. 
Otherwise, system parameters were identical for the reference.

IR spectra, radial distribution functions, diffusivity plots, and diffusion constants were calculated using the open source trajectory analyzer program TRAVIS.~\cite{a:travis-main,a:travis-ir1,a:travis-ir2}

How $\gamma$, $\gamma_L$, and $\gamma_{\Delta}$ are related to GAMMA, NOISY\_GAMMA, and SHADOW\_GAMMA in CP2K?

\begin{align}
\label{eq:rels}
GAMMA + NOISY\_GAMMA &= \gamma_L + \gamma_{\Delta} = \gamma \nonumber \\
GAMMA + SHADOW\_GAMMA &= \gamma_L = \gamma - \gamma_{\Delta} \nonumber
\end{align}
%
One set of useful expressions
%
\begin{align}
\gamma_{\Delta} &= NOISY\_GAMMA - SHADOW\_GAMMA \nonumber \\
\gamma_L &= GAMMA + SHADOW\_GAMMA \nonumber \\
\gamma &= GAMMA + NOISY\_GAMMA \nonumber
\end{align}
%
%and another one
%%
%\begin{align}
%GAMMA &=  \nonumber \\
%SHADOW\_GAMMA &= \nonumber \\
%NOISY\_GAMMA &= \nonumber
%\end{align}
%%

Practical guide:
%
\begin{itemize}
\item The code designed so that any effect can be achieved with positive GAMMAs.
\item Situation 1. NVT with perfect forces. Set NOISY and SHADOW to zeros to produce balanced canonical distributions. Large GAMMA adds lots of noise but leads to fast equilibration. Small GAMMA -- approach NVE. 
\item Situation 2. NVT with imperfect forces. Use either NOISY or SHADOW but not both. The value of GAMMA is chosen as in the previous situation. A. Set SHADOW to finite value to get excessive damping. B. Set NOISY to get excessive stochastic boost.
\end{itemize}

Integrator implemented in \$CP2K/src/motion/integrator.F eqns (19) and (20) from http://dx.doi.org/10.1080/0026897031000108113. Another integrator can be implemented using Eq (15.5.17) from the book of Tuckerman.

\section{Conclusions}

\textbf{Acknowledgments.} The research was funded by the Natural Sciences and Engineering Research Council of Canada through the Discovery Grant. The authors are grateful to Compute Canada and McGill HPC Centre for computer time.

%\textbf{Supporting Information}
%Calculated radial distribution functions of liquid water, comparison of timing benchmarks for the DZVP and TZV2P basis sets, timing benchmarks for systems containing 32,768 water molecules, timing benchmarks for the Kohn-Sham matrix build. This material is available free of charge via the Internet at http://pubs.acs.org.

\bibliography{almo-langevin}

\end{document}

### 2017-08-04 almo-langevin.tex
August 4th, v2
Small changes made throughout.

@@
 % Remove Following text for some journals
-AIMD provides a great advantage over traditional predefined-potential molecular dynamics in that it calculates its PES on-the-fly, rather than needing to be laboriously optimized to the specific system of interest beforehand.
+AIMD provides a great advantage over traditional predefined-potential molecular dynamics in that it calculates its potential energy surface (PES) on-the-fly, rather than needing to be laboriously optimized to the specific system of interest beforehand.
 This allows for exploration of much more chemically complex systems than are possible with traditional MD methods. 
@@
 
-\maketitle
 \section{Theory}
@@
 To optimize $\gamma_L$ given the relative importance of the above properties, we suggest running several short trajectories on the system of interest (or a scaled-down version of it) using a tightly converged PES calculation and varying the value of $\gamma_L$.
-We found that a value of $\gamma_L$ between \SI{1e-5}{\per\fs} and \SI{1e-3}{\per\fs} was optimal for the purposes of the research presented here.
+We found that a value of $\gamma_L$ between $10^{-5}$ and $10^{-3}\ \mathrm{fs^{-1}}$ was optimal for the purposes of the research presented here.
 
 An optimized $\Delta$ is one that, in the absence of $\gamma_L$, perfectly offsets any systematic inaccuracies in the force calculation. 
-A trajectory with inaccurate forces, \mbox{$\gamma_L = \SI{0}{\per\fs}$}, and optimized $\Delta$ should have no long term drift in the total energy of the system.
-Short term energy fluctuations should be expected, and can later be ``smoothed out'' by setting a finite $\gamma_L$.
+A trajectory with inaccurate forces, \mbox{$\gamma_L = 0$}, and optimized $\Delta$ should have no long term drift in the total energy of the system.
+Short term energy fluctuations should be expected, and can later be ``smoothed out'' by setting finite $\gamma_L$.
 
@@
 In other words, we use the integral of the ACF to estimate the strength of the random force. 
-In practice, this method for estimating $\Delta$ is not particularly accurate due to rapidly oscillating correlations between sequential values of $R^{\Delta}_{i\alpha} (t)$, the ACF of which is not possible to accurately integrate without running a reference trajectory with extremely short time-steps (see figure~\ref{fig:randomforce}).
+In practice, this method for estimating $\Delta$ is not particularly accurate due to rapidly oscillating correlations between sequential values of $R^{\Delta}_{i\alpha} (t)$.
+The ACF is not usually possible to accurately integrate without running a reference trajectory with extremely short time-steps (see figure~\ref{fig:randomforce}).
 
@@
 However, the autocorrelation decays rapidly enough that these forces can be considered random and uncorrelated to a good approximation for sufficiently long time scales. 
-Calculations were performed with the PBE $\mathrm{E_{XC}}$ functional and the TZV2P basis set on 128 water molecules.}
+Calculations were performed with the PBE $\mathrm{E_{XC}}$ functional and the TZV2P basis set on 125 water molecules.}
 \end{figure}
 
-\maketitle
 \section{Implementation}
@@
 The computation scheme used is mostly identical to that described in the implementation section of ref.\ \citenum{a:almo-ls}. 
-For the dynamical properties calculations using ALMO DFT calculation methods, a periodic cell containing 128 water molecules was simulated for $\SI{40}{\ns}$ at a constant temperature of $\SI{298}{\K}$ and a constant density of $\SI{0.9966}{\g\per\cm^{3}}$.
-The system was thermostated using the Langevin ensemble (eq.\ \ref{eq:langevin2}) where $\gamma_L$ was set to $\SI{1e-3}{\per\fs}$, and $\Delta$ was optimized to the order of $10^{-5}$ - $10^{-4}\ \mathrm{fs^{-1}}$.
+For the dynamical properties calculations using ALMO DFT calculation methods, a periodic cell containing 125 water molecules was simulated for $\SI{40}{\ns}$ at a constant temperature of $\SI{298}{\K}$ and a constant density of $\SI{0.9966}{\g\per\cm^{3}}$.
+The system was thermostated using the Langevin ensemble (eq.\ \ref{eq:langevin2}) where $\gamma_L$ was set to $\SI{e-5}{\per\fs}$, and $\Delta$ was optimized to the order of $10^{-5}$ - $10^{-4}\, \mathrm{fs^{-1}}$.
 Note that for timing benchmarks, $\Delta$ was not optimized. 
-This should not have a perceptible effect on calculation time.
+This does not have a perceptible effect on calculation time.
+
 Simulations for timing benchmarks were run for ten time steps, which is sufficient for convergence of wavefunctions from an initial guess.
@@
 
-How $\gamma$, $\gamma_L$, and $\gamma_{\Delta}$ are related to GAMMA, NOISY\_GAMMA, and SHADOW\_GAMMA in CP2K?
+\section{Accuracy}
 
-\begin{align}
-\label{eq:rels}
-GAMMA + NOISY\_GAMMA &= \gamma_L + \gamma_{\Delta} = \gamma \nonumber \\
-GAMMA + SHADOW\_GAMMA &= \gamma_L = \gamma - \gamma_{\Delta} \nonumber
-\end{align}
-%
-One set of useful expressions
-%
-\begin{align}
-\gamma_{\Delta} &= NOISY\_GAMMA - SHADOW\_GAMMA \nonumber \\
-\gamma_L &= GAMMA + SHADOW\_GAMMA \nonumber \\
-\gamma &= GAMMA + NOISY\_GAMMA \nonumber
-\end{align}
-%
-%and another one
-%%
-%\begin{align}
-%GAMMA &=  \nonumber \\
-%SHADOW\_GAMMA &= \nonumber \\
-%NOISY\_GAMMA &= \nonumber
-%\end{align}
-%%
+To test the accuracy of ALMO DFT based MD, we computed dynamical properties for 40 ns MD trajectories using fully converged Kohn-Sham DFT forces and forces computed with our ALMO DFT methods (see fig.
+%\ref{fig:dynproperties}).
 
-Practical guide:
-%
-\begin{itemize}
-\item The code designed so that any effect can be achieved with positive GAMMAs.
-\item Situation 1.
-NVT with perfect forces.
-Set NOISY and SHADOW to zeros to produce balanced canonical distributions.
-Large GAMMA adds lots of noise but leads to fast equilibration.
-Small GAMMA -- approach NVE.
-\item Situation 2.
-NVT with imperfect forces.
-Use either NOISY or SHADOW but not both.
-The value of GAMMA is chosen as in the previous situation.
-A.
-Set SHADOW to finite value to get excessive damping.
-B.
-Set NOISY to get excessive stochastic boost.
-\end{itemize}
-
-Integrator implemented in \$CP2K/src/motion/integrator.F eqns (19) and (20) from http://dx.doi.org/10.1080/0026897031000108113.
-Another integrator can be implemented using Eq (15.5.17) from the book of Tuckerman.
+\section{Speed}
 

### 2017-08-10 almo-langevin.tex
August 10
Delta Force Data Completed

@@
 
-\documentclass[10pt,aps,prb,twocolumn,amsmath,amssymb,superscriptaddress,longbibliography]{revtex4-1}
+\documentclass[10pt,aps,prl,twocolumn,amsmath,amssymb,superscriptaddress,longbibliography]{revtex4-1}
 
@@
 % Re-write the following section and include any recent advances
-A number of such alternative methods explore the natural sparsity of the one-electron density matrix (DM)~\cite{a:ls-kohn-1996,a:ls-rev-1999} and are capable of yielding linear scaling (LS) for large systems.~\cite{a:linscale3,a:lee-yang-1996,a:ls-scuseria-1997,a:ls-manolopoulos-1998,a:ls-helgaker-2001,a:ls-niklasson-2003,a:curvy2,a:ls-dm-sign}
-Unfortunately, the variational optimization of the DM is very inefficient for accurate DFT calculations~\cite{a:ls-rev-1999,a:ls-dm-sign}, which require many basis functions per atom. 
+A number of such alternative methods explore the natural sparsity of the one-electron density matrix (DM)~\cite{a:ls-kohn-1996,a:ls-rev-1999} and are capable of yielding LS for large systems.~\cite{a:linscale3,a:lee-yang-1996,a:ls-scuseria-1997,a:ls-manolopoulos-1998,a:ls-helgaker-2001,a:ls-niklasson-2003,a:curvy2,a:ls-dm-sign}
+Unfortunately, the variational optimization of the DM is very inefficient for accurate DFT calculations,~\cite{a:ls-rev-1999,a:ls-dm-sign} which require many basis functions per atom. 
 Therefore, the applications of DM-based LS methods have been limited to minimal-basis tight-binding problems. 
@@
 This is done by balancing a new computationally economical but systemically inaccurate force calculation method with an optimized version of the Langevin Equation of motion. 
-We also introduce an easy to follow algorithmic approach to optimize a system for generating correct dynamical behaviour.
+We also introduce an easy to follow heuristic approach to optimize a system for generating correct dynamical behaviour.
 
@@
 \begin{figure}
-\includegraphics[trim={2.2cm 0.5cm 3cm 0.2cm},clip,width=8cm]{125mol_ALMO_Early_Random_force.eps}
-\caption{\label{fig:randomforce}The average ``Random force'' along a water molecule dimension $\langle R^{\Delta}_{i\alpha} \rangle (t)$ calculated as the difference between the fully converged Born-Oppenheimer forces and unconverged ALMO DFT forces, where $\epsilon_{err} = \num{1e-2}$.
-(a) Plot of $\langle R^{\Delta}_{i\alpha} \rangle (t)$ as a function of time. 
+\includegraphics[trim={0.5cm 0.7cm 0.5cm 0cm},clip,width=8cm]{DeltaForceComparison_with_ACF.eps}
+\caption{\label{fig:randomforce}The average ``Random force'' along a water molecule dimension $\langle R^{\Delta}_{i\alpha} \rangle (t)$ calculated as the difference between the fully converged Born-Oppenheimer forces and ALMO DFT forces.
 Here it is clear that Eq.\ (\ref{eq:stochastic2}) is satisfied.
-(b) Autocovariance function of $\langle R^{\Delta}_{i\alpha} \rangle (t)$. 
-The ACF is clearly not a delta-function, and is characterized by rapid oscillation between positive and negative correlation. 
+(a) Plot of $\langle R^{\Delta}_{i\alpha} \rangle (t)$ as a function of time, where $R_{C} = 1.6$ VdW Radii and $\epsilon_{err} = \num{1e-2}$.
+(b) Plot of $\langle R^{\Delta}_{i\alpha} \rangle (t)$ as a function of time, where $R_{C} = 1.6$ VdW Radii and forces are fully converged.
+(c) Autocorrelation function of plot a, $\langle R^{\Delta}_{i\alpha} (t) R^{\Delta}_{i\alpha} (t+\tau) \rangle $.
+Using Eq.\ \ref{eq:delta}, $\Delta = \SI{2e-5}{\per\fs}$, but its true value is $\SI{8e-5}{\per\fs}$ for this system.
+Note that the  ACF is not truly a delta-function, and is characterized by rapidly oscillating correlations for short lag times. 
 However, the autocorrelation decays rapidly enough that these forces can be considered random and uncorrelated to a good approximation for sufficiently long time scales. 
-Calculations were performed with the PBE $\mathrm{E_{XC}}$ functional and the TZV2P basis set on 125 water molecules.}
+Calculations were performed with the PBE $\mathrm{E_{XC}}$ functional and the TZV2P basis set on 125 water molecules for 10 ps.}
 \end{figure}
@@
 
-To test the accuracy of ALMO DFT based MD, we computed dynamical properties for 40 ns MD trajectories using fully converged Kohn-Sham DFT forces and forces computed with our ALMO DFT methods (see fig.
+In order to confirm that ALMO DFT forces are accurate in the limit of $R_{c} = \infty$, we computed the average difference in force between fully converged Born-Oppenheimer forces and forces calculated from ALMO DFT with varying $R_{c}$.
+Figure \ref{fig:forcecomp} shows that as $R_{c}$ is increased to infinity, forces converge toward the Born-Oppenheimer limit.
+
+\begin{figure}
+\includegraphics[trim={0.1cm 0cm 0cm 0cm},clip,width=8cm]{DeltaForceComparison_ALMO_SCF.eps}
+\caption{\label{fig:forcecomp}The average ``Random force'' along a water molecule dimension $\langle R^{\Delta}_{i\alpha} \rangle (t)$ calculated as the average difference between fully converged Born-Oppenheimer forces and fully converged ALMO DFT forces with varying $R_{c}$.
+Here it is clear that ALMO DFT forces converge to Born-Oppenheimer forces in the limit of $R_{c} = \infty$.
+Calculations were performed with the PBE $\mathrm{E_{XC}}$ functional and the TZV2P basis set on 125 water molecules for 10 ps.}
+\end{figure}
+
+To test the accuracy of ALMO DFT based MD, we computed dynamical properties for 40 ns MD trajectories using fully converged Kohn-Sham DFT forces and compared with equivalent trajectories using forces computed with ALMO DFT methods (see fig.
 %\ref{fig:dynproperties}).

### 2017-08-10 almo-langevin.tex
August 8v3
Added Strong Scaling Data.

@@
 Here it is clear that Eq.\ (\ref{eq:stochastic2}) is satisfied.
-(a) Plot of $\langle R^{\Delta}_{i\alpha} \rangle (t)$ as a function of time, where $R_{C} = 1.6$ VdW Radii and $\epsilon_{err} = \num{1e-2}$.
-(b) Plot of $\langle R^{\Delta}_{i\alpha} \rangle (t)$ as a function of time, where $R_{C} = 1.6$ VdW Radii and forces are fully converged.
+(a) Plot of $\langle R^{\Delta}_{i\alpha} \rangle (t)$ as a function of time, where $R_{C} = 1.6$ VdWR and $\epsilon_{err} = \num{1e-2}$.
+(b) Plot of $\langle R^{\Delta}_{i\alpha} \rangle (t)$ as a function of time, where $R_{C} = 1.6$ VdWR and forces are fully converged.
 (c) Autocorrelation function of plot a, $\langle R^{\Delta}_{i\alpha} (t) R^{\Delta}_{i\alpha} (t+\tau) \rangle $.
@@
 
-In order to confirm that ALMO DFT forces are accurate in the limit of $R_{c} = \infty$, we computed the average difference in force between fully converged Born-Oppenheimer forces and forces calculated from ALMO DFT with varying $R_{c}$.
+In order to confirm that ALMO DFT forces are accurate in the limit where localization radius $R_{c} = \infty$, we computed the average difference in force between fully converged Born-Oppenheimer forces and forces calculated from ALMO DFT with varying $R_{c}$.
 Figure \ref{fig:forcecomp} shows that as $R_{c}$ is increased to infinity, forces converge toward the Born-Oppenheimer limit.
@@
 \includegraphics[trim={0.1cm 0cm 0cm 0cm},clip,width=8cm]{DeltaForceComparison_ALMO_SCF.eps}
-\caption{\label{fig:forcecomp}The average ``Random force'' along a water molecule dimension $\langle R^{\Delta}_{i\alpha} \rangle (t)$ calculated as the average difference between fully converged Born-Oppenheimer forces and fully converged ALMO DFT forces with varying $R_{c}$.
+\caption{\label{fig:forcecomp}The average ``Random force'' $\langle R^{\Delta}_{i\alpha} \rangle (t)$ calculated as the average difference between fully converged Born-Oppenheimer forces and fully converged ALMO DFT forces, averaged across all dimensions and all molecules separately at each time step.
+Plotted are the random forces with varying localization radius $R_{c}$, in units of Van der Waals Radii.
 Here it is clear that ALMO DFT forces converge to Born-Oppenheimer forces in the limit of $R_{c} = \infty$.
@@
 
-To test the accuracy of ALMO DFT based MD, we computed dynamical properties for 40 ns MD trajectories using fully converged Kohn-Sham DFT forces and compared with equivalent trajectories using forces computed with ALMO DFT methods (see fig.
+To test the accuracy of ALMO DFT based MD, we computed dynamical properties for 30 ns MD trajectories using fully converged Kohn-Sham DFT forces and compared with equivalent trajectories using forces computed with ALMO DFT methods (see fig.
 %\ref{fig:dynproperties}).
@@
 \section{Speed}
+
+To show the computational efficiency gained by using unconverged ALMO DFT forces, we compare the average wall-time per MD step over 10 time steps for various values of $\epsilon_{err}$ in figures \ref{fig:strongscaling_linear} and \ref{fig:strongscaling_log}.
+For these simulations, we chose to set the localization radius to 1.6 Van der Waal Radii (VdWR), as was done for calculation of dynamical properties.
+We also show timing of the fully converged OT SCF method, as well as the recently developed ``second generation Car-Parinello'' SCF method, which improves upon the OT computational cost prefactor.~\cite{a:2ndcpmd}
+All plots in figures \ref{fig:strongscaling_linear} and \ref{fig:strongscaling_log} were generating using 256 CPU cores. 
+ALMO methods show excellent LS behavior for all values of $\epsilon_{err}$ tested.
+OT SCF scales approximately cubically, as expected.
+Our findings show that the fastest ALMO SCF method beats out second-generation Car-Parinello using OT SCF at 256 molecules.
+
+\begin{figure}
+\includegraphics[trim={1.9cm 0.4cm 3.5cm 1.7cm},clip,width=8cm]{strongscaling_linear.eps}
+\caption{\label{fig:strongscaling_linear} Linear plots showing LS behaviour of all ALMO methods for liquid water at 298 K. 
+Calculations were performed with the PBE $\mathrm{E_{XC}}$ functional and TZV2P basis set on 256 CPUs. 
+The localization radius was set to $R_{c} = 1.6$ VdWR.}
+\end{figure}
+
+\begin{figure}
+\includegraphics[trim={2.8cm 0.4cm 3.5cm 1.2cm},clip,width=8cm]{strongscaling_log.eps}
+\caption{\label{fig:strongscaling_log} Log-log plots which clearly show the asymptotic LS behaviour of ALMO methods for liqid water at 298 K. 
+Cyan lines represent perfect cubic scaling, whereas gray lines represent perfect linear scaling. 
+Calculations were performed with the PBE $\mathrm{E_{XC}}$ functional and TZV2P basis set on 256 CPUs. 
+The localization radius was set to $R_{c} = 1.6$ VdWR.}
+\end{figure}
 

### 2017-08-11 almo-langevin.tex
August 10th v3
Updated some descriptions.

@@
 (c) Autocorrelation function of plot a, $\langle R^{\Delta}_{i\alpha} (t) R^{\Delta}_{i\alpha} (t+\tau) \rangle $.
-Using Eq.\ \ref{eq:delta}, $\Delta = \SI{2e-5}{\per\fs}$, but its true value is $\SI{8e-5}{\per\fs}$ for this system.
+Using Eq.\ \ref{eq:delta}, we calculated $\Delta \approx \SI{2e-5}{\per\fs}$, but its true value is $\SI{8e-5}{\per\fs}$ for this system.
 Note that the  ACF is not truly a delta-function, and is characterized by rapidly oscillating correlations for short lag times. 
-However, the autocorrelation decays rapidly enough that these forces can be considered random and uncorrelated to a good approximation for sufficiently long time scales. 
+The high frequency correlations made integration of the ACF inaccurate, but provided a decent starting point for further investigation.
+The autocorrelation decays rapidly enough that these forces can be considered random and uncorrelated to a good approximation for sufficiently long time scales ($\gg\SI{300}{\fs}$ for this system). 
 Calculations were performed with the PBE $\mathrm{E_{XC}}$ functional and the TZV2P basis set on 125 water molecules for 10 ps.}

### 2017-08-11 almo-langevin.tex
August 10 v4

@@
 Plotted are the random forces with varying localization radius $R_{c}$, in units of Van der Waals Radii.
-Here it is clear that ALMO DFT forces converge to Born-Oppenheimer forces in the limit of $R_{c} = \infty$.
+Here it is clear that ALMO DFT forces converge to Born-Oppenheimer forces in the limit of $R_{c} \rightarrow \infty$.
 Calculations were performed with the PBE $\mathrm{E_{XC}}$ functional and the TZV2P basis set on 125 water molecules for 10 ps.}
@@
 We also show timing of the fully converged OT SCF method, as well as the recently developed ``second generation Car-Parinello'' SCF method, which improves upon the OT computational cost prefactor.~\cite{a:2ndcpmd}
-All plots in figures \ref{fig:strongscaling_linear} and \ref{fig:strongscaling_log} were generating using 256 CPU cores. 
+All plots in figures \ref{fig:strongscaling_linear} and \ref{fig:strongscaling_log} were generated using 256 CPU cores. 
 ALMO methods show excellent LS behavior for all values of $\epsilon_{err}$ tested.
 OT SCF scales approximately cubically, as expected.
-Our findings show that the fastest ALMO SCF method beats out second-generation Car-Parinello using OT SCF at 256 molecules.
+Our findings show that the fastest ALMO SCF method presented here runs faster than second-generation Car-Parinello using OT SCF for water simulations larger than 256 molecules.
 

### 2017-08-15 almo-langevin.tex
August 15
Updated all figures with matching font type and fontsize. Added Weakscaling plot. Modified plots so they will still be readable without colour. Added more text to SPEED subsection.

@@
 Generally, this step utilizes a iterative search function in order to reduce the absolute norm of the energy gradient to some max allowable value $\epsilon_{err}$. 
-Previously, $\epsilon_{err}$ had to be set very low to ensure sufficient accuracy of the electron structure calculations. 
+Previously, $\epsilon_{err}$ had to be set very low to ensure convergence of the electron structure calculations. 
 By increasing the allowable $\epsilon_{err}$, a great reduction in computational cost can be achieved. 
@@
 \begin{figure}
-\includegraphics[trim={0.5cm 0.7cm 0.5cm 0cm},clip,width=8cm]{DeltaForceComparison_with_ACF.eps}
-\caption{\label{fig:randomforce}The average ``Random force'' along a water molecule dimension $\langle R^{\Delta}_{i\alpha} \rangle (t)$ calculated as the difference between the fully converged Born-Oppenheimer forces and ALMO DFT forces.
+\includegraphics[trim={0.6cm 0.7cm 0.7cm 0.3cm},clip,width=8.6cm]{DeltaForceComparison_with_ACF.eps}
+\caption{\label{fig:randomforce}The ``Random force'' along a water molecule dimension $\langle R^{\Delta}_{i\alpha} \rangle (t)$ calculated as the difference between the fully converged Born-Oppenheimer forces and ALMO DFT forces.
 Here it is clear that Eq.\ (\ref{eq:stochastic2}) is satisfied.
-(a) Plot of $\langle R^{\Delta}_{i\alpha} \rangle (t)$ as a function of time, where $R_{C} = 1.6$ VdWR and $\epsilon_{err} = \num{1e-2}$.
+(a) Plot of $\langle R^{\Delta}_{i\alpha} \rangle (t)$ as a function of time, where $R_{C} = 1.6$ VdWR and $\epsilon_{err} = 10^{-2}$.
 (b) Plot of $\langle R^{\Delta}_{i\alpha} \rangle (t)$ as a function of time, where $R_{C} = 1.6$ VdWR and forces are fully converged.
@@
 \begin{figure}
-\includegraphics[trim={0.1cm 0cm 0cm 0cm},clip,width=8cm]{DeltaForceComparison_ALMO_SCF.eps}
-\caption{\label{fig:forcecomp}The average ``Random force'' $\langle R^{\Delta}_{i\alpha} \rangle (t)$ calculated as the average difference between fully converged Born-Oppenheimer forces and fully converged ALMO DFT forces, averaged across all dimensions and all molecules separately at each time step.
+\includegraphics[trim={0.1cm 0cm 0.2cm 0.1cm},clip,width=8.6cm]{DeltaForceComparison_ALMO_SCF.eps}
+\caption{\label{fig:forcecomp}The ``Random force'' $\langle R^{\Delta}_{i\alpha} \rangle (t)$ calculated as the average difference between fully converged Born-Oppenheimer forces and fully converged ALMO DFT forces, averaged across all dimensions and all molecules separately at each time step.
 Plotted are the random forces with varying localization radius $R_{c}$, in units of Van der Waals Radii.
@@
 
-To show the computational efficiency gained by using unconverged ALMO DFT forces, we compare the average wall-time per MD step over 10 time steps for various values of $\epsilon_{err}$ in figures \ref{fig:strongscaling_linear} and \ref{fig:strongscaling_log}.
+To show the computational efficiency gained by using unconverged ALMO DFT forces, we compare the average wall-time per MD step over 10 time steps for various values of $\epsilon_{err}$ in figure \ref{fig:strongscaling_log}.
 For these simulations, we chose to set the localization radius to 1.6 Van der Waal Radii (VdWR), as was done for calculation of dynamical properties.
-We also show timing of the fully converged OT SCF method, as well as the recently developed ``second generation Car-Parinello'' SCF method, which improves upon the OT computational cost prefactor.~\cite{a:2ndcpmd}
-All plots in figures \ref{fig:strongscaling_linear} and \ref{fig:strongscaling_log} were generated using 256 CPU cores. 
+We also show timing of the fully converged OT SCF method, as well as the recently developed ``second generation Car-Parinello'' SCF method using OT, which improves upon the OT computational cost prefactor.~\cite{a:2ndcpmd}
 ALMO methods show excellent LS behavior for all values of $\epsilon_{err}$ tested.
@@
 
+Figure \ref{fig:weakscaling} shows that ALMO MD code is well optimized for parallelization, and that computational cost grows linearly under a wide range of available computing power. 
+In fact, we were able to successfully simulate systems as large as 32768 water molecules within 25 minutes of wall-clock time using 4096 cores.
+
+%\begin{figure}
+%\includegraphics[trim={1.7cm 0.5cm 3.3cm 0.1cm},clip,width=8.6cm]{strongscaling_linear.eps}
+%\caption{\label{fig:strongscaling_linear} Linear plots showing LS behaviour of all ALMO methods for liquid water at 298 K. 
+%Calculations were performed with the PBE $\mathrm{E_{XC}}$ functional and TZV2P basis set on 256 CPUs. 
+%The localization radius was set to $R_{c} = 1.6$ VdWR.}
+%\end{figure}
+
 \begin{figure}
-\includegraphics[trim={1.9cm 0.4cm 3.5cm 1.7cm},clip,width=8cm]{strongscaling_linear.eps}
-\caption{\label{fig:strongscaling_linear} Linear plots showing LS behaviour of all ALMO methods for liquid water at 298 K. 
-Calculations were performed with the PBE $\mathrm{E_{XC}}$ functional and TZV2P basis set on 256 CPUs. 
+\includegraphics[trim={2.5cm 0.5cm 3.4cm 0.1cm},clip,width=8.6cm]{strongscaling_log.eps}
+\caption{\label{fig:strongscaling_log} Logscale plots which show the asymptotic LS behaviour of ALMO methods for liquid water at 298 K.
+Compared with the orbital transformation using fully converged SCF and second generation Car-Parinello force calculation methods.
+Cyan lines represent perfect cubic scaling, whereas gray lines represent perfect linear scaling. 
+Calculations were performed with the PBE $\mathrm{E_{XC}}$ functional and TZV2P basis set on 256 CPUs, averaged over 10 MD steps. 
 The localization radius was set to $R_{c} = 1.6$ VdWR.}
@@
 \begin{figure}
-\includegraphics[trim={2.8cm 0.4cm 3.5cm 1.2cm},clip,width=8cm]{strongscaling_log.eps}
-\caption{\label{fig:strongscaling_log} Log-log plots which clearly show the asymptotic LS behaviour of ALMO methods for liqid water at 298 K. 
-Cyan lines represent perfect cubic scaling, whereas gray lines represent perfect linear scaling. 
-Calculations were performed with the PBE $\mathrm{E_{XC}}$ functional and TZV2P basis set on 256 CPUs. 
-The localization radius was set to $R_{c} = 1.6$ VdWR.}
+\includegraphics[trim={1.6cm 0cm 4.7cm 0cm},clip,width=8.6cm]{weakscaling.eps}
+\caption{\label{fig:weakscaling} Weak scalability benchmarks which show optimization of code for parallel computing.
+Dashed gray lines connect systems simulated using the same number of CPU cores, showing LS behaviour over a range of configurations.
+Of note is the 32768 molecule system successfully simulated using 4096 cores at around 25 minutes per MD step.
+We also show that systems as large as 8192 molecules can be simulated for 10,000 MD steps in under a month using our method.
+Calculations were performed with the PBE $\mathrm{E_{XC}}$ functional and TZV2P basis set, averaged over 10 MD steps. 
+The localization radius was set to $R_{c} = 1.6$ VdWR and $\epsilon_{err} = 10^{-2}$.}
 \end{figure}

### 2017-08-16 almo-langevin.tex
August 16
New plots, Added text, updated incorrect info!

@@
 During the second step of the ALMO SCF method described above, a self-consistent variational approach is taken to converge the non-orthogonal MOs. 
-Generally, this step utilizes a iterative search function in order to reduce the absolute norm of the energy gradient to some max allowable value $\epsilon_{err}$. 
-Previously, $\epsilon_{err}$ had to be set very low to ensure convergence of the electron structure calculations. 
-By increasing the allowable $\epsilon_{err}$, a great reduction in computational cost can be achieved. 
+Generally, this step utilizes a iterative search function in order to reduce the absolute norm of the energy gradient to some max allowable value $\epsilon_{SCF}$. 
+Previously, $\epsilon_{SCF}$ had to be set very low to ensure convergence of the electron structure calculations. 
+By increasing the allowable $\epsilon_{SCF}$, a great reduction in computational cost can be achieved. 
 Obviously, this introduces a significant error into the calculation of forces during each time step. 
@@
 %
-Where $f^{\text{SCF}}_{i\alpha}$ is the fully converged ALMO DFT force calculation, $f^{\text{APP}}_{i\alpha}$ is the approximate force calculated by allowing a large $\epsilon_{err}$, and $R^{\Delta}_{i\alpha} (t)$ is a stochastic ``white noise'' term that obeys
+Where $f^{\text{SCF}}_{i\alpha}$ is the fully converged ALMO DFT force calculation, $f^{\text{APP}}_{i\alpha}$ is the approximate force calculated by allowing a large $\epsilon_{SCF}$, and $R^{\Delta}_{i\alpha} (t)$ is a stochastic ``white noise'' term that obeys
 %
@@
 Here it is clear that Eq.\ (\ref{eq:stochastic2}) is satisfied.
-(a) Plot of $\langle R^{\Delta}_{i\alpha} \rangle (t)$ as a function of time, where $R_{C} = 1.6$ VdWR and $\epsilon_{err} = 10^{-2}$.
+(a) Plot of $\langle R^{\Delta}_{i\alpha} \rangle (t)$ as a function of time, where $R_{C} = 1.6$ VdWR and $\epsilon_{SCF} = 10^{-2}$.
 (b) Plot of $\langle R^{\Delta}_{i\alpha} \rangle (t)$ as a function of time, where $R_{C} = 1.6$ VdWR and forces are fully converged.
 (c) Autocorrelation function of plot a, $\langle R^{\Delta}_{i\alpha} (t) R^{\Delta}_{i\alpha} (t+\tau) \rangle $.
-Using Eq.\ \ref{eq:delta}, we calculated $\Delta \approx \SI{2e-5}{\per\fs}$, but its true value is $\SI{8e-5}{\per\fs}$ for this system.
+Using Eq.\ \ref{eq:delta}, we calculated $\Delta \approx \SI{2e-5}{\per\fs}$, but its true value is $\SI{6e-5}{\per\fs}$ for this system.
 Note that the  ACF is not truly a delta-function, and is characterized by rapidly oscillating correlations for short lag times. 
@@
 The computation scheme used is mostly identical to that described in the implementation section of ref.\ \citenum{a:almo-ls}. 
-For the dynamical properties calculations using ALMO DFT calculation methods, a periodic cell containing 125 water molecules was simulated for $\SI{40}{\ns}$ at a constant temperature of $\SI{298}{\K}$ and a constant density of $\SI{0.9966}{\g\per\cm^{3}}$.
-The system was thermostated using the Langevin ensemble (eq.\ \ref{eq:langevin2}) where $\gamma_L$ was set to $\SI{e-5}{\per\fs}$, and $\Delta$ was optimized to the order of $10^{-5}$ - $10^{-4}\, \mathrm{fs^{-1}}$.
+For the dynamical properties calculations using ALMO DFT calculation methods, a periodic cell containing 125 water molecules was simulated for $\SI{10}{\ps}$ at a temperature of $\SI{293}{\K}$ and a constant density of $\SI{1.00597}{\g\per\cm^{3}}$.
+The system was thermostated using the Langevin ensemble (eq.\ \ref{eq:langevin2}) where $\gamma_L$ was set to $\SI{e-3}{\per\fs}$, and $\Delta$ was optimized to the order of $10^{-5}\, \mathrm{fs^{-1}}$.
 Note that for timing benchmarks, $\Delta$ was not optimized. 
@@
 
-The reference PES was calculated using the orbital transformation (OT) method~\cite{a:ot,a:ot2} and the canonical sampling through velocity rescaling (CSVR) thermostat~\cite{a:CSVR} with a time constant of $\SI{200}{fs}$.
-Otherwise, system parameters were identical for the reference.
+The reference PES was calculated using the orbital transformation (OT) method~\cite{a:ot,a:ot2} and the Langevin thermostat with $\gamma_L = \SI{e-3}{\per\fs}$. 
+All other system parameters were set identical to the ALMO system.
 
@@
 
-To test the accuracy of ALMO DFT based MD, we computed dynamical properties for 30 ns MD trajectories using fully converged Kohn-Sham DFT forces and compared with equivalent trajectories using forces computed with ALMO DFT methods (see fig.
-%\ref{fig:dynproperties}).
+To test the accuracy of ALMO DFT based MD we computed the Infrared (IR) spectrum, radial distribution function (RDF), kinetic energy distribution, and mean-squared displacement (MSD) function from a 10 ps MD trajectory.
+We compared the IR spectrum, RDF, and MSD to equivalent functions calculated from fully converged Kohn-Sham DFT forces (see fig. \ref{fig:dynproperties} and supporting information).
+From the MSD functions we also calculated diffusion constants for water at approximately 300 K.
+With fully converged forces we calculated D = $\SI{5.47e-9}{\m^{2}\per\s}$.
+With ALMO unconverged forces we calculated D = $\SI{5.55E-09}{\m^{2}\per\s}$.
+
+\begin{figure}
+\includegraphics[trim={1.3cm 0.1cm 3.3cm 1.3cm},clip,width=8.6cm]{Dynamical_Data_Tiled.eps}
+\caption{\label{fig:dynproperties} Calculated dynamic and static properties of water using ALMO DFT with $\epsilon_{SCF} = 10^{-2}$ and $R_{c} = 1.6$ VdWR.
+(a) Radial distribution function of an ALMO simulation compared to OT SCF with fully converged PES. 
+(b) Kinetic Energy distribution of an ALMO simulation compared to the theoretical Maxwell-Boltzmann kinetic energy distribution.
+(c) Calculated Infrared spectrum of an ALMO simulation compared to OT SCF with fully converged PES.
+Calculations were performed with the PBE $\mathrm{E_{XC}}$ functional and TZV2P basis set, using 125 water molecules for 10 ps.}
+\end{figure}
 
@@
 
-To show the computational efficiency gained by using unconverged ALMO DFT forces, we compare the average wall-time per MD step over 10 time steps for various values of $\epsilon_{err}$ in figure \ref{fig:strongscaling_log}.
+To show the computational efficiency gained by using unconverged ALMO DFT forces, we compare the average wall-time per MD step over 10 time steps for various values of $\epsilon_{SCF}$ in figure \ref{fig:strongscaling_log}.
 For these simulations, we chose to set the localization radius to 1.6 Van der Waal Radii (VdWR), as was done for calculation of dynamical properties.
 We also show timing of the fully converged OT SCF method, as well as the recently developed ``second generation Car-Parinello'' SCF method using OT, which improves upon the OT computational cost prefactor.~\cite{a:2ndcpmd}
-ALMO methods show excellent LS behavior for all values of $\epsilon_{err}$ tested.
+ALMO methods show excellent LS behavior for all values of $\epsilon_{SCF}$ tested.
 OT SCF scales approximately cubically, as expected.
@@
 Calculations were performed with the PBE $\mathrm{E_{XC}}$ functional and TZV2P basis set, averaged over 10 MD steps. 
-The localization radius was set to $R_{c} = 1.6$ VdWR and $\epsilon_{err} = 10^{-2}$.}
+The localization radius was set to $R_{c} = 1.6$ VdWR and $\epsilon_{SCF} = 10^{-2}$.}
 \end{figure}

### 2017-08-17 almo-langevin.tex
Intro is being re-written

--- 2017-08-11 almo-langevin.tex
@@
 
-\documentclass[10pt,aps,prl,twocolumn,amsmath,amssymb,superscriptaddress,longbibliography]{revtex4-1}
+\documentclass[aps,prl,reprint,amsmath,amssymb]{revtex4-1}
 
+%\documentclass[aps,prl,,reprint,amsmath,amssymb,superscriptaddress,longbibliography]{revtex4-1}
 %\documentclass[preprint,showpacs,preprintnumbers,amsmath,amssymb]{revtex4-1}
@@
 \usepackage{color}
-\usepackage{graphicx}
+\usepackage{epstopdf} 
+%\graphicspath{{pics}}
 %\usepackage[normalem]{ulem}
@@
 Low-cost linear-scaling \emph{ab initio} molecular dynamics for weakly-interacting systems
+%%for weakly-coupled atoms
+%% for non-covalently bonded atoms
 }
@@
 \maketitle
+
 \section{Introduction}
-Since the unification of molecular dynamics (MD) and density functional theory (DFT) in 1985 by Car and Parrinello,~\cite{a:thecpmd} much research has been focused on reducing the computational overhead of \emph{ab initio} molecular dynamics (AIMD) potential energy calculations.~\cite{b:aimd}
-% Remove Following text for some journals
-AIMD provides a great advantage over traditional predefined-potential molecular dynamics in that it calculates its potential energy surface (PES) on-the-fly, rather than needing to be laboriously optimized to the specific system of interest beforehand.
-This allows for exploration of much more chemically complex systems than are possible with traditional MD methods. 
-AIMD also allows for much greater predictive capability than traditional MD, as a first principles approach is not confined to a predetermined range of system parameters. 
-The tradeoff is the computational complexity involved in calculating an accurate set of forces for each atom at every time step of the simulation.~\cite{a:2gen-cpmd}
-% End remove
 
-The computational cost of the straightforward approach to Kohn-Sham DFT is known to grow (or \emph{scale}) cubically with system size because it involves either diagonalization of the effective Kohn-Sham Hamiltonian or, if the direct optimization of Kohn-Sham orbitals is performed, the equally costly inversion of a preconditioner.~\cite{a:ot} % Add other references here?
-This cubic scaling severely constricts the possible length and time scales accessible by AIMD using such an approach. 
-In an attempt to address this issue, there has been much research into the possibility of linear scaling (LS) DFT methods for the calculation of forces.
+Since the unification of molecular dynamics and density functional theory (DFT)~\cite{a:thecpmd},
+\emph{ab initio} molecular dynamics (AIMD)~\cite{b:aimd} has become an important tool for computational studies of weakly bonded ensembles of molecules such as gas-phase clusters, molecular and ionic liquids, solutions, and molecular solids.
+[of gas-phase molecules and condensed-phase systems.]
+Unfortunately, the computational cost of AIMD based on the conventional Kohn-Sham (KS) DFT grows cubically with the number of atoms in the system and severely limits the \emph{length scale} accessible by AIMD.
+In an attempt to address this issue, substantial efforts have been directed to the development of linear scaling (LS) DFT methods for the on-the-fly calculation of forces.
 
@@
 % Re-write the following section and include any recent advances
-A number of such alternative methods explore the natural sparsity of the one-electron density matrix (DM)~\cite{a:ls-kohn-1996,a:ls-rev-1999} and are capable of yielding LS for large systems.~\cite{a:linscale3,a:lee-yang-1996,a:ls-scuseria-1997,a:ls-manolopoulos-1998,a:ls-helgaker-2001,a:ls-niklasson-2003,a:curvy2,a:ls-dm-sign}
+The steep scaling of the conventional KS DFT originates from the delocalized nature of the canonical KS orbitals, which tend to delocalize over all atoms in the system, with an alternative set of electronic descriptors.
+All LS methods must A number of such methods explore the natural sparsity of the one-electron density matrix (DM)~\cite{a:ls-kohn-1996,a:ls-rev-1999} and are capable of yielding LS for large systems.~\cite{a:linscale3,a:lee-yang-1996,a:ls-scuseria-1997,a:ls-manolopoulos-1998,a:ls-helgaker-2001,a:ls-niklasson-2003,a:curvy2,a:ls-dm-sign}
 Unfortunately, the variational optimization of the DM is very inefficient for accurate DFT calculations,~\cite{a:ls-rev-1999,a:ls-dm-sign} which require many basis functions per atom. 
@@
 
-Here, we introduce a new low-cost linear scaling AIMD method for weakly-interacting molecular systems based on the recently developed absolutely localized molecular orbitals (ALMO) DFT approach.~\cite{a:almo-ls}
-
-This method exploits non-orthogonal molecular orbitals (MOs) in a two-step self-consistent filed (SCF) approach that is designed to avoid the usual convergence issues.~\cite{a:ls-mauri-galli-car-1993,a:ls-ordejon-1995,a:ls-fattebert-2004,a:ls-rev-1999}
+Here, we introduce a new low-cost linear scaling AIMD method for weakly-interacting molecular systems based on the recently developed absolutely localized molecular orbitals (ALMO) DFT approach~\cite{a:almo-ls}.
+ALMO DFT exploits non-orthogonal molecular orbitals (MOs) in a two-step self-consistent filed (SCF) approach that is designed to avoid the usual convergence issues~\cite{a:ls-mauri-galli-car-1993,a:ls-ordejon-1995,a:ls-fattebert-2004,a:ls-rev-1999} and .
 We show that due to its linear scaling nature and relatively low computational prefactor, ALMO-based DFT calculations allow for simulation of very large weakly-coupled molecular systems. 
@@
 \begin{figure}
-\includegraphics[trim={0.5cm 0.7cm 0.5cm 0cm},clip,width=8cm]{DeltaForceComparison_with_ACF.eps}
+\includegraphics[trim={0.5cm 0.7cm 0.5cm 0cm},clip,width=8cm]{DeltaForceComparison_with_ACF}
 \caption{\label{fig:randomforce}The average ``Random force'' along a water molecule dimension $\langle R^{\Delta}_{i\alpha} \rangle (t)$ calculated as the difference between the fully converged Born-Oppenheimer forces and ALMO DFT forces.
@@
 \begin{figure}
-\includegraphics[trim={0.1cm 0cm 0cm 0cm},clip,width=8cm]{DeltaForceComparison_ALMO_SCF.eps}
+\includegraphics[trim={0.1cm 0cm 0cm 0cm},clip,width=8cm]{DeltaForceComparison_ALMO_SCF}
 \caption{\label{fig:forcecomp}The average ``Random force'' $\langle R^{\Delta}_{i\alpha} \rangle (t)$ calculated as the average difference between fully converged Born-Oppenheimer forces and fully converged ALMO DFT forces, averaged across all dimensions and all molecules separately at each time step.
@@
 \begin{figure}
-\includegraphics[trim={1.9cm 0.4cm 3.5cm 1.7cm},clip,width=8cm]{strongscaling_linear.eps}
+\includegraphics[trim={1.9cm 0.4cm 3.5cm 1.7cm},clip,width=8cm]{strongscaling_linear}
 \caption{\label{fig:strongscaling_linear} Linear plots showing LS behaviour of all ALMO methods for liquid water at 298 K. 
@@
 \begin{figure}
-\includegraphics[trim={2.8cm 0.4cm 3.5cm 1.2cm},clip,width=8cm]{strongscaling_log.eps}
+\includegraphics[trim={2.8cm 0.4cm 3.5cm 1.2cm},clip,width=8cm]{strongscaling_log}
 \caption{\label{fig:strongscaling_log} Log-log plots which clearly show the asymptotic LS behaviour of ALMO methods for liqid water at 298 K. 
@@
 \section{Conclusions}
+
+%Such simulations are important for Use if empirical force fields or QM/MM are not applicable:
+%Chemical reactions that happen throughout your system, not in a localized spatial region (e.g. proton transfer in water)
+%Unusual thermodynamic states for which simple analytical potential become inaccurate (e.g. high pressure states, interfaces, heterojunctions)
+%Electron density is an important observable (e.g. X-ray scattering) or information of the electronic wavefunctions is crucial.
 

### 2017-08-17 almo-langevin.tex

--- 2017-08-16 almo-langevin.tex
@@
 
+%Such simulations are important for Use if empirical force fields or QM/MM are not applicable:
+%Chemical reactions that happen throughout your system, not in a localized spatial region (e.g. proton transfer in water)
+%Unusual thermodynamic states for which simple analytical potential become inaccurate (e.g. high pressure states, interfaces, heterojunctions)
+%Electron density is an important observable (e.g. X-ray scattering) or information of the electronic wavefunctions is crucial.
+
 \textbf{Acknowledgments.} The research was funded by the Natural Sciences and Engineering Research Council of Canada through the Discovery Grant.

### 2017-08-17 almo-langevin.tex
August 17
Lots of small edits, removed section titles, added conclusion, modified dynamical properties plot slightly.

--- 2017-08-16 almo-langevin.tex
@@
 \begin{abstract}
-Linear scaling density functional theory based on absolutely localized molecular orbitals is used in conjunction with the Langevin equation of motion to create a low-cost molecular dynamics method for weakly coupled molecular systems which retains sufficient accuracy to generate correct dynamical properties in water.
+Linear scaling density functional theory based on absolutely localized molecular orbitals is used in conjunction with the Langevin equation of motion to create a low-cost molecular dynamics method for weakly coupled molecular systems which retains sufficient accuracy to generate accurate dynamical properties in water.
 
@@
 \maketitle
-\section{Introduction}
+%\section{Introduction}
 Since the unification of molecular dynamics (MD) and density functional theory (DFT) in 1985 by Car and Parrinello,~\cite{a:thecpmd} much research has been focused on reducing the computational overhead of \emph{ab initio} molecular dynamics (AIMD) potential energy calculations.~\cite{b:aimd}
@@
 
-\section{Theory}
+%\section{Theory}
 
@@
 By utilizing Eq.(\ref{eq:langevin}) in conjunction with the ALMO DFT method of force calculation, we will show that discontinuity in the PES can be ignored for sufficiently large $\gamma_L$. 
-Note that increasing $\gamma_L$ increases the sample size required to generate accurate dynamical properties, as random Gaussian noise is introduced.
+Note that increasing $\gamma_L$ increases the sample size required to generate accurate dynamical properties, as more random Gaussian noise is introduced.
 
@@
 During the second step of the ALMO SCF method described above, a self-consistent variational approach is taken to converge the non-orthogonal MOs. 
-Generally, this step utilizes a iterative search function in order to reduce the absolute norm of the energy gradient to some max allowable value $\epsilon_{SCF}$. 
-Previously, $\epsilon_{SCF}$ had to be set very low to ensure convergence of the electron structure calculations. 
+Generally, this step utilizes a iterative search function in order to reduce the absolute norm of the energy gradient to some target value $\epsilon_{SCF}$. 
+Previously, $\epsilon_{SCF}$ had to be set very low to ensure sufficient convergence of the electron structure calculations. 
 By increasing the allowable $\epsilon_{SCF}$, a great reduction in computational cost can be achieved. 
 Obviously, this introduces a significant error into the calculation of forces during each time step. 
-We assume that this error has the form
+We assume that this error has the form:
 %
@@
 %
-Where $f^{\text{SCF}}_{i\alpha}$ is the fully converged ALMO DFT force calculation, $f^{\text{APP}}_{i\alpha}$ is the approximate force calculated by allowing a large $\epsilon_{SCF}$, and $R^{\Delta}_{i\alpha} (t)$ is a stochastic ``white noise'' term that obeys
+where $f^{\text{SCF}}_{i\alpha}$ is the fully converged ALMO DFT force calculation, $f^{\text{APP}}_{i\alpha}$ is the approximate force calculated by allowing a large $\epsilon_{SCF}$, and $R^{\Delta}_{i\alpha} (t)$ is a stochastic ``white noise'' term that obeys
 %
@@
 %
-We will see that this assumption is sufficiently justified later (see fig. \ref{fig:randomforce}). 
-The Langevin equation of motion (\ref{eq:langevin}) can be re-written with the two stochastic terms combined into one $\gamma = \gamma_L+\Delta$.
+It's clear from figure \ref{fig:randomforce} that this assumption is sufficiently justified. 
+The Langevin equation of motion (\ref{eq:langevin}) can be re-written with the two stochastic terms combined into one $\gamma = \gamma_L-\Delta$.
 %
@@
 Note that $\Delta$ can be either positive or negative depending on the way in which errors are introduced in the system of interest. 
-$\Delta$ was found to be negative for all systems considered here.
+$\Delta$ was found to be positive for all systems considered here, reducing the effective random force on the system.
 
-\section{Optimizing the values of $\gamma_L$ and $\Delta$}
+%\section{Optimizing the values of $\gamma_L$ and $\Delta$}
 
@@
 
-An optimized $\gamma_L$ is one that is: large enough to adequately smooth force discontinuities; large enough to correctly thermostat the system of interest; and small enough not to significantly affect dynamical properties.
+An optimized $\gamma_L$ is one that is: large enough to adequately smooth force discontinuities; large enough to correctly thermostat the system of interest; and small enough not to significantly affect dynamical properties on the time scale of interest.
 The relative weight of the three is determined by the specifics of the system and what is most important to the researcher. 
@@
 %
+
 Another approach based on Eq. (\ref{eq:assumption}) can be used to find $\Delta$ within an order of magnitude if one can afford computing tightly converged SCF forces for a short MD trajectory. 
@@
 In other words, we use the integral of the ACF to estimate the strength of the random force. 
-In practice, this method for estimating $\Delta$ is not particularly accurate due to rapidly oscillating correlations between sequential values of $R^{\Delta}_{i\alpha} (t)$.
-The ACF is not usually possible to accurately integrate without running a reference trajectory with extremely short time-steps (see figure~\ref{fig:randomforce}).
-
-Once $\gamma_L$ and $\Delta$ are both optimized, approximate forces can be used to perform long MD simulations efficiently.
-\bigskip
+In practice, we found this method for estimating $\Delta$ was not particularly accurate due to rapidly oscillating correlations between sequential values of $R^{\Delta}_{i\alpha} (t)$. 
+The ACF was not possible to accurately integrate without running a reference trajectory with extremely short time-steps (see figure~\ref{fig:randomforce}).
+Still, this method provided a decent starting point for further investigation.
 
@@
 (c) Autocorrelation function of plot a, $\langle R^{\Delta}_{i\alpha} (t) R^{\Delta}_{i\alpha} (t+\tau) \rangle $.
-Using Eq.\ \ref{eq:delta}, we calculated $\Delta \approx \SI{2e-5}{\per\fs}$, but its true value is $\SI{6e-5}{\per\fs}$ for this system.
-Note that the  ACF is not truly a delta-function, and is characterized by rapidly oscillating correlations for short lag times. 
-The high frequency correlations made integration of the ACF inaccurate, but provided a decent starting point for further investigation.
-The autocorrelation decays rapidly enough that these forces can be considered random and uncorrelated to a good approximation for sufficiently long time scales ($\gg\SI{300}{\fs}$ for this system). 
-Calculations were performed with the PBE $\mathrm{E_{XC}}$ functional and the TZV2P basis set on 125 water molecules for 10 ps.}
+Using Eq.\ \ref{eq:delta}, we calculated $\Delta \approx \SI{2e-5}{\per\fs}$, but value was optimized heuristically to $\SI{6e-5}{\per\fs}$ for this system.
+The autocorrelation decays rapidly enough that these forces can be considered random and uncorrelated to a good approximation for sufficiently long time scales ($\gg\SI{200}{\fs}$). 
+Calculations were performed with the PBE $\mathrm{E_{XC}}$ functional and the TZV2P basis set on 125 water molecules for 10 ps at 298 K.}
 \end{figure}
 
-\section{Implementation}
+%\section{Implementation}
 
@@
 The computation scheme used is mostly identical to that described in the implementation section of ref.\ \citenum{a:almo-ls}. 
-For the dynamical properties calculations using ALMO DFT calculation methods, a periodic cell containing 125 water molecules was simulated for $\SI{10}{\ps}$ at a temperature of $\SI{293}{\K}$ and a constant density of $\SI{1.00597}{\g\per\cm^{3}}$.
+For the dynamical properties calculations using ALMO DFT methods, a periodic cell containing 125 water molecules was simulated for $\SI{10}{\ps}$ at a temperature of $\SI{293}{\K}$ and a constant density of $\SI{1.00597}{\g\per\cm^{3}}$.
 The system was thermostated using the Langevin ensemble (eq.\ \ref{eq:langevin2}) where $\gamma_L$ was set to $\SI{e-3}{\per\fs}$, and $\Delta$ was optimized to the order of $10^{-5}\, \mathrm{fs^{-1}}$.
@@
 
-The reference PES was calculated using the orbital transformation (OT) method~\cite{a:ot,a:ot2} and the Langevin thermostat with $\gamma_L = \SI{e-3}{\per\fs}$. 
+The reference PES was calculated using the orbital transformation (OT) method~\cite{a:ot,a:ot2} at $\SI{302}{\K}$ and the Langevin thermostat with $\gamma_L = \SI{e-3}{\per\fs}$. 
 All other system parameters were set identical to the ALMO system.
@@
 
-\section{Accuracy}
+%\section{Accuracy}
 
-In order to confirm that ALMO DFT forces are accurate in the limit where localization radius $R_{c} = \infty$, we computed the average difference in force between fully converged Born-Oppenheimer forces and forces calculated from ALMO DFT with varying $R_{c}$.
+In order to confirm that ALMO DFT forces are accurate in the limit where localization radius $R_{c} \rightarrow \infty$, we computed the average difference in force between fully converged Born-Oppenheimer forces and forces calculated from ALMO DFT with varying $R_{c}$.
 Figure \ref{fig:forcecomp} shows that as $R_{c}$ is increased to infinity, forces converge toward the Born-Oppenheimer limit.
@@
 To test the accuracy of ALMO DFT based MD we computed the Infrared (IR) spectrum, radial distribution function (RDF), kinetic energy distribution, and mean-squared displacement (MSD) function from a 10 ps MD trajectory.
-We compared the IR spectrum, RDF, and MSD to equivalent functions calculated from fully converged Kohn-Sham DFT forces (see fig. \ref{fig:dynproperties} and supporting information).
-From the MSD functions we also calculated diffusion constants for water at approximately 300 K.
-With fully converged forces we calculated D = $\SI{5.47e-9}{\m^{2}\per\s}$.
-With ALMO unconverged forces we calculated D = $\SI{5.55E-09}{\m^{2}\per\s}$.
+We compared the IR spectrum, RDF, and MSD function to corresponding reference functions calculated with fully converged Kohn-Sham DFT forces (see fig.
+\ref{fig:dynproperties} and supporting information).
+From the MSD functions we also calculated diffusion constants for water at approximately 298 K.
+Using fully converged forces we calculated D = $\SI{5.47e-9}{\m^{2}\per\s}$, whereas with ALMO unconverged forces we calculated D = $\SI{5.55E-09}{\m^{2}\per\s}$.
+
+Although the accuracy of calculated properties in figure \ref{fig:dynproperties} is quite good, it is clear that there are errors present.
+Some of the error may have arisen from the 9 K difference in the thermodynamic temperature between the reference system and our system of interest.
+Other sources of error are: small temporal sample size given our choice of $\gamma_L$; random error due to finite localization radius $R_{c}$, and random error due to high $\epsilon_{SCF}$.
+From figure \ref{fig:randomforce}, it's apparent that high values of $R_{c}$ introduce more random error than high values of $\epsilon_{SCF}$.
+However, raising both $R_{c}$ and $\gamma_L$ decreases computation time significantly, and it is ultimately up to the user to determine their maximum tolerance of force inaccuracy when using ALMO DFT methods.
 
@@
 
-\section{Speed}
+%\section{Speed}
 
 To show the computational efficiency gained by using unconverged ALMO DFT forces, we compare the average wall-time per MD step over 10 time steps for various values of $\epsilon_{SCF}$ in figure \ref{fig:strongscaling_log}.
-For these simulations, we chose to set the localization radius to 1.6 Van der Waal Radii (VdWR), as was done for calculation of dynamical properties.
-We also show timing of the fully converged OT SCF method, as well as the recently developed ``second generation Car-Parinello'' SCF method using OT, which improves upon the OT computational cost prefactor.~\cite{a:2ndcpmd}
-ALMO methods show excellent LS behavior for all values of $\epsilon_{SCF}$ tested.
+For these simulations, we chose to set the localization radius $R_{c} = 1.6$ Van der Waal Radii (VdWR), as was done for calculation of dynamical properties.
+We also show timing of the fully converged OT SCF method, as well as the recently developed ``second generation Car-Parinello'' MD method, which improves upon the OT computational cost prefactor.~\cite{a:2ndcpmd}
+ALMO methods show clear asymptotic LS behavior for all values of $\epsilon_{SCF}$ tested.
 OT SCF scales approximately cubically, as expected.
-Our findings show that the fastest ALMO SCF method presented here runs faster than second-generation Car-Parinello using OT SCF for water simulations larger than 256 molecules.
+Our findings show that the fastest ALMO DFT method presented here runs faster than second-generation Car-Parinello using OT DFT for water simulations larger than 256 molecules.
+
+\begin{figure}
+\includegraphics[trim={2.5cm 0.5cm 3.4cm 0.1cm},clip,width=8.6cm]{strongscaling_log.eps}
+\caption{\label{fig:strongscaling_log} Logscale plots which show the asymptotic LS behaviour of ALMO methods for liquid water at 298 K.
+Compared with the orbital transformation using fully converged SCF and second generation Car-Parinello force calculation methods.~\cite{a:ot,a:ot2,a:2ndcpmd}
+Cyan lines represent perfect cubic scaling, whereas gray lines represent perfect linear scaling. 
+Calculations were performed with the PBE $\mathrm{E_{XC}}$ functional and TZV2P basis set on 256 cores, averaged over 10 MD steps. 
+The localization radius was set to $R_{c} = 1.6$ VdWR.}
+\end{figure}
 
@@
 %\caption{\label{fig:strongscaling_linear} Linear plots showing LS behaviour of all ALMO methods for liquid water at 298 K. 
-%Calculations were performed with the PBE $\mathrm{E_{XC}}$ functional and TZV2P basis set on 256 CPUs. 
+%Calculations were performed with the PBE $\mathrm{E_{XC}}$ functional and TZV2P basis set on 256 cores. 
 %The localization radius was set to $R_{c} = 1.6$ VdWR.}
@@
 \begin{figure}
-\includegraphics[trim={2.5cm 0.5cm 3.4cm 0.1cm},clip,width=8.6cm]{strongscaling_log.eps}
-\caption{\label{fig:strongscaling_log} Logscale plots which show the asymptotic LS behaviour of ALMO methods for liquid water at 298 K.
-Compared with the orbital transformation using fully converged SCF and second generation Car-Parinello force calculation methods.
-Cyan lines represent perfect cubic scaling, whereas gray lines represent perfect linear scaling. 
-Calculations were performed with the PBE $\mathrm{E_{XC}}$ functional and TZV2P basis set on 256 CPUs, averaged over 10 MD steps. 
-The localization radius was set to $R_{c} = 1.6$ VdWR.}
-\end{figure}
-
-\begin{figure}
 \includegraphics[trim={1.6cm 0cm 4.7cm 0cm},clip,width=8.6cm]{weakscaling.eps}
-\caption{\label{fig:weakscaling} Weak scalability benchmarks which show optimization of code for parallel computing.
-Dashed gray lines connect systems simulated using the same number of CPU cores, showing LS behaviour over a range of configurations.
+\caption{\label{fig:weakscaling} Weak scalability benchmarks showing optimization of code for parallel computing.
+Dashed gray lines connect systems simulated using the same number of cores, verifying LS behaviour over a range of configurations.
 Of note is the 32768 molecule system successfully simulated using 4096 cores at around 25 minutes per MD step.
-We also show that systems as large as 8192 molecules can be simulated for 10,000 MD steps in under a month using our method.
+We also show that systems as large as 8192 molecules can be simulated for 10,000 MD steps in under a month using this method.
 Calculations were performed with the PBE $\mathrm{E_{XC}}$ functional and TZV2P basis set, averaged over 10 MD steps. 
@@
 
-\section{Conclusions}
+%\section{Conclusions}
+
+In this article we have clearly demonstrated the utility of ALMO DFT methods, especially when combined with the Langevin Equation of motion.
+ALMO DFT methods are a powerful class of AIMD electronic structure determination techniques capable of efficiently calculating forces for weakly coupled molecular systems.
+It's clear that forces calculated by ALMO DFT can be tuned as accurately as required by the user, up to the Born-Oppenheimer limit.
+Indeed, the ALMO DFT technique highlighted in this article has been shown to be capable of producing potential energy surfaces extremely efficiently for medium to large systems with only a minimal drop in the accuracy of calculated bulk system properties.
+We have also shown that due to its linear scaling nature, the ALMO DFT technique can be pushed to simulate enormous chemical systems accurately in remarkably short calculation times.
 
@@
 %\textbf{Supporting Information}
-%Calculated radial distribution functions of liquid water, comparison of timing benchmarks for the DZVP and TZV2P basis sets, timing benchmarks for systems containing 32,768 water molecules, timing benchmarks for the Kohn-Sham matrix build.
-This material is available free of charge via the Internet at http://pubs.acs.org.
 

### 2017-08-17 almo-langevin.tex
August 17th v2
Corrected a small math mistake

@@
 It's clear from figure \ref{fig:randomforce} that this assumption is sufficiently justified. 
-The Langevin equation of motion (\ref{eq:langevin}) can be re-written with the two stochastic terms combined into one $\gamma = \gamma_L-\Delta$.
+The Langevin equation of motion (\ref{eq:langevin}) can be re-written with the two stochastic terms combined into one $\gamma = \gamma_L+\Delta$.
 %
@@
 Note that $\Delta$ can be either positive or negative depending on the way in which errors are introduced in the system of interest. 
-$\Delta$ was found to be positive for all systems considered here, reducing the effective random force on the system.
+$\Delta$ was found to be positive for all systems considered here, increasing the effective random force on the system.
 

### 2017-08-18 almo-langevin.tex
Intro done

--- 2017-08-17 almo-langevin.tex
@@
 \begin{abstract}
-Linear scaling density functional theory based on absolutely localized molecular orbitals is used in conjunction with the Langevin equation of motion to create a low-cost molecular dynamics method for weakly coupled molecular systems which retains sufficient accuracy to generate correct dynamical properties in water.
+Linear scaling density functional theory based on absolutely localized molecular orbitals is used in conjunction with the Langevin equation of motion to create a low-cost molecular dynamics method for large systems of weakly-interacting molecules.
+which retains sufficient accuracy to generate correct dynamical properties in water.
 
@@
 We also introduce a simple heuristic approach to find the best Langevin Equation parameters to use for balancing system-specific imperfect forces and maintaining correct dynamical behavior.
+
+% We developed a linear-scaling AIMD method with low computational overhead by utilizing compact molecular orbitals strictly localized within predefined radii.
+High efficiency of the method is achieved without sacrificing its accuracy with a combination of two techniques: (1) on-the-fly construction of accurate localized orbitals without lengthy self-consistent optimization and (2) the unbalanced Langevin integrator that is fine-tuned to retain stable dynamics even with imperfect forces.
+A remarkable feature of the implemented method is that it remains efficient even for challenging condensed phase systems even if large diffuse basis sets are used.
+We demonstrated that, for systems well-represented by the compact orbitals (e.g.
+molecular systems, ionic salts), the new AIMD method enables simulations on previously inaccessible time and length scales.
+The first steps towards generalizing the method to more challenging systems of strongly interacting atoms (e.g.
+covalent crystals) will also be discussed.
+
 \end{abstract}
@@
 
-\section{Introduction}
+%\section{Introduction}
 
 Since the unification of molecular dynamics and density functional theory (DFT)~\cite{a:thecpmd},
-\emph{ab initio} molecular dynamics (AIMD)~\cite{b:aimd} has become an important tool for computational studies of weakly bonded ensembles of molecules such as gas-phase clusters, molecular and ionic liquids, solutions, and molecular solids.
-[of gas-phase molecules and condensed-phase systems.]
-Unfortunately, the computational cost of AIMD based on the conventional Kohn-Sham (KS) DFT grows cubically with the number of atoms in the system and severely limits the \emph{length scale} accessible by AIMD.
-In an attempt to address this issue, substantial efforts have been directed to the development of linear scaling (LS) DFT methods for the on-the-fly calculation of forces.
-
-% Most methods are based on localization of electrons into small sub domains of the system, so that long range charge transfer is ignored.~\cite{a:linscale3,a:lee-yang-1996,a:ls-scuseria-1997,a:ls-manolopoulos-1998,a:ls-helgaker-2001,a:ls-niklasson-2003,a:curvy2,a:ls-dm-sign,a:ls-parinello} So far the available methods still suffer from a high computational cost prefactor, even though they scale linearly for large systems.
-This has limited the available time scales for LS AIMD severely.
+\emph{ab initio} molecular dynamics (AIMD)~\cite{b:aimd} has become an important tool for studying of weakly bonded ensembles of molecules such as gas-phase molecular clusters, liquids, solutions, and molecular solids.
+%[of gas-phase molecules and condensed-phase systems.]
+Unfortunately, the computational cost of the conventional Kohn-Sham (KS) DFT grows cubically with the number of atoms, which severely limits the \emph{length scales} accessible by AIMD.
+%and hinders it application to large heterogeneous systems (interfaces, nuclei, defects).
+To address this issue and calculate forces in large systems efficiently, substantial efforts have been directed to the development of linear scaling (LS) DFT methods.
 
 % Re-write the following section and include any recent advances
-The steep scaling of the conventional KS DFT originates from the delocalized nature of the canonical KS orbitals, which tend to delocalize over all atoms in the system, with an alternative set of electronic descriptors.
-All LS methods must A number of such methods explore the natural sparsity of the one-electron density matrix (DM)~\cite{a:ls-kohn-1996,a:ls-rev-1999} and are capable of yielding LS for large systems.~\cite{a:linscale3,a:lee-yang-1996,a:ls-scuseria-1997,a:ls-manolopoulos-1998,a:ls-helgaker-2001,a:ls-niklasson-2003,a:curvy2,a:ls-dm-sign}
-Unfortunately, the variational optimization of the DM is very inefficient for accurate DFT calculations,~\cite{a:ls-rev-1999,a:ls-dm-sign} which require many basis functions per atom. 
-Therefore, the applications of DM-based LS methods have been limited to minimal-basis tight-binding problems. 
-The optimal basis variants of the DM methods~\cite{a:ls-stechel-1994,a:ls-gillan-1995,a:ls-gillan-1996,a:ls-onetep-2003} designed to address this issue contract the large basis set into a small number of new localized basis functions and then optimize the DM in the contracted basis.
-Although such methods have been successfully used for the evaluation of accurate DFT energies of very large systems~\cite{a:ls-onetep-2009,a:ls-conquest-2010,a:ls-onetep-2010-app1,a:ls-rev-2012} their application for AIMD is hampered by the computationally costly optimization of both the contracted orbitals and the density matrix.
-%
-From this point of view, methods based on the direct optimization of localized KS orbitals~\cite{a:ls-galli-parrinello-1992,a:ls-mauri-galli-car-1993,a:ls-ordejon-1993,a:ls-mauri-galli-1994,a:ls-ordejon-1995,a:ls-kim-mauri-galli-1995,a:ls-fattebert-2004,a:ls-fattebert-2006,a:burger-yang-2008} are advantageous since they require only the occupied orbitals and, thus, deal with fewer variational degrees of freedom than the DM methods.
-Unfortunately, the progress in the development of orbital-based LS methods has been hindered by the inherently difficult convergence of the localized-orbital optimization.~\cite{a:ls-mauri-galli-car-1993,a:ls-ordejon-1995,a:ls-fattebert-2004,a:ls-rev-1999,a:weitao-yang-2013}
-%Furthermore, numerous proposed LS functionals exhibit multiple local minima, which limits their applicability in molecular dynamics simulations because of the resulting poor conservation of the total energy~\cite{a:ls-rev-1999}.
+In all LS DFT methods, the \emph{delocalized} eigenstates of the effective KS Hamiltonian, known as canonical KS orbitals, must be replaced with an alternative set of \emph{local} electronic descriptors
+~\footnote{Canonical KS orbitals are delocalized over all atoms in the system. 
+This delocalization is the main reason for the steep growth of the cost of the conventional KS DFT: since both the number of orbitals and the number of localization regions (e.g.
+atomic basis functions) increase linearly with the number of atoms, the total number of the electronic degrees of freedom that describe delocalized orbitals grows quadratically.}.
+Most LS methods~\cite{a:linscale3,a:lee-yang-1996,a:ls-scuseria-1997,a:ls-manolopoulos-1998,a:ls-helgaker-2001,a:ls-niklasson-2003,a:curvy2,a:ls-dm-sign} explore the natural locality of the one-electron density matrix (DM).
+%However, the variational optimization of the DM is very inefficient for accurate DFT calculations that require many basis functions per atom~\cite{a:ls-rev-1999,a:ls-dm-sign}. 
+However, in accurate DFT calculations that require many basis functions per atom, the variational optimization of the DM formulation is advantageous only for impractically large systems~\cite{a:ls-rev-1999,a:ls-dm-sign,a:almo-ls}.
+%However, in accurate DFT calculations with many basis functions per atom are required, the DM becomes sufficiently sparse only for impractically large systems~\cite{a:ls-rev-1999,a:ls-dm-sign,a:almo-ls}.
+Therefore, the application of DM-based LS methods have been restricted to minimal-basis tight-binding problems. 
+The optimal basis variants of the DM methods~\cite{a:ls-stechel-1994,a:ls-gillan-1995,a:ls-gillan-1996,a:ls-onetep-2003} designed to rectify this issue contract large basis sets into a small number of new localized basis functions and then optimize the DM in the contracted basis.
+Although such methods have been successfully used for the evaluation of accurate DFT energies of large static systems~\cite{a:ls-onetep-2009,a:ls-conquest-2010,a:ls-onetep-2010-app1,a:ls-rev-2012} their use in AIMD is hampered by the costly optimization of both the contracted orbitals and the density matrix.
+Alternatively, LS can be achieved using a set of local electronic descriptors constructed by enforcing strict locality on the KS orbitals and, at the same time, by relaxing the orthogonality constraints imposed on them.
+From the computational point of view, LS methods based on the direct optimization of such compact localized nonorthogonal orbitals~\cite{a:ls-galli-parrinello-1992,a:ls-mauri-galli-car-1993,a:ls-ordejon-1993,a:ls-mauri-galli-1994,a:ls-ordejon-1995,a:ls-kim-mauri-galli-1995,a:ls-fattebert-2004,a:ls-fattebert-2006,a:burger-yang-2008} are more promising since they require only the occupied orbitals and, thus, deal with fewer variational degrees of freedom than the DM methods.
+Unfortunately, the progress in the development of orbital-based LS methods has been hindered by the inherently difficult optimization of localized orbitals~\cite{a:ls-mauri-galli-car-1993,a:ls-ordejon-1995,a:ls-fattebert-2004,a:ls-rev-1999,a:weitao-yang-2013}.
 
-Hence, the computational prefactor of the existing LS DFT methods~\cite{a:ls-rev-1995,a:ls-rev-1999,a:ls-rev-2002,a:ls-rev-2012} remains very high and most AIMD simulations are still performed using conventional low-prefactor cubic scaling DFT methods.
+%Thus, despite remarkable recent progress in linear scaling density function theory, the computational cost of existing methods remains too high for routine \emph{ab initio} molecular dynamics (AIMD) simulations.
+Thus, despite remarkable recent progress the high computational overhead of existing asymptotically linear-scaling DFT methods restrict their use in AIMD simulations to very short \emph{time scales}, systems of low dimensions, or problems that can be adequately described with minimal basis sets.
+Otherwise, LS DFT cannot compete with the straightforward cubically scaling KS DFT on the length scales accessible in routine AIMD.
+%Thus,  the computational cost of the existing asymptotically linear-scaling DFT methods remains too high to compete with the straightforward cubically scaling KS DFT on the length scales practically accessible in routine AIMD simulations.
+%Thus, although there exist numerous DFT methods capable of yielding asymptotic LS, their computational cost for systems in/on the practically accessible length scale remains too high to compete with the conventional cubically scaling KS DFT with its highly-optimized numerical algorithms.
 
-Here, we introduce a new low-cost linear scaling AIMD method for weakly-interacting molecular systems based on the recently developed absolutely localized molecular orbitals (ALMO) DFT approach~\cite{a:almo-ls}.
+In this work, we introduce a new AIMD method for weakly-interacting molecular systems that can reach the linear scaling regime for systems of medium size while maintaining a low computational overhead.
+Using liquid water as an example, we demonstrate that the proposed AIMD methodology enables computational studies of molecular systems on previously inaccessible time and length scales.
+
 ALMO DFT exploits non-orthogonal molecular orbitals (MOs) in a two-step self-consistent filed (SCF) approach that is designed to avoid the usual convergence issues~\cite{a:ls-mauri-galli-car-1993,a:ls-ordejon-1995,a:ls-fattebert-2004,a:ls-rev-1999} and .
@@
 For more details on this method, see ref.\ \citenum{a:almo-ls}.
+
+It is important to note that the presented logical partitioning with the subsequent construction of localized MOs is a rather general approach used in a large number of electronic structure theories, which are collectively known as fragmentation methods~\cite{a:fragmentation-rev}.
+These methods vary greatly in how the partitioning and recombination are performed and, therefore, differ in accuracy and computational cost.
+We would like to emphasize that in our approach a proper quantum mechanical description of the entire system is constructed in the form of the total idempotent density matrix:
+%
+\begin{eqnarray}
+\label{eq:properdm}
+\hat{R} = \sum_{x} \vert \psi^{xi} \rangle \langle \psi_{xi}\vert.
+\end{eqnarray}
+%
+This approach provides the most rigorous description of the electronic structure.
 
@@
 
+To summarize, a low-cost linear-scaling AIMD method for weakly-interacting molecular systems was developed by utilizing compact molecular orbitals strictly localized within predefined radii.
+High efficiency of the method is achieved without sacrificing its accuracy with a combination of two techniques: (1) on-the-fly construction of accurate localized orbitals without lengthy self-consistent optimization and (2) a modified Langevin equation that is fine-tuned to compensate retain stable dynamics even with imperfect forces.
+A remarkable feature of the implemented method is that it remains efficient even for challenging condensed phase systems, even with diffuse atom-centered basis sets.
+Using liquid water as an example, it is demonstrated that the new AIMD method enables simulations of molecular systems on previously inaccessible time and length scales.
+
+Generalization of the method to more challenging systems of strongly interacting atoms (e.g. covalent crystals) will also be discussed.
+
 %Such simulations are important for Use if empirical force fields or QM/MM are not applicable:

### 2017-08-18 almo-langevin.tex

@@
 \begin{abstract}
-Linear scaling density functional theory based on absolutely localized molecular orbitals is used in conjunction with the Langevin equation of motion to create a low-cost molecular dynamics method for large systems of weakly-interacting molecules.
+%Linear scaling density functional theory based on absolutely localized molecular orbitals is used in conjunction with the Langevin equation of motion to create a low-cost molecular dynamics method for large systems of weakly-interacting molecules.
 which retains sufficient accuracy to generate correct dynamical properties in water.
+Linear scaling density functional theory based on absolutely localized molecular orbitals is used in conjunction with the Langevin equation of motion to create a low-cost molecular dynamics method for weakly coupled molecular systems which retains sufficient accuracy to generate accurate dynamical properties in water.
 
@@
 %
-We will see that this assumption is sufficiently justified later (see fig. \ref{fig:randomforce}). 
+It's clear from figure \ref{fig:randomforce} that this assumption is sufficiently justified. 
 The Langevin equation of motion (\ref{eq:langevin}) can be re-written with the two stochastic terms combined into one $\gamma = \gamma_L+\Delta$.
@@
 Note that $\Delta$ can be either positive or negative depending on the way in which errors are introduced in the system of interest. 
-$\Delta$ was found to be negative for all systems considered here.
+$\Delta$ was found to be positive for all systems considered here, increasing the effective random force on the system.
 
-\section{Optimizing the values of $\gamma_L$ and $\Delta$}
+%\section{Optimizing the values of $\gamma_L$ and $\Delta$}
 

### 2017-08-18 almo-langevin.tex
August 18th
Merged our versions, changed \gamma_L to \gamma, modified section explaining \delta, various small edits.

@@
 
-%\documentclass[aps,prl,,reprint,amsmath,amssymb,superscriptaddress,longbibliography]{revtex4-1}
+%\documentclass[10pt,aps,prl,twocolumn,amsmath,amssymb,superscriptaddress,longbibliography]{revtex4-1}
 %\documentclass[preprint,showpacs,preprintnumbers,amsmath,amssymb]{revtex4-1}
@@
 \usepackage{color}
-\usepackage{epstopdf} 
-%\graphicspath{{pics}}
+\usepackage{graphicx}
 %\usepackage[normalem]{ulem}
@@
 \begin{abstract}
-%Linear scaling density functional theory based on absolutely localized molecular orbitals is used in conjunction with the Langevin equation of motion to create a low-cost molecular dynamics method for large systems of weakly-interacting molecules.
-which retains sufficient accuracy to generate correct dynamical properties in water.
-Linear scaling density functional theory based on absolutely localized molecular orbitals is used in conjunction with the Langevin equation of motion to create a low-cost molecular dynamics method for weakly coupled molecular systems which retains sufficient accuracy to generate accurate dynamical properties in water.
+A density functional theory approach based on absolutely localized molecular orbitals is used in conjunction with the Langevin equation of motion to create a low-cost molecular dynamics method for weakly coupled molecular systems which retains sufficient accuracy to generate accurate dynamical properties in water.
 
-The approach utilizes a systematically inaccurate linear-scaling potential energy surface (PES) calculation method. 
-Inaccuracies in the calculation of force using this method are balanced by a weighted version of the Langevin Equation of motion. 
-The stochastic ``white noise'' of the Langevin Equation acts to smooth out PES discontinuities while also thermostating the system to maintain a constant thermodynamic temperature. 
-We show that Langevin dynamics can be successfully implemented to reduce computational cost by allowing systematic error in PES calculation to be absorbed into the stochastic and damping forces associated with the Langevin equation.
-We also introduce a simple heuristic approach to find the best Langevin Equation parameters to use for balancing system-specific imperfect forces and maintaining correct dynamical behavior.
+The approach utilizes a systematically inaccurate potential energy surface calculation method that grows linearly with the size of the simulated system. 
+Systematic inaccuracies in the calculation of force using this method are balanced by a weighted version of the Langevin equation of motion. 
+The stochastic ``white noise'' of the Langevin equation acts to absorb force discontinuities while also thermostating the system to maintain the desired thermodynamic temperature. 
+In addition, we introduce a simple heuristic approach to find optimized Langevin equation parameters to use for balancing system-specific imperfect forces and maintaining correct dynamical behavior.
 
@@
 Although such methods have been successfully used for the evaluation of accurate DFT energies of large static systems~\cite{a:ls-onetep-2009,a:ls-conquest-2010,a:ls-onetep-2010-app1,a:ls-rev-2012} their use in AIMD is hampered by the costly optimization of both the contracted orbitals and the density matrix.
+
 Alternatively, LS can be achieved using a set of local electronic descriptors constructed by enforcing strict locality on the KS orbitals and, at the same time, by relaxing the orthogonality constraints imposed on them.
@@
 
-ALMO DFT exploits non-orthogonal molecular orbitals (MOs) in a two-step self-consistent filed (SCF) approach that is designed to avoid the usual convergence issues~\cite{a:ls-mauri-galli-car-1993,a:ls-ordejon-1995,a:ls-fattebert-2004,a:ls-rev-1999} and .
+ALMO DFT exploits non-orthogonal molecular orbitals (MOs) in a two-step self-consistent filed (SCF) approach that is designed to avoid the usual convergence issues.~\cite{a:ls-mauri-galli-car-1993,a:ls-ordejon-1995,a:ls-fattebert-2004,a:ls-rev-1999}
 We show that due to its linear scaling nature and relatively low computational prefactor, ALMO-based DFT calculations allow for simulation of very large weakly-coupled molecular systems. 
@@
 This is akin to calculating the electronic structure of each molecule (with frozen nuclei) within a coulombic field produced by the surrounding molecules. 
-Charge transfer is completely constrained at this step. 
+Charge transfer is completely constrained. 
 In the second step, the variationally optimized but constrained MOs are allowed to delocalize onto nearby groups up to a desired cutoff radius $R_{c}$. 
@@
 
-A known source of error in force calculation within the ALMO SCF paradigm occurs due to the discontinuous jump in potential energy when a neighboring molecule passes beyond the cutoff radius $R_{c}$ and is suddenly ignored in potential energy calculations.
+A known source of error in force calculation within the ALMO SCF paradigm occurs due to the discontinuous jump in potential energy when a neighboring molecule passes beyond the localization radius $R_{c}$ and is suddenly ignored in potential energy calculations.
 Setting $R_{c}$ to a larger value reduces this discontinuity, but greatly increases the computational cost as longer distance (and less important) interactions must be calculated. 
@@
 
-In the present article, we propose a new approach which allows for relaxation of the constraint on $R_{c}$ by introducing related damping and stochastic terms to the force calculation by utilizing the Langevin Equation of motion.~\cite{a:Kubo-1986}
+In the present article, we propose a new approach which allows for relaxation of the constraint on $R_{c}$ by introducing related damping and stochastic terms to the force calculation by utilizing the Langevin equation of motion.~\cite{a:Kubo-1986}
 
@@
 \label{eq:langevin}
-m_i \ddot{r}_{i\alpha} = f^{\text{SCF}}_{i\alpha} - \gamma_L m_i \dot{r}_{i\alpha} + R^{\gamma_L}_{i\alpha} (t),
+m_i \ddot{r}_{i\alpha} = f^{\text{SCF}}_{i\alpha} - \gamma m_i \dot{r}_{i\alpha} + R^{\gamma}_{i\alpha} (t),
 \end{align}
@@
 Where $m_i$ refers to the mass of the i-th particle, and $r_{i\alpha}$ its position along dimension $\alpha$. 
-$f^{\text{SCF}}_{i\alpha}$ is the force calculated using the ALMO SCF method described above for the i-th particle along dimension $\alpha$, $\gamma_L$ is the Langevin scaling factor which sets the strength of both the damping term  $\gamma_L m_i \dot{r}_{i\alpha}$ and the stochastic Gaussian ``white noise'' term $R^{\gamma_L}_{i\alpha} (t)$.
+$f^{\text{SCF}}_{i\alpha}$ is the force calculated using the ALMO SCF method described above for the i-th particle along dimension $\alpha$, $\gamma$ is the Langevin scaling factor which sets the strength of both the damping term  $\gamma m_i \dot{r}_{i\alpha}$ and the stochastic Gaussian ``white noise'' term $R^{\gamma}_{i\alpha} (t)$.
 In order to correctly sample the canonical distribution, the random force must be zero mean. 
@@
 \label{eq:stochastic}
-\langle R^{\gamma_L}_{i\alpha} (t) \rangle &= 0, \\
-\langle R^{\gamma_L}_{i\alpha} (t)  R^{\gamma_L}_{j\beta} (t') \rangle &= 2 k_B T \gamma_L m_i \delta_{ij} \delta_{\alpha\beta} \delta(t-t')
+\langle R^{\gamma}_{i\alpha} (t) \rangle &= 0, \\
+\langle R^{\gamma}_{i\alpha} (t)  R^{\gamma}_{j\beta} (t') \rangle &= 2 k_B T \gamma m_i \delta_{ij} \delta_{\alpha\beta} \delta(t-t')
 \end{align}
@@
 This means the damping and white noise terms will work to maintain a constant temperature while also generating the correct Maxwell-Boltzmann distribution of velocities.
-It has previously been shown that Langevin dynamics can be utilized to correct imperfect force calculations while maintaining accurate dynamical properties.~\cite{a:langevin-why,a:2ndcpmd,b:tuckerman-stat,a:ceriotti}
-By utilizing Eq.(\ref{eq:langevin}) in conjunction with the ALMO DFT method of force calculation, we will show that discontinuity in the PES can be ignored for sufficiently large $\gamma_L$. 
-Note that increasing $\gamma_L$ increases the sample size required to generate accurate dynamical properties, as more random Gaussian noise is introduced.
 
-The concept of balancing an error in force calculation with the Langevin equation can be taken even further, thereby reducing computational overhead even more. 
+It has previously been shown that Langevin dynamics can be utilized to correct inaccurate force calculations to maintain correct dynamical properties.~\cite{a:langevin-why,a:2ndcpmd,b:tuckerman-stat,a:ceriotti}
+By utilizing Eq.\ (\ref{eq:langevin}) in conjunction with the ALMO SCF method of force calculation, we will show that discontinuity in the PES can be ignored for sufficiently large $\gamma$. 
+Note that increasing $\gamma$ increases the sample size required to generate accurate dynamical properties, as more random Gaussian noise is introduced.
+
+The concept of balancing errors in force calculation with the Langevin equation can be taken even further, thereby reducing computational overhead even more. 
 During the second step of the ALMO SCF method described above, a self-consistent variational approach is taken to converge the non-orthogonal MOs. 
 Generally, this step utilizes a iterative search function in order to reduce the absolute norm of the energy gradient to some target value $\epsilon_{SCF}$. 
-Previously, $\epsilon_{SCF}$ had to be set very low to ensure sufficient convergence of the electron structure calculations. 
+Previously, $\epsilon_{SCF}$ had to be set very small to ensure sufficient convergence of the electron structure calculations. 
 By increasing the allowable $\epsilon_{SCF}$, a great reduction in computational cost can be achieved. 
@@
 %
-where $f^{\text{SCF}}_{i\alpha}$ is the fully converged ALMO DFT force calculation, $f^{\text{APP}}_{i\alpha}$ is the approximate force calculated by allowing a large $\epsilon_{SCF}$, and $R^{\Delta}_{i\alpha} (t)$ is a stochastic ``white noise'' term that obeys
+where $f^{\text{SCF}}_{i\alpha}$ is the force calculated from an accurate and fully converged SCF method, $f^{\text{APP}}_{i\alpha}$ is the approximate force calculated by allowing a large $\epsilon_{SCF}$, and $R^{\Delta}_{i\alpha} (t)$ is a stochastic ``white noise'' term that obeys
 %
@@
 It's clear from figure \ref{fig:randomforce} that this assumption is sufficiently justified. 
-The Langevin equation of motion (\ref{eq:langevin}) can be re-written with the two stochastic terms combined into one $\gamma = \gamma_L+\Delta$.
+Hence, the Langevin equation of motion (\ref{eq:langevin}) can be re-written with the two stochastic terms combined into one:
 %
@@
 \label{eq:langevin2}
-m_i \ddot{r}_{i\alpha} = f^{\text{APP}}_{i\alpha} - \gamma_L m_i \dot{r}_{i\alpha} + R^{\gamma}_{i\alpha} (t)
+m_i \ddot{r}_{i\alpha} = f^{\text{APP}}_{i\alpha} - \gamma m_i \dot{r}_{i\alpha} + R^{\gamma + \Delta}_{i\alpha} (t)
 \end{align}
 %
-Note that $\Delta$ can be either positive or negative depending on the way in which errors are introduced in the system of interest. 
-$\Delta$ was found to be positive for all systems considered here, increasing the effective random force on the system.
+where $R^{\gamma + \Delta}_{i\alpha} = R^{\gamma}_{i\alpha} + R^{\Delta}_{i\alpha}$. 
+Note that $\Delta$ can be either positive or negative depending on the way errors are introduced in the system of interest. 
+$\Delta$ was found to be positive for all systems considered here, increasing the amplitude of the effective random force on the system.
 
-%\section{Optimizing the values of $\gamma_L$ and $\Delta$}
+%\section{Optimizing the values of $\gamma$ and $\Delta$}
 
-In theory, both $\gamma_L$ and $\Delta$ should be system-size independent, but in practice only $\gamma_L$ is.
-$\Delta$ is dependent on system size due to finite size effects, which decay for increasing system size.
-Fortunately, $\gamma_L$ and $\Delta$ can be separately optimized for a given system of interest, assuming a method of accurate force calculation can be applied to the system for a short trajectory.
+In theory, both $\gamma$ and $\Delta$ should be system-size independent, but in practice only $\gamma$ is.
+We found that $\Delta$ is dependent on system size due to finite size effects, which decay for increasing system size.
+Fortunately, $\gamma$ and $\Delta$ can be separately optimized for a given system of interest, assuming a method of accurate force calculation can be applied to the system for a short trajectory.
 
-An optimized $\gamma_L$ is one that is: large enough to adequately smooth force discontinuities; large enough to correctly thermostat the system of interest; and small enough not to significantly affect dynamical properties on the time scale of interest.
+An optimized $\gamma$ is one that is: large enough to adequately absorb force discontinuities; large enough to correctly thermostat the system of interest on a reasonable time scale; and small enough not to significantly affect dynamical properties on the time scale of interest.
 The relative weight of the three is determined by the specifics of the system and what is most important to the researcher. 
-To optimize $\gamma_L$ given the relative importance of the above properties, we suggest running several short trajectories on the system of interest (or a scaled-down version of it) using a tightly converged PES calculation and varying the value of $\gamma_L$.
-We found that a value of $\gamma_L$ between $10^{-5}$ and $10^{-3}\ \mathrm{fs^{-1}}$ was optimal for the purposes of the research presented here.
+To optimize $\gamma$ given the relative importance of the above properties, we suggest running several short trajectories on the system of interest (or a scaled-down version of it) using a tightly converged PES calculation and varying the value of $\gamma$.
+We found that a value of $\gamma$ between $10^{-5}$ and $10^{-3}\ \mathrm{fs^{-1}}$ was optimal for the purposes of the research presented here.
 
-An optimized $\Delta$ is one that, in the absence of $\gamma_L$, perfectly offsets any systematic inaccuracies in the force calculation. 
-A trajectory with inaccurate forces, \mbox{$\gamma_L = 0$}, and optimized $\Delta$ should have no long term drift in the total energy of the system.
-Short term energy fluctuations should be expected, and can later be ``smoothed out'' by setting finite $\gamma_L$.
+To correctly thermostat the system of interest, $\Delta$ needs to be reduced to zero.
+This is done by finding the value of $\Delta$ and adding a third random force term $R^{-\Delta}_{i\alpha}$ to equation \ref{eq:langevin2} which cancels out the energy introduced to the system by $R^{\Delta}_{i\alpha}$.
+This means that a trajectory with inaccurate forces and optimized $R^{-\Delta}_{i\alpha}$ should have no long term drift in the total energy of the system. 
+Short term energy fluctuations should be expected, and can later be ``smoothed out'' by setting finite $\gamma$.
 
-$\Delta$ can therefore be optimized simply by trying various values of $\Delta$ (with \mbox{$\gamma_L = 0$}) until the simulation described by Eq.
-(\ref{eq:langevin2}) has long-term stable total energy and satisfies the equipartition theorem.
+$R^{-\Delta}_{i\alpha}$ can therefore be optimized simply by trying various values of $\Delta$ (with \mbox{$\gamma = 0$}) until the simulation has long-term stable total energy and satisfies the equipartition theorem.
 %
@@
 First, run the simulation with $\Delta = 0$ and converged SCF forces. 
-Second, compute approximate forces for snapshots along the trajectory and calculate the random term $R^{\Delta}_{i\alpha} (t)$ using Eq. \ref{eq:assumption}. 
+Second, compute approximate forces for snapshots along the trajectory and calculate the random term $R^{\Delta}_{i\alpha} (t)$ using Eq.\ \ref{eq:assumption}. 
 Third, calculate the autocovariance function (ACF) of the noise $\text{ACF}(t) = \langle R^{\Delta}_{i\alpha} (0)  R^{\Delta}_{i\alpha} (t) \rangle$, which should be sharply peaked around zero lag time if our assumption is correct.
@@
 The ACF was not possible to accurately integrate without running a reference trajectory with extremely short time-steps (see figure~\ref{fig:randomforce}).
-Still, this method provided a decent starting point for further investigation.
+For example, using Eq.\ \ref{eq:delta} we calculated $\Delta \approx \SI{2e-5}{\per\fs}$, but value was optimized heuristically to $\SI{6e-5}{\per\fs}$.
+Still, this method provided a decent order of magnitude starting point for further investigation.
+Note that for our system, the ACF decays rapidly enough that the forces can be considered random and uncorrelated to a good approximation for sufficiently long time scales ($\gg\SI{200}{\fs}$).
 
@@
 (c) Autocorrelation function of plot a, $\langle R^{\Delta}_{i\alpha} (t) R^{\Delta}_{i\alpha} (t+\tau) \rangle $. 
-Using Eq.\ \ref{eq:delta}, we calculated $\Delta \approx \SI{2e-5}{\per\fs}$, but value was optimized heuristically to $\SI{6e-5}{\per\fs}$ for this system.
-The autocorrelation decays rapidly enough that these forces can be considered random and uncorrelated to a good approximation for sufficiently long time scales ($\gg\SI{200}{\fs}$). 
 Calculations were performed with the PBE $\mathrm{E_{XC}}$ functional and the TZV2P basis set on 125 water molecules for 10 ps at 298 K.}
@@
 For the dynamical properties calculations using ALMO DFT methods, a periodic cell containing 125 water molecules was simulated for $\SI{10}{\ps}$ at a temperature of $\SI{293}{\K}$ and a constant density of $\SI{1.00597}{\g\per\cm^{3}}$.
-The system was thermostated using the Langevin ensemble (eq.\ \ref{eq:langevin2}) where $\gamma_L$ was set to $\SI{e-3}{\per\fs}$, and $\Delta$ was optimized to the order of $10^{-5}\, \mathrm{fs^{-1}}$.
+The system was thermostated using the Langevin ensemble (eq.\ \ref{eq:langevin2}) where $\gamma$ was set to $\SI{e-3}{\per\fs}$, and $\Delta$ was optimized to the order of $10^{-5}\, \mathrm{fs^{-1}}$.
 Note that for timing benchmarks, $\Delta$ was not optimized. 
@@
 
-The reference PES was calculated using the orbital transformation (OT) method~\cite{a:ot,a:ot2} at $\SI{302}{\K}$ and the Langevin thermostat with $\gamma_L = \SI{e-3}{\per\fs}$. 
+The reference PES was calculated using the orbital transformation (OT) method~\cite{a:ot,a:ot2} at $\SI{302}{\K}$ and the Langevin thermostat with $\gamma = \SI{e-3}{\per\fs}$. 
 All other system parameters were set identical to the ALMO system.
@@
 
-To test the accuracy of ALMO DFT based MD we computed the Infrared (IR) spectrum, radial distribution function (RDF), kinetic energy distribution, and mean-squared displacement (MSD) function from a 10 ps MD trajectory.
-We compared the IR spectrum, RDF, and MSD function to corresponding reference functions calculated with fully converged Kohn-Sham DFT forces (see fig.
+To test the accuracy of our unconverged ALMO based MD we computed the Infrared (IR) spectrum, radial distribution function (RDF), kinetic energy distribution, and mean-squared displacement (MSD) function from a 10 ps MD trajectory (see fig.
 \ref{fig:dynproperties} and supporting information).
-From the MSD functions we also calculated diffusion constants for water at approximately 298 K.
-Using fully converged forces we calculated D = $\SI{5.47e-9}{\m^{2}\per\s}$, whereas with ALMO unconverged forces we calculated D = $\SI{5.55E-09}{\m^{2}\per\s}$.
+We compared the IR spectrum, RDF, and MSD function to corresponding reference functions calculated with fully converged KS DFT forces.
+The kinetic energy distribution is compared with the theoretical Maxwell-Boltzmann distribution at the average temperature of the system.
+From the slope of the MSD functions we also calculated diffusion constants.
+Using fully converged KS DFT forces we calculated D = $\SI{5.47e-9}{\m^{2}\per\s}$, whereas with unconverged ALMO forces we calculated D = $\SI{5.55E-09}{\m^{2}\per\s}$.
 
 Although the accuracy of calculated properties in figure \ref{fig:dynproperties} is quite good, it is clear that there are errors present.
-Some of the error may have arisen from the 9 K difference in the thermodynamic temperature between the reference system and our system of interest.
-Other sources of error are: small temporal sample size given our choice of $\gamma_L$; random error due to finite localization radius $R_{c}$, and random error due to high $\epsilon_{SCF}$.
+Some of the error can be attributed to the 9 K difference in the thermodynamic temperature between the reference system and our system of interest.
+Other sources of error are: small temporal sample size given our choice of $\gamma$; random error due to finite localization radius $R_{c}$, and random error due to high $\epsilon_{SCF}$.
 From figure \ref{fig:randomforce}, it's apparent that high values of $R_{c}$ introduce more random error than high values of $\epsilon_{SCF}$.
-However, raising both $R_{c}$ and $\gamma_L$ decreases computation time significantly, and it is ultimately up to the user to determine their maximum tolerance of force inaccuracy when using ALMO DFT methods.
+However, raising both $R_{c}$ and $\gamma$ decreases computation time significantly, and it is ultimately up to the user to determine their maximum tolerance of force inaccuracy when using ALMO DFT methods.
 
@@
 \caption{\label{fig:dynproperties} Calculated dynamic and static properties of water using ALMO DFT with $\epsilon_{SCF} = 10^{-2}$ and $R_{c} = 1.6$ VdWR.
-(a) Radial distribution function of an ALMO simulation compared to OT SCF with fully converged PES. 
-(b) Kinetic Energy distribution of an ALMO simulation compared to the theoretical Maxwell-Boltzmann kinetic energy distribution.
-(c) Calculated Infrared spectrum of an ALMO simulation compared to OT SCF with fully converged PES.
+(a) Radial distribution function of an ALMO simulation compared to OT SCF with a fully converged PES. 
+(b) Kinetic energy distribution of an ALMO simulation compared to the theoretical Maxwell-Boltzmann kinetic energy distribution.
+(c) Calculated Infrared spectrum of an ALMO simulation compared to OT SCF with a fully converged PES.
 Calculations were performed with the PBE $\mathrm{E_{XC}}$ functional and TZV2P basis set, using 125 water molecules for 10 ps.}
@@
 To summarize, a low-cost linear-scaling AIMD method for weakly-interacting molecular systems was developed by utilizing compact molecular orbitals strictly localized within predefined radii.
-High efficiency of the method is achieved without sacrificing its accuracy with a combination of two techniques: (1) on-the-fly construction of accurate localized orbitals without lengthy self-consistent optimization and (2) a modified Langevin equation that is fine-tuned to compensate retain stable dynamics even with imperfect forces.
-A remarkable feature of the implemented method is that it remains efficient even for challenging condensed phase systems, even with diffuse atom-centered basis sets.
+High efficiency of the method is achieved without sacrificing its accuracy with a combination of two techniques: (1) on-the-fly construction of accurate localized orbitals without lengthy self-consistent optimization and (2) a modified Langevin equation that is fine-tuned to retain stable dynamics even with imperfect forces.
+A remarkable feature of the implemented method is that it remains efficient for challenging condensed phase systems, even with diffuse atom-centered basis sets.
 Using liquid water as an example, it is demonstrated that the new AIMD method enables simulations of molecular systems on previously inaccessible time and length scales.

### 2017-08-19 almo-langevin.tex
delta f is introduces, questions for first figure

@@
 %However, the variational optimization of the DM is very inefficient for accurate DFT calculations that require many basis functions per atom~\cite{a:ls-rev-1999,a:ls-dm-sign}. 
-However, in accurate DFT calculations that require many basis functions per atom, the variational optimization of the DM formulation is advantageous only for impractically large systems~\cite{a:ls-rev-1999,a:ls-dm-sign,a:almo-ls}.
+However, in accurate DFT calculations that require many basis functions per atom, the variational optimization of the DM is advantageous only for impractically large systems~\cite{a:ls-rev-1999,a:ls-dm-sign,a:almo-ls}.
 %However, in accurate DFT calculations with many basis functions per atom are required, the DM becomes sufficiently sparse only for impractically large systems~\cite{a:ls-rev-1999,a:ls-dm-sign,a:almo-ls}.
@@
 Although such methods have been successfully used for the evaluation of accurate DFT energies of large static systems~\cite{a:ls-onetep-2009,a:ls-conquest-2010,a:ls-onetep-2010-app1,a:ls-rev-2012} their use in AIMD is hampered by the costly optimization of both the contracted orbitals and the density matrix.
-
+%
 Alternatively, LS can be achieved using a set of local electronic descriptors constructed by enforcing strict locality on the KS orbitals and, at the same time, by relaxing the orthogonality constraints imposed on them.
@@
 
+%ZZZ-important addition: we did not include a large family of the so-called fragmentation methods 
+%included only those methods that attempt to construct  set 
+%It is important to note all LS DFT methods reviewed above attempt to obtain the electronic description that do not construct  based on fragmentation methods that do that there  LS DFT methods the presented logical partitioning with the subsequent construction of localized MOs is a rather general approach used in a large number of electronic structure theories, which are collectively known as fragmentation methods~\cite{a:fragmentation-rev}.
+These methods vary greatly in how the partitioning and recombination are performed and, therefore, differ in accuracy and computational cost.
+We would like to emphasize that in our approach a proper quantum mechanical description of the entire system is constructed in the form of the total idempotent density matrix.
+This approach provides the most rigorous description of the electronic structure.
+
 %Thus, despite remarkable recent progress in linear scaling density function theory, the computational cost of existing methods remains too high for routine \emph{ab initio} molecular dynamics (AIMD) simulations.
 Thus, despite remarkable recent progress the high computational overhead of existing asymptotically linear-scaling DFT methods restrict their use in AIMD simulations to very short \emph{time scales}, systems of low dimensions, or problems that can be adequately described with minimal basis sets.
-Otherwise, LS DFT cannot compete with the straightforward cubically scaling KS DFT on the length scales accessible in routine AIMD.
+Otherwise, LS DFT cannot compete with the straightforward cubically scaling KS DFT on the length scales accessible in routine AIMD simulations.
 %Thus,  the computational cost of the existing asymptotically linear-scaling DFT methods remains too high to compete with the straightforward cubically scaling KS DFT on the length scales practically accessible in routine AIMD simulations.
@@
 
-In this work, we introduce a new AIMD method for weakly-interacting molecular systems that can reach the linear scaling regime for systems of medium size while maintaining a low computational overhead.
+In this work, we introduce a new AIMD method for weakly-interacting molecular systems that is not only linear scaling but also has a low computational overhead. 
 Using liquid water as an example, we demonstrate that the proposed AIMD methodology enables computational studies of molecular systems on previously inaccessible time and length scales.
 
-ALMO DFT exploits non-orthogonal molecular orbitals (MOs) in a two-step self-consistent filed (SCF) approach that is designed to avoid the usual convergence issues.~\cite{a:ls-mauri-galli-car-1993,a:ls-ordejon-1995,a:ls-fattebert-2004,a:ls-rev-1999}
-We show that due to its linear scaling nature and relatively low computational prefactor, ALMO-based DFT calculations allow for simulation of very large weakly-coupled molecular systems. 
-To decrease the cost even further, the present paper builds on the ALMO SCF method as well as other recent work involving Brownian dynamics to balance force calculation errors.~\cite{a:ls-parinello,a:2ndcpmd,a:langevin-why}
-We will show that our approach lowers the computation cost pre-factor for the ALMO SCF method, while maintaining LS and accuracy for calculation of dynamical properties. 
-This is done by balancing a new computationally economical but systemically inaccurate force calculation method with an optimized version of the Langevin Equation of motion. 
-We also introduce an easy to follow heuristic approach to optimize a system for generating correct dynamical behaviour.
+The new AIMD method utilizes a recently developed LS DFT~\cite{a:almo-ls} based on absolutely localized molecular orbitals (\mbox{ALMOs}). 
+Unlike delocalized KS orbitals, each \mbox{ALMO} has its own localization center -- atomic orbitals of a molecule are used here -- and a predefined localization radius that typically includes localization centers of neighbor molecules~\cite{a:stoll,a:almo-ls}.
+The key feature of ALMO DFT is that its trial wavefunctions are constructed to circumvent the problem of the sluggish variational optimization that have rendered many previous MO-based LS methods impractical.
+The trial ALMOs are optimized variationally in a two-step self-consistent-field (SCF) procedure~\cite{a:almo-ls}. 
+In the first step, ALMOs are constrained to their localization centers~\cite{a:khal} whereas, in the second step, ALMOs are relaxed to allow delocalization onto the neighbor molecules within their localization radius $R_{c}$.
+To achieve a robust optimization in the problematic second step, it is important to keep the delocalization component of the trial wavefunction orthogonal to the fixed orbitals obtained in the first step.
+For mathematical details, see ALMO SCF method in Ref.~\citenum{a:almo-ls}.
 
-%\section{Theory}
+ALMO constraints restrict/prohibit/forbid electron density transfer between distant molecules but retain all other types of interaction such as long-range electrostatic, exchange, polarization, and, if the exchange-correlation (XC) functional includes it, dispersion interactions~\cite{a:theeda}.
+Since the energetic contribution of the electron transfer decays exponentially with distance, \mbox{ALMO} approximation provides a natural and accurate representation of the electronic structure of molecular systems with just [a?] few local variables if an appropriate, that is sufficiently large, localization radius is chosen.
+%The advantage of ALMO DFT is that its computational complexity is linear with respect to system size because long-range electron delocalization is ignored in both optimization steps.
+%Because of the greatly reduced number of electronic variables and their robust optimization the computational cost of the implemented ALMO DFT is so low that it is [has the potential] capable of extending the range of accurate simulations to systems of thousands of molecules.
+Because of the greatly reduced number of electronic descriptors and the robust optimization, the computational complexity of ALMO DFT is linear with the number of molecules while its computational overhead remains very low.
+These features make ALMO DFT a promising method to perform accurate AIMD simulations of large molecular systems.
 
-Our approach is based on the ALMO SCF two-step variational method.~\cite{a:almo-ls} 
-Initially, the electrons are partitioned into groups based on a logical approach, generally corresponding to molecules. 
-In the first step, the basis atomic orbitals (AOs) are constrained to their parent molecules and a variational optimization is performed under the Born-Oppenheimer approximation using the DFT-SCF method within the non-overlapping domains.
-This is akin to calculating the electronic structure of each molecule (with frozen nuclei) within a coulombic field produced by the surrounding molecules. 
-Charge transfer is completely constrained. 
-In the second step, the variationally optimized but constrained MOs are allowed to delocalize onto nearby groups up to a desired cutoff radius $R_{c}$. 
-This step requires a second variational optimization to be performed. 
-Because of the logical partitioning of the groups and ignoring of long range forces, both steps of the electronic structure calculation are linear scaling with respect to system size. 
-For more details on this method, see ref.\ \citenum{a:almo-ls}.
+%To decrease the cost even further, the present paper builds on the ALMO SCF method as well as other recent work involving Brownian dynamics to compensate errors in the forces~\cite{a:ls-parinello,a:2ndcpmd,a:langevin-why}.
+%We will show that our approach lowers the computation cost pre-factor for the ALMO SCF method, while maintaining LS and accuracy for calculation of dynamical properties. 
+%This is done by balancing a new computationally economical but systemically inaccurate force calculation method with an optimized version of the Langevin Equation of motion. 
+%We also introduce an easy to follow heuristic approach to optimize a system for generating correct dynamical behaviour.
 
-It is important to note that the presented logical partitioning with the subsequent construction of localized MOs is a rather general approach used in a large number of electronic structure theories, which are collectively known as fragmentation methods~\cite{a:fragmentation-rev}.
-These methods vary greatly in how the partitioning and recombination are performed and, therefore, differ in accuracy and computational cost.
-We would like to emphasize that in our approach a proper quantum mechanical description of the entire system is constructed in the form of the total idempotent density matrix:
+To adapt ALMO DFT for AIMD, the forces on atoms were computed \emph{approximately} using a straightforward procedure and the idempotent DM obtained from ALMOs (see the Supplemental Material). 
+The difference between these forces and the perfect forces from perfectly-converged fully-delocalized KS orbitals is denoted $\delta f_{i\alpha}(t)$ 
 %
-\begin{eqnarray}
-\label{eq:properdm}
-\hat{R} = \sum_{x} \vert \psi^{xi} \rangle \langle \psi_{xi}\vert.
-\end{eqnarray}
+\begin{align}
+\label{eq:assumption}
+f^{\text{KS}}_{i\alpha}(t) = f^{\text{ALMO}}_{i\alpha}(t) + \delta f_{i\alpha} (t),
+\end{align}
 %
-This approach provides the most rigorous description of the electronic structure.
+where $\alpha$ is a Cartesian component of the force acting on atom $i$ at time step $t$.
+$\delta f_{i\alpha} (t)$ comprises neglected terms that originate from %fixed terms in the trial ALMOs (i.e.
+orbitals obtained in the first SCF step are not re-optimized in the second step),
+a finite localization radius $R_c$ and incomplete SCF optimization of the electronic degrees of freedom.
+$\delta f_{i\alpha} (t)$ can be reduced \emph{systematically} to zero by increasing $R_c$ (Figure~\ref{fig:forcecomp} in the Supplemental Material) and, at the same time, decreasing $\epsilon_{SCF}$ -- the maximum norm of the electronic energy gradient, below which the electronic ground state is considered sufficiently close to the Born-Oppenheimer state.
 
-A known source of error in force calculation within the ALMO SCF paradigm occurs due to the discontinuous jump in potential energy when a neighboring molecule passes beyond the localization radius $R_{c}$ and is suddenly ignored in potential energy calculations.
+%A known source of error in force calculation within the ALMO SCF paradigm occurs due to the discontinuous jump in potential energy when a neighboring molecule passes beyond the localization radius $R_{c}$ and is suddenly ignored in potential energy calculations.
 Setting $R_{c}$ to a larger value reduces this discontinuity, but greatly increases the computational cost as longer distance (and less important) interactions must be calculated. 
@@
 Previously, this meant that $R_{c}$ had to be kept relatively large to maintain good accuracy of dynamical properties.
+
+\begin{figure}
+\includegraphics[trim={0.1cm 0cm 0.2cm 0.1cm},clip,width=8.6cm]{DeltaForceComparison_ALMO_SCF.eps}
+\caption{\label{fig:forcecomp} {\color{red} All figures: VdWR$\rightarrow$vdWR.
+Hayden, your description states that you average over Cartesian components of forces.
+Unfortunately, averaging over components will give zero for any isotropic system observed for a sufficiently long time (i.e.
+for every component there will be equal but opposite component).
+What we would like to plot is the instantaneous magnitude of the vector force on atom/molecule, $\vert \delta \vec{f}_i(t) \vert$, averaged over all atoms/molecules.} Dependence of the instantaneous magnitude of $\vert \delta \vec{f}_i(t) \vert$ on the ALMO localization radius $R_c$.
+Forces are averaged over all atoms/molecules, $R_c$ is given in units of atomic van der Waals radii.
+%calculated as the average difference between fully converged Born-Oppenheimer forces and fully converged ALMO DFT forces, averaged over all dimensions and all molecules for each time step.
+Plotted are the random forces with varying localization radius $R_{c}$, .
+%Here it is clear that ALMO DFT forces converge to Born-Oppenheimer forces in the limit of $R_{c} \rightarrow \infty$.
+Forces are computed with the PBE XC functional, TZV2P basis set, and $\epsilon_{SCF} =ZZZ$ for configurations from a 10-ps trajectory generated with KS DFT at T=298~K.}
+\end{figure}
 
@@
 \label{eq:langevin}
-m_i \ddot{r}_{i\alpha} = f^{\text{SCF}}_{i\alpha} - \gamma m_i \dot{r}_{i\alpha} + R^{\gamma}_{i\alpha} (t),
+m_i \ddot{r}_{i\alpha} = f^{\text{KS}}_{i\alpha} - \gamma m_i \dot{r}_{i\alpha} + R^{\gamma}_{i\alpha} (t),
 \end{align}
@@
 Where $m_i$ refers to the mass of the i-th particle, and $r_{i\alpha}$ its position along dimension $\alpha$. 
-$f^{\text{SCF}}_{i\alpha}$ is the force calculated using the ALMO SCF method described above for the i-th particle along dimension $\alpha$, $\gamma$ is the Langevin scaling factor which sets the strength of both the damping term  $\gamma m_i \dot{r}_{i\alpha}$ and the stochastic Gaussian ``white noise'' term $R^{\gamma}_{i\alpha} (t)$.
+$f^{\text{KS}}_{i\alpha}$ is the force calculated using the ALMO SCF method described above for the i-th particle along dimension $\alpha$, $\gamma$ is the Langevin scaling factor which sets the strength of both the damping term  $\gamma m_i \dot{r}_{i\alpha}$ and the stochastic Gaussian ``white noise'' term $R^{\gamma}_{i\alpha} (t)$.
 In order to correctly sample the canonical distribution, the random force must be zero mean. 
@@
 \label{eq:assumption}
-f^{\text{SCF}}_{i\alpha} = f^{\text{APP}}_{i\alpha} + R^{\Delta}_{i\alpha} (t)
+f^{\text{KS}}_{i\alpha} = f^{\text{ALMO}}_{i\alpha} + R^{\Delta}_{i\alpha} (t)
 \end{align}
 %
-where $f^{\text{SCF}}_{i\alpha}$ is the force calculated from an accurate and fully converged SCF method, $f^{\text{APP}}_{i\alpha}$ is the approximate force calculated by allowing a large $\epsilon_{SCF}$, and $R^{\Delta}_{i\alpha} (t)$ is a stochastic ``white noise'' term that obeys
+where $f^{\text{KS}}_{i\alpha}$ is the force calculated from an accurate and fully converged SCF method, $f^{\text{ALMO}}_{i\alpha}$ is the approximate force calculated by allowing a large $\epsilon_{SCF}$, and $R^{\Delta}_{i\alpha} (t)$ is a stochastic ``white noise'' term that obeys
 %
@@
 \label{eq:langevin2}
-m_i \ddot{r}_{i\alpha} = f^{\text{APP}}_{i\alpha} - \gamma m_i \dot{r}_{i\alpha} + R^{\gamma + \Delta}_{i\alpha} (t)
+m_i \ddot{r}_{i\alpha} = f^{\text{ALMO}}_{i\alpha} - \gamma m_i \dot{r}_{i\alpha} + R^{\gamma + \Delta}_{i\alpha} (t)
 \end{align}
@@
 
-\begin{figure}
-\includegraphics[trim={0.1cm 0cm 0.2cm 0.1cm},clip,width=8.6cm]{DeltaForceComparison_ALMO_SCF.eps}
-\caption{\label{fig:forcecomp}The ``Random force'' $\langle R^{\Delta}_{i\alpha} \rangle (t)$ calculated as the average difference between fully converged Born-Oppenheimer forces and fully converged ALMO DFT forces, averaged across all dimensions and all molecules separately at each time step.
-Plotted are the random forces with varying localization radius $R_{c}$, in units of Van der Waals Radii.
-Here it is clear that ALMO DFT forces converge to Born-Oppenheimer forces in the limit of $R_{c} \rightarrow \infty$.
-Calculations were performed with the PBE $\mathrm{E_{XC}}$ functional and the TZV2P basis set on 125 water molecules for 10 ps.}
-\end{figure}
-
 To test the accuracy of our unconverged ALMO based MD we computed the Infrared (IR) spectrum, radial distribution function (RDF), kinetic energy distribution, and mean-squared displacement (MSD) function from a 10 ps MD trajectory (see fig.
-\ref{fig:dynproperties} and supporting information).
+\ref{fig:dynproperties} and the Supplemental Material).
 We compared the IR spectrum, RDF, and MSD function to corresponding reference functions calculated with fully converged KS DFT forces.
@@
 
-%\textbf{Supporting Information}
+%\textbf{Supplemental Material}
+
+%ZZZ outline procedure to compute forces using the DM obtained from ALMOs.
+% ZZZ figure to show
 

### 2017-08-20 almo-langevin.tex
Work in progress, some figre modification are required

@@
 
-To adapt ALMO DFT for AIMD, the forces on atoms were computed \emph{approximately} using a straightforward procedure and the idempotent DM obtained from ALMOs (see the Supplemental Material). 
-The difference between these forces and the perfect forces from perfectly-converged fully-delocalized KS orbitals is denoted $\delta f_{i\alpha}(t)$ 
+To adapt ALMO DFT for AIMD, the forces on atoms were computed \emph{approximately} using the idempotent DM obtained from ALMOs in a straightforward procedure that computes only the Hellmann-Feynman and Pulay forces while ignoring non-SCF forces (see the Supplemental Material).
+The difference between these forces and the exact forces obtained using perfectly-converged fully-delocalized KS orbitals is denoted $\delta f_{i\alpha}(t)$ 
 %
 \begin{align}
-\label{eq:assumption}
+\label{eq:deltaf}
 f^{\text{KS}}_{i\alpha}(t) = f^{\text{ALMO}}_{i\alpha}(t) + \delta f_{i\alpha} (t),
@@
 a finite localization radius $R_c$ and incomplete SCF optimization of the electronic degrees of freedom. 
-$\delta f_{i\alpha} (t)$ can be reduced \emph{systematically} to zero by increasing $R_c$ (Figure~\ref{fig:forcecomp} in the Supplemental Material) and, at the same time, decreasing $\epsilon_{SCF}$ -- the maximum norm of the electronic energy gradient, below which the electronic ground state is considered sufficiently close to the Born-Oppenheimer state.
+$\delta f_{i\alpha} (t)$ can be reduced \emph{systematically} to zero by increasing $R_c$ (Figure~\ref{fig:forcecomp} in the Supplemental Material) and, at the same time, decreasing $\epsilon_{\text{SCF}}$ -- the infinity norm of the electronic gradient of the \mbox{ALMO} energy, below which the electronic ground state is considered sufficiently close to the Born-Oppenheimer state.
+%ZZZ For $\epsilon_{\text{SCF}} \leqslant 10^{-7}$~a.u. ALMO SCF is considered converged.
 
-%A known source of error in force calculation within the ALMO SCF paradigm occurs due to the discontinuous jump in potential energy when a neighboring molecule passes beyond the localization radius $R_{c}$ and is suddenly ignored in potential energy calculations.
-Setting $R_{c}$ to a larger value reduces this discontinuity, but greatly increases the computational cost as longer distance (and less important) interactions must be calculated. 
-As long as the discontinuity in force is small compared with the random error of the calculation, it can be safely ignored.
-Previously, this meant that $R_{c}$ had to be kept relatively large to maintain good accuracy of dynamical properties.
-
-\begin{figure}
+\begin{figure}[t]
 \includegraphics[trim={0.1cm 0cm 0.2cm 0.1cm},clip,width=8.6cm]{DeltaForceComparison_ALMO_SCF.eps}
-\caption{\label{fig:forcecomp} {\color{red} All figures: VdWR$\rightarrow$vdWR.
-Hayden, your description states that you average over Cartesian components of forces.
-Unfortunately, averaging over components will give zero for any isotropic system observed for a sufficiently long time (i.e.
-for every component there will be equal but opposite component).
-What we would like to plot is the instantaneous magnitude of the vector force on atom/molecule, $\vert \delta \vec{f}_i(t) \vert$, averaged over all atoms/molecules.} Dependence of the instantaneous magnitude of $\vert \delta \vec{f}_i(t) \vert$ on the ALMO localization radius $R_c$.
-Forces are averaged over all atoms/molecules, $R_c$ is given in units of atomic van der Waals radii.
+\caption{\label{fig:forcecomp} {\color{red} All figures: $\epsilon_{\text{SCF}}$ must have ``a.u.'' units.
+This figure: VdWR$\rightarrow$vdWR.
+Hayden, you state that you average over Cartesian components of forces.
+Unfortunately, averaging over Cartesian components must give zero for any sufficiently large isotropic system (i.e.
+for every component there will be equal  opposite component).
+It would be better to plot the instantaneous \emph{magnitude} of $\delta \vec{f}_i(t)$, averaged over all atoms/molecules $\langle \ldots \rangle_i$.
+Can you make a new plot?
+Replace legend with: $\langle \delta f_i(t) \rangle_i$ -- red, $\langle \delta f_i(t) \rangle_{it}$ -- black, $\sqrt{\langle \delta \vec{f}_i(t) \cdot \delta \vec{f}_i(t) \rangle_{it}}$ -- green.
+Or remove it completely.
+I suggest we plot two green lines: above and below the average.} Dependence of the Euclidean norm of the instantaneous error $\langle \delta f_i(t) \rangle_i$ (red line) on the ALMO localization radius $R_c$ expressed in units of atomic van der Waals radii.
+Black line is the time average of the instantaneous error, green line is the standard deviation from the average.
 %calculated as the average difference between fully converged Born-Oppenheimer forces and fully converged ALMO DFT forces, averaged over all dimensions and all molecules for each time step.
-Plotted are the random forces with varying localization radius $R_{c}$, .
-%Here it is clear that ALMO DFT forces converge to Born-Oppenheimer forces in the limit of $R_{c} \rightarrow \infty$.
-Forces are computed with the PBE XC functional, TZV2P basis set, and $\epsilon_{SCF} =ZZZ$ for configurations from a 10-ps trajectory generated with KS DFT at T=298~K.}
+Forces are computed with the PBE XC functional, TZV2P basis set, and $\epsilon_{\text{SCF}} = 10^{-7}$~a.u. for the configurations from a 10-ps trajectory generated with KS-based AIMD at T=298~K.}
 \end{figure}
 
-In the present article, we propose a new approach which allows for relaxation of the constraint on $R_{c}$ by introducing related damping and stochastic terms to the force calculation by utilizing the Langevin equation of motion.~\cite{a:Kubo-1986}
+%For any finite value of $R_{c}$, forces are undefined at manifold of due to the discontinuous jump in potential energy when a neighboring molecule passes beyond the localization radius $R_{c}$ and is suddenly ignored in potential energy calculations.
+%Setting $R_{c}$ to larger values can reduce the magnitude of the discontinuity, but increases the computational cost as longer distance (but physically less important) interactions must be calculated.
+%As long as the discontinuity in force is small compared with the random numerical noise in the calculation, it can be safely ignored~\cite{ZZZ-any-references}.
+This means that $R_{c}$ has to be kept relatively large to maintain stable and accurate dynamics.
 
+In order to maintain stable and accurate dynamics, $R_{c}$ has to be larger and $\epsilon_{\text{SCF}}$ has to be smaller than the corresponding optimal values used in static energy calculations or Monte-Carlo simulations.
+This, however, greatly increases the computational overhead of the simulation as longer-distance but physically less important interactions must be calculated. 
+Therefore, in this work, we propose a different approach to ALMO-based AIMD, which allows for relaxation of the tight constraints on $R_{c}$ and $\epsilon_{\text{SCF}}$ and great reduction of the simulation cost.
+The approach is inspired by ZZZ and requires the use of the Langevin equations of motion~\cite{a:Kubo-1986}
+%
 \begin{align}
@@
 \end{align}
-
-Where $m_i$ refers to the mass of the i-th particle, and $r_{i\alpha}$ its position along dimension $\alpha$. 
-$f^{\text{KS}}_{i\alpha}$ is the force calculated using the ALMO SCF method described above for the i-th particle along dimension $\alpha$, $\gamma$ is the Langevin scaling factor which sets the strength of both the damping term  $\gamma m_i \dot{r}_{i\alpha}$ and the stochastic Gaussian ``white noise'' term $R^{\gamma}_{i\alpha} (t)$.
-In order to correctly sample the canonical distribution, the random force must be zero mean. 
-As well, the damping and white noise terms must be related by the second fluctuation dissipation theorem.~\cite{a:Kubo-1986,a:langevin-why,b:tuckerman-stat}
+%
+where $m_i$ is the mass of atom $i$, $r_{i\alpha}$ its position along dimension $\alpha$, $\gamma$ is the Langevin scaling factor.
+% which defines the strength of both the damping term and the stochastic term $R^{\gamma}_{i\alpha} (t)$. 
+Langevin thermostat.
+In order to sample the canonical distribution correctly, the stochastic force must have zero mean and, according to the second fluctuation dissipation theorem~\cite{a:Kubo-1986,b:tuckerman-stat} must be related to the damping through $\gamma$:
 %
@@
 %
-This means the damping and white noise terms will work to maintain a constant temperature while also generating the correct Maxwell-Boltzmann distribution of velocities.
+These equations mean that, for any value of $\gamma$, 
+Thus the perfectly balanced damping and stochastic terms, defined in the equations above, will maintain a constant temperature while also generating the correct Maxwell-Boltzmann distribution of velocities.
+In the limit $\gamma \rightarrow 0$, the Newton equations of motion are recovered and the microcanonical ensemble is sampled.
+Compromise between small and large $\gamma$: It has previously been shown that Langevin dynamics can be utilized to correct inaccurate force calculations to maintain correct dynamical properties.~\cite{a:langevin-why,a:2ndcpmd,b:tuckerman-stat,a:ceriotti}
 
-It has previously been shown that Langevin dynamics can be utilized to correct inaccurate force calculations to maintain correct dynamical properties.~\cite{a:langevin-why,a:2ndcpmd,b:tuckerman-stat,a:ceriotti}
 By utilizing Eq.\ (\ref{eq:langevin}) in conjunction with the ALMO SCF method of force calculation, we will show that discontinuity in the PES can be ignored for sufficiently large $\gamma$. 
-Note that increasing $\gamma$ increases the sample size required to generate accurate dynamical properties, as more random Gaussian noise is introduced.
+% Perhaps this statement is not correct: Note that increasing $\gamma$ increases the sample size required to generate accurate dynamical properties, as more random Gaussian noise is introduced.
 
-The concept of balancing errors in force calculation with the Langevin equation can be taken even further, thereby reducing computational overhead even more. 
-During the second step of the ALMO SCF method described above, a self-consistent variational approach is taken to converge the non-orthogonal MOs. 
-Generally, this step utilizes a iterative search function in order to reduce the absolute norm of the energy gradient to some target value $\epsilon_{SCF}$. 
-Previously, $\epsilon_{SCF}$ had to be set very small to ensure sufficient convergence of the electron structure calculations. 
-By increasing the allowable $\epsilon_{SCF}$, a great reduction in computational cost can be achieved. 
-Obviously, this introduces a significant error into the calculation of forces during each time step. 
-We assume that this error has the form:
+%The concept of balancing errors in force calculation with the Langevin equation can be taken even further, thereby reducing computational overhead even more. 
+%%During the second step of the ALMO SCF method described above, a self-consistent variational approach is taken to converge the non-orthogonal MOs. 
+%Generally, this step utilizes a iterative search function in order to reduce the absolute norm of the energy gradient to some target value $\epsilon_{SCF}$. 
+%Previously, $\epsilon_{SCF}$ had to be set very small to ensure sufficient convergence of the electron structure calculations. 
+%By increasing the allowable $\epsilon_{SCF}$, a great reduction in computational cost can be achieved. 
+%Obviously, this introduces a significant error into the calculation of forces during each time step.
+
+The main assumption of ALMO AIMD is that the error in the ALMO forces is well approximated by a stochastic ``white noise'' term $R^{\Delta}_{i\alpha} (t)$
 %
@@
 \label{eq:assumption}
-f^{\text{KS}}_{i\alpha} = f^{\text{ALMO}}_{i\alpha} + R^{\Delta}_{i\alpha} (t)
+\delta f_{i\alpha} = R^{\Delta}_{i\alpha} (t)
 \end{align}
 %
-where $f^{\text{KS}}_{i\alpha}$ is the force calculated from an accurate and fully converged SCF method, $f^{\text{ALMO}}_{i\alpha}$ is the approximate force calculated by allowing a large $\epsilon_{SCF}$, and $R^{\Delta}_{i\alpha} (t)$ is a stochastic ``white noise'' term that obeys
+that obeys
 %
@@
 %
-It's clear from figure \ref{fig:randomforce} that this assumption is sufficiently justified. 
-Hence, the Langevin equation of motion (\ref{eq:langevin}) can be re-written with the two stochastic terms combined into one:
+[ZZZ later?: Figure~\ref{fig:randomforce} shows that this assumption is sufficiently justified for the case of liquid water.] 
+If this assumption holds, the Langevin equations of motion (\ref{eq:langevin}) can be re-written with the two stochastic terms combined into one:
 %
@@
 \includegraphics[trim={0.6cm 0.7cm 0.7cm 0.3cm},clip,width=8.6cm]{DeltaForceComparison_with_ACF.eps}
-\caption{\label{fig:randomforce}The ``Random force'' along a water molecule dimension $\langle R^{\Delta}_{i\alpha} \rangle (t)$ calculated as the difference between the fully converged Born-Oppenheimer forces and ALMO DFT forces.
-Here it is clear that Eq.\ (\ref{eq:stochastic2}) is satisfied.
-(a) Plot of $\langle R^{\Delta}_{i\alpha} \rangle (t)$ as a function of time, where $R_{C} = 1.6$ VdWR and $\epsilon_{SCF} = 10^{-2}$.
-(b) Plot of $\langle R^{\Delta}_{i\alpha} \rangle (t)$ as a function of time, where $R_{C} = 1.6$ VdWR and forces are fully converged.
-(c) Autocorrelation function of plot a, $\langle R^{\Delta}_{i\alpha} (t) R^{\Delta}_{i\alpha} (t+\tau) \rangle $. 
-Calculations were performed with the PBE $\mathrm{E_{XC}}$ functional and the TZV2P basis set on 125 water molecules for 10 ps at 298 K.}
+\caption{\label{fig:randomforce} {\color{red} The top two panels should be changed to plot the Euclidean norm of the error as described in the notes to the previous Figure.
+Do not forget to change the legend.
+For the ACF, try vector autocorrelation (see this caption below).
+If it is ugly, try simple scalar autocorrelation.
+Do not forget to subtract the non-zero mean when computing the autocorrelation.
+Plot (c), new x-axis label: ``Time lag $\tau$ (fs)''.}
+Red line is the Euclidean norm of the instantaneous error $\langle \delta f_i(t) \rangle_i$, black line is the time average of the instantaneous error, green line is the standard deviation from the average.
+Forces are computed with the PBE XC functional and TZV2P basis set for the configurations from a 10-ps trajectory generated with KS-based AIMD at T=298~K.
+% Here it is clear that Eq.\ (\ref{eq:stochastic2}) is satisfied. 
+(a) $R_{c} = 1.6$ vdWR and $\epsilon_{SCF} = 10^{-2}$, 
+(b) $R_{c} = 1.6$ vdWR and fully converged ALMO SCF,
+(c) Autocorrelation function of the instantaneous error in panel (a) $\langle \vec{f}_i (t) \cdot \vec{f}_i(t+\tau) \rangle_t $.
+}
 \end{figure}
@@
 \includegraphics[trim={1.3cm 0.1cm 3.3cm 1.3cm},clip,width=8.6cm]{Dynamical_Data_Tiled.eps}
-\caption{\label{fig:dynproperties} Calculated dynamic and static properties of water using ALMO DFT with $\epsilon_{SCF} = 10^{-2}$ and $R_{c} = 1.6$ VdWR.
+\caption{\label{fig:dynproperties} Calculated dynamic and static properties of water using ALMO DFT with $\epsilon_{SCF} = 10^{-2}$ and $R_{c} = 1.6$ vdWR.
 (a) Radial distribution function of an ALMO simulation compared to OT SCF with a fully converged PES. 
@@
 To show the computational efficiency gained by using unconverged ALMO DFT forces, we compare the average wall-time per MD step over 10 time steps for various values of $\epsilon_{SCF}$ in figure \ref{fig:strongscaling_log}.
-For these simulations, we chose to set the localization radius $R_{c} = 1.6$ Van der Waal Radii (VdWR), as was done for calculation of dynamical properties.
+For these simulations, we chose to set the localization radius $R_{c} = 1.6$ van der Waals radii (vdWR), as was done for calculation of dynamical properties.
 We also show timing of the fully converged OT SCF method, as well as the recently developed ``second generation Car-Parinello'' MD method, which improves upon the OT computational cost prefactor.~\cite{a:2ndcpmd}
@@
 Calculations were performed with the PBE $\mathrm{E_{XC}}$ functional and TZV2P basis set on 256 cores, averaged over 10 MD steps. 
-The localization radius was set to $R_{c} = 1.6$ VdWR.}
+The localization radius was set to $R_{c} = 1.6$ vdWR.}
 \end{figure}
@@
 %Calculations were performed with the PBE $\mathrm{E_{XC}}$ functional and TZV2P basis set on 256 cores. 
-%The localization radius was set to $R_{c} = 1.6$ VdWR.}
+%The localization radius was set to $R_{c} = 1.6$ vdWR.}
 %\end{figure}
@@
 Calculations were performed with the PBE $\mathrm{E_{XC}}$ functional and TZV2P basis set, averaged over 10 MD steps. 
-The localization radius was set to $R_{c} = 1.6$ VdWR and $\epsilon_{SCF} = 10^{-2}$.}
+The localization radius was set to $R_{c} = 1.6$ vdWR and $\epsilon_{SCF} = 10^{-2}$.}
 \end{figure}

### 2017-08-20 almo-langevin.tex
Detailed explanations on ACF

@@
 ALMO constraints restrict/prohibit/forbid electron density transfer between distant molecules but retain all other types of interaction such as long-range electrostatic, exchange, polarization, and, if the exchange-correlation (XC) functional includes it, dispersion interactions~\cite{a:theeda}.
-Since the energetic contribution of the electron transfer decays exponentially with distance, \mbox{ALMO} approximation provides a natural and accurate representation of the electronic structure of molecular systems with just [a?] few local variables if an appropriate, that is sufficiently large, localization radius is chosen.
+Since the energetic contribution of the electron transfer decays exponentially with distance, \mbox{ALMO} approximation provides a natural and accurate representation of the electronic structure of molecular systems with just [a?] few local variables if an appropriate localization radius is chosen.
 %The advantage of ALMO DFT is that its computational complexity is linear with respect to system size because long-range electron delocalization is ignored in both optimization steps.
@@
 %
-where $\alpha$ is a Cartesian component of the force acting on atom $i$ at time step $t$.
+where $\alpha$ is a Cartesian component of the force acting on atom $i$ at time $t$.
 $\delta f_{i\alpha} (t)$ comprises neglected terms that originate from %fixed terms in the trial ALMOs (i.e.
@@
 a finite localization radius $R_c$ and incomplete SCF optimization of the electronic degrees of freedom. 
-$\delta f_{i\alpha} (t)$ can be reduced \emph{systematically} to zero by increasing $R_c$ (Figure~\ref{fig:forcecomp} in the Supplemental Material) and, at the same time, decreasing $\epsilon_{\text{SCF}}$ -- the infinity norm of the electronic gradient of the \mbox{ALMO} energy, below which the electronic ground state is considered sufficiently close to the Born-Oppenheimer state.
+$\delta f_{i\alpha} (t)$ can be reduced \emph{systematically} to zero by increasing $R_c$ (Figure~\ref{fig:forcecomp} in the Supplemental Material) and decreasing $\epsilon_{\text{SCF}}$ -- [ZZZ the value maximum norm of the gradient of the \mbox{ALMO} DFT energy with respect to the ALMO coefficients.
+below which the electronic ground state is considered sufficiently close to the Born-Oppenheimer state.]
 %ZZZ For $\epsilon_{\text{SCF}} \leqslant 10^{-7}$~a.u. ALMO SCF is considered converged.
@@
 for every component there will be equal  opposite component).
+I am afraid these plots underestimate our error and mostly show finite-size effects.
 It would be better to plot the instantaneous \emph{magnitude} of $\delta \vec{f}_i(t)$, averaged over all atoms/molecules $\langle \ldots \rangle_i$.
@@
 Do not forget to change the legend.
-For the ACF, try vector autocorrelation (see this caption below).
-If it is ugly, try simple scalar autocorrelation.
-Do not forget to subtract the non-zero mean when computing the autocorrelation.
-Plot (c), new x-axis label: ``Time lag $\tau$ (fs)''.}
+Plot (c), new x-axis label: ``Time lag $\tau$ (fs)''.
+As to the ACF, one should be extremely careful how the averaging is performed.
+This caption below shows the ACF we need.
+Let me briefly explain it.
+The main idea is that the correct ACF could be computed from only one Cartesian component of one molecule if we could observe it for really really long time and average over many many time origins.
+We perform averaging over components and molecules just to improve our statistics (it is a shame to discard so many data points).
+We, therefore, should not compute ACF from \emph{one} time series that is already averaged over all possible components.
+Instead we compute and ACF for each component and then average over these ACFs.
+Ok, here is the procedure.
+Fix the time lag $\tau$.
+For a given time origin $t$ and molecule $i$, compute: $p_{it}(\tau)=\frac{1}{3}\delta \vec{f}_i (t)\cdot \delta \vec{f}_i (t+\tau)$, $\frac{1}{3}$ averages over components.
+Now, average over time origins and molecules to get $\text{ACF}(\tau) = \langle p_{it}(\tau) \rangle_{it}$.
+I think we do not need to subtract the mean because you already showed that it is zero.
+I hope there is a function in MATLAB that can do this.}
 Red line is the Euclidean norm of the instantaneous error $\langle \delta f_i(t) \rangle_i$, black line is the time average of the instantaneous error, green line is the standard deviation from the average.
@@
 (b) $R_{c} = 1.6$ vdWR and fully converged ALMO SCF,
-(c) Autocorrelation function of the instantaneous error in panel (a) $\langle \vec{f}_i (t) \cdot \vec{f}_i(t+\tau) \rangle_t $.
+(c) Autocorrelation function of the instantaneous error in panel (a) $\frac{1}{3}\langle \delta \vec{f}_i (t) \cdot \delta\vec{f}_i(t+\tau) \rangle_{it} $.
 }

### 2017-08-31 almo-langevin.tex
Thicker lines
Lots of edits August 30th
Went through the whole article and edited all spots that needed it. Updated some data plots (removed italized SCF from epilon_SCF)

--- 2017-08-20 almo-langevin.tex
@@
 Since the unification of molecular dynamics and density functional theory (DFT)~\cite{a:thecpmd},
-\emph{ab initio} molecular dynamics (AIMD)~\cite{b:aimd} has become an important tool for studying of weakly bonded ensembles of molecules such as gas-phase molecular clusters, liquids, solutions, and molecular solids.
+\emph{ab initio} molecular dynamics (AIMD)~\cite{b:aimd} has become an important tool for the study of weakly bonded ensembles of molecules such as gas-phase molecular clusters, liquids, solutions, and molecular solids.
 %[of gas-phase molecules and condensed-phase systems.]
@@
 
-ALMO constraints restrict/prohibit/forbid electron density transfer between distant molecules but retain all other types of interaction such as long-range electrostatic, exchange, polarization, and, if the exchange-correlation (XC) functional includes it, dispersion interactions~\cite{a:theeda}.
-Since the energetic contribution of the electron transfer decays exponentially with distance, \mbox{ALMO} approximation provides a natural and accurate representation of the electronic structure of molecular systems with just [a?] few local variables if an appropriate, that is sufficiently large, localization radius is chosen.
+ALMO constraints prohibit electron density transfer between distant molecules but retain all other types of interaction such as long-range electrostatic, exchange, polarization, and, if the exchange-correlation (XC) functional includes it, dispersion interactions~\cite{a:theeda}.
+Since the energetic contribution of the electron transfer decays exponentially with distance, \mbox{ALMO} approximation provides a natural and accurate representation of the electronic structure of molecular systems with just a few local variables if an appropriate localization radius is chosen.
 %The advantage of ALMO DFT is that its computational complexity is linear with respect to system size because long-range electron delocalization is ignored in both optimization steps.
 %Because of the greatly reduced number of electronic variables and their robust optimization the computational cost of the implemented ALMO DFT is so low that it is [has the potential] capable of extending the range of accurate simulations to systems of thousands of molecules.
-Because of the greatly reduced number of electronic descriptors and the robust optimization, the computational complexity of ALMO DFT is linear with the number of molecules while its computational overhead remains very low.
+Because of the greatly reduced number of electronic descriptors and the robust optimization, the computational complexity of ALMO DFT grows linearly with the number of molecules while its computational overhead remains very low.
 These features make ALMO DFT a promising method to perform accurate AIMD simulations of large molecular systems.
@@
 a finite localization radius $R_c$ and incomplete SCF optimization of the electronic degrees of freedom. 
-$\delta f_{i\alpha} (t)$ can be reduced \emph{systematically} to zero by increasing $R_c$ (Figure~\ref{fig:forcecomp} in the Supplemental Material) and, at the same time, decreasing $\epsilon_{\text{SCF}}$ -- the infinity norm of the electronic gradient of the \mbox{ALMO} energy, below which the electronic ground state is considered sufficiently close to the Born-Oppenheimer state.
+$\delta f_{i\alpha} (t)$ can be reduced \emph{systematically} to zero by increasing $R_c$ (Figure~\ref{fig:forcecomp} in the Supplemental Material) and decreasing $\epsilon_{\text{SCF}}$ -- the maximum norm of the gradient of the \mbox{ALMO} DFT energy with respect to the ALMO coefficients.
+$\epsilon_{\text{SCF}}$ is chosen such that the electronic ground state is considered sufficiently close to the Born-Oppenheimer state for values less than $\epsilon_{\text{SCF}}$.
 %ZZZ For $\epsilon_{\text{SCF}} \leqslant 10^{-7}$~a.u. ALMO SCF is considered converged.
@@
 \begin{figure}[t]
-\includegraphics[trim={0.1cm 0cm 0.2cm 0.1cm},clip,width=8.6cm]{DeltaForceComparison_ALMO_SCF.eps}
-\caption{\label{fig:forcecomp} {\color{red} All figures: $\epsilon_{\text{SCF}}$ must have ``a.u.'' units.
-This figure: VdWR$\rightarrow$vdWR.
-Hayden, you state that you average over Cartesian components of forces.
-Unfortunately, averaging over Cartesian components must give zero for any sufficiently large isotropic system (i.e.
-for every component there will be equal  opposite component).
-It would be better to plot the instantaneous \emph{magnitude} of $\delta \vec{f}_i(t)$, averaged over all atoms/molecules $\langle \ldots \rangle_i$.
-Can you make a new plot?
-Replace legend with: $\langle \delta f_i(t) \rangle_i$ -- red, $\langle \delta f_i(t) \rangle_{it}$ -- black, $\sqrt{\langle \delta \vec{f}_i(t) \cdot \delta \vec{f}_i(t) \rangle_{it}}$ -- green.
-Or remove it completely.
-I suggest we plot two green lines: above and below the average.} Dependence of the Euclidean norm of the instantaneous error $\langle \delta f_i(t) \rangle_i$ (red line) on the ALMO localization radius $R_c$ expressed in units of atomic van der Waals radii.
-Black line is the time average of the instantaneous error, green line is the standard deviation from the average.
+\includegraphics[trim={0cm 0cm 0.1cm 0.1cm},clip,width=8.6cm]{DeltaForceComparison_ALMO_SCF.eps}
+\caption{\label{fig:forcecomp} Dependence of the Euclidean norm of the instantaneous error $\langle \| \delta \vec{f_{i}}(t) \| \rangle_{i}$ (red line) on the ALMO localization radius $R_c$ expressed in units of atomic van der Waals radii.
+Black line is the magnitude of the time average of the instantaneous error vector.
 %calculated as the average difference between fully converged Born-Oppenheimer forces and fully converged ALMO DFT forces, averaged over all dimensions and all molecules for each time step.
@@
 This, however, greatly increases the computational overhead of the simulation as longer-distance but physically less important interactions must be calculated. 
-Therefore, in this work, we propose a different approach to ALMO-based AIMD, which allows for relaxation of the tight constraints on $R_{c}$ and $\epsilon_{\text{SCF}}$ and great reduction of the simulation cost.
-The approach is inspired by ZZZ and requires the use of the Langevin equations of motion~\cite{a:Kubo-1986}
+Therefore, in this work we propose a different approach to ALMO-based AIMD, which allows for relaxation of the tight constraints on $R_{c}$ and $\epsilon_{\text{SCF}}$ and great reduction of the simulation cost.
+The approach is inspired by ``Second Generation Car-Parinello'' MD~\cite{a:2ndcpmd} and requires the use of the Langevin equations of motion~\cite{a:Kubo-1986}
 %
@@
 %
-where $m_i$ is the mass of atom $i$, $r_{i\alpha}$ its position along dimension $\alpha$, $\gamma$ is the Langevin scaling factor.
+where $m_i$ is the mass of atom $i$, $r_{i\alpha}$ its position along dimension $\alpha$, and $\gamma$ is the Langevin scaling factor.
 % which defines the strength of both the damping term and the stochastic term $R^{\gamma}_{i\alpha} (t)$. 
-Langevin thermostat.
 In order to sample the canonical distribution correctly, the stochastic force must have zero mean and, according to the second fluctuation dissipation theorem~\cite{a:Kubo-1986,b:tuckerman-stat} must be related to the damping through $\gamma$:
@@
 %
-These equations mean that, for any value of $\gamma$, 
-Thus the perfectly balanced damping and stochastic terms, defined in the equations above, will maintain a constant temperature while also generating the correct Maxwell-Boltzmann distribution of velocities.
+These equations mean that, for any value of $\gamma$, the perfectly balanced damping and stochastic terms, defined in the equations above, will maintain a constant temperature and also generate the correct Maxwell-Boltzmann distribution of velocities.
 In the limit $\gamma \rightarrow 0$, the Newton equations of motion are recovered and the microcanonical ensemble is sampled.
-Compromise between small and large $\gamma$: It has previously been shown that Langevin dynamics can be utilized to correct inaccurate force calculations to maintain correct dynamical properties.~\cite{a:langevin-why,a:2ndcpmd,b:tuckerman-stat,a:ceriotti}
+Larger values of $\gamma$ thermostat the system more strongly, but introduce more random noise necessitating longer trajectories when calculating system bulk properties.
+It has previously been shown that Langevin dynamics can be utilized to correct inaccurate force calculations to maintain correct dynamical properties.~\cite{a:langevin-why,a:2ndcpmd,b:tuckerman-stat,a:ceriotti}
 
-By utilizing Eq.\ (\ref{eq:langevin}) in conjunction with the ALMO SCF method of force calculation, we will show that discontinuity in the PES can be ignored for sufficiently large $\gamma$. 
+%By utilizing Eq.\ (\ref{eq:langevin}) in conjunction with the ALMO SCF method of force calculation, we will show that discontinuity in the PES can be ignored for sufficiently large $\gamma$. 
 % Perhaps this statement is not correct: Note that increasing $\gamma$ increases the sample size required to generate accurate dynamical properties, as more random Gaussian noise is introduced.
@@
 %%During the second step of the ALMO SCF method described above, a self-consistent variational approach is taken to converge the non-orthogonal MOs. 
-%Generally, this step utilizes a iterative search function in order to reduce the absolute norm of the energy gradient to some target value $\epsilon_{SCF}$. 
-%Previously, $\epsilon_{SCF}$ had to be set very small to ensure sufficient convergence of the electron structure calculations. 
-%By increasing the allowable $\epsilon_{SCF}$, a great reduction in computational cost can be achieved. 
+%Generally, this step utilizes a iterative search function in order to reduce the absolute norm of the energy gradient to some target value $\epsilon_{\text{SCF}}$. 
+%Previously, $\epsilon_{\text{SCF}}$ had to be set very small to ensure sufficient convergence of the electron structure calculations. 
+%By increasing the allowable $\epsilon_{\text{SCF}}$, a great reduction in computational cost can be achieved. 
 %Obviously, this introduces a significant error into the calculation of forces during each time step.
@@
 %
-[ZZZ later?: Figure~\ref{fig:randomforce} shows that this assumption is sufficiently justified for the case of liquid water.] 
-If this assumption holds, the Langevin equations of motion (\ref{eq:langevin}) can be re-written with the two stochastic terms combined into one:
+Figure~\ref{fig:randomforce} shows that this assumption is sufficiently justified for the case of liquid water. 
+Therefore, the Langevin equations of motion (\ref{eq:langevin}) can be re-written with the two stochastic terms combined into one:
 %
@@
 The relative weight of the three is determined by the specifics of the system and what is most important to the researcher. 
-To optimize $\gamma$ given the relative importance of the above properties, we suggest running several short trajectories on the system of interest (or a scaled-down version of it) using a tightly converged PES calculation and varying the value of $\gamma$.
+To optimize $\gamma$ given the relative importance of the above properties, we suggest running several short trajectories on the system of interest (or a scaled-down version of it) using a tightly converged potential energy surface calculation and varying the value of $\gamma$.
 We found that a value of $\gamma$ between $10^{-5}$ and $10^{-3}\ \mathrm{fs^{-1}}$ was optimal for the purposes of the research presented here.
@@
 To correctly thermostat the system of interest, $\Delta$ needs to be reduced to zero.
-This is done by finding the value of $\Delta$ and adding a third random force term $R^{-\Delta}_{i\alpha}$ to equation \ref{eq:langevin2} which cancels out the energy introduced to the system by $R^{\Delta}_{i\alpha}$.
+This is done by finding the value of $\Delta$ and adding a third random force term $R^{-\Delta}_{i\alpha}$ to equation \ref{eq:langevin2} which cancels out the extra forces introduced to the system by $R^{\Delta}_{i\alpha}$.
 This means that a trajectory with inaccurate forces and optimized $R^{-\Delta}_{i\alpha}$ should have no long term drift in the total energy of the system. 
-Short term energy fluctuations should be expected, and can later be ``smoothed out'' by setting finite $\gamma$.
+Short term energy fluctuations should be expected, and can later be ``smoothed out'' by setting $\gamma$ as needed.
 
@@
 The ACF was not possible to accurately integrate without running a reference trajectory with extremely short time-steps (see figure~\ref{fig:randomforce}).
-For example, using Eq.\ \ref{eq:delta} we calculated $\Delta \approx \SI{2e-5}{\per\fs}$, but value was optimized heuristically to $\SI{6e-5}{\per\fs}$.
+For example, using Eq.\ \ref{eq:delta} we calculated $\Delta \approx \SI{2e-5}{\per\fs}$, but the value was optimized heuristically to $\SI{6e-5}{\per\fs}$.
 Still, this method provided a decent order of magnitude starting point for further investigation.
@@
 \begin{figure}
-\includegraphics[trim={0.6cm 0.7cm 0.7cm 0.3cm},clip,width=8.6cm]{DeltaForceComparison_with_ACF.eps}
-\caption{\label{fig:randomforce} {\color{red} The top two panels should be changed to plot the Euclidean norm of the error as described in the notes to the previous Figure.
-Do not forget to change the legend.
-For the ACF, try vector autocorrelation (see this caption below).
-If it is ugly, try simple scalar autocorrelation.
-Do not forget to subtract the non-zero mean when computing the autocorrelation.
-Plot (c), new x-axis label: ``Time lag $\tau$ (fs)''.}
-Red line is the Euclidean norm of the instantaneous error $\langle \delta f_i(t) \rangle_i$, black line is the time average of the instantaneous error, green line is the standard deviation from the average.
+\includegraphics[trim={0.5cm 0cm 0.7cm 0.1cm},clip,width=8.6cm]{DeltaForceComparison_with_ACF.eps}
+\caption{\label{fig:randomforce} Comparison of random force introduced from $R_{c}$ and $\epsilon_{\text{SCF}}$. 
+The red line is the Euclidean norm of the instantaneous error $\langle \| \delta \vec{f_{i}}(t) \| \rangle_{i}$, the black line is the magnitude of the time average of the instantaneous error vector, and the green line is the standard deviation from the average.
+(a) $R_{c} = 1.6$ vdWR and $\epsilon_{\text{SCF}} = 10^{-2}$, using fully converged ALMO SCF reference.
+(b) $R_{c} = 1.6$ vdWR and fully converged ALMO SCF, using fully converged SCF reference.
+(c) $R_{c} = 1.6$ vdWR and $\epsilon_{\text{SCF}} = 10^{-2}$, using fully converged SCF reference.
+(d) Normalized ACF of the instantaneous error in panel (a) $\frac{1}{3}\langle \delta \vec{f}_i (t) \cdot \delta\vec{f}_i(t+\tau) \rangle_{it} $.
 Forces are computed with the PBE XC functional and TZV2P basis set for the configurations from a 10-ps trajectory generated with KS-based AIMD at T=298~K. 
-% Here it is clear that Eq.\ (\ref{eq:stochastic2}) is satisfied. 
-(a) $R_{c} = 1.6$ vdWR and $\epsilon_{SCF} = 10^{-2}$, 
-(b) $R_{c} = 1.6$ vdWR and fully converged ALMO SCF,
-(c) Autocorrelation function of the instantaneous error in panel (a) $\langle \vec{f}_i (t) \cdot \vec{f}_i(t+\tau) \rangle_t $.
 }
@@
 
-The reference PES was calculated using the orbital transformation (OT) method~\cite{a:ot,a:ot2} at $\SI{302}{\K}$ and the Langevin thermostat with $\gamma = \SI{e-3}{\per\fs}$. 
+The reference electronic structure was calculated using the orbital transformation (OT) method~\cite{a:ot,a:ot2} at $\SI{302}{\K}$ and the Langevin thermostat with $\gamma = \SI{e-3}{\per\fs}$. 
 All other system parameters were set identical to the ALMO system.
@@
 Some of the error can be attributed to the 9 K difference in the thermodynamic temperature between the reference system and our system of interest.
-Other sources of error are: small temporal sample size given our choice of $\gamma$; random error due to finite localization radius $R_{c}$, and random error due to high $\epsilon_{SCF}$.
-From figure \ref{fig:randomforce}, it's apparent that high values of $R_{c}$ introduce more random error than high values of $\epsilon_{SCF}$.
+Other sources of error present: small temporal sample size given our choice of $\gamma$; random error due to finite localization radius $R_{c}$; and random error due to high $\epsilon_{\text{SCF}}$.
+From figure \ref{fig:randomforce}, it's apparent that high values of $R_{c}$ and $\epsilon_{\text{SCF}}$ both introduce substantial error in calculation of forces.
 However, raising both $R_{c}$ and $\gamma$ decreases computation time significantly, and it is ultimately up to the user to determine their maximum tolerance of force inaccuracy when using ALMO DFT methods.
@@
 \includegraphics[trim={1.3cm 0.1cm 3.3cm 1.3cm},clip,width=8.6cm]{Dynamical_Data_Tiled.eps}
-\caption{\label{fig:dynproperties} Calculated dynamic and static properties of water using ALMO DFT with $\epsilon_{SCF} = 10^{-2}$ and $R_{c} = 1.6$ vdWR.
-(a) Radial distribution function of an ALMO simulation compared to OT SCF with a fully converged PES. 
+\caption{\label{fig:dynproperties} Calculated dynamic and static properties of water using ALMO DFT with $\epsilon_{\text{SCF}} = 10^{-2}$ and $R_{c} = 1.6$ vdWR.
+(a) Radial distribution function of an ALMO simulation compared to fully converged OT SCF. 
 (b) Kinetic energy distribution of an ALMO simulation compared to the theoretical Maxwell-Boltzmann kinetic energy distribution.
-(c) Calculated Infrared spectrum of an ALMO simulation compared to OT SCF with a fully converged PES.
+(c) Calculated Infrared spectrum of an ALMO simulation compared to fully converged OT SCF.
 Calculations were performed with the PBE $\mathrm{E_{XC}}$ functional and TZV2P basis set, using 125 water molecules for 10 ps.}
@@
 
-To show the computational efficiency gained by using unconverged ALMO DFT forces, we compare the average wall-time per MD step over 10 time steps for various values of $\epsilon_{SCF}$ in figure \ref{fig:strongscaling_log}.
+To show the computational efficiency gained by using unconverged ALMO DFT forces, we compare the average wall-time per MD step over 10 time steps for various values of $\epsilon_{\text{SCF}}$ in figure \ref{fig:strongscaling_log}.
 For these simulations, we chose to set the localization radius $R_{c} = 1.6$ van der Waals radii (vdWR), as was done for calculation of dynamical properties.
 We also show timing of the fully converged OT SCF method, as well as the recently developed ``second generation Car-Parinello'' MD method, which improves upon the OT computational cost prefactor.~\cite{a:2ndcpmd}
-ALMO methods show clear asymptotic LS behavior for all values of $\epsilon_{SCF}$ tested.
+ALMO methods show clear asymptotic LS behavior for all values of $\epsilon_{\text{SCF}}$ tested.
 OT SCF scales approximately cubically, as expected.
@@
 Calculations were performed with the PBE $\mathrm{E_{XC}}$ functional and TZV2P basis set, averaged over 10 MD steps. 
-The localization radius was set to $R_{c} = 1.6$ vdWR and $\epsilon_{SCF} = 10^{-2}$.}
+The localization radius was set to $R_{c} = 1.6$ vdWR and $\epsilon_{\text{SCF}} = 10^{-2}$.}
 \end{figure}

### 2017-08-31 almo-langevin.tex
Small edit
Added "a.u." after espilon_SCF = (number) in 4 places

@@
 The red line is the Euclidean norm of the instantaneous error $\langle \| \delta \vec{f_{i}}(t) \| \rangle_{i}$, the black line is the magnitude of the time average of the instantaneous error vector, and the green line is the standard deviation from the average.
-(a) $R_{c} = 1.6$ vdWR and $\epsilon_{\text{SCF}} = 10^{-2}$, using fully converged ALMO SCF reference.
+(a) $R_{c} = 1.6$ vdWR and $\epsilon_{\text{SCF}} = 10^{-2}$~a.u., using fully converged ALMO SCF reference.
 (b) $R_{c} = 1.6$ vdWR and fully converged ALMO SCF, using fully converged SCF reference.
-(c) $R_{c} = 1.6$ vdWR and $\epsilon_{\text{SCF}} = 10^{-2}$, using fully converged SCF reference.
+(c) $R_{c} = 1.6$ vdWR and $\epsilon_{\text{SCF}} = 10^{-2}$~a.u., using fully converged SCF reference.
 (d) Normalized ACF of the instantaneous error in panel (a) $\frac{1}{3}\langle \delta \vec{f}_i (t) \cdot \delta\vec{f}_i(t+\tau) \rangle_{it} $.
@@
 \includegraphics[trim={1.3cm 0.1cm 3.3cm 1.3cm},clip,width=8.6cm]{Dynamical_Data_Tiled.eps}
-\caption{\label{fig:dynproperties} Calculated dynamic and static properties of water using ALMO DFT with $\epsilon_{\text{SCF}} = 10^{-2}$ and $R_{c} = 1.6$ vdWR.
+\caption{\label{fig:dynproperties} Calculated dynamic and static properties of water using ALMO DFT with $\epsilon_{\text{SCF}} = 10^{-2}$~a.u. and $R_{c} = 1.6$ vdWR.
 (a) Radial distribution function of an ALMO simulation compared to fully converged OT SCF. 
@@
 Calculations were performed with the PBE $\mathrm{E_{XC}}$ functional and TZV2P basis set, averaged over 10 MD steps. 
-The localization radius was set to $R_{c} = 1.6$ vdWR and $\epsilon_{\text{SCF}} = 10^{-2}$.}
+The localization radius was set to $R_{c} = 1.6$ vdWR and $\epsilon_{\text{SCF}} = 10^{-2}$~a.u.}
 \end{figure}

### 2017-09-10 almo-langevin.tex
September 10th, small edits

@@
 
-The approach utilizes a systematically inaccurate potential energy surface calculation method that grows linearly with the size of the simulated system. 
+The approach utilizes a systematically inaccurate potential energy surface calculation method that grows linearly in computational complexity with the size of the simulated system. 
 Systematic inaccuracies in the calculation of force using this method are balanced by a weighted version of the Langevin equation of motion. 
@@
 ALMO constraints prohibit electron density transfer between distant molecules but retain all other types of interaction such as long-range electrostatic, exchange, polarization, and, if the exchange-correlation (XC) functional includes it, dispersion interactions~\cite{a:theeda}.
-Since the energetic contribution of the electron transfer decays exponentially with distance, \mbox{ALMO} approximation provides a natural and accurate representation of the electronic structure of molecular systems with just a few local variables if an appropriate localization radius is chosen.
+Since the energetic contribution of the electron transfer decays exponentially with distance, the \mbox{ALMO} approximation provides a natural and accurate representation of the electronic structure of molecular systems with just a few local variables if an appropriate localization radius is chosen.
 %The advantage of ALMO DFT is that its computational complexity is linear with respect to system size because long-range electron delocalization is ignored in both optimization steps.
@@
 a finite localization radius $R_c$ and incomplete SCF optimization of the electronic degrees of freedom. 
-$\delta f_{i\alpha} (t)$ can be reduced \emph{systematically} to zero by increasing $R_c$ (Figure~\ref{fig:forcecomp} in the Supplemental Material) and decreasing $\epsilon_{\text{SCF}}$ -- the maximum norm of the gradient of the \mbox{ALMO} DFT energy with respect to the ALMO coefficients.
+$\delta f_{i\alpha} (t)$ can be reduced \emph{systematically} to zero by increasing $R_c$ (Figure~\ref{fig:forcecomp}) and decreasing $\epsilon_{\text{SCF}}$ -- the maximum norm of the gradient of the \mbox{ALMO} DFT energy with respect to the ALMO coefficients.
 $\epsilon_{\text{SCF}}$ is chosen such that the electronic ground state is considered sufficiently close to the Born-Oppenheimer state for values less than $\epsilon_{\text{SCF}}$.
@@
 This, however, greatly increases the computational overhead of the simulation as longer-distance but physically less important interactions must be calculated. 
-Therefore, in this work we propose a different approach to ALMO-based AIMD, which allows for relaxation of the tight constraints on $R_{c}$ and $\epsilon_{\text{SCF}}$ and great reduction of the simulation cost.
-The approach is inspired by ``Second Generation Car-Parinello'' MD~\cite{a:2ndcpmd} and requires the use of the Langevin equations of motion~\cite{a:Kubo-1986}
+Therefore, in this work we propose a different approach to ALMO-based AIMD, which allows for relaxation of the tight constraints on $R_{c}$ and $\epsilon_{\text{SCF}}$ resulting in a great reduction in the simulation cost.
+The approach is inspired by ``Second Generation Car-Parinello'' MD~\cite{a:2ndcpmd} and requires the use of the Langevin equations of motion:~\cite{a:Kubo-1986}
 %
@@
 %
-where $m_i$ is the mass of atom $i$, $r_{i\alpha}$ its position along dimension $\alpha$, and $\gamma$ is the Langevin scaling factor.
+where $m_i$ is the mass of atom $i$, $r_{i\alpha}$ is its position along dimension $\alpha$, and $\gamma$ is the Langevin scaling factor.
 % which defines the strength of both the damping term and the stochastic term $R^{\gamma}_{i\alpha} (t)$. 
-In order to sample the canonical distribution correctly, the stochastic force must have zero mean and, according to the second fluctuation dissipation theorem~\cite{a:Kubo-1986,b:tuckerman-stat} must be related to the damping through $\gamma$:
+In order to sample the canonical distribution correctly, the stochastic force must have zero vector average and, according to the second fluctuation dissipation theorem~\cite{a:Kubo-1986,b:tuckerman-stat} must be related to the damping through $\gamma$:
 %
@@
 %
-These equations mean that, for any value of $\gamma$, the perfectly balanced damping and stochastic terms, defined in the equations above, will maintain a constant temperature and also generate the correct Maxwell-Boltzmann distribution of velocities.
+These equations mean that, for any value of $\gamma$, the perfectly balanced damping and stochastic terms, defined in equation \ref{eq:langevin}, will maintain a constant temperature and also generate the correct Maxwell-Boltzmann distribution of velocities.
 In the limit $\gamma \rightarrow 0$, the Newton equations of motion are recovered and the microcanonical ensemble is sampled.
@@
 
-An optimized $\gamma$ is one that is: large enough to adequately absorb force discontinuities; large enough to correctly thermostat the system of interest on a reasonable time scale; and small enough not to significantly affect dynamical properties on the time scale of interest.
+An optimized $\gamma$ is one that is: large enough to adequately absorb force discontinuities introduced by finite $R_{c}$; large enough to correctly thermostat the system of interest on a reasonable time scale; and small enough not to significantly affect dynamical properties on the time scale of interest.
 The relative weight of the three is determined by the specifics of the system and what is most important to the researcher. 
@@
 \caption{\label{fig:randomforce} Comparison of random force introduced from $R_{c}$ and $\epsilon_{\text{SCF}}$. 
-The red line is the Euclidean norm of the instantaneous error $\langle \| \delta \vec{f_{i}}(t) \| \rangle_{i}$, the black line is the magnitude of the time average of the instantaneous error vector, and the green line is the standard deviation from the average.
-(a) $R_{c} = 1.6$ vdWR and $\epsilon_{\text{SCF}} = 10^{-2}$~a.u., using fully converged ALMO SCF reference.
-(b) $R_{c} = 1.6$ vdWR and fully converged ALMO SCF, using fully converged SCF reference.
-(c) $R_{c} = 1.6$ vdWR and $\epsilon_{\text{SCF}} = 10^{-2}$~a.u., using fully converged SCF reference.
+The red line is the Euclidean norm of the instantaneous error $\langle \| \delta \vec{f_{i}}(t) \| \rangle_{i}$, the black line is the magnitude of the time average of the instantaneous error vector, and the green line is the time average of the red line.
+(a) $R_{c} = 1.6$ vdWR and $\epsilon_{\text{SCF}} = 10^{-2}$~a.u., using fully converged ALMO SCF as a reference.
+(b) $R_{c} = 1.6$ vdWR and fully converged ALMO SCF, using fully converged OT SCF as a reference.
+(c) $R_{c} = 1.6$ vdWR and $\epsilon_{\text{SCF}} = 10^{-2}$~a.u., using fully converged OT SCF as a reference.
 (d) Normalized ACF of the instantaneous error in panel (a) $\frac{1}{3}\langle \delta \vec{f}_i (t) \cdot \delta\vec{f}_i(t+\tau) \rangle_{it} $.
@@
 From figure \ref{fig:randomforce}, it's apparent that high values of $R_{c}$ and $\epsilon_{\text{SCF}}$ both introduce substantial error in calculation of forces.
-However, raising both $R_{c}$ and $\gamma$ decreases computation time significantly, and it is ultimately up to the user to determine their maximum tolerance of force inaccuracy when using ALMO DFT methods.
+However, raising both $R_{c}$ and $\gamma$ decreases computation time significantly, and it is ultimately up to the user to determine their maximum tolerance of force inaccuracy when using the ALMO DFT method presented here.
 
@@
 
-To show the computational efficiency gained by using unconverged ALMO DFT forces, we compare the average wall-time per MD step over 10 time steps for various values of $\epsilon_{\text{SCF}}$ in figure \ref{fig:strongscaling_log}.
-For these simulations, we chose to set the localization radius $R_{c} = 1.6$ van der Waals radii (vdWR), as was done for calculation of dynamical properties.
+To demonstrate the computational efficiency gained by using unconverged ALMO DFT forces, we compare the average wall-time per MD step over 10 time steps for various values of $\epsilon_{\text{SCF}}$ in figure \ref{fig:strongscaling_log}.
+For these simulations, we chose to set the localization radius to $R_{c} = 1.6$ van der Waals radii (vdWR), as was done for calculation of dynamical properties.
 We also show timing of the fully converged OT SCF method, as well as the recently developed ``second generation Car-Parinello'' MD method, which improves upon the OT computational cost prefactor.~\cite{a:2ndcpmd}
@@
 OT SCF scales approximately cubically, as expected.
-Our findings show that the fastest ALMO DFT method presented here runs faster than second-generation Car-Parinello using OT DFT for water simulations larger than 256 molecules.
+Our findings show that the fastest ALMO DFT methods presented here runs faster than second-generation Car-Parinello using OT DFT for water simulations larger than 256 molecules.
 
@@
 
-Figure \ref{fig:weakscaling} shows that ALMO MD code is well optimized for parallelization, and that computational cost grows linearly under a wide range of available computing power. 
+Figure \ref{fig:weakscaling} shows that ALMO MD code is well optimized for parallelization, and that the computational cost grows linearly under a wide range of available computing power. 
 In fact, we were able to successfully simulate systems as large as 32768 water molecules within 25 minutes of wall-clock time using 4096 cores.
@@
 
-Generalization of the method to more challenging systems of strongly interacting atoms (e.g. covalent crystals) will also be discussed.
+Generalization of the method to more challenging systems of strongly interacting atoms (e.g. covalent crystals) will also be explored in the future.
 

### 2017-09-15 almo-langevin.tex
Added section: Supplemental Materials and moved ALMO SCF delta force comparison to there. Added linear scale timing plot and mean-squared displacement plot to SI.

@@
 a finite localization radius $R_c$ and incomplete SCF optimization of the electronic degrees of freedom. 
-$\delta f_{i\alpha} (t)$ can be reduced \emph{systematically} to zero by increasing $R_c$ (Figure~\ref{fig:forcecomp}) and decreasing $\epsilon_{\text{SCF}}$ -- the maximum norm of the gradient of the \mbox{ALMO} DFT energy with respect to the ALMO coefficients.
+$\delta f_{i\alpha} (t)$ can be reduced \emph{systematically} to zero by increasing $R_c$ (Figure~\ref{fig:forcecomp} in the Supplemental Material) and decreasing $\epsilon_{\text{SCF}}$ -- the maximum norm of the gradient of the \mbox{ALMO} DFT energy with respect to the ALMO coefficients.
 $\epsilon_{\text{SCF}}$ is chosen such that the electronic ground state is considered sufficiently close to the Born-Oppenheimer state for values less than $\epsilon_{\text{SCF}}$.
 %ZZZ For $\epsilon_{\text{SCF}} \leqslant 10^{-7}$~a.u. ALMO SCF is considered converged.
-
-\begin{figure}[t]
-\includegraphics[trim={0cm 0cm 0.1cm 0.1cm},clip,width=8.6cm]{DeltaForceComparison_ALMO_SCF.eps}
-\caption{\label{fig:forcecomp} Dependence of the Euclidean norm of the instantaneous error $\langle \| \delta \vec{f_{i}}(t) \| \rangle_{i}$ (red line) on the ALMO localization radius $R_c$ expressed in units of atomic van der Waals radii.
-Black line is the magnitude of the time average of the instantaneous error vector.
-%calculated as the average difference between fully converged Born-Oppenheimer forces and fully converged ALMO DFT forces, averaged over all dimensions and all molecules for each time step.
-Forces are computed with the PBE XC functional, TZV2P basis set, and $\epsilon_{\text{SCF}} = 10^{-7}$~a.u. for the configurations from a 10-ps trajectory generated with KS-based AIMD at T=298~K.}
-\end{figure}
 
@@
 In order to confirm that ALMO DFT forces are accurate in the limit where localization radius $R_{c} \rightarrow \infty$, we computed the average difference in force between fully converged Born-Oppenheimer forces and forces calculated from ALMO DFT with varying $R_{c}$.
-Figure \ref{fig:forcecomp} shows that as $R_{c}$ is increased to infinity, forces converge toward the Born-Oppenheimer limit.
+Figure \ref{fig:forcecomp} in the Supplemental Material shows that as $R_{c}$ is increased to infinity, forces converge toward the Born-Oppenheimer limit.
 
 To test the accuracy of our unconverged ALMO based MD we computed the Infrared (IR) spectrum, radial distribution function (RDF), kinetic energy distribution, and mean-squared displacement (MSD) function from a 10 ps MD trajectory (see fig.
-\ref{fig:dynproperties} and the Supplemental Material).
+\ref{fig:dynproperties} and \ref{fig:diffusion} in the Supplemental Material).
 We compared the IR spectrum, RDF, and MSD function to corresponding reference functions calculated with fully converged KS DFT forces.
@@
 
-%\begin{figure}
-%\includegraphics[trim={1.7cm 0.5cm 3.3cm 0.1cm},clip,width=8.6cm]{strongscaling_linear.eps}
-%\caption{\label{fig:strongscaling_linear} Linear plots showing LS behaviour of all ALMO methods for liquid water at 298 K. 
-%Calculations were performed with the PBE $\mathrm{E_{XC}}$ functional and TZV2P basis set on 256 cores. 
-%The localization radius was set to $R_{c} = 1.6$ vdWR.}
-%\end{figure}
-
 \begin{figure}
@@
 
-%\textbf{Supplemental Material}
-
 %ZZZ outline procedure to compute forces using the DM obtained from ALMOs.
@@
 
+\clearpage
+\widetext
+\setcounter{figure}{0}
+\setcounter{page}{1}
+\renewcommand{\thefigure}{S\arabic{figure}}
+\begin{center}
+\textbf{\large Supplemental Material}
+\end{center}
+
+\begin{figure}[!htb]
+\includegraphics[trim={0cm 0cm 0.1cm 0.1cm},clip,width=\textwidth]{DeltaForceComparison_ALMO_SCF.eps}
+\caption{\label{fig:forcecomp} Dependence of the Euclidean norm of the instantaneous error $\langle \| \delta \vec{f_{i}}(t) \| \rangle_{i}$ (red line) on the ALMO localization radius $R_c$ expressed in units of atomic van der Waals radii.
+Black line is the magnitude of the time average of the instantaneous error vector.
+%calculated as the average difference between fully converged Born-Oppenheimer forces and fully converged ALMO DFT forces, averaged over all dimensions and all molecules for each time step.
+Forces are computed with the PBE XC functional, TZV2P basis set, and $\epsilon_{\text{SCF}} = 10^{-7}$~a.u. for the configurations from a 10-ps trajectory generated with KS-based AIMD at T=298~K.}
+\end{figure}
+
+\begin{figure}[!htb]
+\includegraphics[trim={1.7cm 0.5cm 3.3cm 0.1cm},clip,width=\textwidth]{strongscaling_linear.eps}
+\caption{\label{fig:strongscaling_linear} Linear plots showing LS behaviour of all ALMO methods for liquid water at 298 K. 
+Calculations were performed with the PBE $\mathrm{E_{XC}}$ functional and TZV2P basis set, averaged over 10 MD steps. 
+The localization radius was set to $R_{c} = 1.6$ vdWR and $\epsilon_{\text{SCF}} = 10^{-2}$~a.u.}
+\end{figure}
+
+\begin{figure}[!htb]
+\includegraphics[trim={2.5cm 0.2cm 4.5cm 1.2cm},clip,width=\textwidth]{Diffusion_Comparison.eps}
+\caption{\label{fig:diffusion} Calculated mean-squared displacement function of water using ALMO DFT with $\epsilon_{\text{SCF}} = 10^{-2}$~a.u. and $R_{c} = 1.6$ vdWR.
+Compared with equivalent MD run using OT DFT as a reference.
+Calculations were performed at 298 K with the PBE $\mathrm{E_{XC}}$ functional and TZV2P basis set, using 125 water molecules for 10 ps.}
+\end{figure}
+
 \end{document}

### 2017-09-25 almo-langevin.tex
minor modification

--- 2017-08-31 almo-langevin.tex
@@
 Since the unification of molecular dynamics and density functional theory (DFT)~\cite{a:thecpmd},
-\emph{ab initio} molecular dynamics (AIMD)~\cite{b:aimd} has become an important tool for the study of weakly bonded ensembles of molecules such as gas-phase molecular clusters, liquids, solutions, and molecular solids.
+\emph{ab initio} molecular dynamics (AIMD)~\cite{b:aimd} has become an important tool for studying of weakly bonded ensembles of molecules such as gas-phase molecular clusters, liquids, solutions, and molecular solids.
 %[of gas-phase molecules and condensed-phase systems.]
@@
 The new AIMD method utilizes a recently developed LS DFT~\cite{a:almo-ls} based on absolutely localized molecular orbitals (\mbox{ALMOs}). 
-Unlike delocalized KS orbitals, each \mbox{ALMO} has its own localization center -- atomic orbitals of a molecule are used here -- and a predefined localization radius that typically includes localization centers of neighbor molecules~\cite{a:stoll,a:almo-ls}.
+Unlike delocalized KS orbitals, each \mbox{ALMO} has its own localization center -- atomic orbitals of a molecule are used as a delocalization center in the current implementation -- and a predefined localization radius that typically includes localization centers of neighbor molecules~\cite{a:stoll,a:almo-ls}.
 The key feature of ALMO DFT is that its trial wavefunctions are constructed to circumvent the problem of the sluggish variational optimization that have rendered many previous MO-based LS methods impractical.
@@
 
-ALMO constraints prohibit electron density transfer between distant molecules but retain all other types of interaction such as long-range electrostatic, exchange, polarization, and, if the exchange-correlation (XC) functional includes it, dispersion interactions~\cite{a:theeda}.
-Since the energetic contribution of the electron transfer decays exponentially with distance, \mbox{ALMO} approximation provides a natural and accurate representation of the electronic structure of molecular systems with just a few local variables if an appropriate localization radius is chosen.
+ALMO constraints restrict/prohibit/forbid electron density transfer between distant molecules but retain all other types of interaction such as long-range electrostatic, exchange, polarization, and, if the exchange-correlation (XC) functional includes it, dispersion interactions~\cite{a:theeda}.
+Since the energetic contribution of the electron transfer decays exponentially with distance, \mbox{ALMO} approximation provides a natural and accurate representation of the electronic structure of molecular systems with just [a?] few local variables if an appropriate localization radius is chosen.
 %The advantage of ALMO DFT is that its computational complexity is linear with respect to system size because long-range electron delocalization is ignored in both optimization steps.
 %Because of the greatly reduced number of electronic variables and their robust optimization the computational cost of the implemented ALMO DFT is so low that it is [has the potential] capable of extending the range of accurate simulations to systems of thousands of molecules.
-Because of the greatly reduced number of electronic descriptors and the robust optimization, the computational complexity of ALMO DFT grows linearly with the number of molecules while its computational overhead remains very low.
+Because of the greatly reduced number of electronic descriptors and the robust optimization, the computational complexity of ALMO DFT is linear with the number of molecules while its computational overhead remains very low.
 These features make ALMO DFT a promising method to perform accurate AIMD simulations of large molecular systems.
@@
 a finite localization radius $R_c$ and incomplete SCF optimization of the electronic degrees of freedom. 
-$\delta f_{i\alpha} (t)$ can be reduced \emph{systematically} to zero by increasing $R_c$ (Figure~\ref{fig:forcecomp} in the Supplemental Material) and decreasing $\epsilon_{\text{SCF}}$ -- the maximum norm of the gradient of the \mbox{ALMO} DFT energy with respect to the ALMO coefficients.
-$\epsilon_{\text{SCF}}$ is chosen such that the electronic ground state is considered sufficiently close to the Born-Oppenheimer state for values less than $\epsilon_{\text{SCF}}$.
+$\delta f_{i\alpha} (t)$ can be reduced \emph{systematically} to zero by increasing $R_c$ (Figure~\ref{fig:forcecomp} in the Supplemental Material) and decreasing $\epsilon_{\text{SCF}}$ -- [ZZZ the value maximum norm of the gradient of the \mbox{ALMO} DFT energy with respect to the ALMO coefficients.
+below which the electronic ground state is considered sufficiently close to the Born-Oppenheimer state.]
 %ZZZ For $\epsilon_{\text{SCF}} \leqslant 10^{-7}$~a.u. ALMO SCF is considered converged.
@@
 \begin{figure}[t]
-\includegraphics[trim={0cm 0cm 0.1cm 0.1cm},clip,width=8.6cm]{DeltaForceComparison_ALMO_SCF.eps}
-\caption{\label{fig:forcecomp} Dependence of the Euclidean norm of the instantaneous error $\langle \| \delta \vec{f_{i}}(t) \| \rangle_{i}$ (red line) on the ALMO localization radius $R_c$ expressed in units of atomic van der Waals radii.
-Black line is the magnitude of the time average of the instantaneous error vector.
+\includegraphics[trim={0.1cm 0cm 0.2cm 0.1cm},clip,width=8.6cm]{DeltaForceComparison_ALMO_SCF.eps}
+\caption{\label{fig:forcecomp} {\color{red} All figures: $\epsilon_{\text{SCF}}$ must have ``a.u.'' units.
+This figure: VdWR$\rightarrow$vdWR.
+Hayden, you state that you average over Cartesian components of forces.
+Unfortunately, averaging over Cartesian components must give zero for any sufficiently large isotropic system (i.e.
+for every component there will be equal  opposite component).
+I am afraid these plots underestimate our error and mostly show finite-size effects.
+It would be better to plot the instantaneous \emph{magnitude} of $\delta \vec{f}_i(t)$, averaged over all atoms/molecules $\langle \ldots \rangle_i$.
+Can you make a new plot?
+Replace legend with: $\langle \delta f_i(t) \rangle_i$ -- red, $\langle \delta f_i(t) \rangle_{it}$ -- black, $\sqrt{\langle \delta \vec{f}_i(t) \cdot \delta \vec{f}_i(t) \rangle_{it}}$ -- green.
+Or remove it completely.
+I suggest we plot two green lines: above and below the average.} Dependence of the Euclidean norm of the instantaneous error $\langle \delta f_i(t) \rangle_i$ (red line) on the ALMO localization radius $R_c$ expressed in units of atomic van der Waals radii.
+Black line is the time average of the instantaneous error, green line is the standard deviation from the average.
 %calculated as the average difference between fully converged Born-Oppenheimer forces and fully converged ALMO DFT forces, averaged over all dimensions and all molecules for each time step.
@@
 This, however, greatly increases the computational overhead of the simulation as longer-distance but physically less important interactions must be calculated. 
-Therefore, in this work we propose a different approach to ALMO-based AIMD, which allows for relaxation of the tight constraints on $R_{c}$ and $\epsilon_{\text{SCF}}$ and great reduction of the simulation cost.
-The approach is inspired by ``Second Generation Car-Parinello'' MD~\cite{a:2ndcpmd} and requires the use of the Langevin equations of motion~\cite{a:Kubo-1986}
+Therefore, in this work, we propose a different approach to ALMO-based AIMD, which allows for relaxation of the tight constraints on $R_{c}$ and $\epsilon_{\text{SCF}}$ and great reduction of the simulation cost.
+The approach is inspired by ZZZ and requires the use of the Langevin equations of motion~\cite{a:Kubo-1986}
 %
@@
 %
-where $m_i$ is the mass of atom $i$, $r_{i\alpha}$ its position along dimension $\alpha$, and $\gamma$ is the Langevin scaling factor.
+where $m_i$ is the mass of atom $i$, $r_{i\alpha}$ its position along dimension $\alpha$, $\gamma$ is the Langevin scaling factor.
 % which defines the strength of both the damping term and the stochastic term $R^{\gamma}_{i\alpha} (t)$. 
+Langevin thermostat.
 In order to sample the canonical distribution correctly, the stochastic force must have zero mean and, according to the second fluctuation dissipation theorem~\cite{a:Kubo-1986,b:tuckerman-stat} must be related to the damping through $\gamma$:
@@
 %
-These equations mean that, for any value of $\gamma$, the perfectly balanced damping and stochastic terms, defined in the equations above, will maintain a constant temperature and also generate the correct Maxwell-Boltzmann distribution of velocities.
+These equations mean that, for any value of $\gamma$, 
+Thus the perfectly balanced damping and stochastic terms, defined in the equations above, will maintain a constant temperature while also generating the correct Maxwell-Boltzmann distribution of velocities.
 In the limit $\gamma \rightarrow 0$, the Newton equations of motion are recovered and the microcanonical ensemble is sampled.
-Larger values of $\gamma$ thermostat the system more strongly, but introduce more random noise necessitating longer trajectories when calculating system bulk properties.
-It has previously been shown that Langevin dynamics can be utilized to correct inaccurate force calculations to maintain correct dynamical properties.~\cite{a:langevin-why,a:2ndcpmd,b:tuckerman-stat,a:ceriotti}
+Compromise between small and large $\gamma$: It has previously been shown that Langevin dynamics can be utilized to correct inaccurate force calculations to maintain correct dynamical properties.~\cite{a:langevin-why,a:2ndcpmd,b:tuckerman-stat,a:ceriotti}
 
-%By utilizing Eq.\ (\ref{eq:langevin}) in conjunction with the ALMO SCF method of force calculation, we will show that discontinuity in the PES can be ignored for sufficiently large $\gamma$. 
+By utilizing Eq.\ (\ref{eq:langevin}) in conjunction with the ALMO SCF method of force calculation, we will show that discontinuity in the PES can be ignored for sufficiently large $\gamma$. 
 % Perhaps this statement is not correct: Note that increasing $\gamma$ increases the sample size required to generate accurate dynamical properties, as more random Gaussian noise is introduced.
@@
 %%During the second step of the ALMO SCF method described above, a self-consistent variational approach is taken to converge the non-orthogonal MOs. 
-%Generally, this step utilizes a iterative search function in order to reduce the absolute norm of the energy gradient to some target value $\epsilon_{\text{SCF}}$. 
-%Previously, $\epsilon_{\text{SCF}}$ had to be set very small to ensure sufficient convergence of the electron structure calculations. 
-%By increasing the allowable $\epsilon_{\text{SCF}}$, a great reduction in computational cost can be achieved. 
+%Generally, this step utilizes a iterative search function in order to reduce the absolute norm of the energy gradient to some target value $\epsilon_{SCF}$. 
+%Previously, $\epsilon_{SCF}$ had to be set very small to ensure sufficient convergence of the electron structure calculations. 
+%By increasing the allowable $\epsilon_{SCF}$, a great reduction in computational cost can be achieved. 
 %Obviously, this introduces a significant error into the calculation of forces during each time step.
@@
 %
-Figure~\ref{fig:randomforce} shows that this assumption is sufficiently justified for the case of liquid water. 
-Therefore, the Langevin equations of motion (\ref{eq:langevin}) can be re-written with the two stochastic terms combined into one:
+[ZZZ later?: Figure~\ref{fig:randomforce} shows that this assumption is sufficiently justified for the case of liquid water.] 
+If this assumption holds, the Langevin equations of motion (\ref{eq:langevin}) can be re-written with the two stochastic terms combined into one:
 %
@@
 The relative weight of the three is determined by the specifics of the system and what is most important to the researcher. 
-To optimize $\gamma$ given the relative importance of the above properties, we suggest running several short trajectories on the system of interest (or a scaled-down version of it) using a tightly converged potential energy surface calculation and varying the value of $\gamma$.
+To optimize $\gamma$ given the relative importance of the above properties, we suggest running several short trajectories on the system of interest (or a scaled-down version of it) using a tightly converged PES calculation and varying the value of $\gamma$.
 We found that a value of $\gamma$ between $10^{-5}$ and $10^{-3}\ \mathrm{fs^{-1}}$ was optimal for the purposes of the research presented here.
@@
 To correctly thermostat the system of interest, $\Delta$ needs to be reduced to zero.
-This is done by finding the value of $\Delta$ and adding a third random force term $R^{-\Delta}_{i\alpha}$ to equation \ref{eq:langevin2} which cancels out the extra forces introduced to the system by $R^{\Delta}_{i\alpha}$.
+This is done by finding the value of $\Delta$ and adding a third random force term $R^{-\Delta}_{i\alpha}$ to equation \ref{eq:langevin2} which cancels out the energy introduced to the system by $R^{\Delta}_{i\alpha}$.
 This means that a trajectory with inaccurate forces and optimized $R^{-\Delta}_{i\alpha}$ should have no long term drift in the total energy of the system. 
-Short term energy fluctuations should be expected, and can later be ``smoothed out'' by setting $\gamma$ as needed.
+Short term energy fluctuations should be expected, and can later be ``smoothed out'' by setting finite $\gamma$.
 
@@
 The ACF was not possible to accurately integrate without running a reference trajectory with extremely short time-steps (see figure~\ref{fig:randomforce}).
-For example, using Eq.\ \ref{eq:delta} we calculated $\Delta \approx \SI{2e-5}{\per\fs}$, but the value was optimized heuristically to $\SI{6e-5}{\per\fs}$.
+For example, using Eq.\ \ref{eq:delta} we calculated $\Delta \approx \SI{2e-5}{\per\fs}$, but value was optimized heuristically to $\SI{6e-5}{\per\fs}$.
 Still, this method provided a decent order of magnitude starting point for further investigation.
@@
 \begin{figure}
-\includegraphics[trim={0.5cm 0cm 0.7cm 0.1cm},clip,width=8.6cm]{DeltaForceComparison_with_ACF.eps}
-\caption{\label{fig:randomforce} Comparison of random force introduced from $R_{c}$ and $\epsilon_{\text{SCF}}$. 
-The red line is the Euclidean norm of the instantaneous error $\langle \| \delta \vec{f_{i}}(t) \| \rangle_{i}$, the black line is the magnitude of the time average of the instantaneous error vector, and the green line is the standard deviation from the average.
-(a) $R_{c} = 1.6$ vdWR and $\epsilon_{\text{SCF}} = 10^{-2}$~a.u., using fully converged ALMO SCF reference.
-(b) $R_{c} = 1.6$ vdWR and fully converged ALMO SCF, using fully converged SCF reference.
-(c) $R_{c} = 1.6$ vdWR and $\epsilon_{\text{SCF}} = 10^{-2}$~a.u., using fully converged SCF reference.
-(d) Normalized ACF of the instantaneous error in panel (a) $\frac{1}{3}\langle \delta \vec{f}_i (t) \cdot \delta\vec{f}_i(t+\tau) \rangle_{it} $.
+\includegraphics[trim={0.6cm 0.7cm 0.7cm 0.3cm},clip,width=8.6cm]{DeltaForceComparison_with_ACF.eps}
+\caption{\label{fig:randomforce} {\color{red} The top two panels should be changed to plot the Euclidean norm of the error as described in the notes to the previous Figure.
+Do not forget to change the legend.
+Plot (c), new x-axis label: ``Time lag $\tau$ (fs)''.
+As to the ACF, one should be extremely careful how the averaging is performed.
+This caption below shows the ACF we need.
+Let me briefly explain it.
+The main idea is that the correct ACF could be computed from only one Cartesian component of one molecule if we could observe it for really really long time and average over many many time origins.
+We perform averaging over components and molecules just to improve our statistics (it is a shame to discard so many data points).
+We, therefore, should not compute ACF from \emph{one} time series that is already averaged over all possible components.
+Instead we compute and ACF for each component and then average over these ACFs.
+Ok, here is the procedure.
+Fix the time lag $\tau$.
+For a given time origin $t$ and molecule $i$, compute: $p_{it}(\tau)=\frac{1}{3}\delta \vec{f}_i (t)\cdot \delta \vec{f}_i (t+\tau)$, $\frac{1}{3}$ averages over components.
+Now, average over time origins and molecules to get $\text{ACF}(\tau) = \langle p_{it}(\tau) \rangle_{it}$.
+I think we do not need to subtract the mean because you already showed that it is zero.
+I hope there is a function in MATLAB that can do this.}
+Red line is the Euclidean norm of the instantaneous error $\langle \delta f_i(t) \rangle_i$, black line is the time average of the instantaneous error, green line is the standard deviation from the average.
 Forces are computed with the PBE XC functional and TZV2P basis set for the configurations from a 10-ps trajectory generated with KS-based AIMD at T=298~K.
+% Here it is clear that Eq.\ (\ref{eq:stochastic2}) is satisfied. 
+(a) $R_{c} = 1.6$ vdWR and $\epsilon_{SCF} = 10^{-2}$, 
+(b) $R_{c} = 1.6$ vdWR and fully converged ALMO SCF,
+(c) Autocorrelation function of the instantaneous error in panel (a) $\frac{1}{3}\langle \delta \vec{f}_i (t) \cdot \delta\vec{f}_i(t+\tau) \rangle_{it} $.
 }
@@
 
-The reference electronic structure was calculated using the orbital transformation (OT) method~\cite{a:ot,a:ot2} at $\SI{302}{\K}$ and the Langevin thermostat with $\gamma = \SI{e-3}{\per\fs}$. 
+The reference PES was calculated using the orbital transformation (OT) method~\cite{a:ot,a:ot2} at $\SI{302}{\K}$ and the Langevin thermostat with $\gamma = \SI{e-3}{\per\fs}$. 
 All other system parameters were set identical to the ALMO system.
@@
 Some of the error can be attributed to the 9 K difference in the thermodynamic temperature between the reference system and our system of interest.
-Other sources of error present: small temporal sample size given our choice of $\gamma$; random error due to finite localization radius $R_{c}$; and random error due to high $\epsilon_{\text{SCF}}$.
-From figure \ref{fig:randomforce}, it's apparent that high values of $R_{c}$ and $\epsilon_{\text{SCF}}$ both introduce substantial error in calculation of forces.
+Other sources of error are: small temporal sample size given our choice of $\gamma$; random error due to finite localization radius $R_{c}$, and random error due to high $\epsilon_{SCF}$.
+From figure \ref{fig:randomforce}, it's apparent that high values of $R_{c}$ introduce more random error than high values of $\epsilon_{SCF}$.
 However, raising both $R_{c}$ and $\gamma$ decreases computation time significantly, and it is ultimately up to the user to determine their maximum tolerance of force inaccuracy when using ALMO DFT methods.
@@
 \includegraphics[trim={1.3cm 0.1cm 3.3cm 1.3cm},clip,width=8.6cm]{Dynamical_Data_Tiled.eps}
-\caption{\label{fig:dynproperties} Calculated dynamic and static properties of water using ALMO DFT with $\epsilon_{\text{SCF}} = 10^{-2}$~a.u. and $R_{c} = 1.6$ vdWR.
-(a) Radial distribution function of an ALMO simulation compared to fully converged OT SCF. 
+\caption{\label{fig:dynproperties} Calculated dynamic and static properties of water using ALMO DFT with $\epsilon_{SCF} = 10^{-2}$ and $R_{c} = 1.6$ vdWR.
+(a) Radial distribution function of an ALMO simulation compared to OT SCF with a fully converged PES. 
 (b) Kinetic energy distribution of an ALMO simulation compared to the theoretical Maxwell-Boltzmann kinetic energy distribution.
-(c) Calculated Infrared spectrum of an ALMO simulation compared to fully converged OT SCF.
+(c) Calculated Infrared spectrum of an ALMO simulation compared to OT SCF with a fully converged PES.
 Calculations were performed with the PBE $\mathrm{E_{XC}}$ functional and TZV2P basis set, using 125 water molecules for 10 ps.}
@@
 
-To show the computational efficiency gained by using unconverged ALMO DFT forces, we compare the average wall-time per MD step over 10 time steps for various values of $\epsilon_{\text{SCF}}$ in figure \ref{fig:strongscaling_log}.
+To show the computational efficiency gained by using unconverged ALMO DFT forces, we compare the average wall-time per MD step over 10 time steps for various values of $\epsilon_{SCF}$ in figure \ref{fig:strongscaling_log}.
 For these simulations, we chose to set the localization radius $R_{c} = 1.6$ van der Waals radii (vdWR), as was done for calculation of dynamical properties.
 We also show timing of the fully converged OT SCF method, as well as the recently developed ``second generation Car-Parinello'' MD method, which improves upon the OT computational cost prefactor.~\cite{a:2ndcpmd}
-ALMO methods show clear asymptotic LS behavior for all values of $\epsilon_{\text{SCF}}$ tested.
+ALMO methods show clear asymptotic LS behavior for all values of $\epsilon_{SCF}$ tested.
 OT SCF scales approximately cubically, as expected.
@@
 Calculations were performed with the PBE $\mathrm{E_{XC}}$ functional and TZV2P basis set, averaged over 10 MD steps. 
-The localization radius was set to $R_{c} = 1.6$ vdWR and $\epsilon_{\text{SCF}} = 10^{-2}$~a.u.}
+The localization radius was set to $R_{c} = 1.6$ vdWR and $\epsilon_{SCF} = 10^{-2}$.}
 \end{figure}

### 2017-09-26 almo-langevin.tex
2 pages ++ done

--- 2017-09-15 almo-langevin.tex
@@
 Most LS methods~\cite{a:linscale3,a:lee-yang-1996,a:ls-scuseria-1997,a:ls-manolopoulos-1998,a:ls-helgaker-2001,a:ls-niklasson-2003,a:curvy2,a:ls-dm-sign} explore the natural locality of the one-electron density matrix (DM).
-%However, the variational optimization of the DM is very inefficient for accurate DFT calculations that require many basis functions per atom~\cite{a:ls-rev-1999,a:ls-dm-sign}. 
 However, in accurate DFT calculations that require many basis functions per atom, the variational optimization of the DM is advantageous only for impractically large systems~\cite{a:ls-rev-1999,a:ls-dm-sign,a:almo-ls}.
-%However, in accurate DFT calculations with many basis functions per atom are required, the DM becomes sufficiently sparse only for impractically large systems~\cite{a:ls-rev-1999,a:ls-dm-sign,a:almo-ls}.
 Therefore, the application of DM-based LS methods have been restricted to minimal-basis tight-binding problems. 
@@
 Although such methods have been successfully used for the evaluation of accurate DFT energies of large static systems~\cite{a:ls-onetep-2009,a:ls-conquest-2010,a:ls-onetep-2010-app1,a:ls-rev-2012} their use in AIMD is hampered by the costly optimization of both the contracted orbitals and the density matrix.
-%
-Alternatively, LS can be achieved using a set of local electronic descriptors constructed by enforcing strict locality on the KS orbitals and, at the same time, by relaxing the orthogonality constraints imposed on them.
-From the computational point of view, LS methods based on the direct optimization of such compact localized nonorthogonal orbitals~\cite{a:ls-galli-parrinello-1992,a:ls-mauri-galli-car-1993,a:ls-ordejon-1993,a:ls-mauri-galli-1994,a:ls-ordejon-1995,a:ls-kim-mauri-galli-1995,a:ls-fattebert-2004,a:ls-fattebert-2006,a:burger-yang-2008} are more promising since they require only the occupied orbitals and, thus, deal with fewer variational degrees of freedom than the DM methods.
-Unfortunately, the progress in the development of orbital-based LS methods has been hindered by the inherently difficult optimization of localized orbitals~\cite{a:ls-mauri-galli-car-1993,a:ls-ordejon-1995,a:ls-fattebert-2004,a:ls-rev-1999,a:weitao-yang-2013}.
+
+To obviate the DM optimization, another set of natural local electronic descriptors can be constructed by enforcing strict locality on the KS orbitals and, necessarily, by relaxing the orthogonality constraints imposed on them.
+From the computational point of view, LS methods based on the direct optimization of such compact localized nonorthogonal orbitals~\cite{a:ls-galli-parrinello-1992,a:ls-mauri-galli-car-1993,a:ls-ordejon-1993,a:ls-mauri-galli-1994,a:ls-ordejon-1995,a:ls-kim-mauri-galli-1995,a:ls-fattebert-2004,a:ls-fattebert-2006,a:burger-yang-2008} are very promising since they require only the occupied orbitals and, thus, deal with fewer variational degrees of freedom than the DM methods.
+Unfortunately, the development of orbital-based LS methods has been hindered by the inherently difficult optimization of localized orbitals reported by multiple research groups~\cite{a:ls-mauri-galli-car-1993,a:ls-ordejon-1995,a:ls-fattebert-2004,a:ls-rev-1999,a:weitao-yang-2013}.
 
@@
 
-%Thus, despite remarkable recent progress in linear scaling density function theory, the computational cost of existing methods remains too high for routine \emph{ab initio} molecular dynamics (AIMD) simulations.
 Thus, despite remarkable recent progress the high computational overhead of existing asymptotically linear-scaling DFT methods restrict their use in AIMD simulations to very short \emph{time scales}, systems of low dimensions, or problems that can be adequately described with minimal basis sets.
-Otherwise, LS DFT cannot compete with the straightforward cubically scaling KS DFT on the length scales accessible in routine AIMD simulations.
-%Thus,  the computational cost of the existing asymptotically linear-scaling DFT methods remains too high to compete with the straightforward cubically scaling KS DFT on the length scales practically accessible in routine AIMD simulations.
-%Thus, although there exist numerous DFT methods capable of yielding asymptotic LS, their computational cost for systems in/on the practically accessible length scale remains too high to compete with the conventional cubically scaling KS DFT with its highly-optimized numerical algorithms.
+On the length and time scales required for accurate routine AIMD simulations, LS DFT cannot compete with the straightforward low-cost cubically-scaling KS DFT.
 
-In this work, we introduce a new AIMD method for weakly-interacting molecular systems that is not only linear scaling but also has a low computational overhead. 
-Using liquid water as an example, we demonstrate that the proposed AIMD methodology enables computational studies of molecular systems on previously inaccessible time and length scales.
+In this work, we introduce a new orbital-based AIMD method that is not only linear scaling but also has a low computational overhead. 
+Currently, the method is restricted to large ensembles of atoms or small molecules that do not form strong covalent bonds with each other. 
+Using liquid water as an example, we demonstrate that the new AIMD methodology enables computational studies of molecular systems on previously inaccessible time and length scales.
 
 The new AIMD method utilizes a recently developed LS DFT~\cite{a:almo-ls} based on absolutely localized molecular orbitals (\mbox{ALMOs}). 
-Unlike delocalized KS orbitals, each \mbox{ALMO} has its own localization center -- atomic orbitals of a molecule are used here -- and a predefined localization radius that typically includes localization centers of neighbor molecules~\cite{a:stoll,a:almo-ls}.
-The key feature of ALMO DFT is that its trial wavefunctions are constructed to circumvent the problem of the sluggish variational optimization that have rendered many previous MO-based LS methods impractical.
-The trial ALMOs are optimized variationally in a two-step self-consistent-field (SCF) procedure~\cite{a:almo-ls}. 
+Unlike delocalized KS orbitals, each \mbox{ALMO} has its own localization center and a predefined localization radius $R_{c}$ that typically includes nearby atoms or molecules~\cite{a:stoll,a:almo-ls}.
+In the current implementation, a localization center is a set of all local Gaussian atomic orbitals of one molecule. 
+However the approach can be easily generalized to other local and nonlocal basis sets~\cite{a:ls-galli-parrinello-1992, Galerkin}. 
+The key feature of ALMO DFT is that its one-electron wavefunctions are constructed variationally in a two-step self-consistent-field (SCF) procedure~\cite{a:almo-ls} to circumvent the problem of the sluggish variational optimization that have rendered previous MO-based LS methods impractical.
 In the first step, ALMOs are constrained to their localization centers~\cite{a:khal} whereas, in the second step, ALMOs are relaxed to allow delocalization onto the neighbor molecules within their localization radius $R_{c}$.
@@
 
-ALMO constraints prohibit electron density transfer between distant molecules but retain all other types of interaction such as long-range electrostatic, exchange, polarization, and, if the exchange-correlation (XC) functional includes it, dispersion interactions~\cite{a:theeda}.
-Since the energetic contribution of the electron transfer decays exponentially with distance, the \mbox{ALMO} approximation provides a natural and accurate representation of the electronic structure of molecular systems with just a few local variables if an appropriate localization radius is chosen.
-%The advantage of ALMO DFT is that its computational complexity is linear with respect to system size because long-range electron delocalization is ignored in both optimization steps.
-%Because of the greatly reduced number of electronic variables and their robust optimization the computational cost of the implemented ALMO DFT is so low that it is [has the potential] capable of extending the range of accurate simulations to systems of thousands of molecules.
+Once the localization radius is chosen, ALMO constraint prohibits electron density transfer between distant molecules but retains all other types of interaction such as long-range electrostatic, exchange, polarization, and, if the exchange-correlation (XC) functional includes them, dispersion interactions~\cite{a:theeda}.
+Since the importance of electron transfer decays exponentially with distance, the \mbox{ALMO} approximation provides a natural and accurate representation of the electronic structure of molecular systems with just a few local variables.
 Because of the greatly reduced number of electronic descriptors and the robust optimization, the computational complexity of ALMO DFT grows linearly with the number of molecules while its computational overhead remains very low.
@@
 
-To adapt ALMO DFT for AIMD, the forces on atoms were computed \emph{approximately} using the idempotent DM obtained from ALMOs in a straightforward procedure that computes only the Hellmann-Feynman and Pulay forces while ignoring non-SCF forces (see the Supplemental Material).
-The difference between these forces and the exact forces obtained using perfectly-converged fully-delocalized KS orbitals is denoted $\delta f_{i\alpha}(t)$ 
+The challenge of adopting ALMO DFT for dynamical simulations arises from the slightly nonvariational character of the localized orbitals.
+First, while ALMOs are variationally optimized in both SCF steps, the occupied-space projector obtained in the first step must remain fixed during the second step to ensure convergence.
+Second, some neglected electron transfer effects can suddenly become active in the course of a dynamical simulation when a neighboring molecule passes the localization threshold $R_{c}$.
+In addition to the ALMO-specific issues, the variational optimization in any AIMD methods is never truly complete and is interrupted once the norm of the gradient of the energy with respect to the variational electronic degrees of freedom drops below small but nevertheless finite convergence threshold $\epsilon_{\text{SCF}}$.
+While these non-SCF errors are minor and do not affect accuracy of static ALMO DFT calculations, geometry optimization, and ALMO-based Monte-Carlo simulations, they accumulate in molecular dynamics trajectories resulting in non-physical sampling and eventually disrupting them.
+
+%For any finite value of $R_{c}$, forces are undefined at manifold of due to the discontinuous jump in potential energy when a neighboring molecule passes beyond the localization radius $R_{c}$ and is suddenly ignored in potential energy calculations.
+%Setting $R_{c}$ to larger values can reduce the magnitude of the discontinuity, but increases the computational cost as longer distance (but physically less important) interactions must be calculated.
+%As long as the discontinuity in force is small compared with the random numerical noise in the calculation, it can be safely ignored~\cite{ZZZ-any-references}.
+This means that $R_{c}$ has to be kept relatively large to maintain stable and accurate dynamics.
+
+%While these issues can be resolved by  by increasing  to account for non-SCF forces, increasing  and decreasing While  are the known formal strategies for resolving these
+
+In order to maintain stable and accurate dynamics, one could use a set of strategies that includes computing the non-SCF contribution to the forces through a coupled perturbed procedure, increasing $R_{c}$, and decreasing $\epsilon_{\text{SCF}}$.
+%relative to the corresponding optimal values for static energy calculations or Monte-Carlo simulations.
+This however would dramatically increase the computational cost of the ALMO AIMD.
+%as longer-distance but physically unimportant interactions must be calculated with an additional iterative coupled perturbed solver.
+%
+In this work we propose another approach, which obviates the need in a coupled-perturbed solver, allows for relaxation of the tight constraints on $R_{c}$ and $\epsilon_{\text{SCF}}$, and thus keeps the complexity and cost of simulations low.
+
+In our approach, the forces on atoms are calculated \emph{approximately} after the two-step ALMO SCF using a straightforward procedure that computes only the Hellmann-Feynman and Pulay components and neglects the computationally intense non-SCF component of the forces.
+The difference between these approximate ALMO forces and the \emph{reference} forces that would be obtained from perfectly converged fully-delocalized KS orbitals is $\delta f_{i\alpha}(t)$:
 %
@@
 where $\alpha$ is a Cartesian component of the force acting on atom $i$ at time $t$.
-$\delta f_{i\alpha} (t)$ comprises neglected terms that originate from %fixed terms in the trial ALMOs (i.e.
-orbitals obtained in the first SCF step are not re-optimized in the second step),
-a finite localization radius $R_c$ and incomplete SCF optimization of the electronic degrees of freedom. 
-$\delta f_{i\alpha} (t)$ can be reduced \emph{systematically} to zero by increasing $R_c$ (Figure~\ref{fig:forcecomp} in the Supplemental Material) and decreasing $\epsilon_{\text{SCF}}$ -- the maximum norm of the gradient of the \mbox{ALMO} DFT energy with respect to the ALMO coefficients.
-$\epsilon_{\text{SCF}}$ is chosen such that the electronic ground state is considered sufficiently close to the Born-Oppenheimer state for values less than $\epsilon_{\text{SCF}}$.
+$\delta f_{i\alpha} (t)$ comprises all neglected terms that originate from a finite localization radius $R_c$ and incomplete SCF optimization.
+$\delta f_{i\alpha} (t)$ can be reduced \emph{systematically} to zero by increasing $R_c$ (Figure~\ref{fig:forcecomp} in the Supplemental Material) and decreasing $\epsilon_{\text{SCF}}$.
+%$\epsilon_{\text{SCF}}$ is chosen such that the electronic ground state is considered sufficiently close to the Born-Oppenheimer state for values less than $\epsilon_{\text{SCF}}$.
 %ZZZ For $\epsilon_{\text{SCF}} \leqslant 10^{-7}$~a.u. ALMO SCF is considered converged.
 
-%For any finite value of $R_{c}$, forces are undefined at manifold of due to the discontinuous jump in potential energy when a neighboring molecule passes beyond the localization radius $R_{c}$ and is suddenly ignored in potential energy calculations.
-%Setting $R_{c}$ to larger values can reduce the magnitude of the discontinuity, but increases the computational cost as longer distance (but physically less important) interactions must be calculated.
-%As long as the discontinuity in force is small compared with the random numerical noise in the calculation, it can be safely ignored~\cite{ZZZ-any-references}.
-This means that $R_{c}$ has to be kept relatively large to maintain stable and accurate dynamics.
+%The approach is inspired by ``Second Generation Car-Parinello'' MD~\cite{a:2ndcpmd} and
 
-In order to maintain stable and accurate dynamics, $R_{c}$ has to be larger and $\epsilon_{\text{SCF}}$ has to be smaller than the corresponding optimal values used in static energy calculations or Monte-Carlo simulations.
-This, however, greatly increases the computational overhead of the simulation as longer-distance but physically less important interactions must be calculated. 
-Therefore, in this work we propose a different approach to ALMO-based AIMD, which allows for relaxation of the tight constraints on $R_{c}$ and $\epsilon_{\text{SCF}}$ resulting in a great reduction in the simulation cost.
-The approach is inspired by ``Second Generation Car-Parinello'' MD~\cite{a:2ndcpmd} and requires the use of the Langevin equations of motion:~\cite{a:Kubo-1986}
+Our approach to compensate for the missing $\delta f_{i\alpha}(t)$ term follows the methodology that have been introduced into AIMD by Krajewski \emph{et al.}~\cite{PRB-73-041105R}, formalized by K\"uhne \emph{et al.}~\cite{a:2ndcpmd} and rationalized by Dai \emph{et al.}~\cite{a:langevin-why} before becoming informally known as the second generation Car-Parrinello molecular dynamics~\cite{doi:10.1002/wcms.1176}.
+Inspired by this method, ALMO AIMD is chosen by be governed by the Langevin equations of motion, which can be written using the unknown reference forces as:
 %
@@
 %
-where $m_i$ is the mass of atom $i$, $r_{i\alpha}$ is its position along dimension $\alpha$, and $\gamma$ is the Langevin scaling factor.
-% which defines the strength of both the damping term and the stochastic term $R^{\gamma}_{i\alpha} (t)$. 
-In order to sample the canonical distribution correctly, the stochastic force must have zero vector average and, according to the second fluctuation dissipation theorem~\cite{a:Kubo-1986,b:tuckerman-stat} must be related to the damping through $\gamma$:
+where $m_i$ is the mass of atom $i$, $r_{i\alpha}$ is its position along dimension $\alpha$, $\gamma$ is the Langevin scaling factor, and $R^{\gamma}_{i\alpha} (t)$ is stochastic force represented by a white Gaussian noise with the zero mean and the strength related to the damping term through $\gamma$:
 %
@@
 \langle R^{\gamma}_{i\alpha} (t) \rangle &= 0, \\
-\langle R^{\gamma}_{i\alpha} (t)  R^{\gamma}_{j\beta} (t') \rangle &= 2 k_B T \gamma m_i \delta_{ij} \delta_{\alpha\beta} \delta(t-t')
+\langle R^{\gamma}_{i\alpha} (t)  R^{\gamma}_{j\beta} (t') \rangle &= 2 k_B T \gamma m_i \delta_{ij} \delta_{\alpha\beta} \delta(t-t').
 \end{align}
 %
-These equations mean that, for any value of $\gamma$, the perfectly balanced damping and stochastic terms, defined in equation \ref{eq:langevin}, will maintain a constant temperature and also generate the correct Maxwell-Boltzmann distribution of velocities.
+This relation means that, for any value of $\gamma$, the perfectly balanced damping and stochastic terms will maintain a constant average temperature $T$ and generate the correct Maxwell-Boltzmann distribution of velocities~\cite{a:Kubo-1986,b:tuckerman-stat}.
 In the limit $\gamma \rightarrow 0$, the Newton equations of motion are recovered and the microcanonical ensemble is sampled.
-Larger values of $\gamma$ thermostat the system more strongly, but introduce more random noise necessitating longer trajectories when calculating system bulk properties.
-It has previously been shown that Langevin dynamics can be utilized to correct inaccurate force calculations to maintain correct dynamical properties.~\cite{a:langevin-why,a:2ndcpmd,b:tuckerman-stat,a:ceriotti}
+Finite values of $\gamma$ correspond to imposing canonical thermostat on the system, with larger $\gamma$ leading to more random noise in trajectories, faster equilibration but more distorted results for calculated dynamical properties.
+
+%It has previously been shown that Langevin dynamics can be utilized to correct inaccurate force calculations to maintain correct dynamical properties.~\cite{a:langevin-why,a:2ndcpmd,b:tuckerman-stat,a:ceriotti}
 
@@
 
-The main assumption of ALMO AIMD is that the error in the ALMO forces is well approximated by a stochastic ``white noise'' term $R^{\Delta}_{i\alpha} (t)$
+The main assumption of ALMO AIMD is that the error in the ALMO forces is well approximated by a stochastic white Gaussian noise $R^{\Delta}_{i\alpha} (t)$:
 %
@@
 %
-Figure~\ref{fig:randomforce} shows that this assumption is sufficiently justified for the case of liquid water. 
-Therefore, the Langevin equations of motion (\ref{eq:langevin}) can be re-written with the two stochastic terms combined into one:
+This assumption, shown below to be well justified (Figure~\ref{fig:randomforce}), allows us to write the Langevin equations using the approximate ALMO forces:
 %
@@
 %
-where $R^{\gamma + \Delta}_{i\alpha} = R^{\gamma}_{i\alpha} + R^{\Delta}_{i\alpha}$. 
+where the two stochastic terms are combined into one $R^{\gamma + \Delta}_{i\alpha} = R^{\gamma}_{i\alpha} + R^{\Delta}_{i\alpha}$.
+This new Langevin equation means that the value of $\Delta$ must be known to re-balance the damping and stochastic terms and generate the correct Maxwell-Boltzmann distribution in simulations.
+
 Note that $\Delta$ can be either positive or negative depending on the way errors are introduced in the system of interest. 

### 2017-10-01 almo-langevin.tex
No summary, no abstract yet

@@
 \documentclass[aps,prl,reprint,amsmath,amssymb]{revtex4-1}
+
+% define variable that compiles the main text (as opposed to SI text)
+\newcommand*{\MAINTEXT}{}
 
@@
 
+\ifdefined\MAINTEXT
+\else
+	\clearpage
+	\widetext
+	\setcounter{figure}{0}
+	\setcounter{page}{1}
+	\renewcommand{\thefigure}{S\arabic{figure}}
+\fi
+
 \title{
+\ifdefined\MAINTEXT
+\else
+Supplemental Material: \\
+\fi
 Low-cost linear-scaling \emph{ab initio} molecular dynamics for weakly-interacting systems
@@
 
+\ifdefined\MAINTEXT
+
 \begin{abstract}
@@
 
-The approach utilizes a systematically inaccurate potential energy surface calculation method that grows linearly in computational complexity with the size of the simulated system. 
-Systematic inaccuracies in the calculation of force using this method are balanced by a weighted version of the Langevin equation of motion. 
-The stochastic ``white noise'' of the Langevin equation acts to absorb force discontinuities while also thermostating the system to maintain the desired thermodynamic temperature. 
-In addition, we introduce a simple heuristic approach to find optimized Langevin equation parameters to use for balancing system-specific imperfect forces and maintaining correct dynamical behavior.
+%The approach utilizes a systematically inaccurate potential energy surface calculation method that grows linearly in computational complexity with the size of the simulated system. 
+%Systematic inaccuracies in the calculation of force using this method are balanced by a weighted version of the Langevin equation of motion. 
+%The stochastic ``white noise'' of the Langevin equation acts to absorb force discontinuities while also thermostating the system to maintain the desired thermodynamic temperature. 
+%In addition, we introduce a simple heuristic approach to find optimized Langevin equation parameters to use for balancing system-specific imperfect forces and maintaining correct dynamical behavior.
 
@@
 
+\fi
+
 \maketitle
+
+\ifdefined\MAINTEXT
 
@@
 Since the unification of molecular dynamics and density functional theory (DFT)~\cite{a:thecpmd},
-\emph{ab initio} molecular dynamics (AIMD)~\cite{b:aimd} has become an important tool for the study of weakly bonded ensembles of molecules such as gas-phase molecular clusters, liquids, solutions, and molecular solids.
+\emph{ab initio} molecular dynamics (AIMD) has become an important tool for the study of weakly bonded ensembles of molecules such as gas-phase molecular clusters, liquids, solutions, and molecular solids.
 %[of gas-phase molecules and condensed-phase systems.]
@@
 %and hinders it application to large heterogeneous systems (interfaces, nuclei, defects).
-To address this issue and calculate forces in large systems efficiently, substantial efforts have been directed to the development of linear scaling (LS) DFT methods.
+To address this issue, substantial efforts have been directed to the development of linear scaling (LS) DFT methods.
 
 % Re-write the following section and include any recent advances
-In all LS DFT methods, the \emph{delocalized} eigenstates of the effective KS Hamiltonian, known as canonical KS orbitals, must be replaced with an alternative set of \emph{local} electronic descriptors
-~\footnote{Canonical KS orbitals are delocalized over all atoms in the system. 
-This delocalization is the main reason for the steep growth of the cost of the conventional KS DFT: since both the number of orbitals and the number of localization regions (e.g.
-atomic basis functions) increase linearly with the number of atoms, the total number of the electronic degrees of freedom that describe delocalized orbitals grows quadratically.}.
+In all LS DFT methods, the \emph{delocalized} eigenstates of the effective KS Hamiltonian must be replaced with an alternative set of \emph{local} electronic descriptors
+~\footnote{The delocalization of KS orbitals over all atoms is the main reason for the steep growth of the cost of KS DFT: since both the number of orbitals and the number of localization regions (e.g.
+atomic basis functions) increase linearly with the number of atoms, the total number of the electronic degrees of freedom grows quadratically.}.
 Most LS methods~\cite{a:linscale3,a:lee-yang-1996,a:ls-scuseria-1997,a:ls-manolopoulos-1998,a:ls-helgaker-2001,a:ls-niklasson-2003,a:curvy2,a:ls-dm-sign} explore the natural locality of the one-electron density matrix (DM).
-However, in accurate DFT calculations that require many basis functions per atom, the variational optimization of the DM is advantageous only for impractically large systems~\cite{a:ls-rev-1999,a:ls-dm-sign,a:almo-ls}.
+However, in typical DFT calculations that require diffuse basis functions, the DM approach is advantageous only for impractically large systems~\cite{a:ls-rev-1999,a:ls-dm-sign,a:almo-ls}.
 Therefore, the application of DM-based LS methods have been restricted to minimal-basis tight-binding problems. 
-The optimal basis variants of the DM methods~\cite{a:ls-stechel-1994,a:ls-gillan-1995,a:ls-gillan-1996,a:ls-onetep-2003} designed to rectify this issue contract large basis sets into a small number of new localized basis functions and then optimize the DM in the contracted basis.
+The optimal basis variants of the DM methods~\cite{a:ls-stechel-1994,a:ls-gillan-1995,a:ls-gillan-1996,a:ls-onetep-2003, RZK-MAO} designed to rectify this issue contract large basis sets into a small number of new localized basis functions and then optimize the DM in the contracted basis.
 Although such methods have been successfully used for the evaluation of accurate DFT energies of large static systems~\cite{a:ls-onetep-2009,a:ls-conquest-2010,a:ls-onetep-2010-app1,a:ls-rev-2012} their use in AIMD is hampered by the costly optimization of both the contracted orbitals and the density matrix.
 
-To obviate the DM optimization, another set of natural local electronic descriptors can be constructed by enforcing strict locality on the KS orbitals and, necessarily, by relaxing the orthogonality constraints imposed on them.
-From the computational point of view, LS methods based on the direct optimization of such compact localized nonorthogonal orbitals~\cite{a:ls-galli-parrinello-1992,a:ls-mauri-galli-car-1993,a:ls-ordejon-1993,a:ls-mauri-galli-1994,a:ls-ordejon-1995,a:ls-kim-mauri-galli-1995,a:ls-fattebert-2004,a:ls-fattebert-2006,a:burger-yang-2008} are very promising since they require only the occupied orbitals and, thus, deal with fewer variational degrees of freedom than the DM methods.
-Unfortunately, the development of orbital-based LS methods has been hindered by the inherently difficult optimization of localized orbitals reported by multiple research groups~\cite{a:ls-mauri-galli-car-1993,a:ls-ordejon-1995,a:ls-fattebert-2004,a:ls-rev-1999,a:weitao-yang-2013}.
+To obviate the DM optimization, another set of electronic descriptors can be constructed by enforcing strict locality on one-electron orbitals and, necessarily, by relaxing the orthogonality constraints imposed on them.
+From the computational point of view, LS methods based on the direct optimization of such compact localized nonorthogonal orbitals~\cite{a:ls-galli-parrinello-1992,a:ls-mauri-galli-car-1993,a:ls-ordejon-1993,a:ls-mauri-galli-1994,a:ls-ordejon-1995,a:ls-kim-mauri-galli-1995,a:ls-fattebert-2004,a:ls-fattebert-2006,a:burger-yang-2008} are promising since they deal with fewer variational degrees of freedom than the DM methods.
+Unfortunately, the development of orbital-based LS methods has been all but abandoned because of the inherently difficult optimization of localized orbitals~\cite{a:ls-mauri-galli-car-1993,a:ls-ordejon-1995,a:ls-fattebert-2004,a:ls-rev-1999,a:weitao-yang-2013}.
 
-%ZZZ-important addition: we did not include a large family of the so-called fragmentation methods 
+%RZK-important addition: we did not include a large family of the so-called fragmentation methods 
 %included only those methods that attempt to construct  set 
@@
 
-Thus, despite remarkable recent progress the high computational overhead of existing asymptotically linear-scaling DFT methods restrict their use in AIMD simulations to very short \emph{time scales}, systems of low dimensions, or problems that can be adequately described with minimal basis sets.
-On the length and time scales required for accurate routine AIMD simulations, LS DFT cannot compete with the straightforward low-cost cubically-scaling KS DFT.
+Thus, the high computational overhead of existing asymptotically linear-scaling DFT methods restrict their use in AIMD simulations to very short \emph{time scales}, systems of low dimensions, or problems that can be adequately described with minimal basis sets.
+On the length and time scales accessible in typical AIMD simulations, LS DFT cannot compete with the straightforward low-cost cubically-scaling KS DFT.
 
-In this work, we introduce a new orbital-based AIMD method that is not only linear scaling but also has a low computational overhead. 
-Currently, the method is restricted to large ensembles of atoms or small molecules that do not form strong covalent bonds with each other. 
-Using liquid water as an example, we demonstrate that the new AIMD methodology enables computational studies of molecular systems on previously inaccessible time and length scales.
-
+In this work, we developed a new orbital-based AIMD method that is not only linear scaling but also has a low computational overhead. 
+Currently, the method is restricted to systems of small molecules that do not form strong covalent bonds with each other. 
 The new AIMD method utilizes a recently developed LS DFT~\cite{a:almo-ls} based on absolutely localized molecular orbitals (\mbox{ALMOs}). 
@@
 However the approach can be easily generalized to other local and nonlocal basis sets~\cite{a:ls-galli-parrinello-1992, Galerkin}. 
-The key feature of ALMO DFT is that its one-electron wavefunctions are constructed variationally in a two-step self-consistent-field (SCF) procedure~\cite{a:almo-ls} to circumvent the problem of the sluggish variational optimization that have rendered previous MO-based LS methods impractical.
+The key feature of ALMO DFT is that its one-electron wavefunctions are constructed in a two-step self-consistent-field (SCF) procedure~\cite{a:almo-ls} to circumvent the problem of the sluggish variational optimization emphasized above.
 In the first step, ALMOs are constrained to their localization centers~\cite{a:khal} whereas, in the second step, ALMOs are relaxed to allow delocalization onto the neighbor molecules within their localization radius $R_{c}$.
@@
 %Setting $R_{c}$ to larger values can reduce the magnitude of the discontinuity, but increases the computational cost as longer distance (but physically less important) interactions must be calculated.
-%As long as the discontinuity in force is small compared with the random numerical noise in the calculation, it can be safely ignored~\cite{ZZZ-any-references}.
+%As long as the discontinuity in force is small compared with the random numerical noise in the calculation, it can be safely ignored~\cite{RZK-any-references}.
 This means that $R_{c}$ has to be kept relatively large to maintain stable and accurate dynamics.
@@
 $\delta f_{i\alpha} (t)$ comprises all neglected terms that originate from a finite localization radius $R_c$ and incomplete SCF optimization.
-$\delta f_{i\alpha} (t)$ can be reduced \emph{systematically} to zero by increasing $R_c$ (Figure~\ref{fig:forcecomp} in the Supplemental Material) and decreasing $\epsilon_{\text{SCF}}$.
+$\delta f_{i\alpha} (t)$ can be reduced \emph{systematically} to zero by increasing $R_c$ and decreasing $\epsilon_{\text{SCF}}$.
 %$\epsilon_{\text{SCF}}$ is chosen such that the electronic ground state is considered sufficiently close to the Born-Oppenheimer state for values less than $\epsilon_{\text{SCF}}$.
-%ZZZ For $\epsilon_{\text{SCF}} \leqslant 10^{-7}$~a.u. ALMO SCF is considered converged.
+%RZK For $\epsilon_{\text{SCF}} \leqslant 10^{-7}$~a.u. ALMO SCF is considered converged.
 
@@
 \label{eq:langevin}
-m_i \ddot{r}_{i\alpha} = f^{\text{KS}}_{i\alpha} - \gamma m_i \dot{r}_{i\alpha} + R^{\gamma}_{i\alpha} (t),
+m_i \ddot{r}_{i\alpha} = f^{\text{KS}}_{i\alpha}(t) - \gamma m_i \dot{r}_{i\alpha} + R^{\gamma}_{i\alpha} (t),
 \end{align}
@@
 
-The main assumption of ALMO AIMD is that the error in the ALMO forces is well approximated by a stochastic white Gaussian noise $R^{\Delta}_{i\alpha} (t)$:
+The main assumption of ALMO AIMD is that the error in the ALMO forces is well approximated by Gaussian noise $R^{\Delta}_{i\alpha} (t)$:
 %
@@
 \label{eq:assumption}
-\delta f_{i\alpha} = R^{\Delta}_{i\alpha} (t)
+\delta f_{i\alpha}(t) = R^{\Delta}_{i\alpha} (t)
 \end{align}
@@
 %
-This assumption, shown below to be well justified (Figure~\ref{fig:randomforce}), allows us to write the Langevin equations using the approximate ALMO forces:
+This assumption, shown below to be well justified, allows us to rewrite the Langevin equations using the ALMO forces:
 %
@@
 \label{eq:langevin2}
-m_i \ddot{r}_{i\alpha} = f^{\text{ALMO}}_{i\alpha} - \gamma m_i \dot{r}_{i\alpha} + R^{\gamma + \Delta}_{i\alpha} (t)
+m_i \ddot{r}_{i\alpha} = f^{\text{ALMO}}_{i\alpha}(t) - \gamma m_i \dot{r}_{i\alpha} + R^{\gamma + \Delta}_{i\alpha} (t)
 \end{align}
@@
 where the two stochastic terms are combined into one $R^{\gamma + \Delta}_{i\alpha} = R^{\gamma}_{i\alpha} + R^{\Delta}_{i\alpha}$. 
-This new Langevin equation means that the value of $\Delta$ must be known to re-balance the damping and stochastic terms and generate the correct Maxwell-Boltzmann distribution in simulations.
+The only missing piece in the modified Langevin equation is the value of $\Delta$, which describes the strength of the newly introduced stochastic term. 
+This term compensates for the imperfection of ALMO forces and is required to re-balance the damping and stochastic components in the Langevin equation and generate the correct Maxwell-Boltzmann distribution in ALMO AIMD.
 
-Note that $\Delta$ can be either positive or negative depending on the way errors are introduced in the system of interest. 
-$\Delta$ was found to be positive for all systems considered here, increasing the amplitude of the effective random force on the system.
+%%Note that $\Delta$ can be either positive or negative depending on the system of interest. 
+%%$\Delta$ was found to be positive for all simulations considered here. %, increasing the amplitude of the effective random force on the system.
 
-%\section{Optimizing the values of $\gamma$ and $\Delta$}
+%\section{Optimizing $\Delta$}
 
-In theory, both $\gamma$ and $\Delta$ should be system-size independent, but in practice only $\gamma$ is.
-We found that $\Delta$ is dependent on system size due to finite size effects, which decay for increasing system size.
-Fortunately, $\gamma$ and $\Delta$ can be separately optimized for a given system of interest, assuming a method of accurate force calculation can be applied to the system for a short trajectory.
+%To correctly thermostat the system of interest, $\Delta$ needs to be reduced to zero.
+%This is done by finding the value of $\Delta$ and adding a third random force term $R^{-\Delta}_{i\alpha}$ to equation \ref{eq:langevin2} which cancels out the extra forces introduced to the system by $R^{\Delta}_{i\alpha}$.
+%This means that a trajectory with inaccurate forces and optimized $R^{-\Delta}_{i\alpha}$ should have no long term drift in the total energy of the system. 
+%Short term energy fluctuations should be expected, and can later be ``smoothed out'' by setting $\gamma$ as needed.
 
-An optimized $\gamma$ is one that is: large enough to adequately absorb force discontinuities introduced by finite $R_{c}$; large enough to correctly thermostat the system of interest on a reasonable time scale; and small enough not to significantly affect dynamical properties on the time scale of interest.
-The relative weight of the three is determined by the specifics of the system and what is most important to the researcher. 
-To optimize $\gamma$ given the relative importance of the above properties, we suggest running several short trajectories on the system of interest (or a scaled-down version of it) using a tightly converged potential energy surface calculation and varying the value of $\gamma$.
-We found that a value of $\gamma$ between $10^{-5}$ and $10^{-3}\ \mathrm{fs^{-1}}$ was optimal for the purposes of the research presented here.
-
-To correctly thermostat the system of interest, $\Delta$ needs to be reduced to zero.
-This is done by finding the value of $\Delta$ and adding a third random force term $R^{-\Delta}_{i\alpha}$ to equation \ref{eq:langevin2} which cancels out the extra forces introduced to the system by $R^{\Delta}_{i\alpha}$.
-This means that a trajectory with inaccurate forces and optimized $R^{-\Delta}_{i\alpha}$ should have no long term drift in the total energy of the system. 
-Short term energy fluctuations should be expected, and can later be ``smoothed out'' by setting $\gamma$ as needed.
-
-$R^{-\Delta}_{i\alpha}$ can therefore be optimized simply by trying various values of $\Delta$ (with \mbox{$\gamma = 0$}) until the simulation has long-term stable total energy and satisfies the equipartition theorem.
+(RZK the integral contains dummy index $i$ in the r.h.s.) In principle, $\Delta$ can be calculated using the integral of Eq.~(\ref{eq:stochastic3})
+%
+\begin{align}
+\label{eq:delta}
+\Delta &= (2 k_B T m_i )^{-1} \int_{-\infty}^{\infty}\langle \delta f_{i\alpha} (0)  \delta f_{i\alpha} (t) \rangle dt
+\end{align}
+%
+if one can afford computing the reference forces $f^{\text{KS}}_{i\alpha}(t)$ (i.e.
+$R_c \rightarrow \infty$ and $\epsilon_{\text{SCF}} \rightarrow 0$) for a short representative AIMD trajectory (see Supplemental Material).
+In practice, we found (see results below) that this approach is not particularly accurate due to the noisy autocovariance functions (ACF) from short trajectories and perhaps because the $\delta_{ij}\delta_{\alpha\beta}$ assumption in Eq.~(\ref{eq:stochastic3}) does not strictly hold.
+Nevertheless, the ACF integral can provide a reasonable starting value of $\Delta$. 
+This or, in fact, any starting value can be  rapidly fine-tuned in a series of trial-and-error short ALMO AIMD runs until the equipartition theorem is satisfied: 
 %
@@
 %
+or, in other words, until the simulation stabilizes with the correct average temperature.
 
-Another approach based on Eq. (\ref{eq:assumption}) can be used to find $\Delta$ within an order of magnitude if one can afford computing tightly converged SCF forces for a short MD trajectory. 
-First, run the simulation with $\Delta = 0$ and converged SCF forces. 
-Second, compute approximate forces for snapshots along the trajectory and calculate the random term $R^{\Delta}_{i\alpha} (t)$ using Eq.\ \ref{eq:assumption}. 
-Third, calculate the autocovariance function (ACF) of the noise $\text{ACF}(t) = \langle R^{\Delta}_{i\alpha} (0)  R^{\Delta}_{i\alpha} (t) \rangle$, which should be sharply peaked around zero lag time if our assumption is correct.
-Finally, integrate Eq.(\ref{eq:stochastic3}) to obtain the following expression for $\Delta$.
+%\section{Implementation}
+
+ALMO AIMD presented here was implemented in CP2K, an open source materials modeling package~\cite{www:cp2k}. 
+Accuracy and efficiency of ALMO AIMD was tested using liquid water as an example. 
+This system is challenging because intermolecular electron delocalization is a critical component of hydrogen bonding and must be described correctly in approximate ALMO AIMD to reproduce static and dynamical properties of liquid water.
+%To calculate properties of water at ambient conditions, 
+A periodic cell containing 125 molecules was simulated for $\SI{10}{\ps}$ at a temperature of $\SI{298}{\K}$ and a constant density of $\SI{1.006}{\g\cdot\cm^{-3}}$. 
+% RZK: we do not need to specify details on gamma optimization since they are well-known.
+RZK algorithm~\cite{RZK} was used to integrate the Langevin equation of motion.
+We found that $\gamma = \SI{e-3}{\per\fs}$ is large enough to thermostat the system on the time scale of the simulation and small enough not to significantly affect dynamical properties of liquid water.
 %
-\begin{align}
-\label{eq:delta}
-\Delta &= (2 k_B T m_i )^{-1} \int_{-\infty}^{\infty}\langle R^{\Delta}_{i\alpha} (0)  R^{\Delta}_{i\alpha} (t) \rangle dt
-\end{align}
+In the dual-representation Gaussian and plane-wave scheme implemented in CP2K~\cite{a:quickstep}, the TZV2P basis set was used to represent molecular orbitals, and a plane-wave cutoff of $\SI{320}{Ry}$ used to represent electron density.
+The XC energy was approximated using the PBE functional~\cite{a:PBEfunctional}. 
+The Brillouin zone was sampled at the $\Gamma$-point and separable norm-conserving pseudopotentials were used to describe the interactions between the valence electrons and the ionic cores~\cite{a:hgh}.
+
+The \emph{reference} forces were calculated with fully delocalized electrons using tightly converged, $\epsilon_{\text{SCF}}=10^{-6}$~a.u., orbital transformation (OT) method~\cite{a:ot,a:ot2}. 
+In ALMO AIMD, element-specific cutoff radius for electron delocalization $R_c$ was set to 1.6 in units of the elements' van der Waals radii (vdWR).
+This localization radius includes approximately two coordination shells of an average water molecule and was shown to reproduce the KS reference RDF perfectly in Monte-Carlo simulations~\cite{a:almo-ls}.
 %
-In other words, we use the integral of the ACF to estimate the strength of the random force. 
-In practice, we found this method for estimating $\Delta$ was not particularly accurate due to rapidly oscillating correlations between sequential values of $R^{\Delta}_{i\alpha} (t)$. 
-The ACF was not possible to accurately integrate without running a reference trajectory with extremely short time-steps (see figure~\ref{fig:randomforce}).
-For example, using Eq.\ \ref{eq:delta} we calculated $\Delta \approx \SI{2e-5}{\per\fs}$, but the value was optimized heuristically to $\SI{6e-5}{\per\fs}$.
-Still, this method provided a decent order of magnitude starting point for further investigation.
-Note that for our system, the ACF decays rapidly enough that the forces can be considered random and uncorrelated to a good approximation for sufficiently long time scales ($\gg\SI{200}{\fs}$).
+To measure the ability of the $R^{\Delta}$ term to compensate for imperfections in ALMO forces we varied $\epsilon_{\text{SCF}}$ between tight $10^{-6}$~a.u. and loose $10^{-2}$~a.u.
+
+%The computation scheme used is mostly identical to that described in the implementation section of ref.\ \citenum{a:almo-ls}. 
+%at $\SI{302}{\K}$ and  --- RZK: It is customary to report TARGET temperature, not the acutal average temperature.
+The fact the the actual temperature deviates from the target temperature by 5 K even with perfect forces shows that our simulation is still too short.
+%with the Langevin thermostat with $\gamma = \SI{e-3}{\per\fs}$. 
+%All other system parameters were set identical to the ALMO system.
+
+Even with $\epsilon_{\text{SCF}} = 10^{-2}$, we stabilized the simulation at the correct average temperature and perfect Maxwell-Boltzmann distribution (Figure~\ref{fig:dynproperties}b).
+$\Delta$ was initially estimated at $\SI{2e-5}{\per\fs}$ using Eq.~(\ref{eq:delta}) and then refined heuristically to $\SI{6e-5}{\per\fs}$.
+%RZK: factor of 3 bothers me. Check that this is not because we sum over 3 atoms in a molecule (or over 3 dimensions).
+We found that it is easier to optimize $\Delta$ when $\gamma$ is set to zero because unnecessary stochastic noise is eliminated in the trial runs.
+%In theory, $\Delta$ should be independent on system size.
+However, we found that $\Delta$ dependents on system size due to finite size effects, which decay for increasing system size.
+
+\begin{figure}
+\includegraphics[trim={1.3cm 0.1cm 3.3cm 1.3cm},clip,width=8.6cm]{Dynamical_Data_Tiled.eps}
+\caption{\label{fig:dynproperties} RZK:.
+Calculated properties of water using ALMO DFT with $\epsilon_{\text{SCF}} = 10^{-2}$~a.u.
+and $R_{c} = 1.6$ vdWR (red line) and fully converged OT reference (black line).
+(a) radial distribution function, 
+(b) kinetic energy distribution (the black curve shows the theoretical Maxwell-Boltzmann distribution), 
+(c) infrared spectrum.
+%Calculations were performed with the PBE $\mathrm{E_{XC}}$ functional and TZV2P basis set, using 125 water molecules for 10 ps.
+}
+\end{figure}
+
+Analysis of $\delta \vec{f_{i}}(t)$ shows that the error in forces indeed resembles Gaussian white noise.
+The mean of the error is zero (black line in Figure~\ref{fig:randomforce}c) and the ACF decays rapidly (Figure~\ref{fig:randomforce}d) so that the error in forces can be considered uncorrelated on time scale of $\SI{50}{\fs}$.
+Thus the main assumptions behind our approach to ALMO AIMD are justified for liquid water.
+Figure~\ref{fig:randomforce}b shows that the main source of error in forces in this simulation is the loose convergence criterion and not the finite $R_c$: fully converged ALMO SCF calculations remove the rapidly oscillating component of the error.
+We also confirmed that the ALMO forces converge to the reference forces in the limit $R_{c} \rightarrow \infty$ (Figure~\ref{fig:forcecomp} in the Supplemental Material).
+
+%we computed the average difference in force between fully converged Born-Oppenheimer forces and forces calculated from ALMO DFT with varying $R_{c}$.
+Figure \ref{fig:forcecomp} in the Supplemental Material shows that as $R_{c}$ is increased to infinity, forces converge toward the Born-Oppenheimer limit.
 
@@
 \includegraphics[trim={0.5cm 0cm 0.7cm 0.1cm},clip,width=8.6cm]{DeltaForceComparison_with_ACF.eps}
-\caption{\label{fig:randomforce} Comparison of random force introduced from $R_{c}$ and $\epsilon_{\text{SCF}}$. 
+\caption{\label{fig:randomforce}  
 The red line is the Euclidean norm of the instantaneous error $\langle \| \delta \vec{f_{i}}(t) \| \rangle_{i}$, the black line is the magnitude of the time average of the instantaneous error vector, and the green line is the time average of the red line.
-(a) $R_{c} = 1.6$ vdWR and $\epsilon_{\text{SCF}} = 10^{-2}$~a.u., using fully converged ALMO SCF as a reference.
-(b) $R_{c} = 1.6$ vdWR and fully converged ALMO SCF, using fully converged OT SCF as a reference.
-(c) $R_{c} = 1.6$ vdWR and $\epsilon_{\text{SCF}} = 10^{-2}$~a.u., using fully converged OT SCF as a reference.
+(a) $R_{c} = 1.6$ vdWR and $\epsilon_{\text{SCF}} = 10^{-2}$~a.u. relative to the fully converged ALMO reference, 
+(b) $R_{c} = 1.6$ vdWR and fully converged ALMO SCF, relative to the fully converged OT reference, 
+(c) $R_{c} = 1.6$ vdWR and $\epsilon_{\text{SCF}} = 10^{-2}$~a.u., relative to the fully converged OT reference, 
 (d) Normalized ACF of the instantaneous error in panel (a) $\frac{1}{3}\langle \delta \vec{f}_i (t) \cdot \delta\vec{f}_i(t+\tau) \rangle_{it} $.
-Forces are computed with the PBE XC functional and TZV2P basis set for the configurations from a 10-ps trajectory generated with KS-based AIMD at T=298~K. 
+%Forces are computed with the PBE XC functional and TZV2P basis set for the configurations from a 10-ps trajectory generated with KS-based AIMD at T=298~K. 
 }
@@
 
-%\section{Implementation}
-
-The methods presented here are built into CP2K, an open source MD simulation package.~\cite{www:cp2k} 
-The computation scheme used is mostly identical to that described in the implementation section of ref.\ \citenum{a:almo-ls}. 
-For the dynamical properties calculations using ALMO DFT methods, a periodic cell containing 125 water molecules was simulated for $\SI{10}{\ps}$ at a temperature of $\SI{293}{\K}$ and a constant density of $\SI{1.00597}{\g\per\cm^{3}}$.
-The system was thermostated using the Langevin ensemble (eq.\ \ref{eq:langevin2}) where $\gamma$ was set to $\SI{e-3}{\per\fs}$, and $\Delta$ was optimized to the order of $10^{-5}\, \mathrm{fs^{-1}}$.
-Note that for timing benchmarks, $\Delta$ was not optimized. 
-This does not have a perceptible effect on calculation time.
-
-Simulations for timing benchmarks were run for ten time steps, which is sufficient for convergence of wavefunctions from an initial guess.
-The TZV2P basis set was used to represent MOs, and a cutoff of $\SI{320}{Ry}$ used to represent electron density. 
-The exchange-correlation energy was approximated using the PBE functional, which expands upon the local density approximation by considering derivatives of the electron density at each point.~\cite{a:PBEfunctional}
-
-The reference electronic structure was calculated using the orbital transformation (OT) method~\cite{a:ot,a:ot2} at $\SI{302}{\K}$ and the Langevin thermostat with $\gamma = \SI{e-3}{\per\fs}$. 
-All other system parameters were set identical to the ALMO system.
-
-IR spectra, radial distribution functions, diffusivity plots, and diffusion constants were calculated using the open source trajectory analyzer program TRAVIS.~\cite{a:travis-main,a:travis-ir1,a:travis-ir2}
-
 %\section{Accuracy}
 
-In order to confirm that ALMO DFT forces are accurate in the limit where localization radius $R_{c} \rightarrow \infty$, we computed the average difference in force between fully converged Born-Oppenheimer forces and forces calculated from ALMO DFT with varying $R_{c}$.
-Figure \ref{fig:forcecomp} in the Supplemental Material shows that as $R_{c}$ is increased to infinity, forces converge toward the Born-Oppenheimer limit.
-
-To test the accuracy of our unconverged ALMO based MD we computed the Infrared (IR) spectrum, radial distribution function (RDF), kinetic energy distribution, and mean-squared displacement (MSD) function from a 10 ps MD trajectory (see fig.
-\ref{fig:dynproperties} and \ref{fig:diffusion} in the Supplemental Material).
-We compared the IR spectrum, RDF, and MSD function to corresponding reference functions calculated with fully converged KS DFT forces.
-The kinetic energy distribution is compared with the theoretical Maxwell-Boltzmann distribution at the average temperature of the system.
-From the slope of the MSD functions we also calculated diffusion constants.
-Using fully converged KS DFT forces we calculated D = $\SI{5.47e-9}{\m^{2}\per\s}$, whereas with unconverged ALMO forces we calculated D = $\SI{5.55E-09}{\m^{2}\per\s}$.
-
-Although the accuracy of calculated properties in figure \ref{fig:dynproperties} is quite good, it is clear that there are errors present.
-Some of the error can be attributed to the 9 K difference in the thermodynamic temperature between the reference system and our system of interest.
-Other sources of error present: small temporal sample size given our choice of $\gamma$; random error due to finite localization radius $R_{c}$; and random error due to high $\epsilon_{\text{SCF}}$.
-From figure \ref{fig:randomforce}, it's apparent that high values of $R_{c}$ and $\epsilon_{\text{SCF}}$ both introduce substantial error in calculation of forces.
-However, raising both $R_{c}$ and $\gamma$ decreases computation time significantly, and it is ultimately up to the user to determine their maximum tolerance of force inaccuracy when using the ALMO DFT method presented here.
-
-\begin{figure}
-\includegraphics[trim={1.3cm 0.1cm 3.3cm 1.3cm},clip,width=8.6cm]{Dynamical_Data_Tiled.eps}
-\caption{\label{fig:dynproperties} Calculated dynamic and static properties of water using ALMO DFT with $\epsilon_{\text{SCF}} = 10^{-2}$~a.u. and $R_{c} = 1.6$ vdWR.
-(a) Radial distribution function of an ALMO simulation compared to fully converged OT SCF. 
-(b) Kinetic energy distribution of an ALMO simulation compared to the theoretical Maxwell-Boltzmann kinetic energy distribution.
-(c) Calculated Infrared spectrum of an ALMO simulation compared to fully converged OT SCF.
-Calculations were performed with the PBE $\mathrm{E_{XC}}$ functional and TZV2P basis set, using 125 water molecules for 10 ps.}
-\end{figure}
+To test the accuracy of ALMO AIMD we used trajectory analyzer TRAVIS~\cite{a:travis-main} to compute the infrared (IR) spectrum, radial distribution function (RDF), and diffusion coefficient of liquid water from both the ALMO trajectory ($\epsilon_{\text{SCF}} = 10^{-2}$~a.u.
+and $R_{c} = 1.6$ vdWR) and from the reference trajectory.
+The diffusion coefficients $D_{\text{OT}}=\SI{5.47e-9}{\m^{2}\per\s}$ and $D_{\text{ALMO}}=\SI{5.55E-09}{\m^{2}\per\s}$ and RDFs (Figure~\ref{fig:dynproperties}a) are in perfect agreement.
+The quality of the ALMO IR spectrum (Figure~\ref{fig:dynproperties}c) is good despite minor errors in the intensity of the OH stretching mode, which is sensitive to the precise positions of the centers of localized orbitals.
+These stringent tests show that despite noticeable errors in the ALMO forces (Figure \ref{fig:randomforce}) the compensating $R^{\Delta}$ term in the modified Langevin equation makes it possible to recover atomic dynamics properly.
+We would like to note that ALMO AIMD could not be stabilized with $\Delta=0$.
+We were not able to find any values of $\Delta$ that stabilize trajectories generated using perturbative versions of ALMO DFT~\cite{a:almo-ls}.
 
@@
 
-To demonstrate the computational efficiency gained by using unconverged ALMO DFT forces, we compare the average wall-time per MD step over 10 time steps for various values of $\epsilon_{\text{SCF}}$ in figure \ref{fig:strongscaling_log}.
-For these simulations, we chose to set the localization radius to $R_{c} = 1.6$ van der Waals radii (vdWR), as was done for calculation of dynamical properties.
-We also show timing of the fully converged OT SCF method, as well as the recently developed ``second generation Car-Parinello'' MD method, which improves upon the OT computational cost prefactor.~\cite{a:2ndcpmd}
-ALMO methods show clear asymptotic LS behavior for all values of $\epsilon_{\text{SCF}}$ tested.
-OT SCF scales approximately cubically, as expected.
-Our findings show that the fastest ALMO DFT methods presented here runs faster than second-generation Car-Parinello using OT DFT for water simulations larger than 256 molecules.
+%RZK Simulations for timing benchmarks were run for ten time steps, which is sufficient for convergence of wavefunctions from an initial guess.
+
+To demonstrate the computational advantage of ALMO AIMD, we compared the average wall-time per MD step over 10 time steps for a variety of methods in Figure \ref{fig:strongscaling_log}.
+In addition to the fully converged OT reference, we included timing for the second generation Car-Parrinello method, which decreases computational prefactor of the cubically scaling OT method~\cite{a:2ndcpmd}.
+As expected ALMO methods show clear LS behavior for all values of $\epsilon_{\text{SCF}}$ even for medium-size systems. 
+%OT SCF scales approximately cubically, as expected.
+For liquid water, strict localization of molecular orbitals makes ALMO AIMD faster than the second-generation Car-Parinello for larger than 256 molecules.
+It is important to emphasize that comparison is performed for a three-dimensional condensed phase system described with a large diffuse basis set -- a particularly challenging case for LS methods.
 
@@
 \includegraphics[trim={2.5cm 0.5cm 3.4cm 0.1cm},clip,width=8.6cm]{strongscaling_log.eps}
-\caption{\label{fig:strongscaling_log} Logscale plots which show the asymptotic LS behaviour of ALMO methods for liquid water at 298 K.
-Compared with the orbital transformation using fully converged SCF and second generation Car-Parinello force calculation methods.~\cite{a:ot,a:ot2,a:2ndcpmd}
+\caption{\label{fig:strongscaling_log} Timing benchmarks for PBE/TZV2P simulations of liquid water on 256 compute cores. 
+%Logscale plots which show the asymptotic LS behaviour of ALMO methods for liquid water at 298 K.
+%Compared with the orbital transformation using fully converged SCF and second generation Car-Parinello force calculation methods.~\cite{a:ot,a:ot2,a:2ndcpmd}
+For ALMO methods, $R_{c} = 1.6$ vdWR. 
 Cyan lines represent perfect cubic scaling, whereas gray lines represent perfect linear scaling. 
-Calculations were performed with the PBE $\mathrm{E_{XC}}$ functional and TZV2P basis set on 256 cores, averaged over 10 MD steps. 
-The localization radius was set to $R_{c} = 1.6$ vdWR.}
+%RZK: On figure remove 'DFT', 'Wall-clock time' -> 'Wall time'
+}
 \end{figure}
 
-Figure \ref{fig:weakscaling} shows that ALMO MD code is well optimized for parallelization, and that the computational cost grows linearly under a wide range of available computing power. 
-In fact, we were able to successfully simulate systems as large as 32768 water molecules within 25 minutes of wall-clock time using 4096 cores.
+Weak scaling benchmarks show (Figure \ref{fig:weakscaling}) that computation with localized orbitals is naturally suited for parallel execution and that the computational cost of our implementation grows linearly for a wide range of compute cores.
+It also demonstrates that the range of routine AIMD simulations can be extended from $\sim 10^3$ to $\sim 10^4$ atoms making completely new phenomena such as dynamics on interfaces and nucleation processes accessible on HPC platforms today.
+%We also show that systems as large as 8192 molecules can be simulated for 10,000 MD steps in under a month using this method.
+In fact, we were able to successfully simulate systems as large as 32768 water molecules within reasonable wall-clock time using only moderate number of compute cores (Figure \ref{fig:weakscaling}) -- unprecedented feat for AIMD considering that accurate molecular orbitals and the idempotent DM are computed on each step.
 
@@
 \includegraphics[trim={1.6cm 0cm 4.7cm 0cm},clip,width=8.6cm]{weakscaling.eps}
-\caption{\label{fig:weakscaling} Weak scalability benchmarks showing optimization of code for parallel computing.
-Dashed gray lines connect systems simulated using the same number of cores, verifying LS behaviour over a range of configurations.
-Of note is the 32768 molecule system successfully simulated using 4096 cores at around 25 minutes per MD step.
-We also show that systems as large as 8192 molecules can be simulated for 10,000 MD steps in under a month using this method.
-Calculations were performed with the PBE $\mathrm{E_{XC}}$ functional and TZV2P basis set, averaged over 10 MD steps. 
-The localization radius was set to $R_{c} = 1.6$ vdWR and $\epsilon_{\text{SCF}} = 10^{-2}$~a.u.}
+\caption{\label{fig:weakscaling} Weak scalability benchmarks for PBE/TZV2P ALMO AIMD with  $R_{c} = 1.6$ vdWR and $\epsilon_{\text{SCF}} = 10^{-2}$~a.u. 
+Dashed gray lines connect systems simulated using the same number of compute cores to confirm LS behavior. %over a range of configurations.
+The horizontal line corresponds to 10,000 AIMD steps in one wall month and shows a upper-bound guideline for practical AIMD.
+%RZK: On figure 'Wall-clock time' -> 'Wall time', legend - increase font size, decrease the height of the figure(the y-axis label is now shorter)
+}
 \end{figure}
@@
 
+RZK: systematically improvable approximation to DFT-based AIMD.
 In this article we have clearly demonstrated the utility of ALMO DFT methods, especially when combined with the Langevin Equation of motion.
@@
 
-To summarize, a low-cost linear-scaling AIMD method for weakly-interacting molecular systems was developed by utilizing compact molecular orbitals strictly localized within predefined radii.
+%Using liquid water as an example, we demonstrate that the new AIMD methodology enables computational studies of molecular systems on previously inaccessible time and length scales.
+
+%To summarize, a low-cost linear-scaling AIMD method for weakly-interacting molecular systems was developed by utilizing compact molecular orbitals strictly localized within predefined radii.
 High efficiency of the method is achieved without sacrificing its accuracy with a combination of two techniques: (1) on-the-fly construction of accurate localized orbitals without lengthy self-consistent optimization and (2) a modified Langevin equation that is fine-tuned to retain stable dynamics even with imperfect forces.
@@
 
-Generalization of the method to more challenging systems of strongly interacting atoms (e.g. covalent crystals) will also be explored in the future.
-
-%Such simulations are important for Use if empirical force fields or QM/MM are not applicable:
-%Chemical reactions that happen throughout your system, not in a localized spatial region (e.g. proton transfer in water)
-%Unusual thermodynamic states for which simple analytical potential become inaccurate (e.g. high pressure states, interfaces, heterojunctions)
-%Electron density is an important observable (e.g. X-ray scattering) or information of the electronic wavefunctions is crucial.
+Generalization of the method to more challenging systems of strongly interacting atoms such as covalent crystals is underway.
 
@@
 
-%ZZZ outline procedure to compute forces using the DM obtained from ALMOs.
-% ZZZ figure to show
-
 \bibliography{almo-langevin}
 
-\clearpage
-\widetext
-\setcounter{figure}{0}
-\setcounter{page}{1}
-\renewcommand{\thefigure}{S\arabic{figure}}
-\begin{center}
-\textbf{\large Supplemental Material}
-\end{center}
+\else
+
+%\section{Procedure to estimate the value of $\Delta$}
+%
+%Another approach based on Eq. (\ref{eq:assumption}) can be used to find $\Delta$ within an order of magnitude if one can afford computing tightly converged SCF forces for a short MD trajectory. 
+%First, run the simulation with $\Delta = 0$ and converged SCF forces. 
+%Second, compute approximate forces for snapshots along the trajectory and calculate the random term $R^{\Delta}_{i\alpha} (t)$ using Eq.\ \ref{eq:assumption}. 
+%Third, calculate the autocovariance function (ACF) of the noise $\text{ACF}(t) = \langle R^{\Delta}_{i\alpha} (0)  R^{\Delta}_{i\alpha} (t) \rangle$, which should be sharply peaked around zero lag time if our assumption is correct.
+%Finally, integrate Eq.(\ref{eq:stochastic3}) to obtain the following expression for $\Delta$.
+%%
+%\begin{align}
+%\label{eq:delta}
+%\Delta &= (2 k_B T m_i )^{-1} \int_{-\infty}^{\infty}\langle R^{\Delta}_{i\alpha} (0)  R^{\Delta}_{i\alpha} (t) \rangle dt
+%\end{align}
+%%
+%In other words, we use the integral of the ACF to estimate the strength of the random force.
 
@@
 
-\begin{figure}[!htb]
-\includegraphics[trim={1.7cm 0.5cm 3.3cm 0.1cm},clip,width=\textwidth]{strongscaling_linear.eps}
-\caption{\label{fig:strongscaling_linear} Linear plots showing LS behaviour of all ALMO methods for liquid water at 298 K. 
-Calculations were performed with the PBE $\mathrm{E_{XC}}$ functional and TZV2P basis set, averaged over 10 MD steps. 
-The localization radius was set to $R_{c} = 1.6$ vdWR and $\epsilon_{\text{SCF}} = 10^{-2}$~a.u.}
-\end{figure}
+%\begin{figure}[!htb]
+%\includegraphics[trim={1.7cm 0.5cm 3.3cm 0.1cm},clip,width=\textwidth]{strongscaling_linear.eps}
+%\caption{\label{fig:strongscaling_linear} Linear plots showing LS behaviour of all ALMO methods for liquid water at 298 K. 
+%Calculations were performed with the PBE $\mathrm{E_{XC}}$ functional and TZV2P basis set, averaged over 10 MD steps. 
+%The localization radius was set to $R_{c} = 1.6$ vdWR and $\epsilon_{\text{SCF}} = 10^{-2}$~a.u.}
+%\end{figure}
 
-\begin{figure}[!htb]
-\includegraphics[trim={2.5cm 0.2cm 4.5cm 1.2cm},clip,width=\textwidth]{Diffusion_Comparison.eps}
-\caption{\label{fig:diffusion} Calculated mean-squared displacement function of water using ALMO DFT with $\epsilon_{\text{SCF}} = 10^{-2}$~a.u. and $R_{c} = 1.6$ vdWR.
-Compared with equivalent MD run using OT DFT as a reference.
-Calculations were performed at 298 K with the PBE $\mathrm{E_{XC}}$ functional and TZV2P basis set, using 125 water molecules for 10 ps.}
-\end{figure}
+%\begin{figure}[!htb]
+%\includegraphics[trim={2.5cm 0.2cm 4.5cm 1.2cm},clip,width=\textwidth]{Diffusion_Comparison.eps}
+%\caption{\label{fig:diffusion} Calculated mean-squared displacement function of water using ALMO DFT with $\epsilon_{\text{SCF}} = 10^{-2}$~a.u. and $R_{c} = 1.6$ vdWR.
+%Compared with equivalent MD run using OT DFT as a reference.
+%Calculations were performed at 298 K with the PBE $\mathrm{E_{XC}}$ functional and TZV2P basis set, using 125 water molecules for 10 ps.}
+%\end{figure}
+
+\fi % Manuscript <-> SI switch
 

### 2017-10-02 almo-langevin.tex
No abstract

@@
 \author{Hayden Scheiber}
+% RZK update email
 \email{hayden.scheiber@mcgill.ca}
@@
 \begin{abstract} 
-A density functional theory approach based on absolutely localized molecular orbitals is used in conjunction with the Langevin equation of motion to create a low-cost molecular dynamics method for weakly coupled molecular systems which retains sufficient accuracy to generate accurate dynamical properties in water.
+%The utility of density functional theory based on compact localized molecular orbitals in \emph{ab initio} molecular dynamics. To overcome limitations of existing linear-scaling methods
+%%in conjunction with a modified Langevin equation of motion 
+%to create a linear-scaling \emph{ab initio} molecular dynamics method for weakly coupled molecular systems. 
+%To overcome %which retains sufficient accuracy to generate accurate dynamical properties in water. 
+%%Strictly localized 
+%%Low-cost linear-scaling ab initio molecular dynamics for weakly-interacting systems is developed 
+%%for weakly-interacting systems. 
+%%High efficiency of the methodology is achieved without sacrificing its accuracy with a combination of two techniques: (1) on-the-fly calculation of approximate forces without lengthy self-consistent optimization of localized orbitals and (2) integration of a modified Langevin equation of motion that is fine-tuned to retain stable and correct dynamics even with imperfect forces.
+%%A remarkable feature of the method is that it remains efficient for challenging condensed phase systems even with diffuse atom-centered basis sets. 
+%Application of the method to liquid water demonstrates that the new AIMD method enables simulations of molecular systems on previously inaccessible time and length scales.
 
 %The approach utilizes a systematically inaccurate potential energy surface calculation method that grows linearly in computational complexity with the size of the simulated system. 
-%Systematic inaccuracies in the calculation of force using this method are balanced by a weighted version of the Langevin equation of motion. 
+%Systematic inaccuracies in the calculation of force using this method are balanced by a weighted version of the Langevin equation. 
 %The stochastic ``white noise'' of the Langevin equation acts to absorb force discontinuities while also thermostating the system to maintain the desired thermodynamic temperature. 
@@
 
-%RZK-important addition: we did not include a large family of the so-called fragmentation methods 
-%included only those methods that attempt to construct  set 
+%RZK important addition for a more detialed publication: review of a large family of fragmentation methods into the review following these guidelines.
 %It is important to note all LS DFT methods reviewed above attempt to obtain the electronic description that do not construct  based on fragmentation methods that do that there  LS DFT methods the presented logical partitioning with the subsequent construction of localized MOs is a rather general approach used in a large number of electronic structure theories, which are collectively known as fragmentation methods~\cite{a:fragmentation-rev}.
@@
 
-Once the localization radius is chosen, ALMO constraint prohibits electron density transfer between distant molecules but retains all other types of interaction such as long-range electrostatic, exchange, polarization, and, if the exchange-correlation (XC) functional includes them, dispersion interactions~\cite{a:theeda}.
-Since the importance of electron transfer decays exponentially with distance, the \mbox{ALMO} approximation provides a natural and accurate representation of the electronic structure of molecular systems with just a few local variables.
+ALMO constraints imposed by $R_c$ prohibit electron density transfer between distant molecules but retain all other types of interaction such as long-range electrostatic, exchange, polarization, and, if the exchange-correlation (XC) functional includes them, dispersion interactions~\cite{a:theeda}.
+Since the importance of electron transfer decays exponentially with distance, the \mbox{ALMO} approximation is expected to provide a natural and accurate representation of the electronic structure of molecular systems.
+%with just a few local variables.
 Because of the greatly reduced number of electronic descriptors and the robust optimization, the computational complexity of ALMO DFT grows linearly with the number of molecules while its computational overhead remains very low.
-These features make ALMO DFT a promising method to perform accurate AIMD simulations of large molecular systems.
+These features make ALMO DFT a promising method for accurate AIMD simulations of large molecular systems.
 
@@
 %We will show that our approach lowers the computation cost pre-factor for the ALMO SCF method, while maintaining LS and accuracy for calculation of dynamical properties. 
-%This is done by balancing a new computationally economical but systemically inaccurate force calculation method with an optimized version of the Langevin Equation of motion. 
+%This is done by balancing a new computationally economical but systemically inaccurate force calculation method with an optimized version of the Langevin equation. 
 %We also introduce an easy to follow heuristic approach to optimize a system for generating correct dynamical behaviour.
@@
 The challenge of adopting ALMO DFT for dynamical simulations arises from the slightly nonvariational character of the localized orbitals.
-First, while ALMOs are variationally optimized in both SCF steps, the occupied-space projector obtained in the first step must remain fixed during the second step to ensure convergence.
-Second, some neglected electron transfer effects can suddenly become active in the course of a dynamical simulation when a neighboring molecule passes the localization threshold $R_{c}$.
-In addition to the ALMO-specific issues, the variational optimization in any AIMD methods is never truly complete and is interrupted once the norm of the gradient of the energy with respect to the variational electronic degrees of freedom drops below small but nevertheless finite convergence threshold $\epsilon_{\text{SCF}}$.
-While these non-SCF errors are minor and do not affect accuracy of static ALMO DFT calculations, geometry optimization, and ALMO-based Monte-Carlo simulations, they accumulate in molecular dynamics trajectories resulting in non-physical sampling and eventually disrupting them.
+First, while ALMOs are variationally optimized in both SCF steps, the occupied subspace defined in the first step must remain fixed during the second step to ensure convergence.
+Second, neglected electron transfer effects can suddenly become active in the course of a dynamical simulation when a neighboring molecule passes the localization threshold $R_{c}$.
+In addition to these issues characteristic to ALMOs, the variational optimization in any AIMD method is never truly complete and interrupted once the maximum norm of the gradient of the energy with respect to the electronic descriptors drops below small but nevertheless finite convergence threshold $\epsilon_{\text{SCF}}$.
+While these nonvariational errors do not affect the accuracy of static ALMO DFT calculations, geometry optimization, and Monte-Carlo simulations, they accumulate in molecular dynamics trajectories leading to non-physical sampling and eventual failure.
+Traditional strategies to cope with these problems are computationally expensive and include computing the nonvariational contribution to the forces via a variational coupled-perturbed procedure, increasing $R_c$, and decreasing $\epsilon_{\text{SCF}}$.
+%This however would dramatically increase the computational cost of the ALMO AIMD.
+%as longer-distance but physically unimportant interactions must be calculated with an additional iterative coupled perturbed solver.
 
-%For any finite value of $R_{c}$, forces are undefined at manifold of due to the discontinuous jump in potential energy when a neighboring molecule passes beyond the localization radius $R_{c}$ and is suddenly ignored in potential energy calculations.
-%Setting $R_{c}$ to larger values can reduce the magnitude of the discontinuity, but increases the computational cost as longer distance (but physically less important) interactions must be calculated.
-%As long as the discontinuity in force is small compared with the random numerical noise in the calculation, it can be safely ignored~\cite{RZK-any-references}.
-This means that $R_{c}$ has to be kept relatively large to maintain stable and accurate dynamics.
-
-%While these issues can be resolved by  by increasing  to account for non-SCF forces, increasing  and decreasing While  are the known formal strategies for resolving these
-
-In order to maintain stable and accurate dynamics, one could use a set of strategies that includes computing the non-SCF contribution to the forces through a coupled perturbed procedure, increasing $R_{c}$, and decreasing $\epsilon_{\text{SCF}}$.
-%relative to the corresponding optimal values for static energy calculations or Monte-Carlo simulations.
-This however would dramatically increase the computational cost of the ALMO AIMD.
-%as longer-distance but physically unimportant interactions must be calculated with an additional iterative coupled perturbed solver.
+In this work, we propose another approach that obviates the need in a coupled-perturbed solver, relaxes tight constraints on $R_{c}$ and $\epsilon_{\text{SCF}}$, and thus enables us to maintain stable dynamics and to keep the complexity and cost of simulations low.
 %
-In this work we propose another approach, which obviates the need in a coupled-perturbed solver, allows for relaxation of the tight constraints on $R_{c}$ and $\epsilon_{\text{SCF}}$, and thus keeps the complexity and cost of simulations low.
-
-In our approach, the forces on atoms are calculated \emph{approximately} after the two-step ALMO SCF using a straightforward procedure that computes only the Hellmann-Feynman and Pulay components and neglects the computationally intense non-SCF component of the forces.
-The difference between these approximate ALMO forces and the \emph{reference} forces that would be obtained from perfectly converged fully-delocalized KS orbitals is $\delta f_{i\alpha}(t)$:
+In our approach, the forces on atoms are calculated \emph{approximately} after the two-step ALMO SCF using a straightforward procedure that computes only the Hellmann-Feynman and Pulay components and neglects the computationally intense nonvariational component of the forces.
+The difference between these approximate ALMO forces and the \emph{reference} forces that could be obtained from perfectly converged fully-delocalized KS orbitals is $\delta f_{i\alpha}(t)$:
 %
@@
 $\delta f_{i\alpha} (t)$ comprises all neglected terms that originate from a finite localization radius $R_c$ and incomplete SCF optimization.
-$\delta f_{i\alpha} (t)$ can be reduced \emph{systematically} to zero by increasing $R_c$ and decreasing $\epsilon_{\text{SCF}}$.
+$\delta f_{i\alpha} (t)$ can be reduced to zero \emph{systematically} by increasing $R_c$ and decreasing $\epsilon_{\text{SCF}}$.
 %$\epsilon_{\text{SCF}}$ is chosen such that the electronic ground state is considered sufficiently close to the Born-Oppenheimer state for values less than $\epsilon_{\text{SCF}}$.
@@
 Our approach to compensate for the missing $\delta f_{i\alpha}(t)$ term follows the methodology that have been introduced into AIMD by Krajewski \emph{et al.}~\cite{PRB-73-041105R}, formalized by K\"uhne \emph{et al.}~\cite{a:2ndcpmd} and rationalized by Dai \emph{et al.}~\cite{a:langevin-why} before becoming informally known as the second generation Car-Parrinello molecular dynamics~\cite{doi:10.1002/wcms.1176}.
-Inspired by this method, ALMO AIMD is chosen by be governed by the Langevin equations of motion, which can be written using the unknown reference forces as:
+Inspired by this method, ALMO AIMD is chosen to be governed by the Langevin equation of motion that are first written in terms of the unknown reference forces:
 %
@@
 %
-where $m_i$ is the mass of atom $i$, $r_{i\alpha}$ is its position along dimension $\alpha$, $\gamma$ is the Langevin scaling factor, and $R^{\gamma}_{i\alpha} (t)$ is stochastic force represented by a white Gaussian noise with the zero mean and the strength related to the damping term through $\gamma$:
+where $m_i$ is the mass of atom $i$, $r_{i\alpha}$ is its position along dimension $\alpha$, $\gamma$ is the Langevin scaling factor, and $R^{\gamma}_{i\alpha} (t)$ is the stochastic force represented by a zero-mean white Gaussian noise: %with the zero mean and the strength related to the damping term through $\gamma$:
 %
@@
 % 
-This relation means that, for any value of $\gamma$, the perfectly balanced damping and stochastic terms will maintain a constant average temperature $T$ and generate the correct Maxwell-Boltzmann distribution of velocities~\cite{a:Kubo-1986,b:tuckerman-stat}.
-In the limit $\gamma \rightarrow 0$, the Newton equations of motion are recovered and the microcanonical ensemble is sampled. 
-Finite values of $\gamma$ correspond to imposing canonical thermostat on the system, with larger $\gamma$ leading to more random noise in trajectories, faster equilibration but more distorted results for calculated dynamical properties.
+The last relation means that, for any value of $\gamma$, the damping and stochastic terms are in perfect balance and will generate the correct Maxwell-Boltzmann distribution of velocities for a requested temperature $T$~\cite{a:Kubo-1986}.
+In the limit $\gamma \rightarrow 0$, the Newton equations of motion are recovered. %and the microcanonical ensemble is sampled. 
+Finite values of $\gamma$ correspond to imposing canonical thermostat on the system, with larger $\gamma$ leading to faster equilibration but more distorted results for dynamical properties.
 
@@
 %
-This assumption, shown below to be well justified, allows us to rewrite the Langevin equations using the ALMO forces:
+This assumption, shown to be well justified, allows us to rewrite the Langevin equation using the ALMO forces:
 %
@@
 The only missing piece in the modified Langevin equation is the value of $\Delta$, which describes the strength of the newly introduced stochastic term. 
-This term compensates for the imperfection of ALMO forces and is required to re-balance the damping and stochastic components in the Langevin equation and generate the correct Maxwell-Boltzmann distribution in ALMO AIMD.
+This term compensates for imperfections in ALMO forces and must be adjusted to re-balance the damping and stochastic components in ALMO AIMD.
+%and generate the correct Maxwell-Boltzmann distribution in ALMO AIMD.
 
@@
 
-%To correctly thermostat the system of interest, $\Delta$ needs to be reduced to zero.
-%This is done by finding the value of $\Delta$ and adding a third random force term $R^{-\Delta}_{i\alpha}$ to equation \ref{eq:langevin2} which cancels out the extra forces introduced to the system by $R^{\Delta}_{i\alpha}$.
-%This means that a trajectory with inaccurate forces and optimized $R^{-\Delta}_{i\alpha}$ should have no long term drift in the total energy of the system. 
-%Short term energy fluctuations should be expected, and can later be ``smoothed out'' by setting $\gamma$ as needed.
-
-(RZK the integral contains dummy index $i$ in the r.h.s.) In principle, $\Delta$ can be calculated using the integral of Eq.~(\ref{eq:stochastic3})
+%RZK: There is a problem with the integral equation below.
+The r.h.s.
+depends on $i$ but in the l.h.s.
+does not.
+If I remember correctly we evaluate the integral for molecules, not atoms.
+In liquid water all our molecules are the same and we do not have to deal with the issue of different Delta for different molecules.
+Eliminating the problem
+In principle, $\Delta$ can be calculated using the integral of Eq.~(\ref{eq:stochastic3})
 %
@@
 %
-if one can afford computing the reference forces $f^{\text{KS}}_{i\alpha}(t)$ (i.e.
-$R_c \rightarrow \infty$ and $\epsilon_{\text{SCF}} \rightarrow 0$) for a short representative AIMD trajectory (see Supplemental Material).
+if one can afford computing the reference forces $f^{\text{KS}}_{i\alpha}(t)$ (i.e. $R_c \rightarrow \infty$ and $\epsilon_{\text{SCF}} \rightarrow 0$) for a short representative AIMD trajectory. 
 In practice, we found (see results below) that this approach is not particularly accurate due to the noisy autocovariance functions (ACF) from short trajectories and perhaps because the $\delta_{ij}\delta_{\alpha\beta}$ assumption in Eq.~(\ref{eq:stochastic3}) does not strictly hold.
 Nevertheless, the ACF integral can provide a reasonable starting value of $\Delta$.
-This or, in fact, any starting value can be  rapidly fine-tuned in a series of trial-and-error short ALMO AIMD runs until the equipartition theorem is satisfied: 
-%
-\begin{align}
-\label{eq:eqipartition}
-\langle \dfrac{1}{2} m_i \dot{\bm{r}}^{2}_{i} \rangle &= \dfrac{3}{2} k_{B} T
-\end{align}
-%
-or, in other words, until the simulation stabilizes with the correct average temperature.
+This value can be further fine-tuned in a series of short trial-and-error ALMO AIMD runs until the average kinetic energy corresponds to the requested temperature $\langle \frac{1}{2} m_i \dot{\bm{r}}^{2}_{i} \rangle = \frac{3}{2} k_{B} T$ .
+%%
+%\begin{align}
+%\label{eq:eqipartition}
+%\langle \dfrac{1}{2} m_i \dot{\bm{r}}^{2}_{i} \rangle &= \dfrac{3}{2} k_{B} T
+%\end{align}
 
@@
 Accuracy and efficiency of ALMO AIMD was tested using liquid water as an example. 
-This system is challenging because intermolecular electron delocalization is a critical component of hydrogen bonding and must be described correctly in approximate ALMO AIMD to reproduce static and dynamical properties of liquid water.
+This system is challenging because intermolecular electron delocalization is a critical component of hydrogen bonding and must be described correctly to reproduce static and dynamical properties of liquid water.
 %To calculate properties of water at ambient conditions, 
-A periodic cell containing 125 molecules was simulated for $\SI{10}{\ps}$ at a temperature of $\SI{298}{\K}$ and a constant density of $\SI{1.006}{\g\cdot\cm^{-3}}$. 
+A periodic cell containing 125 molecules was simulated for $\SI{10}{\ps}$ at $T=\SI{298}{\K}$ and a constant density of $\SI{1.01}{\g\cdot\cm^{-3}}$. 
 % RZK: we do not need to specify details on gamma optimization since they are well-known.
-RZK algorithm~\cite{RZK} was used to integrate the Langevin equation of motion.
-We found that $\gamma = \SI{e-3}{\per\fs}$ is large enough to thermostat the system on the time scale of the simulation and small enough not to significantly affect dynamical properties of liquid water.
+Ricci-Ciccotti algorithm~\cite{RZK:10.1080/0026897031000108113} was used to integrate the Langevin equation.
+We found that $\gamma = \SI{e-3}{\per\fs}$ is large enough to thermostat the system efficiently and small enough not to significantly affect dynamical properties of liquid water.
 %
-In the dual-representation Gaussian and plane-wave scheme implemented in CP2K~\cite{a:quickstep}, the TZV2P basis set was used to represent molecular orbitals, and a plane-wave cutoff of $\SI{320}{Ry}$ used to represent electron density.
+In the dual Gaussian and plane-wave scheme implemented in CP2K~\cite{a:quickstep}, the TZV2P basis set was used to represent molecular orbitals, and a plane-wave cutoff of $\SI{320}{Ry}$ used to represent electron density.
 The XC energy was approximated using the PBE functional~\cite{a:PBEfunctional}. 
-The Brillouin zone was sampled at the $\Gamma$-point and separable norm-conserving pseudopotentials were used to describe the interactions between the valence electrons and the ionic cores~\cite{a:hgh}.
+Separable norm-conserving pseudopotentials were used~\cite{a:hgh} and the Brillouin zone was sampled at the $\Gamma$-point. 
+%to describe the interactions between the valence electrons and the ionic cores.
 
-The \emph{reference} forces were calculated with fully delocalized electrons using tightly converged, $\epsilon_{\text{SCF}}=10^{-6}$~a.u., orbital transformation (OT) method~\cite{a:ot,a:ot2}. 
+The \emph{reference} forces were calculated with fully delocalized electrons using tightly converged, $\epsilon_{\text{SCF}}=10^{-6}$~a.u., orbital transformation (OT) method~\cite{a:ot}. 
 In ALMO AIMD, element-specific cutoff radius for electron delocalization $R_c$ was set to 1.6 in units of the elements' van der Waals radii (vdWR).
-This localization radius includes approximately two coordination shells of an average water molecule and was shown to reproduce the KS reference RDF perfectly in Monte-Carlo simulations~\cite{a:almo-ls}.
+This localization radius includes approximately two coordination shells of an average water molecule and was shown to reproduce the reference radial distribution function (RDF) perfectly in Monte-Carlo simulations~\cite{a:almo-ls}.
 %
@@
 
-%The computation scheme used is mostly identical to that described in the implementation section of ref.\ \citenum{a:almo-ls}. 
-%at $\SI{302}{\K}$ and  --- RZK: It is customary to report TARGET temperature, not the acutal average temperature.
-The fact the the actual temperature deviates from the target temperature by 5 K even with perfect forces shows that our simulation is still too short.
-%with the Langevin thermostat with $\gamma = \SI{e-3}{\per\fs}$. 
-%All other system parameters were set identical to the ALMO system.
-
-Even with $\epsilon_{\text{SCF}} = 10^{-2}$, we stabilized the simulation at the correct average temperature and perfect Maxwell-Boltzmann distribution (Figure~\ref{fig:dynproperties}b).
-$\Delta$ was initially estimated at $\SI{2e-5}{\per\fs}$ using Eq.~(\ref{eq:delta}) and then refined heuristically to $\SI{6e-5}{\per\fs}$.
-%RZK: factor of 3 bothers me. Check that this is not because we sum over 3 atoms in a molecule (or over 3 dimensions).
-We found that it is easier to optimize $\Delta$ when $\gamma$ is set to zero because unnecessary stochastic noise is eliminated in the trial runs.
-%In theory, $\Delta$ should be independent on system size.
-However, we found that $\Delta$ dependents on system size due to finite size effects, which decay for increasing system size.
-
 \begin{figure}
 \includegraphics[trim={1.3cm 0.1cm 3.3cm 1.3cm},clip,width=8.6cm]{Dynamical_Data_Tiled.eps}
-\caption{\label{fig:dynproperties} RZK:.
-Calculated properties of water using ALMO DFT with $\epsilon_{\text{SCF}} = 10^{-2}$~a.u.
-and $R_{c} = 1.6$ vdWR (red line) and fully converged OT reference (black line).
-(a) radial distribution function, 
+\caption{\label{fig:dynproperties} 
+%RZK: "G_OO -> g_OO", "Intensity (au)" -> "Intensity", remove "DFT" in two legend labels 
+Calculated properties of water using ALMO AIMD with $\epsilon_{\text{SCF}} = 10^{-2}$~a.u. and $R_{c} = 1.6$ vdWR (red line) and fully converged OT reference (black line).
+(a) RDF, 
 (b) kinetic energy distribution (the black curve shows the theoretical Maxwell-Boltzmann distribution), 
-(c) infrared spectrum.
+(c) IR spectrum.
 %Calculations were performed with the PBE $\mathrm{E_{XC}}$ functional and TZV2P basis set, using 125 water molecules for 10 ps.
@@
 
-Analysis of $\delta \vec{f_{i}}(t)$ shows that the error in forces indeed resembles Gaussian white noise.
-The mean of the error is zero (black line in Figure~\ref{fig:randomforce}c) and the ACF decays rapidly (Figure~\ref{fig:randomforce}d) so that the error in forces can be considered uncorrelated on time scale of $\SI{50}{\fs}$.
-Thus the main assumptions behind our approach to ALMO AIMD are justified for liquid water.
-Figure~\ref{fig:randomforce}b shows that the main source of error in forces in this simulation is the loose convergence criterion and not the finite $R_c$: fully converged ALMO SCF calculations remove the rapidly oscillating component of the error.
-We also confirmed that the ALMO forces converge to the reference forces in the limit $R_{c} \rightarrow \infty$ (Figure~\ref{fig:forcecomp} in the Supplemental Material).
+%The computation scheme used is mostly identical to that described in the implementation section of ref.\ \citenum{a:almo-ls}. 
+%at $\SI{302}{\K}$ and  --- RZK: It is customary to report TARGET temperature, not the acutal average temperature. 
+%with the Langevin thermostat with $\gamma = \SI{e-3}{\per\fs}$. 
+%All other system parameters were set identical to the ALMO system.
+
+%RZK: 5K deviation from the target temperature is considered insignificant.
+Even with perfect forces, the actual temperature deviates from the target temperature by 5 K.
+This shows that our simulations are still too short to get the correct average.
+This is fine though.
+Even with $\epsilon_{\text{SCF}} = 10^{-2}$, we stabilized the simulation at the correct average temperature and perfect Maxwell-Boltzmann distribution (Figure~\ref{fig:dynproperties}b).
+$\Delta$ was initially estimated at $\SI{2e-5}{\per\fs}$ using Eq.~(\ref{eq:delta}) and then refined heuristically to $\SI{6e-5}{\per\fs}$.
+%RZK: factor of 3 bothers me. Check that this is not because we sum over 3 atoms in a molecule (or over 3 dimensions).
+We found that it is easier to optimize $\Delta$ when $\gamma$ is set to zero because of reduced noise in the trial runs. 
+%In theory, $\Delta$ should be independent on system size. However, we found that $\Delta$ dependents on system size due to finite size effects, which decay for increasing system size.
+%
+Analysis of $\delta \vec{f_{i}}(t)$ shows that the error indeed resembles Gaussian white noise.
+The mean of the error is zero (black line in Figure~\ref{fig:randomforce}a).
+Its ACF decays rapidly (Figure~\ref{fig:randomforce}c) so that the errors can be considered uncorrelated on time scale of $\SI{50}{\fs}$.
+Thus the main assumption behind our approach to ALMO AIMD is justified for liquid water.
+The main source of error in forces for this system is the loose convergence criterion and not the finite $R_c$: fully converged ALMO SCF calculations remove the rapidly oscillating component of the error (Figure~\ref{fig:randomforce}b).
+We also confirmed that the ALMO forces converge to the reference forces in the limit $R_{c} \rightarrow \infty$
+%(Figure~\ref{fig:forcecomp} in the Supplemental Material).
+(Figure~S1 in the Supplemental Material).
 
@@
 \caption{\label{fig:randomforce}  
+%RZK: Change panels: Remove panel (a), old panel (c) becomes new panel (a), old (b) -> new (b), old (d) -> new (c). The caption and the text has aready been re-written. 
 The red line is the Euclidean norm of the instantaneous error $\langle \| \delta \vec{f_{i}}(t) \| \rangle_{i}$, the black line is the magnitude of the time average of the instantaneous error vector, and the green line is the time average of the red line.
-(a) $R_{c} = 1.6$ vdWR and $\epsilon_{\text{SCF}} = 10^{-2}$~a.u. relative to the fully converged ALMO reference, 
-(b) $R_{c} = 1.6$ vdWR and fully converged ALMO SCF, relative to the fully converged OT reference, 
-(c) $R_{c} = 1.6$ vdWR and $\epsilon_{\text{SCF}} = 10^{-2}$~a.u., relative to the fully converged OT reference, 
-(d) Normalized ACF of the instantaneous error in panel (a) $\frac{1}{3}\langle \delta \vec{f}_i (t) \cdot \delta\vec{f}_i(t+\tau) \rangle_{it} $.
+%RZK: (a) $R_{c} = 1.6$ vdWR and $\epsilon_{\text{SCF}} = 10^{-2}$~a.u. relative to the fully converged ALMO reference, 
+(a) $R_{c} = 1.6$ vdWR and $\epsilon_{\text{SCF}} = 10^{-2}$~a.u., 
+(b) $R_{c} = 1.6$ vdWR and fully converged ALMO SCF, 
+(c) Normalized ACF of the instantaneous error in panel (a) $\frac{1}{3}\langle \delta \vec{f}_i (t) \cdot \delta\vec{f}_i(t+\tau) \rangle_{it} $.
 %Forces are computed with the PBE XC functional and TZV2P basis set for the configurations from a 10-ps trajectory generated with KS-based AIMD at T=298~K. 
@@
 
-To test the accuracy of ALMO AIMD we used trajectory analyzer TRAVIS~\cite{a:travis-main} to compute the infrared (IR) spectrum, radial distribution function (RDF), and diffusion coefficient of liquid water from both the ALMO trajectory ($\epsilon_{\text{SCF}} = 10^{-2}$~a.u.
+To test the accuracy of ALMO AIMD we used trajectory analyzer TRAVIS~\cite{a:travis-main} to compute the infrared (IR) spectrum, RDF, and diffusion coefficient of liquid water from both the ALMO trajectory ($\epsilon_{\text{SCF}} = 10^{-2}$~a.u.
 and $R_{c} = 1.6$ vdWR) and from the reference trajectory.
@@
 The quality of the ALMO IR spectrum (Figure~\ref{fig:dynproperties}c) is good despite minor errors in the intensity of the OH stretching mode, which is sensitive to the precise positions of the centers of localized orbitals.
-These stringent tests show that despite noticeable errors in the ALMO forces (Figure \ref{fig:randomforce}) the compensating $R^{\Delta}$ term in the modified Langevin equation makes it possible to recover atomic dynamics properly.
+These stringent tests show that despite noticeable errors in the ALMO forces (Figure \ref{fig:randomforce}a) the compensating $R^{\Delta}$ term in the modified Langevin equation makes it possible to recover atomic dynamics properly.
 We would like to note that ALMO AIMD could not be stabilized with $\Delta=0$.
-We were not able to find any values of $\Delta$ that stabilize trajectories generated using perturbative versions of ALMO DFT~\cite{a:almo-ls}.
+Neither we were able to find any values of $\Delta$ that stabilize trajectories generated using perturbative versions of ALMO DFT~\cite{a:almo-ls}.
 
@@
 
-%RZK Simulations for timing benchmarks were run for ten time steps, which is sufficient for convergence of wavefunctions from an initial guess.
+%RZK addition for a longer article: extrapolation of WFs on both SCF steps is a crucial component of our fast algorithm.
 
 To demonstrate the computational advantage of ALMO AIMD, we compared the average wall-time per MD step over 10 time steps for a variety of methods in Figure \ref{fig:strongscaling_log}.
-In addition to the fully converged OT reference, we included timing for the second generation Car-Parrinello method, which decreases computational prefactor of the cubically scaling OT method~\cite{a:2ndcpmd}.
+In addition to the fully converged OT reference, we included timing for the second generation Car-Parrinello method designed to decrease the computational prefactor of the cubically scaling OT method~\cite{a:2ndcpmd}.
 As expected ALMO methods show clear LS behavior for all values of $\epsilon_{\text{SCF}}$ even for medium-size systems. 
 %OT SCF scales approximately cubically, as expected.
-For liquid water, strict localization of molecular orbitals makes ALMO AIMD faster than the second-generation Car-Parinello for larger than 256 molecules.
-It is important to emphasize that comparison is performed for a three-dimensional condensed phase system described with a large diffuse basis set -- a particularly challenging case for LS methods.
+For liquid water, strict localization of ALMOs lowers the crossover point between ALMO AIMD and cubically scaling methods to 256 molecules -- length scales routinely accessible with AIMD today.
+It is important to emphasize that the comparison is performed for a three-dimensional condensed phase system described with a large diffuse basis set -- a particularly challenging case for all LS methods.
 
@@
 Cyan lines represent perfect cubic scaling, whereas gray lines represent perfect linear scaling. 
-%RZK: On figure remove 'DFT', 'Wall-clock time' -> 'Wall time'
+%RZK: On figure remove 'DFT', 'Wall-clock time' -> 'Wall time'. Decrease the height of the figure (the y-axis label is now shorter).
 }
@@
 
-Weak scaling benchmarks show (Figure \ref{fig:weakscaling}) that computation with localized orbitals is naturally suited for parallel execution and that the computational cost of our implementation grows linearly for a wide range of compute cores.
-It also demonstrates that the range of routine AIMD simulations can be extended from $\sim 10^3$ to $\sim 10^4$ atoms making completely new phenomena such as dynamics on interfaces and nucleation processes accessible on HPC platforms today.
+Weak scaling benchmarks for very large systems show (Figure~\ref{fig:weakscaling}) that localized orbitals are naturally suited for parallel execution and that the computational cost of our implementation grows linearly for a wide range of compute cores.
+In fact, we were able to successfully simulate systems as large as 32768 water molecules within reasonable wall-clock time using moderate number of compute cores -- unprecedented feat for AIMD considering that accurate molecular orbitals and the idempotent DM are computed on each step.
+The horizontal line in Figure~\ref{fig:weakscaling} is shown as a rough guide to time and length scales accessible in a fixed wall-clock time given various computational resources.
+It indicates that ALMO AIMD can extend the range of routine simulations to $\sim10^4$ atoms on modern HPC platforms.
+
 %We also show that systems as large as 8192 molecules can be simulated for 10,000 MD steps in under a month using this method.
-In fact, we were able to successfully simulate systems as large as 32768 water molecules within reasonable wall-clock time using only moderate number of compute cores (Figure \ref{fig:weakscaling}) -- unprecedented feat for AIMD considering that accurate molecular orbitals and the idempotent DM are computed on each step.
 
@@
 \caption{\label{fig:weakscaling} Weak scalability benchmarks for PBE/TZV2P ALMO AIMD with  $R_{c} = 1.6$ vdWR and $\epsilon_{\text{SCF}} = 10^{-2}$~a.u. 
-Dashed gray lines connect systems simulated using the same number of compute cores to confirm LS behavior. %over a range of configurations.
-The horizontal line corresponds to 10,000 AIMD steps in one wall month and shows a upper-bound guideline for practical AIMD.
-%RZK: On figure 'Wall-clock time' -> 'Wall time', legend - increase font size, decrease the height of the figure(the y-axis label is now shorter)
+Dashed gray lines connect systems simulated on the same number of cores to confirm LS behavior. %over a range of configurations.
+%RZK: On figure 'Wall-clock time' -> 'Wall time', legend - increase font size, decrease the height of the figure(the y-axis label is now shorter).
+This figure will be black-and-white in the printed version and will have color online: replace magenta cross with another "solid" symbol such as inverted triangle.
 }
@@
 
-RZK: systematically improvable approximation to DFT-based AIMD.
-In this article we have clearly demonstrated the utility of ALMO DFT methods, especially when combined with the Langevin Equation of motion.
-ALMO DFT methods are a powerful class of AIMD electronic structure determination techniques capable of efficiently calculating forces for weakly coupled molecular systems.
-It's clear that forces calculated by ALMO DFT can be tuned as accurately as required by the user, up to the Born-Oppenheimer limit.
-Indeed, the ALMO DFT technique highlighted in this article has been shown to be capable of producing potential energy surfaces extremely efficiently for medium to large systems with only a minimal drop in the accuracy of calculated bulk system properties.
-We have also shown that due to its linear scaling nature, the ALMO DFT technique can be pushed to simulate enormous chemical systems accurately in remarkably short calculation times.
-
-%Using liquid water as an example, we demonstrate that the new AIMD methodology enables computational studies of molecular systems on previously inaccessible time and length scales.
-
-%To summarize, a low-cost linear-scaling AIMD method for weakly-interacting molecular systems was developed by utilizing compact molecular orbitals strictly localized within predefined radii.
-High efficiency of the method is achieved without sacrificing its accuracy with a combination of two techniques: (1) on-the-fly construction of accurate localized orbitals without lengthy self-consistent optimization and (2) a modified Langevin equation that is fine-tuned to retain stable dynamics even with imperfect forces.
-A remarkable feature of the implemented method is that it remains efficient for challenging condensed phase systems, even with diffuse atom-centered basis sets.
-Using liquid water as an example, it is demonstrated that the new AIMD method enables simulations of molecular systems on previously inaccessible time and length scales.
-
-Generalization of the method to more challenging systems of strongly interacting atoms such as covalent crystals is underway.
+% systematically improvable approximation to DFT-based AIMD.
+To summarize, we demonstrated, for the first time, the utility of compact localized orbitals for low-cost linear-scaling AIMD for weakly-interacting systems. 
+High efficiency of the methodology is achieved without sacrificing its accuracy with a combination of two techniques: (1) on-the-fly calculation of approximate forces without lengthy self-consistent optimization of localized orbitals and (2) integration of a modified Langevin equation of motion that is fine-tuned to retain stable and correct dynamics even with imperfect forces.
+A remarkable feature of the method is that it remains efficient for challenging condensed phase systems even with diffuse atom-centered basis sets. 
+Using liquid water as an example, we showed that the new AIMD method enables simulations of molecular systems on previously inaccessible time and length scales. 
+The developed methods will have a significant impact on modeling of large heterogeneous molecular systems (e.g. interfaces, nuclei) making completely new phenomena accessible to AIMD. 
+Generalization of the methodology to systems of strongly interacting atoms (e.g. covalent crystals) is underway.
 
@@
 
+%RZK: journal abbreviation in Ref.  a:almo-ls
 \bibliography{almo-langevin}
 
+%RZK: resolve SI problem
 \else
@@
 
-\begin{figure}[!htb]
+\begin{figure*}[h!]
 \includegraphics[trim={0cm 0cm 0.1cm 0.1cm},clip,width=\textwidth]{DeltaForceComparison_ALMO_SCF.eps}
@@
 Forces are computed with the PBE XC functional, TZV2P basis set, and $\epsilon_{\text{SCF}} = 10^{-7}$~a.u. for the configurations from a 10-ps trajectory generated with KS-based AIMD at T=298~K.}
-\end{figure}
+\end{figure*}
 
@@
 
-\fi % Manuscript <-> SI switch
+\fi % SI is always created to
 

### 2017-10-02 almo-langevin.tex
PRL version created, minor changes are still necessary

@@
 
-\bibliographystyle{achemso}
+%\bibliographystyle{achemso}
 %\bibliographystyle{naturemag}
@@
 \begin{abstract} 
-%The utility of density functional theory based on compact localized molecular orbitals in \emph{ab initio} molecular dynamics. To overcome limitations of existing linear-scaling methods
-%%in conjunction with a modified Langevin equation of motion 
-%to create a linear-scaling \emph{ab initio} molecular dynamics method for weakly coupled molecular systems. 
-%To overcome %which retains sufficient accuracy to generate accurate dynamical properties in water. 
-%%Strictly localized 
-%%Low-cost linear-scaling ab initio molecular dynamics for weakly-interacting systems is developed 
-%%for weakly-interacting systems. 
-%%High efficiency of the methodology is achieved without sacrificing its accuracy with a combination of two techniques: (1) on-the-fly calculation of approximate forces without lengthy self-consistent optimization of localized orbitals and (2) integration of a modified Langevin equation of motion that is fine-tuned to retain stable and correct dynamics even with imperfect forces.
-%%A remarkable feature of the method is that it remains efficient for challenging condensed phase systems even with diffuse atom-centered basis sets. 
-%Application of the method to liquid water demonstrates that the new AIMD method enables simulations of molecular systems on previously inaccessible time and length scales.
-
-%The approach utilizes a systematically inaccurate potential energy surface calculation method that grows linearly in computational complexity with the size of the simulated system. 
-%Systematic inaccuracies in the calculation of force using this method are balanced by a weighted version of the Langevin equation. 
-%The stochastic ``white noise'' of the Langevin equation acts to absorb force discontinuities while also thermostating the system to maintain the desired thermodynamic temperature. 
-%In addition, we introduce a simple heuristic approach to find optimized Langevin equation parameters to use for balancing system-specific imperfect forces and maintaining correct dynamical behavior.
-
-% We developed a linear-scaling AIMD method with low computational overhead by utilizing compact molecular orbitals strictly localized within predefined radii.
-High efficiency of the method is achieved without sacrificing its accuracy with a combination of two techniques: (1) on-the-fly construction of accurate localized orbitals without lengthy self-consistent optimization and (2) the unbalanced Langevin integrator that is fine-tuned to retain stable dynamics even with imperfect forces.
-A remarkable feature of the implemented method is that it remains efficient even for challenging condensed phase systems even if large diffuse basis sets are used.
-We demonstrated that, for systems well-represented by the compact orbitals (e.g.
-molecular systems, ionic salts), the new AIMD method enables simulations on previously inaccessible time and length scales.
-The first steps towards generalizing the method to more challenging systems of strongly interacting atoms (e.g.
-covalent crystals) will also be discussed.
-
+A robust method for constructing compact localized orbitals in density functional theory enables us to create low-cost linear-scaling \emph{ab initio} molecular dynamics for weakly-interacting systems.
+High efficiency of simulations is achieved by calculating the forces approximately without lengthy optimization of localized orbitals and by adjusting the Langevin equation of motion to stabilize the dynamics with the imperfect forces.
+An application to liquid water demonstrates that extremely low computational overhead of this procedure makes it suitable for routine medium-size simulations whereas its linear scaling cost allows to extend accurate first-principle molecular dynamics to previously inaccessible time and length scales.
 \end{abstract}
@@
 
-% Re-write the following section and include any recent advances
+% RZK: search and include recent references
 In all LS DFT methods, the \emph{delocalized} eigenstates of the effective KS Hamiltonian must be replaced with an alternative set of \emph{local} electronic descriptors
@@
 atomic basis functions) increase linearly with the number of atoms, the total number of the electronic degrees of freedom grows quadratically.}.
-Most LS methods~\cite{a:linscale3,a:lee-yang-1996,a:ls-scuseria-1997,a:ls-manolopoulos-1998,a:ls-helgaker-2001,a:ls-niklasson-2003,a:curvy2,a:ls-dm-sign} explore the natural locality of the one-electron density matrix (DM).
+Most LS methods~\cite{a:ls-rev-1999,a:ls-rev-2012} explore the natural locality of the one-electron density matrix (DM). 
 However, in typical DFT calculations that require diffuse basis functions, the DM approach is advantageous only for impractically large systems~\cite{a:ls-rev-1999,a:ls-dm-sign,a:almo-ls}.
 Therefore, the application of DM-based LS methods have been restricted to minimal-basis tight-binding problems. 
-The optimal basis variants of the DM methods~\cite{a:ls-stechel-1994,a:ls-gillan-1995,a:ls-gillan-1996,a:ls-onetep-2003, RZK-MAO} designed to rectify this issue contract large basis sets into a small number of new localized basis functions and then optimize the DM in the contracted basis.
+The optimal basis variants of the DM methods~\cite{a:ls-stechel-1994,a:ls-gillan-1995,a:ls-gillan-1996,a:ls-onetep-2003} designed to rectify this issue contract large basis sets into a small number of new localized basis functions and then optimize the DM in the contracted basis.
 Although such methods have been successfully used for the evaluation of accurate DFT energies of large static systems~\cite{a:ls-onetep-2009,a:ls-conquest-2010,a:ls-onetep-2010-app1,a:ls-rev-2012} their use in AIMD is hampered by the costly optimization of both the contracted orbitals and the density matrix.
+
+% DM-based LS methods: a:linscale3,a:lee-yang-1996,a:ls-scuseria-1997,a:ls-manolopoulos-1998,a:ls-helgaker-2001,a:ls-niklasson-2003,a:curvy2,a:ls-dm-sign
 
@@
 In the current implementation, a localization center is a set of all local Gaussian atomic orbitals of one molecule. 
-However the approach can be easily generalized to other local and nonlocal basis sets~\cite{a:ls-galli-parrinello-1992, Galerkin}. 
+However the approach can be easily generalized to other local and nonlocal basis sets~\cite{a:ls-galli-parrinello-1992, RZK:10.1016/j.jcp.2011.11.032}. 
 The key feature of ALMO DFT is that its one-electron wavefunctions are constructed in a two-step self-consistent-field (SCF) procedure~\cite{a:almo-ls} to circumvent the problem of the sluggish variational optimization emphasized above.
@@
 $\delta f_{i\alpha} (t)$ can be reduced to zero \emph{systematically} by increasing $R_c$ and decreasing $\epsilon_{\text{SCF}}$.
-%$\epsilon_{\text{SCF}}$ is chosen such that the electronic ground state is considered sufficiently close to the Born-Oppenheimer state for values less than $\epsilon_{\text{SCF}}$.
-%RZK For $\epsilon_{\text{SCF}} \leqslant 10^{-7}$~a.u. ALMO SCF is considered converged.
 
-%The approach is inspired by ``Second Generation Car-Parinello'' MD~\cite{a:2ndcpmd} and
-
-Our approach to compensate for the missing $\delta f_{i\alpha}(t)$ term follows the methodology that have been introduced into AIMD by Krajewski \emph{et al.}~\cite{PRB-73-041105R}, formalized by K\"uhne \emph{et al.}~\cite{a:2ndcpmd} and rationalized by Dai \emph{et al.}~\cite{a:langevin-why} before becoming informally known as the second generation Car-Parrinello molecular dynamics~\cite{doi:10.1002/wcms.1176}.
+Our approach to compensate for the missing $\delta f_{i\alpha}(t)$ term follows the methodology that have been introduced into AIMD by Krajewski \emph{et al.}~\cite{RZK:PRB-73-041105R}, formalized by K\"uhne \emph{et al.}~\cite{a:2ndcpmd} and rationalized by Dai \emph{et al.}~\cite{a:langevin-why} before becoming informally known as the second generation Car-Parrinello molecular dynamics~\cite{RZK:10.1002/wcms.1176}.
 Inspired by this method, ALMO AIMD is chosen to be governed by the Langevin equation of motion that are first written in terms of the unknown reference forces:
@@
 
+\begin{figure}
+\includegraphics[trim={1.3cm 0.1cm 3.3cm 1.3cm},clip,width=8.6cm]{Dynamical_Data_Tiled.eps}
+\caption{\label{fig:dynproperties} 
+%RZK: "G_OO -> g_OO", "Intensity (au)" -> "Intensity", remove "DFT" in two legend labels 
+Calculated properties of water using ALMO AIMD with $\epsilon_{\text{SCF}} = 10^{-2}$~a.u. and $R_{c} = 1.6$ vdWR (red line) and fully converged OT reference (black line).
+(a) RDF, 
+(b) kinetic energy distribution (the black curve shows the theoretical Maxwell-Boltzmann distribution), 
+(c) IR spectrum.
+%Calculations were performed with the PBE $\mathrm{E_{XC}}$ functional and TZV2P basis set, using 125 water molecules for 10 ps.
+}
+\end{figure}
+
 ALMO AIMD presented here was implemented in CP2K, an open source materials modeling package~\cite{www:cp2k}. 
@@
 \begin{figure}
-\includegraphics[trim={1.3cm 0.1cm 3.3cm 1.3cm},clip,width=8.6cm]{Dynamical_Data_Tiled.eps}
-\caption{\label{fig:dynproperties} 
-%RZK: "G_OO -> g_OO", "Intensity (au)" -> "Intensity", remove "DFT" in two legend labels 
-Calculated properties of water using ALMO AIMD with $\epsilon_{\text{SCF}} = 10^{-2}$~a.u. and $R_{c} = 1.6$ vdWR (red line) and fully converged OT reference (black line).
-(a) RDF, 
-(b) kinetic energy distribution (the black curve shows the theoretical Maxwell-Boltzmann distribution), 
-(c) IR spectrum.
-%Calculations were performed with the PBE $\mathrm{E_{XC}}$ functional and TZV2P basis set, using 125 water molecules for 10 ps.
+\includegraphics[trim={0.5cm 0cm 0.7cm 0.1cm},clip,width=8.6cm]{DeltaForceComparison_with_ACF.eps}
+\caption{\label{fig:randomforce}  
+%RZK: Change panels: Remove panel (a), old panel (c) becomes new panel (a), old (b) -> new (b), old (d) -> new (c). The caption and the text has aready been re-written. 
+The red line is the Euclidean norm of the instantaneous error $\langle \| \delta \vec{f_{i}}(t) \| \rangle_{i}$, the black line is the magnitude of the time average of the instantaneous error vector, and the green line is the time average of the red line.
+%(a) $R_{c} = 1.6$ vdWR and $\epsilon_{\text{SCF}} = 10^{-2}$~a.u. relative to the fully converged ALMO reference, 
+(a) $R_{c} = 1.6$ vdWR and $\epsilon_{\text{SCF}} = 10^{-2}$~a.u., 
+(b) $R_{c} = 1.6$ vdWR and fully converged ALMO SCF, 
+(c) Normalized ACF of the instantaneous error in panel (a) $\frac{1}{3}\langle \delta \vec{f}_i (t) \cdot \delta\vec{f}_i(t+\tau) \rangle_{it} $.
+%Forces are computed with the PBE XC functional and TZV2P basis set for the configurations from a 10-ps trajectory generated with KS-based AIMD at T=298~K. 
 }
 \end{figure}
-
-%The computation scheme used is mostly identical to that described in the implementation section of ref.\ \citenum{a:almo-ls}. 
-%at $\SI{302}{\K}$ and  --- RZK: It is customary to report TARGET temperature, not the acutal average temperature. 
-%with the Langevin thermostat with $\gamma = \SI{e-3}{\per\fs}$. 
-%All other system parameters were set identical to the ALMO system.
 
@@
 $\Delta$ was initially estimated at $\SI{2e-5}{\per\fs}$ using Eq.~(\ref{eq:delta}) and then refined heuristically to $\SI{6e-5}{\per\fs}$.
-%RZK: factor of 3 bothers me. Check that this is not because we sum over 3 atoms in a molecule (or over 3 dimensions).
+%RZK: factor of 3 bothers me. We need to check that this is not because we sum over 3 atoms in a molecule. See previous comments on the ACF integral.
 We found that it is easier to optimize $\Delta$ when $\gamma$ is set to zero because of reduced noise in the trial runs. 
@@
 Figure \ref{fig:forcecomp} in the Supplemental Material shows that as $R_{c}$ is increased to infinity, forces converge toward the Born-Oppenheimer limit.
-
-\begin{figure}
-\includegraphics[trim={0.5cm 0cm 0.7cm 0.1cm},clip,width=8.6cm]{DeltaForceComparison_with_ACF.eps}
-\caption{\label{fig:randomforce}  
-%RZK: Change panels: Remove panel (a), old panel (c) becomes new panel (a), old (b) -> new (b), old (d) -> new (c). The caption and the text has aready been re-written. 
-The red line is the Euclidean norm of the instantaneous error $\langle \| \delta \vec{f_{i}}(t) \| \rangle_{i}$, the black line is the magnitude of the time average of the instantaneous error vector, and the green line is the time average of the red line.
-%RZK: (a) $R_{c} = 1.6$ vdWR and $\epsilon_{\text{SCF}} = 10^{-2}$~a.u. relative to the fully converged ALMO reference, 
-(a) $R_{c} = 1.6$ vdWR and $\epsilon_{\text{SCF}} = 10^{-2}$~a.u., 
-(b) $R_{c} = 1.6$ vdWR and fully converged ALMO SCF, 
-(c) Normalized ACF of the instantaneous error in panel (a) $\frac{1}{3}\langle \delta \vec{f}_i (t) \cdot \delta\vec{f}_i(t+\tau) \rangle_{it} $.
-%Forces are computed with the PBE XC functional and TZV2P basis set for the configurations from a 10-ps trajectory generated with KS-based AIMD at T=298~K. 
-}
-\end{figure}
 
@@
 For liquid water, strict localization of ALMOs lowers the crossover point between ALMO AIMD and cubically scaling methods to 256 molecules -- length scales routinely accessible with AIMD today.
-It is important to emphasize that the comparison is performed for a three-dimensional condensed phase system described with a large diffuse basis set -- a particularly challenging case for all LS methods.
+It is important to emphasize that the comparison is performed for a three-dimensional condensed phase system described with a large diffuse basis set -- a particularly challenging case for LS methods.
 
@@
 \end{figure}
+
+% RZK addition for a longer article: parallel performance of the algorithm. Vertical line at drawn at 8k molecules shows peculiar behavior of the code. preferred Ncores = integer^2 ?
 
@@
 %RZK: On figure 'Wall-clock time' -> 'Wall time', legend - increase font size, decrease the height of the figure(the y-axis label is now shorter).
-This figure will be black-and-white in the printed version and will have color online: replace magenta cross with another "solid" symbol such as inverted triangle.
+Replace cross symbols on magenta line with another "solid" symbol such as diamon or inverse triangle.
 }
@@
 To summarize, we demonstrated, for the first time, the utility of compact localized orbitals for low-cost linear-scaling AIMD for weakly-interacting systems. 
-High efficiency of the methodology is achieved without sacrificing its accuracy with a combination of two techniques: (1) on-the-fly calculation of approximate forces without lengthy self-consistent optimization of localized orbitals and (2) integration of a modified Langevin equation of motion that is fine-tuned to retain stable and correct dynamics even with imperfect forces.
+High efficiency is achieved without sacrificing accuracy with a combination of two techniques: (1) on-the-fly calculation of approximate forces without lengthy self-consistent optimization of localized orbitals and (2) integration of a modified Langevin equation of motion that is fine-tuned to retain stable and correct dynamics even with imperfect forces.
 A remarkable feature of the method is that it remains efficient for challenging condensed phase systems even with diffuse atom-centered basis sets. 
 Using liquid water as an example, we showed that the new AIMD method enables simulations of molecular systems on previously inaccessible time and length scales. 
-The developed methods will have a significant impact on modeling of large heterogeneous molecular systems (e.g. interfaces, nuclei) making completely new phenomena accessible to AIMD. 
+The developed methods will have a significant impact on modeling of complex molecular systems (e.g. interfaces, nuclei) making completely new phenomena accessible to AIMD. 
 Generalization of the methodology to systems of strongly interacting atoms (e.g. covalent crystals) is underway.
@@
 
-%RZK: journal abbreviation in Ref.  a:almo-ls
+%RZK: incorrect journal abbreviation in a:almo-ls, formatting of all references must be checked
+
 \bibliography{almo-langevin}
 
-%RZK: resolve SI problem
 \else
@@
 
-\fi % SI is always created to
+\fi % SI - main text switch
 

### 2017-10-03 almo-langevin.tex

@@
 % RZK update email
-\email{hayden.scheiber@mcgill.ca}
+\email{scheiber@chem.ubc.ca}
 \author{Yifei Shi}
@@
 \caption{\label{fig:randomforce}  
-%RZK: Change panels: Remove panel (a), old panel (c) becomes new panel (a), old (b) -> new (b), old (d) -> new (c). The caption and the text has aready been re-written. 
 The red line is the Euclidean norm of the instantaneous error $\langle \| \delta \vec{f_{i}}(t) \| \rangle_{i}$, the black line is the magnitude of the time average of the instantaneous error vector, and the green line is the time average of the red line.
@@
 \begin{figure}
-\includegraphics[trim={2.5cm 0.5cm 3.4cm 0.1cm},clip,width=8.6cm]{strongscaling_log.eps}
+\includegraphics[trim={2.5cm 0.5cm 3.4cm 0.9cm},clip,width=8.6cm]{strongscaling_log.eps}
 \caption{\label{fig:strongscaling_log} Timing benchmarks for PBE/TZV2P simulations of liquid water on 256 compute cores. 
@@
 Cyan lines represent perfect cubic scaling, whereas gray lines represent perfect linear scaling. 
-%RZK: On figure remove 'DFT', 'Wall-clock time' -> 'Wall time'. Decrease the height of the figure (the y-axis label is now shorter).
 }
@@
 \begin{figure}
-\includegraphics[trim={1.6cm 0cm 4.7cm 0cm},clip,width=8.6cm]{weakscaling.eps}
+\includegraphics[trim={1.6cm 0.8cm 4.7cm 0cm},clip,width=8.6cm]{weakscaling.eps}
 \caption{\label{fig:weakscaling} Weak scalability benchmarks for PBE/TZV2P ALMO AIMD with  $R_{c} = 1.6$ vdWR and $\epsilon_{\text{SCF}} = 10^{-2}$~a.u. 
 Dashed gray lines connect systems simulated on the same number of cores to confirm LS behavior. %over a range of configurations.
-%RZK: On figure 'Wall-clock time' -> 'Wall time', legend - increase font size, decrease the height of the figure(the y-axis label is now shorter).
-Replace cross symbols on magenta line with another "solid" symbol such as diamon or inverse triangle.
 }

### 2017-10-05 almo-langevin.tex
Analytical expression for Delta is described in SI

@@
 % define variable that compiles the main text (as opposed to SI text)
-\newcommand*{\MAINTEXT}{}
+%\newcommand*{\MAINTEXT}{}
 
@@
 	\clearpage
-	\widetext
+	%\widetext
 	\setcounter{figure}{0}
@@
 \author{Hayden Scheiber}
-% RZK update email
 \email{scheiber@chem.ubc.ca}
@@
 Our approach to compensate for the missing $\delta f_{i\alpha}(t)$ term follows the methodology that have been introduced into AIMD by Krajewski \emph{et al.}~\cite{RZK:PRB-73-041105R}, formalized by K\"uhne \emph{et al.}~\cite{a:2ndcpmd} and rationalized by Dai \emph{et al.}~\cite{a:langevin-why} before becoming informally known as the second generation Car-Parrinello molecular dynamics~\cite{RZK:10.1002/wcms.1176}.
-Inspired by this method, ALMO AIMD is chosen to be governed by the Langevin equation of motion that are first written in terms of the unknown reference forces:
+Inspired by this method, ALMO AIMD is chosen to be governed by the Langevin equation of motion that is first written in terms of the unknown reference forces:
 %
@@
 % 
-The last relation means that, for any value of $\gamma$, the damping and stochastic terms are in perfect balance and will generate the correct Maxwell-Boltzmann distribution of velocities for a requested temperature $T$~\cite{a:Kubo-1986}.
-In the limit $\gamma \rightarrow 0$, the Newton equations of motion are recovered. %and the microcanonical ensemble is sampled. 
-Finite values of $\gamma$ correspond to imposing canonical thermostat on the system, with larger $\gamma$ leading to faster equilibration but more distorted results for dynamical properties.
+The last relation means that, for any value of $\gamma$, the damping and stochastic terms are in perfect balance and trajectories generated with Eq.~(\ref{eq:langevin}) will sample the canonical ensemble at specified temperature $T$~\cite{a:Kubo-1986}.
+%will generate the correct Maxwell-Boltzmann distribution of velocities for a desired temperature $T$~\cite{a:Kubo-1986}. 
+Larger $\gamma$ lead to faster equilibration but more distorted results for dynamical properties.
+%In the limit $\gamma \rightarrow 0$, the Newton equation is recovered and the microcanonical ensemble is sampled.
 
@@
 Eliminating the problem
-In principle, $\Delta$ can be calculated using the integral of Eq.~(\ref{eq:stochastic3})
+In principle, $\Delta$ can be calculated using the integral of Eq.~(\ref{eq:stochastic3}) weighed over all types of atoms with different $m_i$
 %
@@
 \label{eq:delta}
-\Delta &= (2 k_B T m_i )^{-1} \int_{-\infty}^{\infty}\langle \delta f_{i\alpha} (0)  \delta f_{i\alpha} (t) \rangle dt
+\Delta &= (6 k_B T m_i )^{-1} \int_{-\infty}^{\infty} \langle \delta \vec{f}_i (0) \cdot \delta\vec{f}_i(\tau) \rangle d\tau
 \end{align}
@@
 \caption{\label{fig:dynproperties} 
-%RZK: "G_OO -> g_OO", "Intensity (au)" -> "Intensity", remove "DFT" in two legend labels 
 Calculated properties of water using ALMO AIMD with $\epsilon_{\text{SCF}} = 10^{-2}$~a.u. and $R_{c} = 1.6$ vdWR (red line) and fully converged OT reference (black line).
@@
 A periodic cell containing 125 molecules was simulated for $\SI{10}{\ps}$ at $T=\SI{298}{\K}$ and a constant density of $\SI{1.01}{\g\cdot\cm^{-3}}$. 
-% RZK: we do not need to specify details on gamma optimization since they are well-known.
 Ricci-Ciccotti algorithm~\cite{RZK:10.1080/0026897031000108113} was used to integrate the Langevin equation.
@@
 \caption{\label{fig:randomforce} 
+%RZK: label of the third panel: "d" -> "c" 
 The red line is the Euclidean norm of the instantaneous error $\langle \| \delta \vec{f_{i}}(t) \| \rangle_{i}$, the black line is the magnitude of the time average of the instantaneous error vector, and the green line is the time average of the red line.
@@
 (b) $R_{c} = 1.6$ vdWR and fully converged ALMO SCF, 
-(c) Normalized ACF of the instantaneous error in panel (a) $\frac{1}{3}\langle \delta \vec{f}_i (t) \cdot \delta\vec{f}_i(t+\tau) \rangle_{it} $.
+(c) Normalized ACF $\frac{1}{3}\langle \delta \vec{f}_i (t) \cdot \delta\vec{f}_i(t+\tau) \rangle_{it}$ of the instantaneous error in panel (a).
 %Forces are computed with the PBE XC functional and TZV2P basis set for the configurations from a 10-ps trajectory generated with KS-based AIMD at T=298~K. 
@@
 
-%RZK: 5K deviation from the target temperature is considered insignificant.
-Even with perfect forces, the actual temperature deviates from the target temperature by 5 K.
-This shows that our simulations are still too short to get the correct average.
-This is fine though.
 Even with $\epsilon_{\text{SCF}} = 10^{-2}$, we stabilized the simulation at the correct average temperature and perfect Maxwell-Boltzmann distribution (Figure~\ref{fig:dynproperties}b).
@@
 The main source of error in forces for this system is the loose convergence criterion and not the finite $R_c$: fully converged ALMO SCF calculations remove the rapidly oscillating component of the error (Figure~\ref{fig:randomforce}b).
-We also confirmed that the ALMO forces converge to the reference forces in the limit $R_{c} \rightarrow \infty$
+We also verified that the ALMO forces converge to the reference forces in the limit $R_{c} \rightarrow \infty$
 %(Figure~\ref{fig:forcecomp} in the Supplemental Material).
@@
 
-%\section{Procedure to estimate the value of $\Delta$}
+\subsection{Analytical expression for $\Delta$}
 %
@@
 
-\begin{figure*}[h!]
-\includegraphics[trim={0cm 0cm 0.1cm 0.1cm},clip,width=\textwidth]{DeltaForceComparison_ALMO_SCF.eps}
+Among several possible expression for $\Delta$ we consider the following two options.
+
+\textbf{Option 1.} $\Delta$ is calculated as a weighted average of atom TYPES (e.g. oxygen and hydrogen). Subindex $\text{T}$ refers to atom type.
+
+\textbf{\emph{Step 1.}} Compute the average ACF for each atom type.
+In this step averaging is done over all time origins $t$, over all atoms of THIS type $i$, and over all dimensions $\alpha$.
+The ACF includes the mass-dependent factor $(2 k_B T m_\text{T} )^{-1}$ where $T$ is the requested temperature.
+%
+\begin{align}
+\label{eq:deltaS1}
+\text{ACF}_{\text{T}}(\tau) = (2 k_B T m_\text{T} )^{-1} \, \langle \delta f_{i\alpha} (t) \, \delta f_{i\alpha}(t + \tau) \rangle_{t i\alpha}
+\end{align}
+%
+\textbf{\emph{Step 2.}} Integrate the ACF for each atom type.
+%
+\begin{align}
+\label{eq:deltaS2}
+\Delta_{\text{T}} = \int_{-\infty}^{\infty} \text{ACF}_{T}(\tau) \, d\tau
+\end{align}
+%
+\textbf{\emph{Step 3.}} Average all over all atom types.
+%
+\begin{align}
+\label{eq:deltaS3}
+\Delta = \frac{1}{N} \sum_{\text{T}} N_{\text{T}} \Delta_{\text{T}}
+\end{align}
+%
+where $N$ is the number of atoms.
+
+If all these step are combined the following expression for $\Delta$ is obtained:
+%
+\begin{align}
+\label{eq:deltaTOT}
+\Delta = \int \, d\tau \left[ \frac{1}{3N} \sum_{i=1}^{N} \sum_{\alpha=1}^{3} \frac{ \langle \delta f_{i\alpha} (t) \, \delta f_{i\alpha}(t + \tau) \rangle_{t} }{2 k_B T m_i} \right]
+\end{align}
+%
+Thus the weight of each of the $3N$ degrees of freedom in this scheme is proportional to the inverse atomic mass.
+
+\textbf{Option 2.} If the same averaging scheme is applied to MOLECULE types (not ATOM types) then the value of $\Delta$ is given by:
+%
+\begin{align}
+\label{eq:O2deltaTOT}
+\Delta = \int \, d\tau \left[ \frac{1}{3M} \sum_{p=1}^{M} \sum_{\alpha=1}^{3} \frac{ \sum_i^{N_p}\sum_j^{N_p} \langle \delta f_{i\alpha} (t) \, \delta f_{j\alpha}(t + \tau) \rangle_{t} }{2 k_B T \sum_k^{N_p}m_k} \right]
+\end{align}
+%
+where $M$ is the number of molecules, $N_p$ is the number of atoms in molecule $p$.
+The value of this $\Delta$ might be different from Option 1 because (a) all correlations between two atoms inside one molecule are taken into account and (b) the weight of each degree of freedom is different.
+The difference between Option 1 and 2 can be clearly seen for a special case of identical molecules:
+%
+\begin{align}
+\label{eq:O2deltaSPECIAL}
+\Delta = \int \, d\tau \left[ \frac{1}{3N} \sum_{i=1}^{N} \sum_{\alpha=1}^{3} \frac{ \sum_j^{N_p} \langle \delta f_{i\alpha} (t) \, \delta f_{j\alpha}(t + \tau) \rangle_{t} }{2 k_B T \langle m \rangle } \right]
+\end{align}
+%
+where $\langle m \rangle = (\sum_k^{N_p} m_k )/N_p$ is the average atom mass in a molecule.
+
+\subsection{Dependence of the forces on the electron localization radius}
+
+The ALMO forces converge to the reference forces as the electron localization radius increases (Figure~\ref{fig:forcecomp}).
+
+\begin{figure}[h!]
+\includegraphics[trim={0cm 0cm 0.1cm 0.1cm},clip,width=8.6cm]{DeltaForceComparison_ALMO_SCF.eps}
 \caption{\label{fig:forcecomp} Dependence of the Euclidean norm of the instantaneous error $\langle \| \delta \vec{f_{i}}(t) \| \rangle_{i}$ (red line) on the ALMO localization radius $R_c$ expressed in units of atomic van der Waals radii.
@@
 Forces are computed with the PBE XC functional, TZV2P basis set, and $\epsilon_{\text{SCF}} = 10^{-7}$~a.u. for the configurations from a 10-ps trajectory generated with KS-based AIMD at T=298~K.}
-\end{figure*}
+\end{figure}
 

### 2017-10-26 almo-langevin.tex
Changed 3N -> 3M in SI equation 6
Major modifications of the main text to take into account recent literature

@@
 % define variable that compiles the main text (as opposed to SI text)
-%\newcommand*{\MAINTEXT}{}
+\newcommand*{\MAINTEXT}{}
 
@@
 %% for non-covalently bonded atoms
+%RZK: without one-electron density matrix
 }
@@
 \begin{abstract} 
-A robust method for constructing compact localized orbitals in density functional theory enables us to create low-cost linear-scaling \emph{ab initio} molecular dynamics for weakly-interacting systems.
-High efficiency of simulations is achieved by calculating the forces approximately without lengthy optimization of localized orbitals and by adjusting the Langevin equation of motion to stabilize the dynamics with the imperfect forces.
-An application to liquid water demonstrates that extremely low computational overhead of this procedure makes it suitable for routine medium-size simulations whereas its linear scaling cost allows to extend accurate first-principle molecular dynamics to previously inaccessible time and length scales.
+We coupled a robust method for constructing compact localized orbitals within the framework of Kohn-Sham density functional theory with a tunable Langevin equation of motion to obviate the commonly employed optimization of the one-electron density matrix and to create an efficient \emph{ab initio} molecular dynamics method for weakly-interacting systems.
+%An application to liquid water demonstrates that the linear scaling complexity and low computational overhead of the developed method make it suitable for accurate first-principle simulation of a wide range of systems: from routine medium-size systems to previously inaccessible length scales.
+%well suited for routine first-principle modeling as well as 
+%for a wide range of length scales -- from routinely accessible to previously inaccessible.
+%In contrast to linear-scaling first-principle simulation methods that exploit the locality of the one-electron density matrix we coupled a robust method for constructing compact localized orbitals with a modified Langevin equation of motion to create an efficient \emph{ab initio} molecular dynamics method for weakly-interacting systems.
+%A robust method for constructing compact localized orbitals in density functional theory is coupled with a tunable Langevin equation of motion to create low-cost linear-scaling \emph{ab initio} molecular dynamics for weakly-interacting systems.
+An application to liquid water demonstrates that the low computational overhead of the developed method makes it well-suited for routine simulations whereas its linear scaling complexity allows to extend first-principle dynamical studies of molecular systems to previously inaccessible length scales.
+%
+%%%A robust method for constructing compact localized orbitals in density functional theory enables us to create low-cost linear-scaling \emph{ab initio} molecular dynamics for weakly-interacting systems.
+%%%High efficiency of simulations is achieved by calculating the forces approximately without lengthy optimization of localized orbitals and by adjusting the Langevin equation of motion to stabilize the dynamics with the imperfect forces.
+%%%An application to liquid water demonstrates that a low computational overhead of this procedure makes it suitable for routine medium-size simulations whereas its linear scaling cost allows to extend accurate first-principle molecular dynamics
+%%%%[RZK: of molecular systems] 
+%%%to previously inaccessible length scales. 
+%%%% RZK: no DM, only orbitals. Make it sound like a theory development not implementation focused.
 \end{abstract}
@@
 Since the unification of molecular dynamics and density functional theory (DFT)~\cite{a:thecpmd},
-\emph{ab initio} molecular dynamics (AIMD) has become an important tool for the study of weakly bonded ensembles of molecules such as gas-phase molecular clusters, liquids, solutions, and molecular solids.
-%[of gas-phase molecules and condensed-phase systems.]
+\emph{ab initio} molecular dynamics (AIMD) has become an important tool to study processes in molecules and materials.
+%weakly bonded ensembles of molecules such as gas-phase molecular clusters, liquids, solutions, and molecular solids. 
 Unfortunately, the computational cost of the conventional Kohn-Sham (KS) DFT grows cubically with the number of atoms, which severely limits the \emph{length scales} accessible by AIMD.
 %and hinders it application to large heterogeneous systems (interfaces, nuclei, defects).
-To address this issue, substantial efforts have been directed to the development of linear scaling (LS) DFT methods.
+To address this issue, substantial efforts have been directed to the development of linear-scaling (LS) DFT.
 
-% RZK: search and include recent references
-In all LS DFT methods, the \emph{delocalized} eigenstates of the effective KS Hamiltonian must be replaced with an alternative set of \emph{local} electronic descriptors
-~\footnote{The delocalization of KS orbitals over all atoms is the main reason for the steep growth of the cost of KS DFT: since both the number of orbitals and the number of localization regions (e.g.
+In all LS DFT methods, the \emph{delocalized} eigenstates of the effective KS Hamiltonian must be replaced with an alternative set of \emph{local} electronic descriptors. 
+%~\footnote{The delocalization of KS orbitals over all atoms is the main reason for the steep growth of the cost of KS DFT: since both the number of orbitals and the number of localization regions (e.g.
 atomic basis functions) increase linearly with the number of atoms, the total number of the electronic degrees of freedom grows quadratically.}.
-Most LS methods~\cite{a:ls-rev-1999,a:ls-rev-2012} explore the natural locality of the one-electron density matrix (DM). 
-However, in typical DFT calculations that require diffuse basis functions, the DM approach is advantageous only for impractically large systems~\cite{a:ls-rev-1999,a:ls-dm-sign,a:almo-ls}.
-Therefore, the application of DM-based LS methods have been restricted to minimal-basis tight-binding problems. 
-The optimal basis variants of the DM methods~\cite{a:ls-stechel-1994,a:ls-gillan-1995,a:ls-gillan-1996,a:ls-onetep-2003} designed to rectify this issue contract large basis sets into a small number of new localized basis functions and then optimize the DM in the contracted basis.
-Although such methods have been successfully used for the evaluation of accurate DFT energies of large static systems~\cite{a:ls-onetep-2009,a:ls-conquest-2010,a:ls-onetep-2010-app1,a:ls-rev-2012} their use in AIMD is hampered by the costly optimization of both the contracted orbitals and the density matrix.
-
-% DM-based LS methods: a:linscale3,a:lee-yang-1996,a:ls-scuseria-1997,a:ls-manolopoulos-1998,a:ls-helgaker-2001,a:ls-niklasson-2003,a:curvy2,a:ls-dm-sign
-
-To obviate the DM optimization, another set of electronic descriptors can be constructed by enforcing strict locality on one-electron orbitals and, necessarily, by relaxing the orthogonality constraints imposed on them.
-From the computational point of view, LS methods based on the direct optimization of such compact localized nonorthogonal orbitals~\cite{a:ls-galli-parrinello-1992,a:ls-mauri-galli-car-1993,a:ls-ordejon-1993,a:ls-mauri-galli-1994,a:ls-ordejon-1995,a:ls-kim-mauri-galli-1995,a:ls-fattebert-2004,a:ls-fattebert-2006,a:burger-yang-2008} are promising since they deal with fewer variational degrees of freedom than the DM methods.
-Unfortunately, the development of orbital-based LS methods has been all but abandoned because of the inherently difficult optimization of localized orbitals~\cite{a:ls-mauri-galli-car-1993,a:ls-ordejon-1995,a:ls-fattebert-2004,a:ls-rev-1999,a:weitao-yang-2013}.
+Most LS methods~\cite{a:ls-rev-1999, a:ls-rev-2012, Kussmann2013, Aarons2016, optimal-basis-too} explore the natural locality of the one-electron density matrix (DM). 
+However, with accurate multifunction basis sets the DM DFT becomes advantageous only for impractically large systems~\cite{a:ls-dm-sign, Arita2014, a:ls-rev-2012, a:ls-almo}.
+%Therefore, the application of DM-based LS methods have been restricted to minimal-basis tight-binding problems~\cite{Richters2014, more}. 
+This issue is rectified in optimal-basis DM methods~\cite{Skylaris2005, Nakata2015, Mohr2015} that contract large basis sets into a small number of new localized functions and then optimize the DM in the contracted basis.
+%Later: Although such methods have been successfully used for the evaluation of accurate DFT energies of large static systems~\cite{a:ls-onetep-2009,a:ls-conquest-2010,a:ls-onetep-2010-app1,a:ls-rev-2012}
+%While most large-scale DFT calculations are performed using this class of methods
+Despite becoming the most popular approach to LS DFT the efficiency of these methods is hampered by the costly optimization of both the contracted orbitals and the DM~\cite{convergence-gilan}. 
+%
+%Richters2014 -- DOI:10.1063/1.4869865
+%Aarons2016 -- DOI:10.1063/1.4972007
+%Kussmann2013 -- DOI:10.1002/wcms.1138
+%Vitale2017 -- DOI:10.1063/1.4978684
+%Skylaris2005 -- DOI:10.1063/1.1839852
+%Hine2011 -- DOI:10.1103/PhysRevB.83.195102
+%Bowler2010 -- DOI:10.1088/0953-8984/22/7/074207
+%Nakata2015 -- DOI:10.1039/C5CP00934K
+%Krajewski -- PRB, 73, 041105R
+%Kolafa2003 -- J Comput Chem 25: 335–342, 2004
+%
+From this point of view, a direct variation of molecular orbitals that are strictly localized within predefined regions is preferable because LS can be achieved without costly DM optimization and with significantly fewer variables.
+%especially in the case of large accurate basis sets.
+Advantages of the orbitals-only DFT are especially pronounced in accurate calculations that require many basis functions per atom.
+%From this point of view, it is desirable to obviate the DM optimization and use another set of local electronic descriptors can be constructed by enforcing strict locality on one-electron orbitals and, necessarily, by relaxing the orthogonality constraints imposed on them.
+%
+% direct optimization of such compact localized nonorthogonal orbitals
+%
+%To obviate the DM optimization, another set of local electronic descriptors can be constructed by enforcing strict locality on one-electron orbitals and, necessarily, by relaxing the orthogonality constraints imposed on them.
+%From the computational point of view, LS methods based on the direct optimization of such compact localized nonorthogonal orbitals~\cite{a:ls-galli-parrinello-1992,a:ls-mauri-galli-car-1993,a:ls-ordejon-1993,a:ls-mauri-galli-1994,a:ls-ordejon-1995,a:ls-kim-mauri-galli-1995,a:ls-fattebert-2004,a:ls-fattebert-2006,a:burger-yang-2008} are promising since they deal with fewer variational degrees of freedom than the DM methods.
+Unfortunately, the development of promising orbital-based LS methods has been all but abandoned~\cite{a:review1999, a:weitao-yang-2013,japanese-guys} because of the inherently difficult optimization of localized orbitals~\cite{a:ls-mauri-galli-car-1993,a:ls-ordejon-1995,a:ls-fattebert-2004,a:ls-rev-1999,a:weitao-yang-2013,japanese-guys-citedinmy2013paper}.
 
@@
 
-Thus, the high computational overhead of existing asymptotically linear-scaling DFT methods restrict their use in AIMD simulations to very short \emph{time scales}, systems of low dimensions, or problems that can be adequately described with minimal basis sets.
-On the length and time scales accessible in typical AIMD simulations, LS DFT cannot compete with the straightforward low-cost cubically-scaling KS DFT.
+Thus, despite impressive progress of LS description of the electronic and atomic structure of large static systems~\cite{Bowler2010,a:ls-dm-sign} the high computational overhead of existing LS methods restrict their use in dynamical simulations to very short \emph{time scales}, systems of low dimensions, and low-quality minimal basis sets~\cite{a:ls-dm-sign, RZK:10.7567/JJAP.55.1102B1,Hine2011, Bowler2010}.
+%RZK: Although such methods have been successfully used for the evaluation of accurate DFT energies [electronic structure and atomic configurations] of large static systems~\cite{Bowler2010,a:ls-dm-sign} ...
+%Thus, the high computational overhead of existing LS DFT methods restrict their use in [dynamical] AIMD simulations to very short \emph{time scales}, systems of low dimensions, and low-quality minimal basis sets~\cite{a:ls-dm-sign, RZK:10.7567/JJAP.55.1102B1,Hine2011, Bowler2010}.
+On typical length and time scales required in practical and accurate AIMD simulations, LS DFT still cannot compete with the straightforward low-cost cubically-scaling KS DFT.
 
-In this work, we developed a new orbital-based AIMD method that is not only linear scaling but also has a low computational overhead. 
-Currently, the method is restricted to systems of small molecules that do not form strong covalent bonds with each other. 
+In this work, we present an AIMD method that overcomes difficulties of orbital-only local DFT to achieve LS with extremely low computational overhead. 
+%Currently, the method is restricted to systems of small molecules that do not form strong covalent bonds with each other. 
+To demonstrate advantages of the new method we applied it here to systems of weakly-interacting molecules. 
+However, the same approach is readily applicable to systems of strongly-interacting fragments that do not form strong covalent bonds such as ionic salts, liquids, and semiconductors. 
+A generalization of the method to all finite-gap systems, including covalently bonded atoms, will be reported later. 
+%It is important to emphasize that delocalized electrons in metals, which are particularly challenging for all LS DFT~\cite{Kuehne,recent-Skylaris}, cannot be described efficiently within the presented framework.
+
 The new AIMD method utilizes a recently developed LS DFT~\cite{a:almo-ls} based on absolutely localized molecular orbitals (\mbox{ALMOs}). 
-Unlike delocalized KS orbitals, each \mbox{ALMO} has its own localization center and a predefined localization radius $R_{c}$ that typically includes nearby atoms or molecules~\cite{a:stoll,a:almo-ls}.
-In the current implementation, a localization center is a set of all local Gaussian atomic orbitals of one molecule. 
+Unlike delocalized KS orbitals, each \mbox{ALMO} has its own \emph{localization center} and a predefined \emph{localization radius} $R_{c}$ that typically includes nearby atoms or molecules~\cite{a:stoll,a:almo-ls}.
+In the current implementation, a localization center is defined as a set of all Gaussian atomic orbitals of one molecule. 
 However the approach can be easily generalized to other local and nonlocal basis sets~\cite{a:ls-galli-parrinello-1992, RZK:10.1016/j.jcp.2011.11.032}. 
-The key feature of ALMO DFT is that its one-electron wavefunctions are constructed in a two-step self-consistent-field (SCF) procedure~\cite{a:almo-ls} to circumvent the problem of the sluggish variational optimization emphasized above.
-In the first step, ALMOs are constrained to their localization centers~\cite{a:khal} whereas, in the second step, ALMOs are relaxed to allow delocalization onto the neighbor molecules within their localization radius $R_{c}$.
-To achieve a robust optimization in the problematic second step, it is important to keep the delocalization component of the trial wavefunction orthogonal to the fixed orbitals obtained in the first step.
+The key feature of ALMO DFT is that its one-electron wavefunctions are constructed in a two-stage self-consistent-field (SCF) procedure~\cite{a:almo-ls} to circumvent the problem of the sluggish variational optimization emphasized above.
+In the first stage, ALMOs are constrained to their localization centers~\cite{a:khal} whereas, in the second stage, ALMOs are relaxed to allow delocalization onto the neighbor molecules within their localization radius $R_{c}$.
+To achieve a robust optimization in the problematic second stage, it is important to keep the delocalization component of the trial wavefunction orthogonal to the fixed orbitals obtained in the first stage.
 For mathematical details, see ALMO SCF method in Ref.~\citenum{a:almo-ls}.
@@
 ALMO constraints imposed by $R_c$ prohibit electron density transfer between distant molecules but retain all other types of interaction such as long-range electrostatic, exchange, polarization, and, if the exchange-correlation (XC) functional includes them, dispersion interactions~\cite{a:theeda}.
-Since the importance of electron transfer decays exponentially with distance, the \mbox{ALMO} approximation is expected to provide a natural and accurate representation of the electronic structure of molecular systems.
+Since the importance of electron transfer decays exponentially with distance in finite-gap materials~\cite{a:ls-rev-1999}, the \mbox{ALMO} approximation is expected to provide a natural and accurate representation of the electronic structure of molecular systems.
 %with just a few local variables.
@@
 The challenge of adopting ALMO DFT for dynamical simulations arises from the slightly nonvariational character of the localized orbitals.
-First, while ALMOs are variationally optimized in both SCF steps, the occupied subspace defined in the first step must remain fixed during the second step to ensure convergence.
+First, while ALMOs are variationally optimized in both SCF stages, the occupied subspace defined in the first stage must remain fixed during the second stage to ensure convergence.
 Second, neglected electron transfer effects can suddenly become active in the course of a dynamical simulation when a neighboring molecule passes the localization threshold $R_{c}$.
-In addition to these issues characteristic to ALMOs, the variational optimization in any AIMD method is never truly complete and interrupted once the maximum norm of the gradient of the energy with respect to the electronic descriptors drops below small but nevertheless finite convergence threshold $\epsilon_{\text{SCF}}$.
-While these nonvariational errors do not affect the accuracy of static ALMO DFT calculations, geometry optimization, and Monte-Carlo simulations, they accumulate in molecular dynamics trajectories leading to non-physical sampling and eventual failure.
-Traditional strategies to cope with these problems are computationally expensive and include computing the nonvariational contribution to the forces via a variational coupled-perturbed procedure, increasing $R_c$, and decreasing $\epsilon_{\text{SCF}}$.
+In addition to these issues specific to ALMO DFT, the variational optimization in any AIMD method is never truly complete in practice
+%[exact convergence is never achieved in practical calculation] 
+and interrupted once the maximum norm of the gradient of the energy with respect to the electronic descriptors drops below small but nevertheless finite convergence threshold $\epsilon_{\text{SCF}}$.
+These nonvariational errors do not affect the accuracy of static ALMO DFT calculations, geometry optimization, and Monte-Carlo simulations.
+Unfortunately they tend to accumulate in molecular dynamics trajectories leading to non-physical sampling and eventual failure.
+Traditional strategies to cope with these problems are computationally expensive and include computing the nonvariational contribution to the forces via a variational coupled-perturbed procedure~\cite{Kussmann2013}, increasing $R_c$, and decreasing $\epsilon_{\text{SCF}}$.
 %This however would dramatically increase the computational cost of the ALMO AIMD.
@@
 
-In this work, we propose another approach that obviates the need in a coupled-perturbed solver, relaxes tight constraints on $R_{c}$ and $\epsilon_{\text{SCF}}$, and thus enables us to maintain stable dynamics and to keep the complexity and cost of simulations low.
+In this work, we propose another approach that obviates the need in a coupled-perturbed solver, relaxes tight constraints on $R_{c}$ and $\epsilon_{\text{SCF}}$, and thus enables us to maintain stable dynamics and to keep the algorithmic complexity and cost of simulations low.
 %
-In our approach, the forces on atoms are calculated \emph{approximately} after the two-step ALMO SCF using a straightforward procedure that computes only the Hellmann-Feynman and Pulay components and neglects the computationally intense nonvariational component of the forces.
+In our approach, the forces on atoms are calculated \emph{approximately} after the two-stage ALMO SCF using a straightforward procedure that computes only the Hellmann-Feynman and Pulay components and neglects the computationally intense nonvariational component of the forces.
 The difference between these approximate ALMO forces and the \emph{reference} forces that could be obtained from perfectly converged fully-delocalized KS orbitals is $\delta f_{i\alpha}(t)$:
@@
 
-Our approach to compensate for the missing $\delta f_{i\alpha}(t)$ term follows the methodology that have been introduced into AIMD by Krajewski \emph{et al.}~\cite{RZK:PRB-73-041105R}, formalized by K\"uhne \emph{et al.}~\cite{a:2ndcpmd} and rationalized by Dai \emph{et al.}~\cite{a:langevin-why} before becoming informally known as the second generation Car-Parrinello molecular dynamics~\cite{RZK:10.1002/wcms.1176}.
-Inspired by this method, ALMO AIMD is chosen to be governed by the Langevin equation of motion that is first written in terms of the unknown reference forces:
+Our approach to compensate for the missing $\delta f_{i\alpha}(t)$ term is inspired by the methodology introduced into AIMD by Krajewski \emph{et al.}~\cite{Krajewski}, formalized by K\"uhne \emph{et al.}~\cite{a:2ndcpmd} and rationalized by Dai \emph{et al.}~\cite{a:langevin-why} before becoming informally known as the second generation Car-Parrinello molecular dynamics~\cite{RZK:10.1002/wcms.1176}.
+Adopting the principle of Refs.~\onlinecite{Krajewski} and \onlinecite{a:2ndcpmd}, ALMO AIMD is chosen to be governed by the Langevin equation of motion that can be written in terms of the unknown reference forces:
 %
@@
 % 
-The last relation means that, for any value of $\gamma$, the damping and stochastic terms are in perfect balance and trajectories generated with Eq.~(\ref{eq:langevin}) will sample the canonical ensemble at specified temperature $T$~\cite{a:Kubo-1986}.
-%will generate the correct Maxwell-Boltzmann distribution of velocities for a desired temperature $T$~\cite{a:Kubo-1986}. 
-Larger $\gamma$ lead to faster equilibration but more distorted results for dynamical properties.
-%In the limit $\gamma \rightarrow 0$, the Newton equation is recovered and the microcanonical ensemble is sampled.
-
-%It has previously been shown that Langevin dynamics can be utilized to correct inaccurate force calculations to maintain correct dynamical properties.~\cite{a:langevin-why,a:2ndcpmd,b:tuckerman-stat,a:ceriotti}
-
-%By utilizing Eq.\ (\ref{eq:langevin}) in conjunction with the ALMO SCF method of force calculation, we will show that discontinuity in the PES can be ignored for sufficiently large $\gamma$. 
-% Perhaps this statement is not correct: Note that increasing $\gamma$ increases the sample size required to generate accurate dynamical properties, as more random Gaussian noise is introduced.
-
-%The concept of balancing errors in force calculation with the Langevin equation can be taken even further, thereby reducing computational overhead even more. 
-%%During the second step of the ALMO SCF method described above, a self-consistent variational approach is taken to converge the non-orthogonal MOs. 
-%Generally, this step utilizes a iterative search function in order to reduce the absolute norm of the energy gradient to some target value $\epsilon_{\text{SCF}}$. 
-%Previously, $\epsilon_{\text{SCF}}$ had to be set very small to ensure sufficient convergence of the electron structure calculations. 
-%By increasing the allowable $\epsilon_{\text{SCF}}$, a great reduction in computational cost can be achieved. 
-%Obviously, this introduces a significant error into the calculation of forces during each time step.
+The last relation means that, for any value of $\gamma$, the damping and stochastic terms are in perfect balance and trajectories generated with Eq.~(\ref{eq:langevin}) will sample the canonical ensemble at a specified temperature $T$~\cite{a:Kubo-1986}.
+In the limit $\gamma \rightarrow 0$, the Newton equation is recovered and the microcanonical ensemble is sampled. 
+%Increasing $\gamma$ leads to faster equilibration but more distorted results for dynamical properties.
 
@@
 This term compensates for imperfections in ALMO forces and must be adjusted to re-balance the damping and stochastic components in ALMO AIMD.
-%and generate the correct Maxwell-Boltzmann distribution in ALMO AIMD.
 
@@
 Eliminating the problem
-In principle, $\Delta$ can be calculated using the integral of Eq.~(\ref{eq:stochastic3}) weighed over all types of atoms with different $m_i$
+In principle, $\Delta$ can be calculated using the integral of Eq.~(\ref{eq:stochastic3}) averaged over atoms with different $m_i$
 %
@@
 \label{eq:delta}
-\Delta &= (6 k_B T m_i )^{-1} \int_{-\infty}^{\infty} \langle \delta \vec{f}_i (0) \cdot \delta\vec{f}_i(\tau) \rangle d\tau
+\Delta &= (2 k_B T m_i )^{-1} \int_{-\infty}^{\infty} \frac{1}{3} \langle \delta \vec{f}_i (0) \cdot \delta\vec{f}_i(\tau) \rangle d\tau
 \end{align}
@@
 if one can afford computing the reference forces $f^{\text{KS}}_{i\alpha}(t)$ (i.e. $R_c \rightarrow \infty$ and $\epsilon_{\text{SCF}} \rightarrow 0$) for a short representative AIMD trajectory. 
-In practice, we found (see results below) that this approach is not particularly accurate due to the noisy autocovariance functions (ACF) from short trajectories and perhaps because the $\delta_{ij}\delta_{\alpha\beta}$ assumption in Eq.~(\ref{eq:stochastic3}) does not strictly hold.
+In practice, we found (see results below) that this approach is not particularly accurate because the $\delta_{ij}\delta_{\alpha\beta}$ assumption in Eq.~(\ref{eq:stochastic3}) does not strictly hold.
 Nevertheless, the ACF integral can provide a reasonable starting value of $\Delta$.
-This value can be further fine-tuned in a series of short trial-and-error ALMO AIMD runs until the average kinetic energy corresponds to the requested temperature $\langle \frac{1}{2} m_i \dot{\bm{r}}^{2}_{i} \rangle = \frac{3}{2} k_{B} T$ .
+This value can be further fine-tuned in a series of short trial-and-error ALMO AIMD runs until the average kinetic energy corresponds to the requested temperature $\langle \frac{1}{2} m_i \dot{\bm{r}}^{2}_{i} \rangle = \frac{3}{2} k_{B} T$.
 %%
@@
 %\end{align}
+
+Thus, the inherently stochastic approach presented here does not aim to produce fully time-reversible dynamics for atomic nuclei. 
+However, with $\gamma=0$, the nuclei will follow time-reversible Newtonian trajectories sufficiently closely as long as partially optimized ALMOs remain close to the ground state and keep $\Delta$ small.
+%
+%(for systems with slight discontinuities in the potential energy surfaces)
+%
+Although not a required component of ALMO AIMD, the predictor part of the Kolafa scheme~\cite{kolafa2003} adopted to localized orbitals was found to generate a highly accurate initial ALMOs in both SCF stages, which can be brought close to the ground state with just a few SCF steps.
+
+%RZK: The time-reversibility of ASPC is not a critical reqirement of the proposed method.
 
@@
 %
-To measure the ability of the $R^{\Delta}$ term to compensate for imperfections in ALMO forces we varied $\epsilon_{\text{SCF}}$ between tight $10^{-6}$~a.u. and loose $10^{-2}$~a.u.
+To check the ability of the $R^{\Delta}(t)$ term to compensate for imperfections in ALMO forces we varied $\epsilon_{\text{SCF}}$ between tight $10^{-6}$~a.u. and loose $10^{-2}$~a.u.
 
@@
 $\Delta$ was initially estimated at $\SI{2e-5}{\per\fs}$ using Eq.~(\ref{eq:delta}) and then refined heuristically to $\SI{6e-5}{\per\fs}$.
-%RZK: factor of 3 bothers me. We need to check that this is not because we sum over 3 atoms in a molecule. See previous comments on the ACF integral.
 We found that it is easier to optimize $\Delta$ when $\gamma$ is set to zero because of reduced noise in the trial runs. 
-%In theory, $\Delta$ should be independent on system size. However, we found that $\Delta$ dependents on system size due to finite size effects, which decay for increasing system size.
+%RZK addition for a longer article: In theory, $\Delta$ should be independent on system size.
+However, we found that $\Delta$ dependents on system size due to finite size effects, which decay for increasing system size.
 %
@@
 Thus the main assumption behind our approach to ALMO AIMD is justified for liquid water.
-The main source of error in forces for this system is the loose convergence criterion and not the finite $R_c$: fully converged ALMO SCF calculations remove the rapidly oscillating component of the error (Figure~\ref{fig:randomforce}b).
+We established that the main source of error in forces for this system is the loose convergence criterion and not the finite $R_c$: fully converged ALMO SCF calculations remove the oscillating component of $\delta f$ (Figure~\ref{fig:randomforce}b).
 We also verified that the ALMO forces converge to the reference forces in the limit $R_{c} \rightarrow \infty$
@@
 The quality of the ALMO IR spectrum (Figure~\ref{fig:dynproperties}c) is good despite minor errors in the intensity of the OH stretching mode, which is sensitive to the precise positions of the centers of localized orbitals.
-These stringent tests show that despite noticeable errors in the ALMO forces (Figure \ref{fig:randomforce}a) the compensating $R^{\Delta}$ term in the modified Langevin equation makes it possible to recover atomic dynamics properly.
+These stringent tests show that despite noticeable errors in the ALMO forces (Figure \ref{fig:randomforce}a) the compensating $R^{\Delta}(t)$ term in the modified Langevin equation makes it possible to recover atomic dynamics properly.
 We would like to note that ALMO AIMD could not be stabilized with $\Delta=0$.
@@
 
-To demonstrate the computational advantage of ALMO AIMD, we compared the average wall-time per MD step over 10 time steps for a variety of methods in Figure \ref{fig:strongscaling_log}.
-In addition to the fully converged OT reference, we included timing for the second generation Car-Parrinello method designed to decrease the computational prefactor of the cubically scaling OT method~\cite{a:2ndcpmd}.
-As expected ALMO methods show clear LS behavior for all values of $\epsilon_{\text{SCF}}$ even for medium-size systems. 
-%OT SCF scales approximately cubically, as expected.
-For liquid water, strict localization of ALMOs lowers the crossover point between ALMO AIMD and cubically scaling methods to 256 molecules -- length scales routinely accessible with AIMD today.
-It is important to emphasize that the comparison is performed for a three-dimensional condensed phase system described with a large diffuse basis set -- a particularly challenging case for LS methods.
+% RZK longer article: averaging is done over 10 time steps 
+To demonstrate computational efficiency of ALMO AIMD, we compared the average wall-time per MD step for a variety of methods in Figure \ref{fig:strongscaling_log}.
+%In addition to the fully converged OT reference, we included timing for the second generation Car-Parrinello method designed to decrease the computational prefactor of the cubically-scaling DFT~\cite{a:2ndcpmd}.
+It is important to emphasize that the comparison is performed for a three-dimensional condensed phase system described with an accurate triple-$\zeta$ basis set with polarized functions -- a particularly challenging case for DM-based LS methods.
+ALMO AIMD shows clear LS behavior for all values of $\epsilon_{\text{SCF}}$ even for medium-size systems.
+While the second generation Car-Parrinello method decreases the computational overhead of the cubically-scaling AIMD for small systems~\cite{a:2ndcpmd}, ALMO AIMD exploits the modified Langevin concept to substantially reduce the simulation cost for systems of all sizes.
+The crossover point between ALMO AIMD and cubically scaling methods lies in the region of 256 molecules -- length scales routinely accessible with AIMD today.
+
+%RZK: time-reversibility must be mentioned or even addressed
+
+%RZK longer article: fully converged ALMO calculations must be reported: Delta=0, gamma=0, is total energy constant at R=1.6 vdW?
 
@@
 
-Weak scaling benchmarks for very large systems show (Figure~\ref{fig:weakscaling}) that localized orbitals are naturally suited for parallel execution and that the computational cost of our implementation grows linearly for a wide range of compute cores.
-In fact, we were able to successfully simulate systems as large as 32768 water molecules within reasonable wall-clock time using moderate number of compute cores -- unprecedented feat for AIMD considering that accurate molecular orbitals and the idempotent DM are computed on each step.
+%RZK: number of atoms is 98,304 with very accurate basis set.
+
+Weak scaling benchmarks for very large systems show (Figure~\ref{fig:weakscaling}) that localized orbitals are naturally suited for parallel execution: LS is retained for a wide range of systems and compute cores.
+We were able to successfully simulate systems as large as $\sim10^{5}$ atoms 
+%(32768 water molecules) 
+within reasonable wall-clock time using only moderate number of compute cores -- an unprecedented feat for AIMD considering that accurate molecular orbitals and the idempotent DM are computed on each step.
 The horizontal line in Figure~\ref{fig:weakscaling} is shown as a rough guide to time and length scales accessible in a fixed wall-clock time given various computational resources.
 It indicates that ALMO AIMD can extend the range of routine simulations to $\sim10^4$ atoms on modern HPC platforms.
+
+%CONQUEST code: 32768-atom crystalline silicon, 1 MD step is 1085 s using 1024 CPUs (with 16 cores/CPU) of Fujitsu FX10
 
@@
 % systematically improvable approximation to DFT-based AIMD.
-To summarize, we demonstrated, for the first time, the utility of compact localized orbitals for low-cost linear-scaling AIMD for weakly-interacting systems. 
-High efficiency is achieved without sacrificing accuracy with a combination of two techniques: (1) on-the-fly calculation of approximate forces without lengthy self-consistent optimization of localized orbitals and (2) integration of a modified Langevin equation of motion that is fine-tuned to retain stable and correct dynamics even with imperfect forces.
-A remarkable feature of the method is that it remains efficient for challenging condensed phase systems even with diffuse atom-centered basis sets. 
-Using liquid water as an example, we showed that the new AIMD method enables simulations of molecular systems on previously inaccessible time and length scales. 
-The developed methods will have a significant impact on modeling of complex molecular systems (e.g. interfaces, nuclei) making completely new phenomena accessible to AIMD. 
+To summarize, we demonstrated, for the first time, that compact localized orbitals can be utilized to perform accurate and efficient LS AIMD without concomitant optimization of the DM.
+%for weakly-interacting systems.
+High efficiency of the presented method is achieved without sacrificing accuracy with a combination of two techniques: (1) on-the-fly calculation of approximate forces without lengthy self-consistent optimization of localized orbitals and (2) integration of a modified Langevin equation of motion that is fine-tuned to retain stable dynamics even with imperfect forces.
+By obviating the optimization of the DM the method remains remarkably efficient even with large localized basis sets. 
+%A remarkable feature of the method is that it remains efficient for challenging condensed phase systems even with large localized basis sets. 
+Using liquid water as an example, we showed that the new approach enables simulations of molecular systems on previously inaccessible time and length scales. 
+The developed method will have a significant impact on modeling of complex molecular systems (e.g. interfaces, nuclei) making completely new phenomena accessible to AIMD. 
 Generalization of the methodology to systems of strongly interacting atoms (e.g. covalent crystals) is underway.
 
-\textbf{Acknowledgments.} The research was funded by the Natural Sciences and Engineering Research Council of Canada through the Discovery Grant.
+\textbf{Acknowledgments.} The authors are grateful to Thomas K\"uhne for insightful discussions.
+The research was funded by the Natural Sciences and Engineering Research Council of Canada through the Discovery Grant.
 The authors are grateful to Compute Canada and McGill HPC Centre for computer time.
@@
 %calculated as the average difference between fully converged Born-Oppenheimer forces and fully converged ALMO DFT forces, averaged over all dimensions and all molecules for each time step.
-Forces are computed with the PBE XC functional, TZV2P basis set, and $\epsilon_{\text{SCF}} = 10^{-7}$~a.u. for the configurations from a 10-ps trajectory generated with KS-based AIMD at T=298~K.}
+Forces are computed with the PBE XC functional, TZV2P basis set, and $\epsilon_{\text{SCF}} = 10^{-7}$~a.u. for the configurations from a 10-ps trajectory generated with delocalized AIMD at T=298~K.}
 \end{figure}

### 2017-11-16 almo-langevin.tex
Revision to include most recent literature, minor issues + diff coeff is off

@@
 Low-cost linear-scaling \emph{ab initio} molecular dynamics for weakly-interacting systems
-%%for weakly-coupled atoms
-%% for non-covalently bonded atoms
 %RZK: without one-electron density matrix
@@
 
+%RZK: one order of magnitude error in the PBE+D3 diffusion constant
+
 \begin{abstract}
-We coupled a robust method for constructing compact localized orbitals within the framework of Kohn-Sham density functional theory with a tunable Langevin equation of motion to obviate the commonly employed optimization of the one-electron density matrix and to create an efficient \emph{ab initio} molecular dynamics method for weakly-interacting systems.
-%An application to liquid water demonstrates that the linear scaling complexity and low computational overhead of the developed method make it suitable for accurate first-principle simulation of a wide range of systems: from routine medium-size systems to previously inaccessible length scales.
-%well suited for routine first-principle modeling as well as 
-%for a wide range of length scales -- from routinely accessible to previously inaccessible.
-%In contrast to linear-scaling first-principle simulation methods that exploit the locality of the one-electron density matrix we coupled a robust method for constructing compact localized orbitals with a modified Langevin equation of motion to create an efficient \emph{ab initio} molecular dynamics method for weakly-interacting systems.
-%A robust method for constructing compact localized orbitals in density functional theory is coupled with a tunable Langevin equation of motion to create low-cost linear-scaling \emph{ab initio} molecular dynamics for weakly-interacting systems.
-An application to liquid water demonstrates that the low computational overhead of the developed method makes it well-suited for routine simulations whereas its linear scaling complexity allows to extend first-principle dynamical studies of molecular systems to previously inaccessible length scales.
-%
-%%%A robust method for constructing compact localized orbitals in density functional theory enables us to create low-cost linear-scaling \emph{ab initio} molecular dynamics for weakly-interacting systems.
-%%%High efficiency of simulations is achieved by calculating the forces approximately without lengthy optimization of localized orbitals and by adjusting the Langevin equation of motion to stabilize the dynamics with the imperfect forces.
-%%%An application to liquid water demonstrates that a low computational overhead of this procedure makes it suitable for routine medium-size simulations whereas its linear scaling cost allows to extend accurate first-principle molecular dynamics
-%%%%[RZK: of molecular systems] 
-%%%to previously inaccessible length scales. 
-%%%% RZK: no DM, only orbitals. Make it sound like a theory development not implementation focused.
+Within the framework of linear-scaling Kohn-Sham density functional theory, a robust method for maintaining compact localized orbitals close to the ground state is coupled with nuclear dynamics to obviate the commonly employed optimization of the one-electron density matrix and to create an efficient \emph{ab initio} molecular dynamics method for weakly-interacting systems.
+An application to liquid water demonstrates that the low computational overhead of the developed method makes it well-suited for routine simulations whereas its linear scaling complexity allows to extend first-principle dynamical studies of molecular systems to previously inaccessible time and length scales.
+%Within the framework of linear-scaling Kohn-Sham density functional theory, a robust method for maintaining compact localized orbitals close to the ground state coupled with nuclear dynamics enables us to create low-cost linear-scaling \emph{ab initio} molecular dynamics for weakly-interacting systems.
+High efficiency of simulations is achieved by calculating the forces approximately without lengthy optimization of localized orbitals and by adjusting the Langevin equation of motion to stabilize the dynamics with the imperfect forces.
+An application to liquid water demonstrates that low computational overhead of this procedure makes it well-suited for routine simulations whereas its linear scaling complexity allows to extend first-principle molecular dynamical studies to previously inaccessible time and length scales.
 \end{abstract}
@@
 \ifdefined\MAINTEXT
+
+% NEW REFERENCES:
+%Richters2014 -- DOI:10.1063/1.4869865
+%Aarons2016 -- DOI:10.1063/1.4972007
+%Kussmann2013 -- DOI:10.1002/wcms.1138
+%Vitale2017 -- DOI:10.1063/1.4978684
+%Skylaris2005 -- DOI:10.1063/1.1839852
+%Hine2011 -- DOI:10.1103/PhysRevB.83.195102
+%Bowler2010 -- DOI:10.1088/0953-8984/22/7/074207
+%Nakata2015 -- DOI:10.1039/C5CP00934K
+%Krajewski -- PRB, 73, 041105R
+%Kolafa2003 -- J Comput Chem 25: 335–342, 2004
+%Benoit2001 -- DOI:10.1103/PhysRevLett.87.226401
+%Otsuka2016 -- DOI:10.7567/JJAP.55.1102B1
+%Lin2012 -- DOI:10.1016/j.jcp.2011.11.032
+%Kuhne2013 -- DOI:10.1002/wcms.1176
+%Ricci2003 -- DOI:10.1080/0026897031000108113
 
@@
 atomic basis functions) increase linearly with the number of atoms, the total number of the electronic degrees of freedom grows quadratically.}.
-Most LS methods~\cite{a:ls-rev-1999, a:ls-rev-2012, Kussmann2013, Aarons2016, optimal-basis-too} explore the natural locality of the one-electron density matrix (DM). 
-However, with accurate multifunction basis sets the DM DFT becomes advantageous only for impractically large systems~\cite{a:ls-dm-sign, Arita2014, a:ls-rev-2012, a:ls-almo}.
+Most LS methods~\cite{a:ls-rev-1999, a:ls-rev-2012, Kussmann2013, Aarons2016} explore the natural locality of the one-electron density matrix (DM). 
+However, the DM DFT becomes advantageous only for impractically large systems when accurate multifunction basis sets are used~\cite{a:ls-dm-sign, Arita2014, a:ls-rev-2012, a:almo-ls}.
 %Therefore, the application of DM-based LS methods have been restricted to minimal-basis tight-binding problems~\cite{Richters2014, more}. 
@@
 %While most large-scale DFT calculations are performed using this class of methods
-Despite becoming the most popular approach to LS DFT the efficiency of these methods is hampered by the costly optimization of both the contracted orbitals and the DM~\cite{convergence-gilan}. 
+Despite becoming the most popular approach to LS DFT the efficiency of these methods is hampered by the costly optimization of both the contracted orbitals and the DM~\cite{a:ls-onetep-2003}. 
 %
-%Richters2014 -- DOI:10.1063/1.4869865
-%Aarons2016 -- DOI:10.1063/1.4972007
-%Kussmann2013 -- DOI:10.1002/wcms.1138
-%Vitale2017 -- DOI:10.1063/1.4978684
-%Skylaris2005 -- DOI:10.1063/1.1839852
-%Hine2011 -- DOI:10.1103/PhysRevB.83.195102
-%Bowler2010 -- DOI:10.1088/0953-8984/22/7/074207
-%Nakata2015 -- DOI:10.1039/C5CP00934K
-%Krajewski -- PRB, 73, 041105R
-%Kolafa2003 -- J Comput Chem 25: 335–342, 2004
-%
-From this point of view, a direct variation of molecular orbitals that are strictly localized within predefined regions is preferable because LS can be achieved without costly DM optimization and with significantly fewer variables.
-%especially in the case of large accurate basis sets.
-Advantages of the orbitals-only DFT are especially pronounced in accurate calculations that require many basis functions per atom.
+From this point of view, a direct variation of molecular orbitals that are strictly localized within predefined regions is preferable because LS can be achieved with significantly fewer variables. 
+Advantages of the orbitals-only LS DFT are especially pronounced in accurate calculations that require many basis functions per atom.
 %From this point of view, it is desirable to obviate the DM optimization and use another set of local electronic descriptors can be constructed by enforcing strict locality on one-electron orbitals and, necessarily, by relaxing the orthogonality constraints imposed on them.
@@
 %From the computational point of view, LS methods based on the direct optimization of such compact localized nonorthogonal orbitals~\cite{a:ls-galli-parrinello-1992,a:ls-mauri-galli-car-1993,a:ls-ordejon-1993,a:ls-mauri-galli-1994,a:ls-ordejon-1995,a:ls-kim-mauri-galli-1995,a:ls-fattebert-2004,a:ls-fattebert-2006,a:burger-yang-2008} are promising since they deal with fewer variational degrees of freedom than the DM methods.
-Unfortunately, the development of promising orbital-based LS methods has been all but abandoned~\cite{a:review1999, a:weitao-yang-2013,japanese-guys} because of the inherently difficult optimization of localized orbitals~\cite{a:ls-mauri-galli-car-1993,a:ls-ordejon-1995,a:ls-fattebert-2004,a:ls-rev-1999,a:weitao-yang-2013,japanese-guys-citedinmy2013paper}.
+Unfortunately, the development of promising orbital-based LS methods has been all but abandoned~\cite{a:weitao-yang-2013,a:ls-tsuchida-aomm} because of the inherently difficult optimization of localized orbitals~\cite{a:ls-rev-1999,a:ls-mauri-galli-car-1993,a:ls-ordejon-1995,a:ls-fattebert-2004,a:weitao-yang-2013,a:ls-tsuchida-aomm}.
 
@@
 
-Thus, despite impressive progress of LS description of the electronic and atomic structure of large static systems~\cite{Bowler2010,a:ls-dm-sign} the high computational overhead of existing LS methods restrict their use in dynamical simulations to very short \emph{time scales}, systems of low dimensions, and low-quality minimal basis sets~\cite{a:ls-dm-sign, RZK:10.7567/JJAP.55.1102B1,Hine2011, Bowler2010}.
-%RZK: Although such methods have been successfully used for the evaluation of accurate DFT energies [electronic structure and atomic configurations] of large static systems~\cite{Bowler2010,a:ls-dm-sign} ...
-%Thus, the high computational overhead of existing LS DFT methods restrict their use in [dynamical] AIMD simulations to very short \emph{time scales}, systems of low dimensions, and low-quality minimal basis sets~\cite{a:ls-dm-sign, RZK:10.7567/JJAP.55.1102B1,Hine2011, Bowler2010}.
+Thus, despite impressive progress of LS description of the electronic and atomic structure of large static systems~\cite{Bowler2010,a:ls-dm-sign} the high computational overhead of existing LS methods restrict their use in dynamical simulations to very short \emph{time scales}, systems of low dimensions, and low-quality minimal basis sets~\cite{a:ls-dm-sign, Otsuka2016, Hine2011, Bowler2010}.
+%RZK: Although such methods have been successfully used for the evaluation of accurate DFT energies [electronic structure and atomic configurations] of large static systems~\cite{Bowler2010,a:ls-dm-sign}
 On typical length and time scales required in practical and accurate AIMD simulations, LS DFT still cannot compete with the straightforward low-cost cubically-scaling KS DFT.
@@
 To demonstrate advantages of the new method we applied it here to systems of weakly-interacting molecules. 
-However, the same approach is readily applicable to systems of strongly-interacting fragments that do not form strong covalent bonds such as ionic salts, liquids, and semiconductors. 
+However, the same approach is readily applicable to systems of strongly-interacting fragments that do not form strong covalent bonds such as ionic materials: salts, liquids, and semiconductors. 
 A generalization of the method to all finite-gap systems, including covalently bonded atoms, will be reported later. 
@@
 In the current implementation, a localization center is defined as a set of all Gaussian atomic orbitals of one molecule. 
-However the approach can be easily generalized to other local and nonlocal basis sets~\cite{a:ls-galli-parrinello-1992, RZK:10.1016/j.jcp.2011.11.032}. 
+However the approach can use other local and nonlocal basis sets~\cite{a:ls-galli-parrinello-1992, Lin2012}. 
 The key feature of ALMO DFT is that its one-electron wavefunctions are constructed in a two-stage self-consistent-field (SCF) procedure~\cite{a:almo-ls} to circumvent the problem of the sluggish variational optimization emphasized above.
@@
 The challenge of adopting ALMO DFT for dynamical simulations arises from the slightly nonvariational character of the localized orbitals.
-First, while ALMOs are variationally optimized in both SCF stages, the occupied subspace defined in the first stage must remain fixed during the second stage to ensure convergence.
-Second, neglected electron transfer effects can suddenly become active in the course of a dynamical simulation when a neighboring molecule passes the localization threshold $R_{c}$.
-In addition to these issues specific to ALMO DFT, the variational optimization in any AIMD method is never truly complete in practice
+While ALMOs are variationally optimized in both SCF stages, the occupied subspace defined in the first stage must remain fixed during the second stage to ensure convergence.
+In addition, electron transfer effects can suddenly become inactive in the course of a dynamical simulation when a neighboring molecule crosses the localization threshold $R_{c}$.
+Futhermore, the variational optimization in any AIMD method is never truly complete in practice
 %[exact convergence is never achieved in practical calculation] 
 and interrupted once the maximum norm of the gradient of the energy with respect to the electronic descriptors drops below small but nevertheless finite convergence threshold $\epsilon_{\text{SCF}}$.
-These nonvariational errors do not affect the accuracy of static ALMO DFT calculations, geometry optimization, and Monte-Carlo simulations.
+These errors do not affect the accuracy of static ALMO DFT calculations, geometry optimization, and Monte-Carlo simulations.
 Unfortunately they tend to accumulate in molecular dynamics trajectories leading to non-physical sampling and eventual failure.
-Traditional strategies to cope with these problems are computationally expensive and include computing the nonvariational contribution to the forces via a variational coupled-perturbed procedure~\cite{Kussmann2013}, increasing $R_c$, and decreasing $\epsilon_{\text{SCF}}$.
+Traditional strategies to cope with these problems are computationally expensive and include computing the nonvariational contribution to the forces via a variational coupled-perturbed procedure~\cite{Kussmann2013,Benoit2001}, increasing $R_c$, and decreasing $\epsilon_{\text{SCF}}$.
 %This however would dramatically increase the computational cost of the ALMO AIMD.
@@
 
-Our approach to compensate for the missing $\delta f_{i\alpha}(t)$ term is inspired by the methodology introduced into AIMD by Krajewski \emph{et al.}~\cite{Krajewski}, formalized by K\"uhne \emph{et al.}~\cite{a:2ndcpmd} and rationalized by Dai \emph{et al.}~\cite{a:langevin-why} before becoming informally known as the second generation Car-Parrinello molecular dynamics~\cite{RZK:10.1002/wcms.1176}.
+Our approach to compensate for the missing $\delta f_{i\alpha}(t)$ term is inspired by the methodology introduced into AIMD by Krajewski \emph{et al.}~\cite{Krajewski}, formalized by K\"uhne \emph{et al.}~\cite{a:2ndcpmd} and rationalized by Dai \emph{et al.}~\cite{a:langevin-why} before becoming informally known as the second generation Car-Parrinello molecular dynamics~\cite{Kuhne2013}.
 Adopting the principle of Refs.~\onlinecite{Krajewski} and \onlinecite{a:2ndcpmd}, ALMO AIMD is chosen to be governed by the Langevin equation of motion that can be written in terms of the unknown reference forces:
@@
 
-Thus, the inherently stochastic approach presented here does not aim to produce fully time-reversible dynamics for atomic nuclei. 
-However, with $\gamma=0$, the nuclei will follow time-reversible Newtonian trajectories sufficiently closely as long as partially optimized ALMOs remain close to the ground state and keep $\Delta$ small.
-%
-%(for systems with slight discontinuities in the potential energy surfaces)
-%
-Although not a required component of ALMO AIMD, the predictor part of the Kolafa scheme~\cite{kolafa2003} adopted to localized orbitals was found to generate a highly accurate initial ALMOs in both SCF stages, which can be brought close to the ground state with just a few SCF steps.
-
-%RZK: The time-reversibility of ASPC is not a critical reqirement of the proposed method.
+The inherently stochastic approach presented here does not aim to produce fully time-reversible dynamics for atomic nuclei. 
+Nevertheless, it is capable to reproduce correct dynamical properties of a system as long as $\gamma$ is set to a small value and  partially optimized ALMOs remain close to the ground state resulting in $\Delta \ll \gamma$.
+%With $\gamma$ set to zero, the nuclei will follow time-reversible Newtonian trajectories closely enough to reproduce dynamical properties of the ensemble as long as partially optimized ALMOs remain close to the ground state and $\Delta$ is small.
 
@@
 
-ALMO AIMD presented here was implemented in CP2K, an open source materials modeling package~\cite{www:cp2k}. 
+ALMO AIMD was implemented in CP2K, an open source materials modeling package~\cite{www:cp2k}. 
 Accuracy and efficiency of ALMO AIMD was tested using liquid water as an example. 
@@
 A periodic cell containing 125 molecules was simulated for $\SI{10}{\ps}$ at $T=\SI{298}{\K}$ and a constant density of $\SI{1.01}{\g\cdot\cm^{-3}}$. 
-Ricci-Ciccotti algorithm~\cite{RZK:10.1080/0026897031000108113} was used to integrate the Langevin equation.
+Ricci-Ciccotti algorithm~\cite{Ricci2003} was used to integrate the Langevin equation.
 We found that $\gamma = \SI{e-3}{\per\fs}$ is large enough to thermostat the system efficiently and small enough not to significantly affect dynamical properties of liquid water.
@@
 %to describe the interactions between the valence electrons and the ionic cores. 
+%
+%Although not a required component of ALMO AIMD, 
+The predictor of the Kolafa scheme~\cite{Kolafa2003} was adopted to localized orbitals~\cite{a:2ndcpmd} to generate a highly accurate initial ALMOs in both SCF stages, which can be brought close to the ground state with just a few SCF steps of the robust two-stage optimization procedure.
+%RZK: The time-reversibility of ASPC is not a critical reqirement of the proposed method.
 
@@
 and $R_{c} = 1.6$ vdWR) and from the reference trajectory.
-The diffusion coefficients $D_{\text{OT}}=\SI{5.47e-9}{\m^{2}\per\s}$ and $D_{\text{ALMO}}=\SI{5.55E-09}{\m^{2}\per\s}$ and RDFs (Figure~\ref{fig:dynproperties}a) are in perfect agreement.
+The diffusion coefficients $D_{\text{OT}}=\SI{5.47e-9}{\m^{2}\cdot\s^{-1}}$ and $D_{\text{ALMO}}=\SI{5.55E-09}{\m^{2}\cdot\s^{-1}}$ and RDFs (Figure~\ref{fig:dynproperties}a) are in perfect agreement.
 The quality of the ALMO IR spectrum (Figure~\ref{fig:dynproperties}c) is good despite minor errors in the intensity of the OH stretching mode, which is sensitive to the precise positions of the centers of localized orbitals.
@@
 
-%RZK: time-reversibility must be mentioned or even addressed
-
 %RZK longer article: fully converged ALMO calculations must be reported: Delta=0, gamma=0, is total energy constant at R=1.6 vdW?
@@
 
-%RZK: number of atoms is 98,304 with very accurate basis set.
-
 Weak scaling benchmarks for very large systems show (Figure~\ref{fig:weakscaling}) that localized orbitals are naturally suited for parallel execution: LS is retained for a wide range of systems and compute cores.
@@
 %(32768 water molecules) 
-within reasonable wall-clock time using only moderate number of compute cores -- an unprecedented feat for AIMD considering that accurate molecular orbitals and the idempotent DM are computed on each step.
+within reasonable wall-clock time using only moderate number of compute cores -- an impressive feat for AIMD considering that accurate molecular orbitals and the idempotent DM are computed on each step.
 The horizontal line in Figure~\ref{fig:weakscaling} is shown as a rough guide to time and length scales accessible in a fixed wall-clock time given various computational resources.
@@
 
-\subsection{Analytical expression for $\Delta$}
+%\subsection{Analytical expression for $\Delta$}
 %
@@
 
-Among several possible expression for $\Delta$ we consider the following two options.
-
-\textbf{Option 1.} $\Delta$ is calculated as a weighted average of atom TYPES (e.g. oxygen and hydrogen). Subindex $\text{T}$ refers to atom type.
-
-\textbf{\emph{Step 1.}} Compute the average ACF for each atom type.
+%%% %% This was a summary of the two options for Delta
+%%%Among several possible expression for $\Delta$ we consider the following two options.
+%%%
+%%%\textbf{Option 1.} $\Delta$ is calculated as a weighted average of atom TYPES (e.g. oxygen and hydrogen). Subindex $\text{T}$ refers to atom type.
+%%%
+%%%\textbf{\emph{Step 1.}} Compute the average ACF for each atom type.
 In this step averaging is done over all time origins $t$, over all atoms of THIS type $i$, and over all dimensions $\alpha$.
 The ACF includes the mass-dependent factor $(2 k_B T m_\text{T} )^{-1}$ where $T$ is the requested temperature.
-%
-\begin{align}
-\label{eq:deltaS1}
-\text{ACF}_{\text{T}}(\tau) = (2 k_B T m_\text{T} )^{-1} \, \langle \delta f_{i\alpha} (t) \, \delta f_{i\alpha}(t + \tau) \rangle_{t i\alpha}
-\end{align}
-%
-\textbf{\emph{Step 2.}} Integrate the ACF for each atom type.
-%
-\begin{align}
-\label{eq:deltaS2}
-\Delta_{\text{T}} = \int_{-\infty}^{\infty} \text{ACF}_{T}(\tau) \, d\tau
-\end{align}
-%
-\textbf{\emph{Step 3.}} Average all over all atom types.
-%
-\begin{align}
-\label{eq:deltaS3}
-\Delta = \frac{1}{N} \sum_{\text{T}} N_{\text{T}} \Delta_{\text{T}}
-\end{align}
-%
-where $N$ is the number of atoms.
+%%%%
+%%%\begin{align}
+%%%\label{eq:deltaS1}
+%%%\text{ACF}_{\text{T}}(\tau) = (2 k_B T m_\text{T} )^{-1} \, \langle \delta f_{i\alpha} (t) \, \delta f_{i\alpha}(t + \tau) \rangle_{t i\alpha}
+%%%\end{align}
+%%%%
+%%%\textbf{\emph{Step 2.}} Integrate the ACF for each atom type.
+%%%%
+%%%\begin{align}
+%%%\label{eq:deltaS2}
+%%%\Delta_{\text{T}} = \int_{-\infty}^{\infty} \text{ACF}_{T}(\tau) \, d\tau
+%%%\end{align}
+%%%%
+%%%\textbf{\emph{Step 3.}} Average all over all atom types.
+%%%%
+%%%\begin{align}
+%%%\label{eq:deltaS3}
+%%%\Delta = \frac{1}{N} \sum_{\text{T}} N_{\text{T}} \Delta_{\text{T}}
+%%%\end{align}
+%%%%
+%%%where $N$ is the number of atoms.
+%%%
+%%%If all these step are combined the following expression for $\Delta$ is obtained:
+%%%%
+%%%\begin{align}
+%%%\label{eq:deltaTOT}
+%%%\Delta = \int \, d\tau \left[ \frac{1}{3N} \sum_{i=1}^{N} \sum_{\alpha=1}^{3} \frac{ \langle \delta f_{i\alpha} (t) \, \delta f_{i\alpha}(t + \tau) \rangle_{t} }{2 k_B T m_i} \right]
+%%%\end{align}
+%%%%
+%%%Thus the weight of each of the $3N$ degrees of freedom in this scheme is proportional to the inverse atomic mass.
 
-If all these step are combined the following expression for $\Delta$ is obtained:
-%
-\begin{align}
-\label{eq:deltaTOT}
-\Delta = \int \, d\tau \left[ \frac{1}{3N} \sum_{i=1}^{N} \sum_{\alpha=1}^{3} \frac{ \langle \delta f_{i\alpha} (t) \, \delta f_{i\alpha}(t + \tau) \rangle_{t} }{2 k_B T m_i} \right]
-\end{align}
-%
-Thus the weight of each of the $3N$ degrees of freedom in this scheme is proportional to the inverse atomic mass.
-
-\textbf{Option 2.} If the same averaging scheme is applied to MOLECULE types (not ATOM types) then the value of $\Delta$ is given by:
-%
-\begin{align}
-\label{eq:O2deltaTOT}
-\Delta = \int \, d\tau \left[ \frac{1}{3M} \sum_{p=1}^{M} \sum_{\alpha=1}^{3} \frac{ \sum_i^{N_p}\sum_j^{N_p} \langle \delta f_{i\alpha} (t) \, \delta f_{j\alpha}(t + \tau) \rangle_{t} }{2 k_B T \sum_k^{N_p}m_k} \right]
-\end{align}
-%
-where $M$ is the number of molecules, $N_p$ is the number of atoms in molecule $p$.
+%%%\textbf{Option 2.} If the same averaging scheme is applied to MOLECULE types (not ATOM types) then the value of $\Delta$ is given by:
+%%%%
+%%%\begin{align}
+%%%\label{eq:O2deltaTOT}
+%%%\Delta = \int \, d\tau \left[ \frac{1}{3M} \sum_{p=1}^{M} \sum_{\alpha=1}^{3} \frac{ \sum_i^{N_p}\sum_j^{N_p} \langle \delta f_{i\alpha} (t) \, \delta f_{j\alpha}(t + \tau) \rangle_{t} }{2 k_B T \sum_k^{N_p}m_k} \right]
+%%%\end{align}
+%%%%
+%%%where $M$ is the number of molecules, $N_p$ is the number of atoms in molecule $p$.
 The value of this $\Delta$ might be different from Option 1 because (a) all correlations between two atoms inside one molecule are taken into account and (b) the weight of each degree of freedom is different.
 The difference between Option 1 and 2 can be clearly seen for a special case of identical molecules:
-%
-\begin{align}
-\label{eq:O2deltaSPECIAL}
-\Delta = \int \, d\tau \left[ \frac{1}{3N} \sum_{i=1}^{N} \sum_{\alpha=1}^{3} \frac{ \sum_j^{N_p} \langle \delta f_{i\alpha} (t) \, \delta f_{j\alpha}(t + \tau) \rangle_{t} }{2 k_B T \langle m \rangle } \right]
-\end{align}
-%
-where $\langle m \rangle = (\sum_k^{N_p} m_k )/N_p$ is the average atom mass in a molecule.
+%%%%
+%%%\begin{align}
+%%%\label{eq:O2deltaSPECIAL}
+%%%\Delta = \int \, d\tau \left[ \frac{1}{3N} \sum_{i=1}^{N} \sum_{\alpha=1}^{3} \frac{ \sum_j^{N_p} \langle \delta f_{i\alpha} (t) \, \delta f_{j\alpha}(t + \tau) \rangle_{t} }{2 k_B T \langle m \rangle } \right]
+%%%\end{align}
+%%%%
+%%%where $\langle m \rangle = (\sum_k^{N_p} m_k )/N_p$ is the average atom mass in a molecule.
+
+%$\Delta$ is calculated by averaging over molecule types:
+%%
+%\begin{align}
+%\label{eq:O2deltaTOT}
+%\Delta = \int \, d\tau \left[ \frac{1}{3M} \sum_{p=1}^{M} \sum_{\alpha=1}^{3} \frac{ \sum_i^{N_p}\sum_j^{N_p} \langle \delta f_{i\alpha} (t) \, \delta f_{j\alpha}(t + \tau) \rangle_{t} }{2 k_B T \sum_k^{N_p}m_k} \right]
+%\end{align}
+%%
+%where $M$ is the number of molecules, $N_p$ is the number of atoms in molecule $p$.
+The value of this $\Delta$ might be different from Option 1 because (a) all correlations between two atoms inside one molecule are taken into account and (b) the weight of each degree of freedom is different.
+The difference between Option 1 and 2 can be clearly seen for a special case of identical molecules:
+%%
+%\begin{align}
+%\label{eq:O2deltaSPECIAL}
+%\Delta = \int \, d\tau \left[ \frac{1}{3N} \sum_{i=1}^{N} \sum_{\alpha=1}^{3} \frac{ \sum_j^{N_p} \langle \delta f_{i\alpha} (t) \, \delta f_{j\alpha}(t + \tau) \rangle_{t} }{2 k_B T \langle m \rangle } \right]
+%\end{align}
+%%
+%where $\langle m \rangle = (\sum_k^{N_p} m_k )/N_p$ is the average atom mass in a molecule.
 
@@
 
-%\begin{figure}[!htb]
-%\includegraphics[trim={2.5cm 0.2cm 4.5cm 1.2cm},clip,width=\textwidth]{Diffusion_Comparison.eps}
-%\caption{\label{fig:diffusion} Calculated mean-squared displacement function of water using ALMO DFT with $\epsilon_{\text{SCF}} = 10^{-2}$~a.u. and $R_{c} = 1.6$ vdWR.
-%Compared with equivalent MD run using OT DFT as a reference.
-%Calculations were performed at 298 K with the PBE $\mathrm{E_{XC}}$ functional and TZV2P basis set, using 125 water molecules for 10 ps.}
-%\end{figure}
+\begin{figure}[!htb]
+\includegraphics[trim={0cm 0cm 0.1cm 0.1cm},clip,width=8.6cm]{Diffusion_Comparison.eps}
+\caption{\label{fig:diffusion} Calculated mean-squared displacement function of water using ALMO DFT with $\epsilon_{\text{SCF}} = 10^{-2}$~a.u. and $R_{c} = 1.6$ vdWR.
+Compared with equivalent MD run using OT DFT as a reference.
+Calculations were performed at 298 K with the dispersion-corrected PBE $\mathrm{E_{XC}}$ functional and TZV2P basis set, using 125 water molecules for 10 ps.}
+\end{figure}
 

### 2017-11-21 almo-langevin.tex
Dispersion-corrected XC
Updated MSD and diffusion constants.

@@
 and $R_{c} = 1.6$ vdWR) and from the reference trajectory.
-The diffusion coefficients $D_{\text{OT}}=\SI{5.47e-9}{\m^{2}\cdot\s^{-1}}$ and $D_{\text{ALMO}}=\SI{5.55E-09}{\m^{2}\cdot\s^{-1}}$ and RDFs (Figure~\ref{fig:dynproperties}a) are in perfect agreement.
+The diffusion coefficients $D_{\text{OT}}=\SI{4.61E-10}{\m^{2}\cdot\s^{-1}}$ and $D_{\text{ALMO}}=\SI{3.19E-10}{\m^{2}\cdot\s^{-1}}$ and RDFs (Figure~\ref{fig:dynproperties}a) are in good agreement.
 The quality of the ALMO IR spectrum (Figure~\ref{fig:dynproperties}c) is good despite minor errors in the intensity of the OH stretching mode, which is sensitive to the precise positions of the centers of localized orbitals.

### 2018-03-05 almo-langevin.tex
missing references
diff coeff values updated

@@
+\documentclass[aps,prl,reprint,amsmath,amssymb]{revtex4-1}
 %\documentclass[journal=jacsat]{achemso}
-
-\documentclass[aps,prl,reprint,amsmath,amssymb]{revtex4-1}
 
@@
 
-%RZK: one order of magnitude error in the PBE+D3 diffusion constant
-
 \begin{abstract}
-Within the framework of linear-scaling Kohn-Sham density functional theory, a robust method for maintaining compact localized orbitals close to the ground state is coupled with nuclear dynamics to obviate the commonly employed optimization of the one-electron density matrix and to create an efficient \emph{ab initio} molecular dynamics method for weakly-interacting systems.
-An application to liquid water demonstrates that the low computational overhead of the developed method makes it well-suited for routine simulations whereas its linear scaling complexity allows to extend first-principle dynamical studies of molecular systems to previously inaccessible time and length scales.
+Within the framework of linear-scaling Kohn-Sham density functional theory, a robust method for maintaining compact orbitals close to the ground state is coupled with nuclear dynamics to obviate the commonly employed optimization of the one-electron density matrix and to create an efficient orbital-only molecular dynamics method for weakly-interacting systems.
+An application to liquid water demonstrates that the low computational overhead of the method makes it well-suited for routine simulations whereas its linear scaling complexity allows to extend first-principle dynamical studies of molecular systems to previously inaccessible time and length scales.
 %Within the framework of linear-scaling Kohn-Sham density functional theory, a robust method for maintaining compact localized orbitals close to the ground state coupled with nuclear dynamics enables us to create low-cost linear-scaling \emph{ab initio} molecular dynamics for weakly-interacting systems.
@@
 To demonstrate advantages of the new method we applied it here to systems of weakly-interacting molecules. 
-However, the same approach is readily applicable to systems of strongly-interacting fragments that do not form strong covalent bonds such as ionic materials: salts, liquids, and semiconductors. 
+However, the same approach is readily applicable to systems of strongly-interacting fragments that do not form strong covalent bonds such as ionic materials---salts, liquids, and semiconductors. 
 A generalization of the method to all finite-gap systems, including covalently bonded atoms, will be reported later. 
@@
 %To calculate properties of water at ambient conditions, 
-A periodic cell containing 125 molecules was simulated for $\SI{10}{\ps}$ at $T=\SI{298}{\K}$ and a constant density of $\SI{1.01}{\g\cdot\cm^{-3}}$. 
+A periodic cell containing 125 molecules was simulated for $\SI{30}{\ps}$ at $T=\SI{298}{\K}$ and a constant density of $\SI{1.01}{\g\cdot\cm^{-3}}$. 
 Ricci-Ciccotti algorithm~\cite{Ricci2003} was used to integrate the Langevin equation.
@@
 In the dual Gaussian and plane-wave scheme implemented in CP2K~\cite{a:quickstep}, the TZV2P basis set was used to represent molecular orbitals, and a plane-wave cutoff of $\SI{320}{Ry}$ used to represent electron density.
-The XC energy was approximated using the PBE functional~\cite{a:PBEfunctional}. 
+The XC energy was approximated using the dispersion-corrected PBE functional~\cite{a:PBEfunctional,Grimme2010}. 
 Separable norm-conserving pseudopotentials were used~\cite{a:hgh} and the Brillouin zone was sampled at the $\Gamma$-point. 
@@
 and $R_{c} = 1.6$ vdWR) and from the reference trajectory.
-The diffusion coefficients $D_{\text{OT}}=\SI{4.61E-10}{\m^{2}\cdot\s^{-1}}$ and $D_{\text{ALMO}}=\SI{3.19E-10}{\m^{2}\cdot\s^{-1}}$ and RDFs (Figure~\ref{fig:dynproperties}a) are in good agreement.
+The diffusion coefficients $D_{\text{OT}}=\SI{1.9\pm0.9E-10}{\m^{2}\cdot\s^{-1}}$ and $D_{\text{ALMO}}=\SI{1.8\pm0.4E-10}{\m^{2}\cdot\s^{-1}}$ and RDFs (Figure~\ref{fig:dynproperties}a) are in good agreement.
 The quality of the ALMO IR spectrum (Figure~\ref{fig:dynproperties}c) is good despite minor errors in the intensity of the OH stretching mode, which is sensitive to the precise positions of the centers of localized orbitals.

### 2018-03-05 almo-langevin.tex
Edited References
Merged some identical references; Generated abbreviated journal names; removed links; etc. References imported from Mendeley, should be in PRL format.

@@
 In the current implementation, a localization center is defined as a set of all Gaussian atomic orbitals of one molecule. 
-However the approach can use other local and nonlocal basis sets~\cite{a:ls-galli-parrinello-1992, Lin2012}. 
+However the approach can use other local and nonlocal basis sets~\cite{a:galli_loc, Lin2012}. 
 The key feature of ALMO DFT is that its one-electron wavefunctions are constructed in a two-stage self-consistent-field (SCF) procedure~\cite{a:almo-ls} to circumvent the problem of the sluggish variational optimization emphasized above.
@@
 In the dual Gaussian and plane-wave scheme implemented in CP2K~\cite{a:quickstep}, the TZV2P basis set was used to represent molecular orbitals, and a plane-wave cutoff of $\SI{320}{Ry}$ used to represent electron density.
-The XC energy was approximated using the dispersion-corrected PBE functional~\cite{a:PBEfunctional,Grimme2010}. 
+The XC energy was approximated using the dispersion-corrected PBE functional~\cite{a:pbe,Grimme2010}. 
 Separable norm-conserving pseudopotentials were used~\cite{a:hgh} and the Brillouin zone was sampled at the $\Gamma$-point. 

### 2018-03-06 almo-langevin.tex
Small edits
Went through and edited, made small edits in grammar.

@@
 %While most large-scale DFT calculations are performed using this class of methods
-Despite becoming the most popular approach to LS DFT the efficiency of these methods is hampered by the costly optimization of both the contracted orbitals and the DM~\cite{a:ls-onetep-2003}. 
+Despite becoming the most popular approach to LS DFT, the efficiency of these methods is hampered by the costly optimization of both the contracted orbitals and the DM~\cite{a:ls-onetep-2003}. 
 %
@@
 
-Thus, despite impressive progress of LS description of the electronic and atomic structure of large static systems~\cite{Bowler2010,a:ls-dm-sign} the high computational overhead of existing LS methods restrict their use in dynamical simulations to very short \emph{time scales}, systems of low dimensions, and low-quality minimal basis sets~\cite{a:ls-dm-sign, Otsuka2016, Hine2011, Bowler2010}.
+Thus, despite impressive progress of the LS description of the electronic and atomic structure of large static systems~\cite{Bowler2010,a:ls-dm-sign}, the high computational overhead of existing LS methods restrict their use in dynamical simulations to very short \emph{time scales}, systems of low dimensions, and low-quality minimal basis sets~\cite{a:ls-dm-sign, Otsuka2016, Hine2011, Bowler2010}.
 %RZK: Although such methods have been successfully used for the evaluation of accurate DFT energies [electronic structure and atomic configurations] of large static systems~\cite{Bowler2010,a:ls-dm-sign}
@@
 In the current implementation, a localization center is defined as a set of all Gaussian atomic orbitals of one molecule. 
-However the approach can use other local and nonlocal basis sets~\cite{a:galli_loc, Lin2012}. 
+However, the approach can use other local and nonlocal basis sets~\cite{a:galli_loc, Lin2012}. 
 The key feature of ALMO DFT is that its one-electron wavefunctions are constructed in a two-stage self-consistent-field (SCF) procedure~\cite{a:almo-ls} to circumvent the problem of the sluggish variational optimization emphasized above.
@@
 To achieve a robust optimization in the problematic second stage, it is important to keep the delocalization component of the trial wavefunction orthogonal to the fixed orbitals obtained in the first stage.
-For mathematical details, see ALMO SCF method in Ref.~\citenum{a:almo-ls}.
+For mathematical details, see the ALMO SCF method in Ref.~\citenum{a:almo-ls}.
 
-ALMO constraints imposed by $R_c$ prohibit electron density transfer between distant molecules but retain all other types of interaction such as long-range electrostatic, exchange, polarization, and, if the exchange-correlation (XC) functional includes them, dispersion interactions~\cite{a:theeda}.
+ALMO constraints imposed by $R_c$ prohibit electron density transfer between distant molecules, but retain all other types of interaction such as long-range electrostatic, exchange, polarization, and---if the exchange-correlation (XC) functional includes them---dispersion interactions~\cite{a:theeda}.
 Since the importance of electron transfer decays exponentially with distance in finite-gap materials~\cite{a:ls-rev-1999}, the \mbox{ALMO} approximation is expected to provide a natural and accurate representation of the electronic structure of molecular systems.
@@
 Our approach to compensate for the missing $\delta f_{i\alpha}(t)$ term is inspired by the methodology introduced into AIMD by Krajewski \emph{et al.}~\cite{Krajewski}, formalized by K\"uhne \emph{et al.}~\cite{a:2ndcpmd} and rationalized by Dai \emph{et al.}~\cite{a:langevin-why} before becoming informally known as the second generation Car-Parrinello molecular dynamics~\cite{Kuhne2013}.
-Adopting the principle of Refs.~\onlinecite{Krajewski} and \onlinecite{a:2ndcpmd}, ALMO AIMD is chosen to be governed by the Langevin equation of motion that can be written in terms of the unknown reference forces:
+Adopting the principle of Refs.~\onlinecite{Krajewski} and \onlinecite{a:2ndcpmd}, ALMO AIMD is chosen to be governed by the Langevin equation of motion that can be written in terms of the unknown reference forces
 %
@@
 %
-where $m_i$ is the mass of atom $i$, $r_{i\alpha}$ is its position along dimension $\alpha$, $\gamma$ is the Langevin scaling factor, and $R^{\gamma}_{i\alpha} (t)$ is the stochastic force represented by a zero-mean white Gaussian noise: %with the zero mean and the strength related to the damping term through $\gamma$:
+where $m_i$ is the mass of atom $i$, $r_{i\alpha}$ is its position along dimension $\alpha$, $\gamma$ is the Langevin scaling factor, and $R^{\gamma}_{i\alpha} (t)$ is the stochastic force represented by a zero-mean white Gaussian noise %with the zero mean and the strength related to the damping term through $\gamma$:
 %
@@
 \label{eq:stochastic3}
-\langle R^{\Delta}_{i\alpha} (t)  R^{\Delta}_{j\beta} (t') \rangle &= 2 k_B T \Delta m_i \delta_{ij} \delta_{\alpha\beta} \delta(t-t')
+\langle R^{\Delta}_{i\alpha} (t)  R^{\Delta}_{j\beta} (t') \rangle &= 2 k_B T \Delta m_i \delta_{ij} \delta_{\alpha\beta} \delta(t-t') .
 \end{align}
 %
-This assumption, shown to be well justified, allows us to rewrite the Langevin equation using the ALMO forces:
+This assumption, shown to be well justified, allows us to rewrite the Langevin equation using the ALMO forces
 %
@@
 \label{eq:langevin2}
-m_i \ddot{r}_{i\alpha} = f^{\text{ALMO}}_{i\alpha}(t) - \gamma m_i \dot{r}_{i\alpha} + R^{\gamma + \Delta}_{i\alpha} (t)
+m_i \ddot{r}_{i\alpha} = f^{\text{ALMO}}_{i\alpha}(t) - \gamma m_i \dot{r}_{i\alpha} + R^{\gamma + \Delta}_{i\alpha} (t),
 \end{align}
@@
 (a) RDF, 
-(b) kinetic energy distribution (the black curve shows the theoretical Maxwell-Boltzmann distribution), 
+(b) kinetic energy distribution (the gray curve shows the theoretical Maxwell-Boltzmann distribution), 
 (c) IR spectrum.
@@
 
-The \emph{reference} forces were calculated with fully delocalized electrons using tightly converged, $\epsilon_{\text{SCF}}=10^{-6}$~a.u., orbital transformation (OT) method~\cite{a:ot}. 
-In ALMO AIMD, element-specific cutoff radius for electron delocalization $R_c$ was set to 1.6 in units of the elements' van der Waals radii (vdWR).
+The \emph{reference} forces were calculated with fully delocalized electrons using the tightly converged, $\epsilon_{\text{SCF}}=10^{-6}$~a.u., orbital transformation (OT) method~\cite{a:ot}. 
+In ALMO AIMD, the element-specific cutoff radius for electron delocalization $R_c$ was set to 1.6 in units of the elements' van der Waals radii (vdWR).
 This localization radius includes approximately two coordination shells of an average water molecule and was shown to reproduce the reference radial distribution function (RDF) perfectly in Monte-Carlo simulations~\cite{a:almo-ls}.
 %
-To check the ability of the $R^{\Delta}(t)$ term to compensate for imperfections in ALMO forces we varied $\epsilon_{\text{SCF}}$ between tight $10^{-6}$~a.u. and loose $10^{-2}$~a.u.
+To check the ability of the $R^{\Delta}(t)$ term to compensate for imperfections in ALMO forces, we varied $\epsilon_{\text{SCF}}$ between tight $10^{-6}$~a.u. and loose $10^{-2}$~a.u.
 
@@
 
-To test the accuracy of ALMO AIMD we used trajectory analyzer TRAVIS~\cite{a:travis-main} to compute the infrared (IR) spectrum, RDF, and diffusion coefficient of liquid water from both the ALMO trajectory ($\epsilon_{\text{SCF}} = 10^{-2}$~a.u.
+To test the accuracy of ALMO AIMD we used the trajectory analyzer TRAVIS~\cite{a:travis-main} to compute the infrared (IR) spectrum, RDF, and diffusion coefficient of liquid water from both the ALMO trajectory ($\epsilon_{\text{SCF}} = 10^{-2}$~a.u.
 and $R_{c} = 1.6$ vdWR) and from the reference trajectory.
@@
 The quality of the ALMO IR spectrum (Figure~\ref{fig:dynproperties}c) is good despite minor errors in the intensity of the OH stretching mode, which is sensitive to the precise positions of the centers of localized orbitals.
-These stringent tests show that despite noticeable errors in the ALMO forces (Figure \ref{fig:randomforce}a) the compensating $R^{\Delta}(t)$ term in the modified Langevin equation makes it possible to recover atomic dynamics properly.
+These stringent tests show that despite noticeable errors in the ALMO forces (Figure \ref{fig:randomforce}a), the compensating $R^{\Delta}(t)$ term in the modified Langevin equation makes it possible to recover atomic dynamics properly.
 We would like to note that ALMO AIMD could not be stabilized with $\Delta=0$.
-Neither we were able to find any values of $\Delta$ that stabilize trajectories generated using perturbative versions of ALMO DFT~\cite{a:almo-ls}.
+Neither were we able to find any values of $\Delta$ that stabilize trajectories generated using perturbative versions of ALMO DFT~\cite{a:almo-ls}.
 
@@
 % RZK longer article: averaging is done over 10 time steps 
-To demonstrate computational efficiency of ALMO AIMD, we compared the average wall-time per MD step for a variety of methods in Figure \ref{fig:strongscaling_log}.
+To demonstrate the computational efficiency of ALMO AIMD, we compared the average wall-time per MD step for a variety of methods in Figure \ref{fig:strongscaling_log}.
 %In addition to the fully converged OT reference, we included timing for the second generation Car-Parrinello method designed to decrease the computational prefactor of the cubically-scaling DFT~\cite{a:2ndcpmd}.
-It is important to emphasize that the comparison is performed for a three-dimensional condensed phase system described with an accurate triple-$\zeta$ basis set with polarized functions -- a particularly challenging case for DM-based LS methods.
-ALMO AIMD shows clear LS behavior for all values of $\epsilon_{\text{SCF}}$ even for medium-size systems.
+It is important to emphasize that the comparison is performed for a three-dimensional condensed phase system described with an accurate triple-$\zeta$ basis set with polarized functions---a particularly challenging case for DM-based LS methods.
+ALMO AIMD shows clear LS behavior for all values of $\epsilon_{\text{SCF}}$, even for medium-size systems.
 While the second generation Car-Parrinello method decreases the computational overhead of the cubically-scaling AIMD for small systems~\cite{a:2ndcpmd}, ALMO AIMD exploits the modified Langevin concept to substantially reduce the simulation cost for systems of all sizes.
-The crossover point between ALMO AIMD and cubically scaling methods lies in the region of 256 molecules -- length scales routinely accessible with AIMD today.
+The crossover point between ALMO AIMD and cubically scaling methods lies in the region of 256 molecules---length scales routinely accessible with AIMD today.
 
@@
 %(32768 water molecules) 
-within reasonable wall-clock time using only moderate number of compute cores -- an impressive feat for AIMD considering that accurate molecular orbitals and the idempotent DM are computed on each step.
+within reasonable wall-clock time using only moderate number of compute cores---an impressive feat for AIMD considering that accurate molecular orbitals and the idempotent DM are computed on each step.
 The horizontal line in Figure~\ref{fig:weakscaling} is shown as a rough guide to time and length scales accessible in a fixed wall-clock time given various computational resources.
@@
 % systematically improvable approximation to DFT-based AIMD.
-To summarize, we demonstrated, for the first time, that compact localized orbitals can be utilized to perform accurate and efficient LS AIMD without concomitant optimization of the DM.
+To summarize, we demonstrated---for the first time---that compact localized orbitals can be utilized to perform accurate and efficient LS AIMD without concomitant optimization of the DM.
 %for weakly-interacting systems.
 High efficiency of the presented method is achieved without sacrificing accuracy with a combination of two techniques: (1) on-the-fly calculation of approximate forces without lengthy self-consistent optimization of localized orbitals and (2) integration of a modified Langevin equation of motion that is fine-tuned to retain stable dynamics even with imperfect forces.
-By obviating the optimization of the DM the method remains remarkably efficient even with large localized basis sets. 
+By obviating the optimization of the DM, the method remains remarkably efficient even with large localized basis sets. 
 %A remarkable feature of the method is that it remains efficient for challenging condensed phase systems even with large localized basis sets. 
 Using liquid water as an example, we showed that the new approach enables simulations of molecular systems on previously inaccessible time and length scales. 
-The developed method will have a significant impact on modeling of complex molecular systems (e.g. interfaces, nuclei) making completely new phenomena accessible to AIMD. 
+The developed method will have a significant impact on modeling of complex molecular systems (e.g. interfaces or nuclei) making completely new phenomena accessible to AIMD. 
 Generalization of the methodology to systems of strongly interacting atoms (e.g. covalent crystals) is underway.

### 2018-03-06 almo-langevin.tex
Changed figures
Used best data to update figures.
Edited Diff. Constant to match.

@@
 and $R_{c} = 1.6$ vdWR) and from the reference trajectory.
-The diffusion coefficients $D_{\text{OT}}=\SI{1.9\pm0.9E-10}{\m^{2}\cdot\s^{-1}}$ and $D_{\text{ALMO}}=\SI{1.8\pm0.4E-10}{\m^{2}\cdot\s^{-1}}$ and RDFs (Figure~\ref{fig:dynproperties}a) are in good agreement.
+The diffusion coefficients $D_{\text{OT}}=\SI{1.7\pm0.1E-10}{\m^{2}\cdot\s^{-1}}$ and $D_{\text{ALMO}}=\SI{1.8\pm0.4E-10}{\m^{2}\cdot\s^{-1}}$ and RDFs (Figure~\ref{fig:dynproperties}a) are in good agreement.
 The quality of the ALMO IR spectrum (Figure~\ref{fig:dynproperties}c) is good despite minor errors in the intensity of the OH stretching mode, which is sensitive to the precise positions of the centers of localized orbitals.
@@
 \begin{figure}[!htb]
-\includegraphics[trim={0cm 0cm 0.1cm 0.1cm},clip,width=8.6cm]{Diffusion_Comparison.eps}
+\includegraphics[trim={2.4cm 0cm 3.0cm 1cm},clip,width=8.6cm]{Diffusion_Comparison.eps}
 \caption{\label{fig:diffusion} Calculated mean-squared displacement function of water using ALMO DFT with $\epsilon_{\text{SCF}} = 10^{-2}$~a.u. and $R_{c} = 1.6$ vdWR.

### 2018-03-06 almo-langevin.tex
title

--- 2018-03-05 almo-langevin.tex
@@
 \fi
-Low-cost linear-scaling \emph{ab initio} molecular dynamics for weakly-interacting systems
-%RZK: without one-electron density matrix
+Low-cost orbital-based linear-scaling \emph{ab initio} molecular dynamics for weakly-interacting systems
 }
@@
 In the current implementation, a localization center is defined as a set of all Gaussian atomic orbitals of one molecule. 
-However the approach can use other local and nonlocal basis sets~\cite{a:galli_loc, Lin2012}. 
+However the approach can use other local and nonlocal basis sets~\cite{a:ls-galli-parrinello-1992, Lin2012}. 
 The key feature of ALMO DFT is that its one-electron wavefunctions are constructed in a two-stage self-consistent-field (SCF) procedure~\cite{a:almo-ls} to circumvent the problem of the sluggish variational optimization emphasized above.
@@
 In the dual Gaussian and plane-wave scheme implemented in CP2K~\cite{a:quickstep}, the TZV2P basis set was used to represent molecular orbitals, and a plane-wave cutoff of $\SI{320}{Ry}$ used to represent electron density.
-The XC energy was approximated using the dispersion-corrected PBE functional~\cite{a:pbe,Grimme2010}. 
+The XC energy was approximated using the dispersion-corrected PBE functional~\cite{a:PBEfunctional,Grimme2010}. 
 Separable norm-conserving pseudopotentials were used~\cite{a:hgh} and the Brillouin zone was sampled at the $\Gamma$-point. 

### 2018-03-06 almo-langevin.tex
switched to using bbl directly

--- 2018-03-06 almo-langevin.tex
@@
 
-Even with $\epsilon_{\text{SCF}} = 10^{-2}$, we stabilized the simulation at the correct average temperature and perfect Maxwell-Boltzmann distribution (Figure~\ref{fig:dynproperties}b).
+Even with $\epsilon_{\text{SCF}} = 10^{-2}$, the simulation is stable with the correct average temperature and perfect Maxwell-Boltzmann distribution (Figure~\ref{fig:dynproperties}b).
 $\Delta$ was initially estimated at $\SI{2e-5}{\per\fs}$ using Eq.~(\ref{eq:delta}) and then refined heuristically to $\SI{6e-5}{\per\fs}$.
@@
 \textbf{Acknowledgments.} The authors are grateful to Thomas K\"uhne for insightful discussions.
-The research was funded by the Natural Sciences and Engineering Research Council of Canada through the Discovery Grant.
+The research was funded by the Natural Sciences and Engineering Research Council of Canada through the Discovery Grant (RGPIN-2016-05059).
 The authors are grateful to Compute Canada and McGill HPC Centre for computer time.
 
-%RZK: incorrect journal abbreviation in a:almo-ls, formatting of all references must be checked
-
-\bibliography{almo-langevin}
+%\bibliography{almo-langevin}
+\input{almo-langevin.bbl}
 
@@
 %calculated as the average difference between fully converged Born-Oppenheimer forces and fully converged ALMO DFT forces, averaged over all dimensions and all molecules for each time step.
-Forces are computed with the PBE XC functional, TZV2P basis set, and $\epsilon_{\text{SCF}} = 10^{-7}$~a.u. for the configurations from a 10-ps trajectory generated with delocalized AIMD at T=298~K.}
+Forces are computed with the PBE XC functional, TZV2P basis set, and $\epsilon_{\text{SCF}} = 10^{-7}$~a.u. for the configurations from a 30-ps trajectory generated with delocalized AIMD at T=298~K.}
 \end{figure}
@@
 
-\begin{figure}[!htb]
-\includegraphics[trim={2.4cm 0cm 3.0cm 1cm},clip,width=8.6cm]{Diffusion_Comparison.eps}
-\caption{\label{fig:diffusion} Calculated mean-squared displacement function of water using ALMO DFT with $\epsilon_{\text{SCF}} = 10^{-2}$~a.u. and $R_{c} = 1.6$ vdWR.
-Compared with equivalent MD run using OT DFT as a reference.
-Calculations were performed at 298 K with the dispersion-corrected PBE $\mathrm{E_{XC}}$ functional and TZV2P basis set, using 125 water molecules for 10 ps.}
-\end{figure}
+%\begin{figure}[!htb]
+%\includegraphics[trim={2.4cm 0cm 3.0cm 1cm},clip,width=8.6cm]{Diffusion_Comparison.eps}
+%\caption{\label{fig:diffusion} Calculated mean-squared displacement function of water using ALMO DFT with $\epsilon_{\text{SCF}} = 10^{-2}$~a.u. and $R_{c} = 1.6$ vdWR.
+%Compared with equivalent MD run using OT DFT as a reference.
+%Calculations were performed at 298 K with the dispersion-corrected PBE $\mathrm{E_{XC}}$ functional and TZV2P basis set, using 125 water molecules for 30 ps.}
+%\end{figure}
 

### 2018-03-06 almo-langevin.tex
files organized
Comments: cleaned up

@@
 An application to liquid water demonstrates that the low computational overhead of the method makes it well-suited for routine simulations whereas its linear scaling complexity allows to extend first-principle dynamical studies of molecular systems to previously inaccessible time and length scales.
-%Within the framework of linear-scaling Kohn-Sham density functional theory, a robust method for maintaining compact localized orbitals close to the ground state coupled with nuclear dynamics enables us to create low-cost linear-scaling \emph{ab initio} molecular dynamics for weakly-interacting systems.
-High efficiency of simulations is achieved by calculating the forces approximately without lengthy optimization of localized orbitals and by adjusting the Langevin equation of motion to stabilize the dynamics with the imperfect forces.
-An application to liquid water demonstrates that low computational overhead of this procedure makes it well-suited for routine simulations whereas its linear scaling complexity allows to extend first-principle molecular dynamical studies to previously inaccessible time and length scales.
 \end{abstract}
@@
 \ifdefined\MAINTEXT
-
-% NEW REFERENCES:
-%Richters2014 -- DOI:10.1063/1.4869865
-%Aarons2016 -- DOI:10.1063/1.4972007
-%Kussmann2013 -- DOI:10.1002/wcms.1138
-%Vitale2017 -- DOI:10.1063/1.4978684
-%Skylaris2005 -- DOI:10.1063/1.1839852
-%Hine2011 -- DOI:10.1103/PhysRevB.83.195102
-%Bowler2010 -- DOI:10.1088/0953-8984/22/7/074207
-%Nakata2015 -- DOI:10.1039/C5CP00934K
-%Krajewski -- PRB, 73, 041105R
-%Kolafa2003 -- J Comput Chem 25: 335–342, 2004
-%Benoit2001 -- DOI:10.1103/PhysRevLett.87.226401
-%Otsuka2016 -- DOI:10.7567/JJAP.55.1102B1
-%Lin2012 -- DOI:10.1016/j.jcp.2011.11.032
-%Kuhne2013 -- DOI:10.1002/wcms.1176
-%Ricci2003 -- DOI:10.1080/0026897031000108113
 
@@
 However, the DM DFT becomes advantageous only for impractically large systems when accurate multifunction basis sets are used~\cite{a:ls-dm-sign, Arita2014, a:ls-rev-2012, a:almo-ls}.
-%Therefore, the application of DM-based LS methods have been restricted to minimal-basis tight-binding problems~\cite{Richters2014, more}. 
+%Therefore, the application of DM-based LS methods have been restricted to minimal-basis tight-binding problems~\cite{Richters2014}. 
 This issue is rectified in optimal-basis DM methods~\cite{Skylaris2005, Nakata2015, Mohr2015} that contract large basis sets into a small number of new localized functions and then optimize the DM in the contracted basis.
-%Later: Although such methods have been successfully used for the evaluation of accurate DFT energies of large static systems~\cite{a:ls-onetep-2009,a:ls-conquest-2010,a:ls-onetep-2010-app1,a:ls-rev-2012}
-%While most large-scale DFT calculations are performed using this class of methods
 Despite becoming the most popular approach to LS DFT, the efficiency of these methods is hampered by the costly optimization of both the contracted orbitals and the DM~\cite{a:ls-onetep-2003}. 
-%
 From this point of view, a direct variation of molecular orbitals that are strictly localized within predefined regions is preferable because LS can be achieved with significantly fewer variables. 
 Advantages of the orbitals-only LS DFT are especially pronounced in accurate calculations that require many basis functions per atom.
-%From this point of view, it is desirable to obviate the DM optimization and use another set of local electronic descriptors can be constructed by enforcing strict locality on one-electron orbitals and, necessarily, by relaxing the orthogonality constraints imposed on them.
-%
-% direct optimization of such compact localized nonorthogonal orbitals
-%
-%To obviate the DM optimization, another set of local electronic descriptors can be constructed by enforcing strict locality on one-electron orbitals and, necessarily, by relaxing the orthogonality constraints imposed on them.
-%From the computational point of view, LS methods based on the direct optimization of such compact localized nonorthogonal orbitals~\cite{a:ls-galli-parrinello-1992,a:ls-mauri-galli-car-1993,a:ls-ordejon-1993,a:ls-mauri-galli-1994,a:ls-ordejon-1995,a:ls-kim-mauri-galli-1995,a:ls-fattebert-2004,a:ls-fattebert-2006,a:burger-yang-2008} are promising since they deal with fewer variational degrees of freedom than the DM methods.
 Unfortunately, the development of promising orbital-based LS methods has been all but abandoned~\cite{a:weitao-yang-2013,a:ls-tsuchida-aomm} because of the inherently difficult optimization of localized orbitals~\cite{a:ls-rev-1999,a:ls-mauri-galli-car-1993,a:ls-ordejon-1995,a:ls-fattebert-2004,a:weitao-yang-2013,a:ls-tsuchida-aomm}.
 
-%RZK important addition for a more detialed publication: review of a large family of fragmentation methods into the review following these guidelines.
+%RZK important addition for a more detialed publication: review of a large family of fragmentation methods following these guidelines.
 %It is important to note all LS DFT methods reviewed above attempt to obtain the electronic description that do not construct  based on fragmentation methods that do that there  LS DFT methods the presented logical partitioning with the subsequent construction of localized MOs is a rather general approach used in a large number of electronic structure theories, which are collectively known as fragmentation methods~\cite{a:fragmentation-rev}.
@@
 Thus, despite impressive progress of the LS description of the electronic and atomic structure of large static systems~\cite{Bowler2010,a:ls-dm-sign}, the high computational overhead of existing LS methods restrict their use in dynamical simulations to very short \emph{time scales}, systems of low dimensions, and low-quality minimal basis sets~\cite{a:ls-dm-sign, Otsuka2016, Hine2011, Bowler2010}.
-%RZK: Although such methods have been successfully used for the evaluation of accurate DFT energies [electronic structure and atomic configurations] of large static systems~\cite{Bowler2010,a:ls-dm-sign}
 On typical length and time scales required in practical and accurate AIMD simulations, LS DFT still cannot compete with the straightforward low-cost cubically-scaling KS DFT.
@@
 In this work, we present an AIMD method that overcomes difficulties of orbital-only local DFT to achieve LS with extremely low computational overhead. 
-%Currently, the method is restricted to systems of small molecules that do not form strong covalent bonds with each other. 
 To demonstrate advantages of the new method we applied it here to systems of weakly-interacting molecules. 
@@
 A generalization of the method to all finite-gap systems, including covalently bonded atoms, will be reported later. 
-%It is important to emphasize that delocalized electrons in metals, which are particularly challenging for all LS DFT~\cite{Kuehne,recent-Skylaris}, cannot be described efficiently within the presented framework.
+%It is important to emphasize that delocalized electrons in metals, which are particularly challenging for all LS DFT~\cite{Kuehne}, cannot be described efficiently within the presented framework.
 
@@
 
-%To decrease the cost even further, the present paper builds on the ALMO SCF method as well as other recent work involving Brownian dynamics to compensate errors in the forces~\cite{a:ls-parinello,a:2ndcpmd,a:langevin-why}.
-%We will show that our approach lowers the computation cost pre-factor for the ALMO SCF method, while maintaining LS and accuracy for calculation of dynamical properties. 
-%This is done by balancing a new computationally economical but systemically inaccurate force calculation method with an optimized version of the Langevin equation. 
-%We also introduce an easy to follow heuristic approach to optimize a system for generating correct dynamical behaviour.
-
 The challenge of adopting ALMO DFT for dynamical simulations arises from the slightly nonvariational character of the localized orbitals.
@@
 In addition, electron transfer effects can suddenly become inactive in the course of a dynamical simulation when a neighboring molecule crosses the localization threshold $R_{c}$.
-Futhermore, the variational optimization in any AIMD method is never truly complete in practice
-%[exact convergence is never achieved in practical calculation] 
-and interrupted once the maximum norm of the gradient of the energy with respect to the electronic descriptors drops below small but nevertheless finite convergence threshold $\epsilon_{\text{SCF}}$.
+Futhermore, the variational optimization in any AIMD method is never complete in practice and interrupted once the maximum norm of the gradient of the energy with respect to the electronic descriptors drops below small but nevertheless finite convergence threshold $\epsilon_{\text{SCF}}$.
 These errors do not affect the accuracy of static ALMO DFT calculations, geometry optimization, and Monte-Carlo simulations.
@@
 
-%RZK: There is a problem with the integral equation below.
-The r.h.s.
-depends on $i$ but in the l.h.s.
-does not.
-If I remember correctly we evaluate the integral for molecules, not atoms.
-In liquid water all our molecules are the same and we do not have to deal with the issue of different Delta for different molecules.
-Eliminating the problem
 In principle, $\Delta$ can be calculated using the integral of Eq.~(\ref{eq:stochastic3}) averaged over atoms with different $m_i$
@@
 This value can be further fine-tuned in a series of short trial-and-error ALMO AIMD runs until the average kinetic energy corresponds to the requested temperature $\langle \frac{1}{2} m_i \dot{\bm{r}}^{2}_{i} \rangle = \frac{3}{2} k_{B} T$.
-%%
-%\begin{align}
-%\label{eq:eqipartition}
-%\langle \dfrac{1}{2} m_i \dot{\bm{r}}^{2}_{i} \rangle &= \dfrac{3}{2} k_{B} T
-%\end{align}
 
@@
 Separable norm-conserving pseudopotentials were used~\cite{a:hgh} and the Brillouin zone was sampled at the $\Gamma$-point. 
-%to describe the interactions between the valence electrons and the ionic cores. 
-%
 %Although not a required component of ALMO AIMD, 
@@
 \caption{\label{fig:randomforce} 
-%RZK: label of the third panel: "d" -> "c" 
 The red line is the Euclidean norm of the instantaneous error $\langle \| \delta \vec{f_{i}}(t) \| \rangle_{i}$, the black line is the magnitude of the time average of the instantaneous error vector, and the green line is the time average of the red line.
@@
 %\bibliography{almo-langevin}
-\input{almo-langevin.bbl}
+\input{almo-langevin.ref}
 
 \else
-
-%\subsection{Analytical expression for $\Delta$}
-%
-%Another approach based on Eq. (\ref{eq:assumption}) can be used to find $\Delta$ within an order of magnitude if one can afford computing tightly converged SCF forces for a short MD trajectory. 
-%First, run the simulation with $\Delta = 0$ and converged SCF forces. 
-%Second, compute approximate forces for snapshots along the trajectory and calculate the random term $R^{\Delta}_{i\alpha} (t)$ using Eq.\ \ref{eq:assumption}. 
-%Third, calculate the autocovariance function (ACF) of the noise $\text{ACF}(t) = \langle R^{\Delta}_{i\alpha} (0)  R^{\Delta}_{i\alpha} (t) \rangle$, which should be sharply peaked around zero lag time if our assumption is correct.
-%Finally, integrate Eq.(\ref{eq:stochastic3}) to obtain the following expression for $\Delta$.
-%%
-%\begin{align}
-%\label{eq:delta}
-%\Delta &= (2 k_B T m_i )^{-1} \int_{-\infty}^{\infty}\langle R^{\Delta}_{i\alpha} (0)  R^{\Delta}_{i\alpha} (t) \rangle dt
-%\end{align}
-%%
-%In other words, we use the integral of the ACF to estimate the strength of the random force.
-
-%%% %% This was a summary of the two options for Delta
-%%%Among several possible expression for $\Delta$ we consider the following two options.
-%%%
-%%%\textbf{Option 1.} $\Delta$ is calculated as a weighted average of atom TYPES (e.g. oxygen and hydrogen). Subindex $\text{T}$ refers to atom type.
-%%%
-%%%\textbf{\emph{Step 1.}} Compute the average ACF for each atom type.
-In this step averaging is done over all time origins $t$, over all atoms of THIS type $i$, and over all dimensions $\alpha$.
-The ACF includes the mass-dependent factor $(2 k_B T m_\text{T} )^{-1}$ where $T$ is the requested temperature.
-%%%%
-%%%\begin{align}
-%%%\label{eq:deltaS1}
-%%%\text{ACF}_{\text{T}}(\tau) = (2 k_B T m_\text{T} )^{-1} \, \langle \delta f_{i\alpha} (t) \, \delta f_{i\alpha}(t + \tau) \rangle_{t i\alpha}
-%%%\end{align}
-%%%%
-%%%\textbf{\emph{Step 2.}} Integrate the ACF for each atom type.
-%%%%
-%%%\begin{align}
-%%%\label{eq:deltaS2}
-%%%\Delta_{\text{T}} = \int_{-\infty}^{\infty} \text{ACF}_{T}(\tau) \, d\tau
-%%%\end{align}
-%%%%
-%%%\textbf{\emph{Step 3.}} Average all over all atom types.
-%%%%
-%%%\begin{align}
-%%%\label{eq:deltaS3}
-%%%\Delta = \frac{1}{N} \sum_{\text{T}} N_{\text{T}} \Delta_{\text{T}}
-%%%\end{align}
-%%%%
-%%%where $N$ is the number of atoms.
-%%%
-%%%If all these step are combined the following expression for $\Delta$ is obtained:
-%%%%
-%%%\begin{align}
-%%%\label{eq:deltaTOT}
-%%%\Delta = \int \, d\tau \left[ \frac{1}{3N} \sum_{i=1}^{N} \sum_{\alpha=1}^{3} \frac{ \langle \delta f_{i\alpha} (t) \, \delta f_{i\alpha}(t + \tau) \rangle_{t} }{2 k_B T m_i} \right]
-%%%\end{align}
-%%%%
-%%%Thus the weight of each of the $3N$ degrees of freedom in this scheme is proportional to the inverse atomic mass.
-
-%%%\textbf{Option 2.} If the same averaging scheme is applied to MOLECULE types (not ATOM types) then the value of $\Delta$ is given by:
-%%%%
-%%%\begin{align}
-%%%\label{eq:O2deltaTOT}
-%%%\Delta = \int \, d\tau \left[ \frac{1}{3M} \sum_{p=1}^{M} \sum_{\alpha=1}^{3} \frac{ \sum_i^{N_p}\sum_j^{N_p} \langle \delta f_{i\alpha} (t) \, \delta f_{j\alpha}(t + \tau) \rangle_{t} }{2 k_B T \sum_k^{N_p}m_k} \right]
-%%%\end{align}
-%%%%
-%%%where $M$ is the number of molecules, $N_p$ is the number of atoms in molecule $p$.
-The value of this $\Delta$ might be different from Option 1 because (a) all correlations between two atoms inside one molecule are taken into account and (b) the weight of each degree of freedom is different.
-The difference between Option 1 and 2 can be clearly seen for a special case of identical molecules:
-%%%%
-%%%\begin{align}
-%%%\label{eq:O2deltaSPECIAL}
-%%%\Delta = \int \, d\tau \left[ \frac{1}{3N} \sum_{i=1}^{N} \sum_{\alpha=1}^{3} \frac{ \sum_j^{N_p} \langle \delta f_{i\alpha} (t) \, \delta f_{j\alpha}(t + \tau) \rangle_{t} }{2 k_B T \langle m \rangle } \right]
-%%%\end{align}
-%%%%
-%%%where $\langle m \rangle = (\sum_k^{N_p} m_k )/N_p$ is the average atom mass in a molecule.
-
-%$\Delta$ is calculated by averaging over molecule types:
-%%
-%\begin{align}
-%\label{eq:O2deltaTOT}
-%\Delta = \int \, d\tau \left[ \frac{1}{3M} \sum_{p=1}^{M} \sum_{\alpha=1}^{3} \frac{ \sum_i^{N_p}\sum_j^{N_p} \langle \delta f_{i\alpha} (t) \, \delta f_{j\alpha}(t + \tau) \rangle_{t} }{2 k_B T \sum_k^{N_p}m_k} \right]
-%\end{align}
-%%
-%where $M$ is the number of molecules, $N_p$ is the number of atoms in molecule $p$.
-The value of this $\Delta$ might be different from Option 1 because (a) all correlations between two atoms inside one molecule are taken into account and (b) the weight of each degree of freedom is different.
-The difference between Option 1 and 2 can be clearly seen for a special case of identical molecules:
-%%
-%\begin{align}
-%\label{eq:O2deltaSPECIAL}
-%\Delta = \int \, d\tau \left[ \frac{1}{3N} \sum_{i=1}^{N} \sum_{\alpha=1}^{3} \frac{ \sum_j^{N_p} \langle \delta f_{i\alpha} (t) \, \delta f_{j\alpha}(t + \tau) \rangle_{t} }{2 k_B T \langle m \rangle } \right]
-%\end{align}
-%%
-%where $\langle m \rangle = (\sum_k^{N_p} m_k )/N_p$ is the average atom mass in a molecule.
 

### 2018-03-13 almo-langevin.tex
Title and abstract updated for JPCL

@@
 \fi
-Low-cost orbital-based linear-scaling \emph{ab initio} molecular dynamics for weakly-interacting systems
+Compact orbitals enable low-cost linear-scaling \emph{ab initio} molecular dynamics for weakly-interacting systems
 }
@@
 \email{scheiber@chem.ubc.ca}
+\affiliation{Department of Chemistry, McGill University, 801 Sherbrooke St. West, Montreal, QC H3A 0B8, Canada}
 \author{Yifei Shi}
+\affiliation{Department of Chemistry, McGill University, 801 Sherbrooke St. West, Montreal, QC H3A 0B8, Canada}
 \author{Rustam Z. Khaliullin}
@@
 \begin{abstract}
-Within the framework of linear-scaling Kohn-Sham density functional theory, a robust method for maintaining compact orbitals close to the ground state is coupled with nuclear dynamics to obviate the commonly employed optimization of the one-electron density matrix and to create an efficient orbital-only molecular dynamics method for weakly-interacting systems.
-An application to liquid water demonstrates that the low computational overhead of the method makes it well-suited for routine simulations whereas its linear scaling complexity allows to extend first-principle dynamical studies of molecular systems to previously inaccessible time and length scales.
+Today, \emph{ab initio} molecular dynamics (AIMD) relies on the locality of one-electron density matrices to achieve linear growth of computation time with systems size, crucial in large-scale simulations.
+While Kohn-Sham orbitals strictly localized within predefined radii can offer substantial computational advantages over density matrices, such compact orbitals are not used in AIMD because a compact representation of the electronic ground state is difficult to find.
+Here, a robust method for maintaining compact orbitals close to the ground state is coupled with a modified Langevin integrator to produce stable nuclear dynamics for molecular and ionic systems.
+This eliminates inefficient density matrix optimization and enables first orbital-only linear-scaling AIMD.
+An application to liquid water demonstrates that low computational overhead of the new method makes it ideal for routine medium-scale simulations while its linear-scaling complexity allows to extend first-principle studies of molecular systems to completely new chemical phenomena on previously inaccessible length scales.
+%Within the framework of linear-scaling Kohn-Sham density functional theory, a robust method for maintaining compact localized orbitals close to the ground state is coupled with nuclear dynamics.
+This allows to obviate the commonly employed optimization of the one-electron density matrix and thus create an efficient orbital-only molecular dynamics method for weakly-interacting systems.
+%An application to liquid water demonstrates that the low computational overhead of the method makes it well-suited for routine simulations whereas its linear-scaling complexity allows to extend first-principle dynamical studies of molecular systems to previously inaccessible length scales.
 \end{abstract}
@@
 Accuracy and efficiency of ALMO AIMD was tested using liquid water as an example. 
+%RZK: cite our own water paper
 This system is challenging because intermolecular electron delocalization is a critical component of hydrogen bonding and must be described correctly to reproduce static and dynamical properties of liquid water.
@@
 %A remarkable feature of the method is that it remains efficient for challenging condensed phase systems even with large localized basis sets. 
-Using liquid water as an example, we showed that the new approach enables simulations of molecular systems on previously inaccessible time and length scales. 
+Using liquid water as an example, we showed that the new approach enables simulations of molecular systems on previously inaccessible length scales. 
 The developed method will have a significant impact on modeling of complex molecular systems (e.g. interfaces or nuclei) making completely new phenomena accessible to AIMD. 
@@
 %\bibliography{almo-langevin}
-\input{almo-langevin.ref}
+%\input{almo-langevin.ref}
+\begin{thebibliography}{40}%
+\makeatletter
+\providecommand \@ifxundefined [1]{%
+ \@ifx{#1\undefined}
+}%
+\providecommand \@ifnum [1]{%
+ \ifnum #1\expandafter \@firstoftwo
+ \else \expandafter \@secondoftwo
+ \fi
+}%
+\providecommand \@ifx [1]{%
+ \ifx #1\expandafter \@firstoftwo
+ \else \expandafter \@secondoftwo
+ \fi
+}%
+\providecommand \natexlab [1]{#1}%
+\providecommand \enquote  [1]{``#1''}%
+\providecommand \bibnamefont  [1]{#1}%
+\providecommand \bibfnamefont [1]{#1}%
+\providecommand \citenamefont [1]{#1}%
+\providecommand \href@noop [0]{\@secondoftwo}%
+\providecommand \href [0]{\begingroup \@sanitize@url \@href}%
+\providecommand \@href[1]{\@@startlink{#1}\@@href}%
+\providecommand \@@href[1]{\endgroup#1\@@endlink}%
+\providecommand \@sanitize@url [0]{\catcode `\\12\catcode `\$12\catcode
+  `\&12\catcode `\#12\catcode `\^12\catcode `\_12\catcode `\%12\relax}%
+\providecommand \@@startlink[1]{}%
+\providecommand \@@endlink[0]{}%
+\providecommand \url  [0]{\begingroup\@sanitize@url \@url }%
+\providecommand \@url [1]{\endgroup\@href {#1}{\urlprefix }}%
+\providecommand \urlprefix  [0]{URL }%
+\providecommand \Eprint [0]{\href }%
+\providecommand \doibase [0]{http://dx.doi.org/}%
+\providecommand \selectlanguage [0]{\@gobble}%
+\providecommand \bibinfo  [0]{\@secondoftwo}%
+\providecommand \bibfield  [0]{\@secondoftwo}%
+\providecommand \translation [1]{[#1]}%
+\providecommand \BibitemOpen [0]{}%
+\providecommand \bibitemStop [0]{}%
+\providecommand \bibitemNoStop [0]{.\EOS\space}%
+\providecommand \EOS [0]{\spacefactor3000\relax}%
+\providecommand \BibitemShut  [1]{\csname bibitem#1\endcsname}%
+\let\auto@bib@innerbib\@empty
+%</preamble>
+\bibitem [{\citenamefont {Car}\ and\ \citenamefont
+  {Parrinello}(1985)}]{a:thecpmd}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {R.}~\bibnamefont
+  {Car}}\ and\ \bibinfo {author} {\bibfnamefont {M.}~\bibnamefont
+  {Parrinello}},\ }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal}
+  {Phys. Rev. Lett.}\ }\textbf {\bibinfo {volume} {55}},\ \bibinfo {pages}
+  {2471} (\bibinfo {year} {1985})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Goedecker}(1999)}]{a:ls-rev-1999}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {S.}~\bibnamefont
+  {Goedecker}},\ }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal}
+  {Rev. Mod. Phys.}\ }\textbf {\bibinfo {volume} {71}},\ \bibinfo {pages}
+  {1085} (\bibinfo {year} {1999})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Bowler}\ and\ \citenamefont
+  {Miyazaki}(2012)}]{a:ls-rev-2012}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {D.~R.}\ \bibnamefont
+  {Bowler}}\ and\ \bibinfo {author} {\bibfnamefont {T.}~\bibnamefont
+  {Miyazaki}},\ }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal} {Rep.
+  Prog. Phys.}\ }\textbf {\bibinfo {volume} {75}},\ \bibinfo {pages} {36503}
+  (\bibinfo {year} {2012})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Kussmann}\ \emph {et~al.}(2013)\citenamefont
+  {Kussmann}, \citenamefont {Beer},\ and\ \citenamefont
+  {Ochsenfeld}}]{Kussmann2013}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {J.}~\bibnamefont
+  {Kussmann}}, \bibinfo {author} {\bibfnamefont {M.}~\bibnamefont {Beer}}, \
+  and\ \bibinfo {author} {\bibfnamefont {C.}~\bibnamefont {Ochsenfeld}},\
+  }\href {\doibase 10.1002/wcms.1138} {\bibfield  {journal} {\bibinfo
+  {journal} {Wiley Interdiscip. Rev. Comput. Mol. Sci.}\
+  }\textbf {\bibinfo {volume} {3}},\ \bibinfo {pages} {614} (\bibinfo {year}
+  {2013})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Aarons}\ \emph {et~al.}(2016)\citenamefont {Aarons},
+  \citenamefont {Sarwar}, \citenamefont {Thompsett},\ and\ \citenamefont
+  {Skylaris}}]{Aarons2016}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {J.}~\bibnamefont
+  {Aarons}}, \bibinfo {author} {\bibfnamefont {M.}~\bibnamefont {Sarwar}},
+  \bibinfo {author} {\bibfnamefont {D.}~\bibnamefont {Thompsett}}, \ and\
+  \bibinfo {author} {\bibfnamefont {C.-K.}\ \bibnamefont {Skylaris}},\ }\href
+  {\doibase 10.1063/1.4972007} {\bibfield  {journal} {\bibinfo  {journal} {J. 
+  Chem. Phys.}\ }\textbf {\bibinfo {volume} {145}},\ \bibinfo
+  {pages} {220901} (\bibinfo {year} {2016})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Vandevondele}\ \emph {et~al.}(2012)\citenamefont
+  {Vandevondele}, \citenamefont {Borstnik},\ and\ \citenamefont
+  {Hutter}}]{a:ls-dm-sign}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {J.}~\bibnamefont
+  {Vandevondele}}, \bibinfo {author} {\bibfnamefont {U.}~\bibnamefont
+  {Borstnik}}, \ and\ \bibinfo {author} {\bibfnamefont {J.}~\bibnamefont
+  {Hutter}},\ }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal} {J.
+  Chem. Theory Comput.}\ }\textbf {\bibinfo {volume} {8}},\ \bibinfo {pages}
+  {3565} (\bibinfo {year} {2012})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Arita}\ \emph {et~al.}(2014)\citenamefont {Arita},
+  \citenamefont {Bowler},\ and\ \citenamefont {Miyazaki}}]{Arita2014}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {M.}~\bibnamefont
+  {Arita}}, \bibinfo {author} {\bibfnamefont {D.~R.}\ \bibnamefont {Bowler}}, \
+  and\ \bibinfo {author} {\bibfnamefont {T.}~\bibnamefont {Miyazaki}},\
+  }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal} {J.
+  Chem. Theory Comput.}\ }\textbf {\bibinfo {volume} {10}},\
+  \bibinfo {pages} {5419} (\bibinfo {year} {2014})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Khaliullin}\ \emph {et~al.}(2013)\citenamefont
+  {Khaliullin}, \citenamefont {VandeVondele},\ and\ \citenamefont
+  {Hutter}}]{a:almo-ls}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {R.~Z.}\ \bibnamefont
+  {Khaliullin}}, \bibinfo {author} {\bibfnamefont {J.}~\bibnamefont
+  {VandeVondele}}, \ and\ \bibinfo {author} {\bibfnamefont {J.}~\bibnamefont
+  {Hutter}},\ }\href {\doibase 10.1021/ct400595k} {\bibfield  {journal}
+  {\bibinfo  {journal} {J. Chem. Theory Comput.}\ }\textbf
+  {\bibinfo {volume} {9}},\ \bibinfo {pages} {4421} (\bibinfo {year}
+  {2013})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Skylaris}\ \emph {et~al.}(2005)\citenamefont
+  {Skylaris}, \citenamefont {Haynes}, \citenamefont {Mostofi},\ and\
+  \citenamefont {Payne}}]{Skylaris2005}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {C.-K.}\ \bibnamefont
+  {Skylaris}}, \bibinfo {author} {\bibfnamefont {P.~D.}\ \bibnamefont
+  {Haynes}}, \bibinfo {author} {\bibfnamefont {A.~A.}\ \bibnamefont {Mostofi}},
+  \ and\ \bibinfo {author} {\bibfnamefont {M.~C.}\ \bibnamefont {Payne}},\
+  }\href {\doibase 10.1063/1.1839852} {\bibfield  {journal} {\bibinfo
+  {journal} {J. Chem. Phys.}\ }\textbf {\bibinfo {volume}
+  {122}},\ \bibinfo {pages} {84119} (\bibinfo {year} {2005})}\BibitemShut
+  {NoStop}%
+\bibitem [{\citenamefont {Nakata}\ \emph {et~al.}(2015)\citenamefont {Nakata},
+  \citenamefont {Bowler},\ and\ \citenamefont {Miyazaki}}]{Nakata2015}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {A.}~\bibnamefont
+  {Nakata}}, \bibinfo {author} {\bibfnamefont {D.~R.}\ \bibnamefont {Bowler}},
+  \ and\ \bibinfo {author} {\bibfnamefont {T.}~\bibnamefont {Miyazaki}},\
+  }\href {\doibase 10.1039/C5CP00934K} {\bibfield  {journal} {\bibinfo
+  {journal} {Phys. Chem. Chem. Phys.}\ }\textbf {\bibinfo {volume} {17}},\
+  \bibinfo {pages} {31427} (\bibinfo {year} {2015})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Mohr}\ \emph {et~al.}(2015)\citenamefont {Mohr},
+  \citenamefont {Ratcliff}, \citenamefont {Genovese}, \citenamefont {Caliste},
+  \citenamefont {Boulanger}, \citenamefont {Goedecker},\ and\ \citenamefont
+  {Deutsch}}]{Mohr2015}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {S.}~\bibnamefont
+  {Mohr}}, \bibinfo {author} {\bibfnamefont {L.~E.}\ \bibnamefont {Ratcliff}},
+  \bibinfo {author} {\bibfnamefont {L.}~\bibnamefont {Genovese}}, \bibinfo
+  {author} {\bibfnamefont {D.}~\bibnamefont {Caliste}}, \bibinfo {author}
+  {\bibfnamefont {P.}~\bibnamefont {Boulanger}}, \bibinfo {author}
+  {\bibfnamefont {S.}~\bibnamefont {Goedecker}}, \ and\ \bibinfo {author}
+  {\bibfnamefont {T.}~\bibnamefont {Deutsch}},\ }\href {\doibase
+  10.1039/C5CP00437C} {\bibfield  {journal} {\bibinfo  {journal} {Phys. Chem.
+  Chem. Phys.}\ }\textbf {\bibinfo {volume} {17}},\ \bibinfo {pages} {31360}
+  (\bibinfo {year} {2015})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Mostofi}\ \emph {et~al.}(2003)\citenamefont
+  {Mostofi}, \citenamefont {Haynes}, \citenamefont {Skylaris},\ and\
+  \citenamefont {Payne}}]{a:ls-onetep-2003}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {A.~A.}\ \bibnamefont
+  {Mostofi}}, \bibinfo {author} {\bibfnamefont {P.~D.}\ \bibnamefont {Haynes}},
+  \bibinfo {author} {\bibfnamefont {C.-K.}\ \bibnamefont {Skylaris}}, \ and\
+  \bibinfo {author} {\bibfnamefont {M.~C.}\ \bibnamefont {Payne}},\ }\href@noop
+  {} {\bibfield  {journal} {\bibinfo  {journal} {J. Chem. Phys.}\ }\textbf
+  {\bibinfo {volume} {119}},\ \bibinfo {pages} {8842} (\bibinfo {year}
+  {2003})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {van Aggelen}\ \emph {et~al.}(2013)\citenamefont {van
+  Aggelen}, \citenamefont {Yang},\ and\ \citenamefont
+  {Yang}}]{a:weitao-yang-2013}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {H.}~\bibnamefont {van
+  Aggelen}}, \bibinfo {author} {\bibfnamefont {Y.}~\bibnamefont {Yang}}, \ and\
+  \bibinfo {author} {\bibfnamefont {W.}~\bibnamefont {Yang}},\ }\href {\doibase
+  10.1103/PhysRevA.88.030501} {\bibfield  {journal} {\bibinfo  {journal} {Phys.
+  Rev. A}\ }\textbf {\bibinfo {volume} {88}},\ \bibinfo {pages} {030501}
+  (\bibinfo {year} {2013})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Tsuchida}(2007)}]{a:ls-tsuchida-aomm}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {E.}~\bibnamefont
+  {Tsuchida}},\ }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal} {J.
+  Phys. Soc. Jpn.}\ }\textbf {\bibinfo {volume} {76}},\ \bibinfo {pages}
+  {34708} (\bibinfo {year} {2007})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Mauri}\ \emph {et~al.}(1993)\citenamefont {Mauri},
+  \citenamefont {Galli},\ and\ \citenamefont
+  {Car}}]{a:ls-mauri-galli-car-1993}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {F.}~\bibnamefont
+  {Mauri}}, \bibinfo {author} {\bibfnamefont {G.}~\bibnamefont {Galli}}, \ and\
+  \bibinfo {author} {\bibfnamefont {R.}~\bibnamefont {Car}},\ }\href@noop {}
+  {\bibfield  {journal} {\bibinfo  {journal} {Phys. Rev. B}\ }\textbf {\bibinfo
+  {volume} {47}},\ \bibinfo {pages} {9973} (\bibinfo {year}
+  {1993})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Ordejon}\ \emph {et~al.}(1995)\citenamefont
+  {Ordejon}, \citenamefont {Drabold}, \citenamefont {Martin},\ and\
+  \citenamefont {Grumbach}}]{a:ls-ordejon-1995}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {P.}~\bibnamefont
+  {Ordejon}}, \bibinfo {author} {\bibfnamefont {D.~A.}\ \bibnamefont
+  {Drabold}}, \bibinfo {author} {\bibfnamefont {R.~M.}\ \bibnamefont {Martin}},
+  \ and\ \bibinfo {author} {\bibfnamefont {M.~P.}\ \bibnamefont {Grumbach}},\
+  }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal} {Phys. Rev. B}\
+  }\textbf {\bibinfo {volume} {51}},\ \bibinfo {pages} {1456} (\bibinfo {year}
+  {1995})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Fattebert}\ and\ \citenamefont
+  {Gygi}(2004)}]{a:ls-fattebert-2004}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {J.~L.}\ \bibnamefont
+  {Fattebert}}\ and\ \bibinfo {author} {\bibfnamefont {F.}~\bibnamefont
+  {Gygi}},\ }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal} {Comput.
+  Phys. Commun.}\ }\textbf {\bibinfo {volume} {162}},\ \bibinfo {pages} {24}
+  (\bibinfo {year} {2004})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Bowler}\ and\ \citenamefont
+  {Miyazaki}(2010)}]{Bowler2010}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {D.~R.}\ \bibnamefont
+  {Bowler}}\ and\ \bibinfo {author} {\bibfnamefont {T.}~\bibnamefont
+  {Miyazaki}},\ }\href {http://stacks.iop.org/0953-8984/22/i=7/a=074207}
+  {\bibfield  {journal} {\bibinfo  {journal} {J. Phys. Condens.
+  Matter}\ }\textbf {\bibinfo {volume} {22}},\ \bibinfo {pages} {74207}
+  (\bibinfo {year} {2010})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Otsuka}\ \emph {et~al.}(2016)\citenamefont {Otsuka},
+  \citenamefont {Taiji}, \citenamefont {Bowler},\ and\ \citenamefont
+  {Miyazaki}}]{Otsuka2016}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {T.}~\bibnamefont
+  {Otsuka}}, \bibinfo {author} {\bibfnamefont {M.}~\bibnamefont {Taiji}},
+  \bibinfo {author} {\bibfnamefont {D.~R.}\ \bibnamefont {Bowler}}, \ and\
+  \bibinfo {author} {\bibfnamefont {T.}~\bibnamefont {Miyazaki}},\ }\href
+  {http://stacks.iop.org/1347-4065/55/i=11/a=1102B1} {\bibfield  {journal}
+  {\bibinfo  {journal} {Jpn. J. Appl. Phys.}\ }\textbf
+  {\bibinfo {volume} {55}},\ \bibinfo {pages} {1102B1} (\bibinfo {year}
+  {2016})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Hine}\ \emph {et~al.}(2011)\citenamefont {Hine},
+  \citenamefont {Robinson}, \citenamefont {Haynes}, \citenamefont {Skylaris},
+  \citenamefont {Payne},\ and\ \citenamefont {Mostofi}}]{Hine2011}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {N.~D.~M.}\
+  \bibnamefont {Hine}}, \bibinfo {author} {\bibfnamefont {M.}~\bibnamefont
+  {Robinson}}, \bibinfo {author} {\bibfnamefont {P.~D.}\ \bibnamefont
+  {Haynes}}, \bibinfo {author} {\bibfnamefont {C.-K.}\ \bibnamefont
+  {Skylaris}}, \bibinfo {author} {\bibfnamefont {M.~C.}\ \bibnamefont {Payne}},
+  \ and\ \bibinfo {author} {\bibfnamefont {A.~A.}\ \bibnamefont {Mostofi}},\
+  }\href {\doibase 10.1103/PhysRevB.83.195102} {\bibfield  {journal} {\bibinfo
+  {journal} {Phys. Rev. B}\ }\textbf {\bibinfo {volume} {83}},\ \bibinfo
+  {pages} {195102} (\bibinfo {year} {2011})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Stoll}\ \emph {et~al.}(1980)\citenamefont {Stoll},
+  \citenamefont {Wagenblast},\ and\ \citenamefont {Preuss}}]{a:stoll}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {H.}~\bibnamefont
+  {Stoll}}, \bibinfo {author} {\bibfnamefont {G.}~\bibnamefont {Wagenblast}}, \
+  and\ \bibinfo {author} {\bibfnamefont {H.}~\bibnamefont {Preuss}},\
+  }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal} {Theor. Chim.
+  Acta}\ }\textbf {\bibinfo {volume} {57}},\ \bibinfo {pages} {169} (\bibinfo
+  {year} {1980})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Galli}\ and\ \citenamefont
+  {Parrinello}(1992)}]{a:galli_loc}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {G.}~\bibnamefont
+  {Galli}}\ and\ \bibinfo {author} {\bibfnamefont {M.}~\bibnamefont
+  {Parrinello}},\ }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal}
+  {Phys. Rev. Lett.}\ }\textbf {\bibinfo {volume} {69}},\ \bibinfo {pages}
+  {3547} (\bibinfo {year} {1992})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Lin}\ \emph {et~al.}(2012)\citenamefont {Lin},
+  \citenamefont {Lu}, \citenamefont {Ying},\ and\ \citenamefont {E}}]{Lin2012}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {L.}~\bibnamefont
+  {Lin}}, \bibinfo {author} {\bibfnamefont {J.}~\bibnamefont {Lu}}, \bibinfo
+  {author} {\bibfnamefont {L.}~\bibnamefont {Ying}}, \ and\ \bibinfo {author}
+  {\bibfnamefont {W.}~\bibnamefont {E}},\ }\href {\doibase
+  https://doi.org/10.1016/j.jcp.2011.11.032} {\bibfield  {journal} {\bibinfo
+  {journal} {J. Comput. Phys.}\ }\textbf {\bibinfo {volume}
+  {231}},\ \bibinfo {pages} {2140} (\bibinfo {year} {2012})}\BibitemShut
+  {NoStop}%
+\bibitem [{\citenamefont {Khaliullin}\ \emph {et~al.}(2006)\citenamefont
+  {Khaliullin}, \citenamefont {Head-Gordon},\ and\ \citenamefont
+  {Bell}}]{a:khal}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {R.~Z.}\ \bibnamefont
+  {Khaliullin}}, \bibinfo {author} {\bibfnamefont {M.}~\bibnamefont
+  {Head-Gordon}}, \ and\ \bibinfo {author} {\bibfnamefont {A.~T.}\ \bibnamefont
+  {Bell}},\ }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal} {J. Chem.
+  Phys.}\ }\textbf {\bibinfo {volume} {124}},\ \bibinfo {pages} {204105}
+  (\bibinfo {year} {2006})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Khaliullin}\ \emph {et~al.}(2007)\citenamefont
+  {Khaliullin}, \citenamefont {Cobar}, \citenamefont {Lochan}, \citenamefont
+  {Bell},\ and\ \citenamefont {Head-Gordon}}]{a:theeda}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {R.~Z.}\ \bibnamefont
+  {Khaliullin}}, \bibinfo {author} {\bibfnamefont {E.~A.}\ \bibnamefont
+  {Cobar}}, \bibinfo {author} {\bibfnamefont {R.~C.}\ \bibnamefont {Lochan}},
+  \bibinfo {author} {\bibfnamefont {A.~T.}\ \bibnamefont {Bell}}, \ and\
+  \bibinfo {author} {\bibfnamefont {M.}~\bibnamefont {Head-Gordon}},\
+  }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal} {J. Phys. Chem. A}\
+  }\textbf {\bibinfo {volume} {111}},\ \bibinfo {pages} {8753} (\bibinfo {year}
+  {2007})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Benoit}\ \emph {et~al.}(2001)\citenamefont {Benoit},
+  \citenamefont {Sebastiani},\ and\ \citenamefont {Parrinello}}]{Benoit2001}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {D.~M.}\ \bibnamefont
+  {Benoit}}, \bibinfo {author} {\bibfnamefont {D.}~\bibnamefont {Sebastiani}},
+  \ and\ \bibinfo {author} {\bibfnamefont {M.}~\bibnamefont {Parrinello}},\
+  }\href {\doibase 10.1103/PhysRevLett.87.226401} {\bibfield  {journal}
+  {\bibinfo  {journal} {Phys. Rev. Lett.}\ }\textbf {\bibinfo {volume} {87}},\
+  \bibinfo {pages} {226401} (\bibinfo {year} {2001})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Krajewski}\ and\ \citenamefont
+  {Parrinello}(2006)}]{Krajewski}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {F.~R.}\ \bibnamefont
+  {Krajewski}}\ and\ \bibinfo {author} {\bibfnamefont {M.}~\bibnamefont
+  {Parrinello}},\ }\href {\doibase 10.1103/PhysRevB.73.041105} {\bibfield
+  {journal} {\bibinfo  {journal} {Phys. Rev. B}\ }\textbf {\bibinfo {volume}
+  {73}},\ \bibinfo {pages} {041105} (\bibinfo {year} {2006})}\BibitemShut
+  {NoStop}%
+\bibitem [{\citenamefont {K{\"{u}}hne}\ \emph {et~al.}(2007)\citenamefont
+  {K{\"{u}}hne}, \citenamefont {Krack}, \citenamefont {Mohamed},\ and\
+  \citenamefont {Parrinello}}]{a:2ndcpmd}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {T.~D.}\ \bibnamefont
+  {K{\"{u}}hne}}, \bibinfo {author} {\bibfnamefont {M.}~\bibnamefont {Krack}},
+  \bibinfo {author} {\bibfnamefont {F.~R.}\ \bibnamefont {Mohamed}}, \ and\
+  \bibinfo {author} {\bibfnamefont {M.}~\bibnamefont {Parrinello}},\
+  }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal} {Phys. Rev. Lett.}\
+  }\textbf {\bibinfo {volume} {98}},\ \bibinfo {pages} {066401} (\bibinfo {year}
+  {2007})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Dai}\ and\ \citenamefont
+  {Yuan}(2009)}]{a:langevin-why}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {J.}~\bibnamefont
+  {Dai}}\ and\ \bibinfo {author} {\bibfnamefont {J.}~\bibnamefont {Yuan}},\
+  }\href {http://stacks.iop.org/0295-5075/88/i=2/a=20001} {\bibfield  {journal}
+  {\bibinfo  {journal} {EPL}\ }\textbf {\bibinfo {volume}
+  {88}},\ \bibinfo {pages} {20001} (\bibinfo {year} {2009})}\BibitemShut
+  {NoStop}%
+\bibitem [{\citenamefont {K{\"{u}}hne}(2014)}]{Kuhne2013}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {T.~D.}\ \bibnamefont
+  {K{\"{u}}hne}},\ }\href {\doibase 10.1002/wcms.1176} {\bibfield  {journal}
+  {\bibinfo  {journal} {Wiley Interdiscip. Rev. Comput. Mol. Sci.
+  }\ }\textbf {\bibinfo {volume} {4}},\ \bibinfo {pages} {391}
+  (\bibinfo {year} {2014})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Kubo}(1986)}]{a:Kubo-1986}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {R.}~\bibnamefont
+  {Kubo}},\ }\href {\doibase 10.1126/science.233.4761.330} {\bibfield
+  {journal} {\bibinfo  {journal} {Science}\ }\textbf {\bibinfo {volume}
+  {233}},\ \bibinfo {pages} {330} (\bibinfo {year} {1986})}\BibitemShut
+  {NoStop}%
+\bibitem [{\citenamefont {www.cp2k.org}(2017)}]{www:cp2k}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibnamefont {Www.cp2k.org}},\
+  }\href@noop {} {\enquote {\bibinfo {title} {{CP2K Open Source Molecular
+  Dynamics}},}\ } (\bibinfo {year} {Accessed August 18, 2017})\BibitemShut
+  {NoStop}%
+\bibitem [{\citenamefont {Ricci}\ and\ \citenamefont
+  {Ciccotti}(2003)}]{Ricci2003}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {A.}~\bibnamefont
+  {Ricci}}\ and\ \bibinfo {author} {\bibfnamefont {G.}~\bibnamefont
+  {Ciccotti}},\ }\href {\doibase 10.1080/0026897031000108113} {\bibfield
+  {journal} {\bibinfo  {journal} {Mol. Phys.}\ }\textbf {\bibinfo
+  {volume} {101}},\ \bibinfo {pages} {1927} (\bibinfo {year}
+  {2003})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Vandevondele}\ \emph {et~al.}(2005)\citenamefont
+  {Vandevondele}, \citenamefont {Krack}, \citenamefont {Mohamed}, \citenamefont
+  {Parrinello}, \citenamefont {Chassaing},\ and\ \citenamefont
+  {Hutter}}]{a:quickstep}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {J.}~\bibnamefont
+  {Vandevondele}}, \bibinfo {author} {\bibfnamefont {M.}~\bibnamefont {Krack}},
+  \bibinfo {author} {\bibfnamefont {F.}~\bibnamefont {Mohamed}}, \bibinfo
+  {author} {\bibfnamefont {M.}~\bibnamefont {Parrinello}}, \bibinfo {author}
+  {\bibfnamefont {T.}~\bibnamefont {Chassaing}}, \ and\ \bibinfo {author}
+  {\bibfnamefont {J.}~\bibnamefont {Hutter}},\ }\href@noop {} {\bibfield
+  {journal} {\bibinfo  {journal} {Comput. Phys. Commun.}\ }\textbf {\bibinfo
+  {volume} {167}},\ \bibinfo {pages} {103} (\bibinfo {year}
+  {2005})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Perdew}\ \emph {et~al.}(1996)\citenamefont {Perdew},
+  \citenamefont {Burke},\ and\ \citenamefont {Ernzerhof}}]{a:pbe}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {J.~P.}\ \bibnamefont
+  {Perdew}}, \bibinfo {author} {\bibfnamefont {K.}~\bibnamefont {Burke}}, \
+  and\ \bibinfo {author} {\bibfnamefont {M.}~\bibnamefont {Ernzerhof}},\
+  }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal} {Phys. Rev. Lett.}\
+  }\textbf {\bibinfo {volume} {77}},\ \bibinfo {pages} {3865} (\bibinfo {year}
+  {1996})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Grimme}\ \emph {et~al.}(2010)\citenamefont {Grimme},
+  \citenamefont {Antony}, \citenamefont {Ehrlich},\ and\ \citenamefont
+  {Krieg}}]{Grimme2010}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {S.}~\bibnamefont
+  {Grimme}}, \bibinfo {author} {\bibfnamefont {J.}~\bibnamefont {Antony}},
+  \bibinfo {author} {\bibfnamefont {S.}~\bibnamefont {Ehrlich}}, \ and\
+  \bibinfo {author} {\bibfnamefont {H.}~\bibnamefont {Krieg}},\ }\href
+  {\doibase 10.1063/1.3382344} {\bibfield  {journal} {\bibinfo  {journal} {
+  J. Chem. Phys.}\ }\textbf {\bibinfo {volume} {132}},\ \bibinfo
+  {pages} {154104} (\bibinfo {year} {2010})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Hartwigsen}\ \emph {et~al.}(1998)\citenamefont
+  {Hartwigsen}, \citenamefont {Goedecker},\ and\ \citenamefont
+  {Hutter}}]{a:hgh}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {C.}~\bibnamefont
+  {Hartwigsen}}, \bibinfo {author} {\bibfnamefont {S.}~\bibnamefont
+  {Goedecker}}, \ and\ \bibinfo {author} {\bibfnamefont {J.}~\bibnamefont
+  {Hutter}},\ }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal} {Phys.
+  Rev. B}\ }\textbf {\bibinfo {volume} {58}},\ \bibinfo {pages} {3641}
+  (\bibinfo {year} {1998})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Kolafa}(2004)}]{Kolafa2003}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {J.}~\bibnamefont
+  {Kolafa}},\ }\href {\doibase 10.1002/jcc.10385} {\bibfield  {journal}
+  {\bibinfo  {journal} {J. Comput. Chem.}\ }\textbf {\bibinfo
+  {volume} {25}},\ \bibinfo {pages} {335} (\bibinfo {year} {2004})}\BibitemShut
+  {NoStop}%
+\bibitem [{\citenamefont {Vandevondele}\ and\ \citenamefont
+  {Hutter}(2003)}]{a:ot}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {J.}~\bibnamefont
+  {Vandevondele}}\ and\ \bibinfo {author} {\bibfnamefont {J.}~\bibnamefont
+  {Hutter}},\ }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal} {J.
+  Chem. Phys.}\ }\textbf {\bibinfo {volume} {118}},\ \bibinfo {pages} {4365}
+  (\bibinfo {year} {2003})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Brehm}\ and\ \citenamefont
+  {Kirchner}(2011)}]{a:travis-main}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {M.}~\bibnamefont
+  {Brehm}}\ and\ \bibinfo {author} {\bibfnamefont {B.}~\bibnamefont
+  {Kirchner}},\ }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal} {J.
+  Chem. Inf. Model.}\ }\textbf {\bibinfo {volume} {51}},\ \bibinfo {pages}
+  {2007} (\bibinfo {year} {2011})}\BibitemShut {NoStop}%
+\end{thebibliography}%
 

### 2018-03-14 almo-langevin.tex
JPCL ready to submit

--- 2018-03-06 almo-langevin.tex
@@
-\documentclass[aps,prl,reprint,amsmath,amssymb]{revtex4-1}
-%\documentclass[journal=jacsat]{achemso}
-
-% define variable that compiles the main text (as opposed to SI text)
-\newcommand*{\MAINTEXT}{}
-
+%\documentclass[aps,prl,reprint,amsmath,amssymb]{revtex4-1}
 %\documentclass[10pt,aps,prl,twocolumn,amsmath,amssymb,superscriptaddress,longbibliography]{revtex4-1}
@@
 %\documentclass[prb,amsmath,amssymb]{revtex4}% Physical Review B
+
+% define variable that compiles the main text (as opposed to SI text)
+\newcommand*{\MAINTEXT}{}
+
+%%% JPCL
+\documentclass[journal=jpclcd,manuscript=letter]{achemso}
+\usepackage{achemso}
+%\usepackage[T1]{fontenc} % Use modern font encodings
+%%% end JPCL
 
@@
 
+\title{
+\ifdefined\MAINTEXT
+\else
+Supplemental Material: \\
+\fi
+Compact orbitals enable low-cost linear-scaling \emph{ab initio} molecular dynamics for weakly-interacting systems
+}
+
+\author{Hayden Scheiber}
+\email{scheiber@chem.ubc.ca}
+\affiliation{Department of Chemistry, McGill University, 801 Sherbrooke St. West, Montreal, QC~H3A~0B8, Canada}
+\author{Yifei Shi}
+\affiliation{Department of Chemistry, McGill University, 801 Sherbrooke St. West, Montreal, QC~H3A~0B8, Canada}
+\author{Rustam Z. Khaliullin}
+\email{rustam.khaliullin@mcgill.ca}
+\affiliation{Department of Chemistry, McGill University, 801 Sherbrooke St. West, Montreal, QC~H3A~0B8, Canada}
+
 \begin{document}
@@
 
-\title{
-\ifdefined\MAINTEXT
-\else
-Supplemental Material: \\
-\fi
-Low-cost orbital-based linear-scaling \emph{ab initio} molecular dynamics for weakly-interacting systems
-}
-
-\author{Hayden Scheiber}
-\email{scheiber@chem.ubc.ca}
-\author{Yifei Shi}
-\author{Rustam Z. Khaliullin}
-\email{rustam.khaliullin@mcgill.ca}
-\affiliation{Department of Chemistry, McGill University, 801 Sherbrooke St. West, Montreal, QC H3A 0B8, Canada}
-
-\date{\today}
+%\date{\today}
 
@@
 
+\begin{tocentry}
+%\begin{figure}
+\includegraphics[width=2in]{TOCG.eps}
+%\end{figure}
+\end{tocentry}
+
 \begin{abstract}
-Within the framework of linear-scaling Kohn-Sham density functional theory, a robust method for maintaining compact orbitals close to the ground state is coupled with nuclear dynamics to obviate the commonly employed optimization of the one-electron density matrix and to create an efficient orbital-only molecular dynamics method for weakly-interacting systems.
-An application to liquid water demonstrates that the low computational overhead of the method makes it well-suited for routine simulations whereas its linear scaling complexity allows to extend first-principle dynamical studies of molecular systems to previously inaccessible time and length scales.
+Today, \emph{ab initio} molecular dynamics (AIMD) relies on the locality of one-electron density matrices to achieve linear growth of computation time with systems size, crucial in large-scale simulations.
+While Kohn-Sham orbitals strictly localized within predefined radii can offer substantial computational advantages over density matrices, such compact orbitals are not used in AIMD because a compact representation of the electronic ground state is difficult to find.
+Here, a robust method for maintaining compact orbitals close to the ground state is coupled with a modified Langevin integrator to produce stable nuclear dynamics for molecular and ionic systems.
+This eliminates a density matrix optimization and enables first orbital-only linear-scaling AIMD.
+An application to liquid water demonstrates that low computational overhead of the new method makes it ideal for routine medium-scale simulations while its linear-scaling complexity allows to extend first-principle studies of molecular systems to completely new physical phenomena on previously inaccessible length scales.
+%Within the framework of linear-scaling Kohn-Sham density functional theory, a robust method for maintaining compact localized orbitals close to the ground state is coupled with nuclear dynamics.
+This allows to obviate the commonly employed optimization of the one-electron density matrix and thus create an efficient orbital-only molecular dynamics method for weakly-interacting systems.
+%An application to liquid water demonstrates that the low computational overhead of the method makes it well-suited for routine simulations whereas its linear-scaling complexity allows to extend first-principle dynamical studies of molecular systems to previously inaccessible length scales.
 \end{abstract}
@@
 
-\ifdefined\MAINTEXT
+%\ifdefined\MAINTEXT
 
@@
 Despite becoming the most popular approach to LS DFT, the efficiency of these methods is hampered by the costly optimization of both the contracted orbitals and the DM~\cite{a:ls-onetep-2003}. 
-From this point of view, a direct variation of molecular orbitals that are strictly localized within predefined regions is preferable because LS can be achieved with significantly fewer variables. 
+From this point of view, a direct variation of compact molecular orbitals---orbitals that are strictly localized within predefined regions---is preferable because LS can be achieved with significantly fewer variables.
 Advantages of the orbitals-only LS DFT are especially pronounced in accurate calculations that require many basis functions per atom.
@@
 
-In this work, we present an AIMD method that overcomes difficulties of orbital-only local DFT to achieve LS with extremely low computational overhead. 
+In this work, we present an AIMD method that overcomes difficulties of DFT based on compact orbitals to achieve LS with extremely low computational overhead. 
 To demonstrate advantages of the new method we applied it here to systems of weakly-interacting molecules. 
@@
 
-The new AIMD method utilizes a recently developed LS DFT~\cite{a:almo-ls} based on absolutely localized molecular orbitals (\mbox{ALMOs}). 
+The new AIMD method utilizes a recently developed LS DFT~\cite{a:almo-ls} based on absolutely localized molecular orbitals (\mbox{ALMOs}) -- compact orbitals first described in Ref.~\citenum{a:stoll}.
 Unlike delocalized KS orbitals, each \mbox{ALMO} has its own \emph{localization center} and a predefined \emph{localization radius} $R_{c}$ that typically includes nearby atoms or molecules~\cite{a:stoll,a:almo-ls}.
@@
 Our approach to compensate for the missing $\delta f_{i\alpha}(t)$ term is inspired by the methodology introduced into AIMD by Krajewski \emph{et al.}~\cite{Krajewski}, formalized by K\"uhne \emph{et al.}~\cite{a:2ndcpmd} and rationalized by Dai \emph{et al.}~\cite{a:langevin-why} before becoming informally known as the second generation Car-Parrinello molecular dynamics~\cite{Kuhne2013}.
-Adopting the principle of Refs.~\onlinecite{Krajewski} and \onlinecite{a:2ndcpmd}, ALMO AIMD is chosen to be governed by the Langevin equation of motion that can be written in terms of the unknown reference forces
+Adopting the principle of 
+%Refs.~\onlinecite{Krajewski} and \onlinecite{a:2ndcpmd}, 
+Refs.~\citenum{Krajewski} and~\citenum{a:2ndcpmd}, 
+ALMO AIMD is chosen to be governed by the Langevin equation of motion that can be written in terms of the unknown reference forces
 %
@@
 \begin{figure}
-\includegraphics[trim={1.3cm 0.1cm 3.3cm 1.3cm},clip,width=8.6cm]{Dynamical_Data_Tiled.eps}
+\includegraphics[trim={1.3cm 0.1cm 3.3cm 1.3cm},clip,width=13cm]{Dynamical_Data_Tiled.eps}
 \caption{\label{fig:dynproperties} 
@@
 \begin{figure}
-\includegraphics[trim={0.5cm 0cm 0.7cm 0.1cm},clip,width=8.6cm]{DeltaForceComparison_with_ACF.eps}
+\includegraphics[trim={0.5cm 0cm 0.7cm 0.1cm},clip,width=13cm]{DeltaForceComparison_with_ACF.eps}
 \caption{\label{fig:randomforce} 
@@
 \begin{figure}
-\includegraphics[trim={2.5cm 0.5cm 3.4cm 0.9cm},clip,width=8.6cm]{strongscaling_log.eps}
+\includegraphics[trim={2.5cm 0.5cm 3.4cm 0.9cm},clip,width=13cm]{strongscaling_log.eps}
 \caption{\label{fig:strongscaling_log} Timing benchmarks for PBE/TZV2P simulations of liquid water on 256 compute cores. 
@@
 \begin{figure}
-\includegraphics[trim={1.6cm 0.8cm 4.7cm 0cm},clip,width=8.6cm]{weakscaling.eps}
+\includegraphics[trim={1.6cm 0.8cm 4.7cm 0cm},clip,width=13cm]{weakscaling.eps}
 \caption{\label{fig:weakscaling} Weak scalability benchmarks for PBE/TZV2P ALMO AIMD with  $R_{c} = 1.6$ vdWR and $\epsilon_{\text{SCF}} = 10^{-2}$~a.u. 
@@
 
-\textbf{Acknowledgments.} The authors are grateful to Thomas K\"uhne for insightful discussions.
+\begin{acknowledgement}
+
+The authors are grateful to Thomas K\"uhne for insightful discussions.
 The research was funded by the Natural Sciences and Engineering Research Council of Canada through the Discovery Grant (RGPIN-2016-05059).
@@
 
-%\bibliography{almo-langevin}
-\input{almo-langevin.ref}
+\end{acknowledgement}
 
-\else
+\bibliography{almo-langevin}
+%\input{almo-langevin.ref}
+
+%\else
+
+\begin{suppinfo}
 
@@
 \begin{figure}[h!]
-\includegraphics[trim={0cm 0cm 0.1cm 0.1cm},clip,width=8.6cm]{DeltaForceComparison_ALMO_SCF.eps}
+\includegraphics[trim={0cm 0cm 0.1cm 0.1cm},clip,width=13cm]{DeltaForceComparison_ALMO_SCF.eps}
 \caption{\label{fig:forcecomp} Dependence of the Euclidean norm of the instantaneous error $\langle \| \delta \vec{f_{i}}(t) \| \rangle_{i}$ (red line) on the ALMO localization radius $R_c$ expressed in units of atomic van der Waals radii.
@@
 
-\fi % SI - main text switch
+\end{suppinfo}
+
+%\fi % SI - main text switch
 

### 2018-05-26 JCP/SI.tex
JCP Communication directory with the manuscript is added

\documentclass[aip,jcp,reprint,amsmath,amssymb]{revtex4-1}

\usepackage{graphicx}% Include figure files
\usepackage{dcolumn}% Align table columns on decimal point
\usepackage{bm}% bold math
\usepackage{amsmath}% bold math
\usepackage{siunitx}% si units
\usepackage{color}
\usepackage{graphicx}
%\usepackage{algorithmic}
\usepackage{algorithm}
\usepackage{algpseudocode}
%\usepackage[normalem]{ulem}
%\nofiles

%\bibliographystyle{achemso}
%\bibliographystyle{naturemag}
%\bibliographystyle{abbrv}

\begin{document}

\renewcommand{\thefigure}{S\arabic{figure}}

\title{
Supporting Information: \\
Compact orbitals enable low-cost linear-scaling \emph{ab initio} molecular dynamics for weakly-interacting systems
}

\author{Hayden Scheiber}
\email{scheiber@chem.ubc.ca}
\affiliation{Department of Chemistry, McGill University, 801 Sherbrooke St. West, Montreal, QC~H3A~0B8, Canada}
\author{Yifei Shi}
\affiliation{Department of Chemistry, McGill University, 801 Sherbrooke St. West, Montreal, QC~H3A~0B8, Canada}
\author{Rustam Z. Khaliullin}
\email{rustam.khaliullin@mcgill.ca}
\affiliation{Department of Chemistry, McGill University, 801 Sherbrooke St. West, Montreal, QC~H3A~0B8, Canada}

\maketitle
%\date{\today}

\subsection{Algorithms}

The pseudocode for the new \emph{ab initio} molecular dynamics (AIMD) method is shown in Figure~\ref{fig:ricci}. As described in the main text, the algorithm represents the Ricci-Ciccotti integrator~\cite{Ricci2003} modified (Lines~\ref{line:delta1} and \ref{line:delta2}) to take into account the imperfect forces computed using a constrained density functional theory (DFT) based on absolutely localized molecular orbitals (ALMOs) (Figure~\ref{fig:force}).

\begin{figure}
\begin{algorithm}[H]
  \caption{Modified Ricci-Ciccotti integrator}
  \label{alg:ricci}
   \begin{algorithmic}[1]
   	\State Input $t$ \Comment{Time step}
   	\State Input $N$ \Comment{Number of AIMD steps}
   	\State Input $T$ \Comment{Temperature}
   	\State Input $\gamma$ \Comment{Langevin damping coefficient}
   	\State Input $\Delta$ \Comment{Compensation for imperfect forces} \label{line:delta1}
   	\State Input $A$ \Comment{Number of atoms}
	\State $\tau\gets 0$ \Comment{Current time step}
	\State Allocate $\mathbf{r}, \mathbf{p}, \mathbf{f}$ \Comment{Matrices: positions, momenta, forces}
   	\State Input $\mathbf{r}$
	\State $\mathbf{p} \gets \text{Maxwell-Boltzmann}(T)$ \Comment{Generate initial momenta}
	\State $\mathbf{f} \gets \text{ALMO.Force}(\mathbf{r},\tau)$ \Comment{Compute forces (Algorithm~\ref{alg:force})}
	\For{$\tau \gets 1$ to $N$} \Comment{Time loop}
		\For{$i \gets 1$ to $A$} \Comment{Loop over atoms}
			\State $\sigma_i \gets 2 k_B T m_i t (\gamma + \Delta)$  \label{line:delta2}
			\State $\vec{w}_{i} \gets \mathcal{N}(0,\sigma_i)$ \Comment{Vector of random numbers}
			\State $\vec{r}_i \gets \vec{r}_i + \mathrm{e}^{-\gamma t/2} \frac{t}{m_i}\vec{p}_i + \mathrm{e}^{-\gamma t/4} \frac{t}{2 m_i} \left( t \vec{f}_i + \vec{w}_i \right) $
			\State $\vec{p}_i \gets \vec{p}_i + \mathrm{e}^{\gamma t/2} \frac{t}{2} \vec{f}_i$ \Comment{Start updating momenta}
		\EndFor
		\State $\mathbf{f} \gets \text{ALMO.Force}(\mathbf{r},\tau)$ \Comment{Update forces}
		\For{$i \gets 1$ to $A$}
%			\State $\vec{p}_i \gets \mathrm{e}^{-\gamma t} \left( \vec{p}_i + \mathrm{e}^{\gamma t/2} \frac{t}{2} \vec{f}_i \right) + \mathrm{e}^{-\gamma t/2} \vec{w}_i $   \Comment{Update momenta}
			\State $\vec{p}_i \gets \vec{p}_i + \mathrm{e}^{\gamma t/2} \frac{t}{2} \vec{f}_i $
			\State $\vec{p}_i \gets \mathrm{e}^{-\gamma t} \vec{p}_i + \mathrm{e}^{-\gamma t/2} \vec{w}_i $   \Comment{Update momenta}
		\EndFor
	\EndFor
   \end{algorithmic}
\end{algorithm}
\caption{\label{fig:ricci} ALMO AIMD algorithm.}
\end{figure}

\begin{figure}
\begin{algorithm}[H]
  \caption{Force evaluation}
  \label{alg:force}
   \begin{algorithmic}[1]
   	\Procedure{ALMO.Force}{$\mathbf{r}, \tau$}
   	\State Global $\mathbf{S}$ \Comment{Compute AO overlap}
	\State $\mathbf{T}_0 \gets \text{ALMO.Stage1}(\mathbf{r}, \tau)$ \Comment{SCF, Stage 1}
	\State $\mathbf{T}_{R_c} \gets \text{ALMO.Stage2}(\mathbf{r}, \tau, \mathbf{T}_0)$ \Comment{SCF, Stage 2}
	\State $\mathbf{\sigma}, \mathbf{R} \gets \text{Eq.}\ref{eq:dm}(\mathbf{T}_{R_c})$ \Comment{Compute DM}
	\State $\mathbf{F} \gets \text{DFT.BuildKS}(\mathbf{R})$ \Comment{Kohn-Sham matrix}
	\State $\mathbf{W} \gets \mathbf{RFR}$ \Comment{Energy-weighted DM}
	\State $\mathbf{f} \gets \text{DFT.HellmannFeynmanForce}(\mathbf{R},\mathbf{F}) + \text{DFT.PulayForce}(\mathbf{W},\mathbf{S}) $ \Comment{Atomic forces}
	\State $\mathbf{return}$ $\mathbf{f}$
	\EndProcedure
   \end{algorithmic}
\end{algorithm}
\caption{\label{fig:force} ALMO DFT algorithm to compute atomic forces.}
\end{figure}

As described in Refs.~\citenum{almo-ls}, ALMO DFT is designed to be computationally efficient for large systems that can be logically partitioned into small fragments---molecules, atoms, ions---not connected to each other by strong covalent bonds. ALMO DFT differs from the conventional DFT only by constraints imposed on electrons: all electrons and AOs in the system are assigned to fragments and each electron is allowed to delocalize only over its own fragment and all neighboring fragments within a pre-defined cutoff radius $R_c$ (Figure~\ref{fig:blocks}A). This restriction imposes the blocked pattern on electron's molecular orbital and the molecular orbital coefficient matrix $\mathbf{T}$ (Figure~\ref{fig:blocks}B).

To simplify equations that involve blocked matrices, the following notation is used. Square brackets $[\mathbf{M}]_{R_c}$ represent an operator that enforces the block structure on matrix $\mathbf{M}$ for a specified $R_c$ (Figure~\ref{fig:blocks}). Once the block structure is imposed any $\text{N}_{\text{AO}} \times \text{N}_{\text{MO}}$ matrix can be represented by two interconvertible formats shown in Figure~\ref{fig:blocks}B. The two formats, denoted by subscript or superscript $R_c$, are necessary to make matrix equations consistent throughout the pseudocode. If $R_c = 0$ the two formats are equivalent. For an $\text{N}_{\text{AO}} \times \text{N}_{\text{AO}}$ matrix, only one blocked format is defined. It is denoted by superscript $R_c$. When used without the superscript, an $\text{N}_{\text{AO}} \times \text{N}_{\text{AO}}$ matrix is not assumed to be blocked, which means that all its blocks are computed. Nevertheless, all algorithms presented here are linear scaling because of the natural sparsity of all $\text{N}_{\text{AO}} \times \text{N}_{\text{AO}}$ matrices. It is important to remind that the square bracket operator and blocked structures in Figure~\ref{fig:blocks} are just a convenient notation. The actual data structures and the implementation of matrix handling subroutines, designed to achieve linear scaling with low computational overhead, are significantly more complex.

\begin{figure*}
\includegraphics[trim={0cm 0cm 0.1cm 0.1cm},clip,width=17.5cm]{ALMO.eps}
\caption{\label{fig:blocks} Partitioning of a 4-molecule system and the resulting blocked structure imposed on matrices. A. Four molecules are represented by different colors. Solid lines connect neighbors lying within $R_c$ from each other. B. The corresponding blocked structure imposed on an $\text{N}_{\text{AO}} \times \text{N}_{\text{MO}}$ matrix (e.g. molecular orbital coefficient matrix, gradient matrix) and conversion between the two different formats denoted by subscript and superscript $R_c$. White blocks are zero. C. The blocked structure imposed on an $\text{N}_{\text{AO}}\times\text{N}_{\text{AO}}$ matrix (e.g. Kohn-Sham, density matrices).}
\end{figure*}

The algorithms for the force evaluation rely on the following conventions. Functions with names starting with \texttt{ALMO} are written out explicitly. Functions with names starting with \texttt{DFT} are conventional and are not shown. There are two functions that refer to the direct inversion in the iterative subspace (DIIS) method~\cite{ZZZ}. \texttt{DIIS.push} updates the DIIS matrix and enables \texttt{DIIS.extrapolate} to generate a new variable that, hopefully, minimizes the error. These functions are not written out explicitly. Functions with names starting with \texttt{HISTORY} store and retrieve the contravariant-covariant density matrices from recent AIMD steps. These matrices are necessary to generate a good initial guess for the electronic descriptors in the two self-consistent field (SCF) procedures. The atomic orbital (AO) overlap matrix $\mathbf{S}$ remains constant within one AIMD step and, therefore, is conveniently treated as a global variable visible in all procedures.

As shown in Figure~\ref{fig:force}, ALMO DFT evaluates the final density matrix (DM) in two SCF stages. In both stages, the overlap of ALMOs $\mathbf{\sigma}$ and the density matrix (DM) $\mathbf{R}$ is computed using Eq.~\ref{eq:dm}. Note that these two matrices are naturally sparse, that is, the block structure is not imposed on them.
%
\begin{eqnarray}\label{eq:dm}
\mathbf{\sigma} &=& \mathbf{T}_{R_c}^{\dagger}\mathbf{S} \mathbf{T}_{R_c} \nonumber \\
\mathbf{R} &=& \mathbf{T}_{R_c} \mathbf{\sigma}^{-1} \mathbf{T}_{R_c}^{\dagger}
\end{eqnarray}
%

In the first stage, molecular orbitals are constrained to their fragments and optimized using the DIIS accelerated algorithm in Figure~\ref{fig:almo0}. This algorithm is almost identical to the Stoll algorithm~\cite{a:stoll} of Ref.~\cite{a:khal} and refers to the following equations. The locally projected Hamiltonian~\cite{a:stoll} is computed using Eq.~\ref{eq:lph}. This equation is the same as Eq.~(51) in Ref.~\onlinecite{a:khal}.
%
\begin{eqnarray}\label{eq:lph}
\mathbf{L}_0 &=& [ (\mathbf{1} - \mathbf{SR}) \mathbf{F} (\mathbf{1} - \mathbf{RS}) ]_{0} + \nonumber\\
&+& [ (\mathbf{1} - \mathbf{SR}) \mathbf{FT}_0 \sigma^{-1}]_0 [ \mathbf{T}_0^{\dagger}\mathbf{S} ]_{0} \nonumber\\
&+& [ \mathbf{ST}_0 ]_{0} [ \sigma^{-1}\mathbf{T}_0^{\dagger}\mathbf{F}(\mathbf{1} - \mathbf{RS}) ]_0 + \nonumber\\
&+& [ \mathbf{ST}_0]_0 [ \sigma^{-1}\mathbf{T}_0^{\dagger}\mathbf{FT}_0 \sigma^{-1}]_{0} [ \mathbf{T}_0^{\dagger}\mathbf{S} ]_{0}
\end{eqnarray}
%
The DIIS error in the first stage is computed using Eq.~(\ref{eq:diis-error}) below, which is the same as Eq.~(39) in Ref.~\onlinecite{a:khal}.
%
\begin{eqnarray}\label{eq:diis-error}
\mathbf{G}_0 = [\mathbf{S}]_0 [\mathbf{RF}(\mathbf{RS-1})]_0 - [(\mathbf{SR-1})\mathbf{FR}]_0 [\mathbf{S}]_0
\end{eqnarray}

\begin{figure}
\begin{algorithm}[H]
  \caption{DIIS optimization of block-diagonal ALMOs}
  \label{alg:almo0}
   \begin{algorithmic}[1]
   	\Procedure{ALMO.Stage1}{$\mathbf{r}, \tau$}
	\State Input $\epsilon_{\text{SCF}}$ \Comment{Convergence threshold}
	\State Input $N_{\text{SCF}}$ \Comment{Max SCF iterations}
	\State $\mathbf{K}_0 \gets ([\mathbf{S}]_0)^{-1/2}$ \Comment{Orthogonalization matrix}
	\State $\mathbf{T}_0 \gets \text{ALMO.Guess}(1, 0.0)$ \Comment{MO guess}
	\State $\mathbf{\sigma,R} \gets \text{Eq.\ref{eq:dm}}(\mathbf{T}_0)$ \Comment{Compute DM}
	\State $\mathbf{F} \gets \text{DFT.BuildKS}(\mathbf{R})$ \Comment{Kohn-Sham matrix}

	\State StopSCF $\gets$ False \Comment{Prepare to exit SCF loop}
	\State $i_{\text{SCF}} \gets 0$ \Comment{Iteration counter}
	\Repeat \Comment{SCF loop}
		\State $i_{\text{SCF}} \gets i_{\text{SCF}} + 1$ 
		\State $\mathbf{L}_{0} \gets \text{Eq.}\ref{eq:lph}(\mathbf{F},\mathbf{R},\mathbf{T}_0,\mathbf{\sigma})$ \Comment{Projected KS matrix}
		\State $\mathbf{G}_0 \gets \text{Eq.\ref{eq:diis-error}}(\mathbf{F},\mathbf{R})$ \Comment{DIIS error}
		\State DIIS.push($\mathbf{L}_{0}, \mathbf{G}_{0}$)
		\State $\text{Error}_\text{SCF} \gets \vert\vert \mathbf{G}_0 \vert \vert_{\text{max}}$
		\If{$\text{Error}_\text{SCF} > \epsilon_{\text{SCF}}$}
			\State Converged $\gets$ False
		\Else
			\State Converged $\gets$ True
		\EndIf
		\If{$i_{\text{SCF}} > N_{\text{SCF}}$ \textbf{Or} Converged}
			\State StopSCF $\gets$ True
		\EndIf
		\If{$i_{\text{SCF}}=1$}
			\State StopSCF $\gets$ False \Comment{Do at least one iteration}
		\EndIf
		\If{Not StopSCF}
			\State $\mathbf{L}_{0}\gets$DIIS.extrapolate()
			\State $\mathbf{\tilde{M}}_{0} \gets \mathbf{K}_0 \mathbf{L}_{0} \mathbf{K}_0$ \Comment{KS in orthogonal basis}
			\State $\mathbf{\tilde{C}}_0 \gets \text{Solve}[\mathbf{\tilde{M}}_{0} \mathbf{\tilde{C}}_{0} = \mathbf{\tilde{C}}_{0} \mathbf{e}_{0}] $ \Comment{Eigensystem}
			\State $\mathbf{\tilde{T}}_{0} \gets \text{occ}(\mathbf{\tilde{C}}_{0})$ \Comment{Keep occupied orbitals}
			\State $\mathbf{T}_{0} \gets \mathbf{K}_0 \mathbf{\tilde{T}}_{0}$ \Comment{Into nonorthogonal basis}
			\State $\mathbf{\sigma,R} \gets \text{Eq.\ref{eq:dm}}(\mathbf{T}_0)$ \Comment{Compute DM}
			\State $\mathbf{F} \gets \text{DFT.BuildKS}(\mathbf{R})$ \Comment{Kohn-Sham matrix}
		\EndIf
	\Until{StopSCF}
	\State HISTORY.Stage1.Save($\mathbf{RS}$) \Comment{Needed in \texttt{ALMO.Guess}}
	\State HISTORY.Stage1.Save($\mathbf{T}_0$) \Comment{Needed in \texttt{ALMO.Guess}}
	\State $\mathbf{return}$ $\mathbf{T}_0$
	\EndProcedure
   \end{algorithmic}
\end{algorithm}
\caption{\label{fig:almo0} First stage of ALMO DFT computes block-diagonal orbitals ($R_c=0$) using DIIS accelerator.}
\end{figure}

In the second stage, ALMOs are relaxed to allow delocalization onto the neighboring fragments within localization radius $R_{c}$. In this stage, ALMOs are completely defined using the block-diagonal reference from the first stage and the delocalization amplitudes $\mathbf{X}_{R_c}$ (Line~\ref{line:almos} in Figure~\ref{fig:almoR}). It is the delocalization amplitudes that serve as the variational degrees of freedom (DOFs) in the optimization SCF procedure. The optimization is based on a preconditioned conjugate gradient algorithm shown in Figure~\ref{fig:almoR} and described in Ref.~\citenum{a:almo-ls}. The preconditioner in the second stage is computed using Eq.~(\ref{eq:prec}) below, which is the same as Eq.~(10) in Ref.~\onlinecite{a:almo-ls}.
%
\begin{eqnarray}\label{eq:prec}
\mathbf{P}^{R_c} =  (\mathbf{I} - \mathbf{M}^{R_c} \mathbf{K}^{R_c}) [\mathbf{S+F}]^{R_c} (\mathbf{I} - \mathbf{K}^{R_c} \mathbf{M}^{R_c})
\end{eqnarray}

\begin{figure}
\begin{algorithm}[H]
  \caption{PCG optimization of ALMOs}
  \label{alg:almoR}
   \begin{algorithmic}[1]
   	\Procedure{ALMO.Stage2}{$\mathbf{r}, \tau, \mathbf{T}_0$}
	\State Input $R_c$ \Comment{Electron localization radius}
	\State Input $\epsilon_{\text{SCF}}$ \Comment{Convergence threshold}
	\State Input $N_{\text{PCG}}$, $N_{\text{Outer}}$ \Comment{Max PCG, outer iterations}
	\State $\mathbf{K}^{R_c} \gets ([\mathbf{S}]^{R_c})^{-1}$ \Comment{Inverted blocked AO overlap}
	\State $\mathbf{\sigma,R} \gets \text{Eq.\ref{eq:dm}}(\mathbf{T}_{0})$ \Comment{DM}
	\State $\mathbf{M}^{R_c} \gets [\mathbf{SRS}]^{R_c}$ \Comment{Blocked covariant projector}

	\State $\mathbf{X}_{R_c} \gets \text{ALMO.Guess}(2, R_c, \mathbf{T}_0)$ \Comment{Variational DOFs}

	\State Converged $\gets$ False
	\State StopOuter $\gets$ False \Comment{Flag to exit the outer loop}
	\State $i_{\text{Outer}} \gets 0$ \Comment{Iteration counter}
	\Repeat \Comment{Outer loop restarts PCG}
		\State $i_{\text{Outer}} \gets i_{\text{Outer}} + 1$ 
		\State StopPCG $\gets$ False \Comment{Flag to exit the PCG loop}
		\State $i_{\text{PCG}} \gets 0$ \Comment{Iteration counter}
		\State $\beta \gets 0$ \Comment{Reset conjugation}
		\Repeat \Comment{PCG loop}
			\State $i_{\text{PCG}} \gets i_{\text{PCG}} + 1$ 
			\State $\mathbf{T}_{R_c} \gets \mathbf{T}_{0} + [(\mathbf{I} - \mathbf{K}^{R_c} \mathbf{M}^{R_c}) \mathbf{X}^{R_c}]_{R_c}$ \Comment{ALMO} \label{line:almos}
			\State $\mathbf{\sigma,R} \gets \text{Eq.\ref{eq:dm}}(\mathbf{T}_{R_c})$ \Comment{DM}
			\State $\mathbf{F} \gets \text{DFT.BuildKS}(\mathbf{R})$ \Comment{Kohn-Sham matrix}
			\If{$i_{\text{PCG}}>1$}
				\State $\mathbf{\Gamma}_{R_c} \gets \mathbf{G}_{R_c}$ \Comment{Save old gradient}
			\EndIf 
			\State $\mathbf{G}_{R_c} \gets [ (\mathbf{I} - \mathbf{SR}) \mathbf{FT}_{R_c} \mathbf{\sigma}^{-1}]_{R_c}$ \Comment{Gradient}
			\State $\text{Error}_\text{PCG} \gets \vert\vert \mathbf{G}_{R_c} \vert \vert_{\text{max}}$
			\If{$\text{Error}_\text{PCG} < \epsilon_{\text{SCF}}$}
				\State Converged $\gets$ True
			\EndIf
			\If{$i_{\text{PCG}} > N_{\text{PCG}}$ \textbf{Or} Converged}
				\State StopPCG $\gets$ True
			\EndIf
			\If{$i_{\text{PCG}}=1$ \textbf{And} $i_{\text{Outer}}=1$}
				\State StopPCG $\gets$ False \Comment{Do first iteration}
			\EndIf
			\If{\textbf{Not} StopPCG}
				\If{$i_{\text{PCG}} = 1$}
					\State $\mathbf{P}^{R_c} \gets \text{Eq.\ref{eq:prec}}(\mathbf{F},\mathbf{M}^{R_c},\mathbf{K}^{R_c}) $\Comment{Precon.}
				\Else
					\State $\mathbf{O}_{R_c} \gets \mathbf{D}_{R_c}$ \Comment{Save old direction}
				\EndIf
				\State $\mathbf{D}_{R_c} \gets - [(\mathbf{P}^{R_c})^{-1} \mathbf{G}^{R_c}]_{R_c}$ \Comment{Precon. grad.}
				\If{$i_{\text{PCG}}>1$}
					\State $\beta \gets \text{Tr}(\mathbf{G}^{\dagger}_{R_c} \mathbf{D}_{R_c})/\text{Tr}(\mathbf{\Gamma}^{\dagger}_{R_c}\mathbf{O}_{R_c})$
				\EndIf 
				\State $\mathbf{D}_{R_c} \gets \mathbf{D}_{R_c} + \beta \mathbf{O}_{R_c}$ \Comment{Search direction}
				\State $\alpha \gets \text{LineSearch}(\mathbf{D}_{R_c})$ \Comment{Step size}
				\State $\mathbf{X}_{R_c} \gets \mathbf{X}_{R_c} + \alpha \mathbf{D}_{R_c}$ \Comment{Update DOFs}
			\EndIf
		\Until{StopPCG} 
		\If{$i_{\text{Outer}} > N_{\text{Outer}}$ \textbf{Or} Converged}
			\State StopOuter $\gets$ True
		\EndIf
	\Until{StopOuter}
	\State HISTORY.Stage2.Save($\mathbf{RS}$) \Comment{Needed in \texttt{ALMO.Guess}}
	\State $\mathbf{return}$ $\mathbf{T}_{R_c}$
	\EndProcedure
   \end{algorithmic}
\end{algorithm}
\caption{\label{fig:almoR} Second stage of ALMO DFT computes orbitals localized on molecules and their neighbors ($R_c$ is finite). The SCF procedure is based on the preconditioned conjugate gradient (PCG) method.}
\end{figure}

The efficiency of both SCF stages is greatly improved by the availability of a good initial guess for the localized electronic descriptors. The initial guess is generated using an algorithm (Figure~\ref{fig:guess}) inspired by the always stable predictor-corrector method of Kolafa~\cite{Kolafa2004}.

\begin{figure}
\begin{algorithm}[H]
  \caption{Initial guess}
  \label{alg:guess}
   \begin{algorithmic}[1]
   	\Procedure{ALMO.Guess}{Stage, $R_c$, optional $\mathbf{T}_{0}$}

	\State $N \gets \text{HISTORY.AccumulatedLength()}$ \Comment{Max 3--5}

	\If{$N = 0$} \Comment{Generate new guess}

		\If{Stage = 1}
			\State $\mathbf{T}_{R_c} \gets \text{Random()}$ %[\text{DFT.SumOfAtomicDensities}]_0
			\State $\mathbf{T}_{R_c} \gets \mathbf{T}_{Rc} ([\mathbf{T}_{Rc}^{\dagger}\mathbf{S}\mathbf{T}_{Rc}]_0)^{-1/2}$ \Comment{Orthogonalize within blocks to preserve locality}
		\Else
			\State $\mathbf{T}_{R_c} \gets 0 $
		\EndIf

	\Else \Comment{Extrapolate}

		\State Allocate $\mathbf{RS}(N)$ \Comment{Array of contravariant-covariant DMs from previous AIMD steps, 1st element of the array is the most recent}	
		\If{Stage = 1}
			\State $\mathbf{RS} \gets \text{HISTORY.Stage1.Load()}$ %\Comment{Converged DMs from SCF Stage1}
		\Else
			\State $\mathbf{RS} \gets \text{HISTORY.Stage2.Load()}$ %\Comment{Converged DMs from SCF Stage2}
		\EndIf
		\State $\mathbf{T}_{0}(1) \gets \text{HISTORY.Stage1.Load()}$ \Comment{Converged block-diagonal ALMOs from the previous AIMD step}
		\State $\mathbf{T}_{R_c} \gets 0$

		\For{	$m \gets 1, N$}
			\State $\alpha \gets (-1)^{m+1} m \binom{2N}{ N-m} / \binom{2N-2}{N-1}$ 
			\State $\mathbf{T}_{R_c} \gets \mathbf{T}_{R_c} + \alpha [\mathbf{RS}(m) \mathbf{T}_{0}(1)]_{R_c} $ 
			%\State $\mathbf{E}_{R_c} \gets [\mathbf{R}(m) \mathbf{S}(m) \mathbf{T}_{0}(1)]_{R_c}$
			%\State $\mathbf{E}_{R_c} \gets \mathbf{E}_{Rc} ([\mathbf{E}_{Rc}^{\dagger}\mathbf{S}\mathbf{E}_{Rc}]_0)^{-1/2}$
			%\State $\mathbf{T}_{R_c} \gets \mathbf{T}_{R_c} + \frac{\alpha}{N} \mathbf{E}_{Rc}$
		\EndFor
		\State $\mathbf{T}_{R_c} \gets \mathbf{T}_{Rc} ([\mathbf{T}_{Rc}^{\dagger}\mathbf{S}\mathbf{T}_{Rc}]_0)^{-1/2}$ \Comment{Orthogonalize within blocks to preserve locality}

		\If{ Stage = 2}
			\State $\mathbf{T}_{R_c} \gets \mathbf{T}_{R_c}  - \mathbf{T}_{0}$
		\EndIf

	\EndIf

	\State $\mathbf{return}$ $\mathbf{T}_{R_c}$
	\EndProcedure
   \end{algorithmic}
\end{algorithm}
\caption{\label{fig:guess} Initial guess for the localized electronic descriptors. The algorithm is inspired by always stable predictor-corrector method of Kolafa~\cite{Kolafa2004}.}
\end{figure}

\subsection{Dependence of the forces on the electron localization radius}

The ALMO forces converge to the reference forces as the electron localization radius increases (Figure~\ref{fig:forcecomp}).

\begin{figure}[h!]
\includegraphics[trim={0cm 0cm 0.1cm 0.1cm},clip,width=8.6cm]{5.eps}
\caption{\label{fig:forcecomp} Dependence of the Euclidean norm of the instantaneous error $\langle \| \delta \vec{f_{i}}(t) \| \rangle_{i}$ (red line) on the ALMO localization radius $R_c$ expressed in units of atomic van der Waals radii. Black line is the magnitude of the time average of the instantaneous error vector. %calculated as the average difference between fully converged Born-Oppenheimer forces and fully converged ALMO DFT forces, averaged over all dimensions and all molecules for each time step. 
Forces are computed with the PBE exchange-correlation functional, TZV2P basis set, and $\epsilon_{\text{SCF}} = 10^{-7}$~a.u. for the configurations from a 30-ps trajectory generated with delocalized AIMD at T=298~K.}
\end{figure}

\end{document}

### 2018-05-26 JCP/manuscript.tex
JCP Communication directory with the manuscript is added

--- 2018-03-14 almo-langevin.tex
@@
-%\documentclass[aps,prl,reprint,amsmath,amssymb]{revtex4-1}
-%\documentclass[10pt,aps,prl,twocolumn,amsmath,amssymb,superscriptaddress,longbibliography]{revtex4-1}
-%\documentclass[preprint,showpacs,preprintnumbers,amsmath,amssymb]{revtex4-1}
-%\documentclass[twocolumn,showpacs,preprintnumbers,amsmath,amssymb]{revtex4}
-% Some other (several out of many) possibilities
-%\documentclass[preprint,aps]{revtex4}
-%\documentclass[preprint,aps,draft]{revtex4}
-%\documentclass[prb,amsmath,amssymb]{revtex4}% Physical Review B
-
-% define variable that compiles the main text (as opposed to SI text)
-\newcommand*{\MAINTEXT}{}
-
-%%% JPCL
-\documentclass[journal=jpclcd,manuscript=letter]{achemso}
-\usepackage{achemso}
-%\usepackage[T1]{fontenc} % Use modern font encodings
-%%% end JPCL
+\documentclass[aip,jcp,reprint,amsmath,amssymb]{revtex4-1}
 
@@
 
+\begin{document}
+
 \title{
-\ifdefined\MAINTEXT
-\else
-Supplemental Material: \\
-\fi
 Compact orbitals enable low-cost linear-scaling \emph{ab initio} molecular dynamics for weakly-interacting systems
@@
 
-\begin{document}
-
-\ifdefined\MAINTEXT
-\else
-	\clearpage
-	%\widetext
-	\setcounter{figure}{0}
-	\setcounter{page}{1}
-	\renewcommand{\thefigure}{S\arabic{figure}}
-\fi
-
 %\date{\today}
 
-\ifdefined\MAINTEXT
-
-\begin{tocentry}
-%\begin{figure}
-\includegraphics[width=2in]{TOCG.eps}
-%\end{figure}
-\end{tocentry}
+%\begin{tocentry}
+%%\begin{figure}
+%\includegraphics[width=2in]{TOCG.eps}
+%%\end{figure}
+%\end{tocentry}
 
@@
 
-\fi
-
 \maketitle
-
-%\ifdefined\MAINTEXT
 
@@
 To achieve a robust optimization in the problematic second stage, it is important to keep the delocalization component of the trial wavefunction orthogonal to the fixed orbitals obtained in the first stage.
-For mathematical details, see the ALMO SCF method in Ref.~\citenum{a:almo-ls}.
+For mathematical details, see the ALMO SCF method in Ref.~\citenum{a:almo-ls}. ALMO DFT agorithms are presented in the Supporting Information.
 
@@
 if one can afford computing the reference forces $f^{\text{KS}}_{i\alpha}(t)$ (i.e. $R_c \rightarrow \infty$ and $\epsilon_{\text{SCF}} \rightarrow 0$) for a short representative AIMD trajectory. 
-In practice, we found (see results below) that this approach is not particularly accurate because the $\delta_{ij}\delta_{\alpha\beta}$ assumption in Eq.~(\ref{eq:stochastic3}) does not strictly hold.
-Nevertheless, the ACF integral can provide a reasonable starting value of $\Delta$.
-This value can be further fine-tuned in a series of short trial-and-error ALMO AIMD runs until the average kinetic energy corresponds to the requested temperature $\langle \frac{1}{2} m_i \dot{\bm{r}}^{2}_{i} \rangle = \frac{3}{2} k_{B} T$.
+In practice, we found (see results below) that this approach is not sufficiently accurate because the $\delta_{ij}\delta_{\alpha\beta}$ assumption in Eq.~(\ref{eq:stochastic3}) does not strictly hold.
+Nevertheless, the integral of the autocorrelation function (ACF) in Eq.~\ref{eq:delta} can provide a reasonable starting value of $\Delta$.
+%If the reference forces cannot be computed even for a short simulations, then a starting calue of $\Delta$ can be determined .
+The estimated $\Delta$ can be further fine-tuned in a series of short trial-and-error ALMO AIMD runs until the average kinetic energy corresponds to the requested temperature $\langle \frac{1}{2} m_i \dot{\bm{r}}^{2}_{i} \rangle = \frac{3}{2} k_{B} T$.
 
@@
 Nevertheless, it is capable to reproduce correct dynamical properties of a system as long as $\gamma$ is set to a small value and  partially optimized ALMOs remain close to the ground state resulting in $\Delta \ll \gamma$.
-%With $\gamma$ set to zero, the nuclei will follow time-reversible Newtonian trajectories closely enough to reproduce dynamical properties of the ensemble as long as partially optimized ALMOs remain close to the ground state and $\Delta$ is small.
+An alternative approach to performing microcanonical MO-based LS AIMD would be to combine ALMO DFT with the extended Lagrangian method described in Ref.~\onlinecite{Niklasson2008} that has been shown to conserve the total energy in DM-based simulations~\cite{Cawkwell2012}.
+
+% RZK
+% Niklasson2008  --> DOI: 10.1103/PhysRevLett.100.123004
+% Cawkwell2012 -->  doi: 10.1063/1.4755991
 
@@
 \begin{figure}
-\includegraphics[trim={1.3cm 0.1cm 3.3cm 1.3cm},clip,width=13cm]{Dynamical_Data_Tiled.eps}
+\includegraphics[trim={1.3cm 0.1cm 3.3cm 1.3cm},clip,width=8.6cm]{1.eps}
 \caption{\label{fig:dynproperties} 
@@
 A periodic cell containing 125 molecules was simulated for $\SI{30}{\ps}$ at $T=\SI{298}{\K}$ and a constant density of $\SI{1.01}{\g\cdot\cm^{-3}}$. 
-Ricci-Ciccotti algorithm~\cite{Ricci2003} was used to integrate the Langevin equation.
+The initial configurations of the water boxes were created by equilibrating the system for 20~ps using the fully converged orbitals. 
+Ricci-Ciccotti algorithm~\cite{Ricci2003} was used to integrate the Langevin equation (see Supporting Information).
 We found that $\gamma = \SI{e-3}{\per\fs}$ is large enough to thermostat the system efficiently and small enough not to significantly affect dynamical properties of liquid water.
@@
 \begin{figure}
-\includegraphics[trim={0.5cm 0cm 0.7cm 0.1cm},clip,width=13cm]{DeltaForceComparison_with_ACF.eps}
+\includegraphics[trim={0.5cm 0cm 0.7cm 0.1cm},clip,width=8.6cm]{2.eps}
 \caption{\label{fig:randomforce} 
@@
 %(Figure~\ref{fig:forcecomp} in the Supplemental Material).
-(Figure~S1 in the Supplemental Material).
+(Figure~SZZZ in the Supporting Information).
 
@@
 The crossover point between ALMO AIMD and cubically scaling methods lies in the region of 256 molecules---length scales routinely accessible with AIMD today.
+For comparison, the crossover point between the cubically-scaling DFT and a representative LS method that computes a sparse DM as the matrix sign function of the effective Hamiltonian~\cite{a:ls-dm-sign} is expected to lie well above 10,000 water molecules (Figures 5a and S5 in Ref.~\onlinecite{:a:almo-ls}).
 
@@
 \begin{figure}
-\includegraphics[trim={2.5cm 0.5cm 3.4cm 0.9cm},clip,width=13cm]{strongscaling_log.eps}
+\includegraphics[trim={2.5cm 0.5cm 3.4cm 0.9cm},clip,width=8.6cm]{3.eps}
 \caption{\label{fig:strongscaling_log} Timing benchmarks for PBE/TZV2P simulations of liquid water on 256 compute cores. 
@@
 \begin{figure}
-\includegraphics[trim={1.6cm 0.8cm 4.7cm 0cm},clip,width=13cm]{weakscaling.eps}
+\includegraphics[trim={1.6cm 0.8cm 4.7cm 0cm},clip,width=8.6cm]{4.eps}
 \caption{\label{fig:weakscaling} Weak scalability benchmarks for PBE/TZV2P ALMO AIMD with  $R_{c} = 1.6$ vdWR and $\epsilon_{\text{SCF}} = 10^{-2}$~a.u. 
@@
 The developed method will have a significant impact on modeling of complex molecular systems (e.g. interfaces or nuclei) making completely new phenomena accessible to AIMD. 
-Generalization of the methodology to systems of strongly interacting atoms (e.g. covalent crystals) is underway.
+A generalization of the methodology to systems of strongly interacting atoms (e.g. covalent crystals) is underway.
 
-\begin{acknowledgement}
+%\begin{suppinfo}
 
-The authors are grateful to Thomas K\"uhne for insightful discussions.
+%Convergence of approximate forces to the reference forces with the electron localization radius.
+
+%\end{suppinfo}
+
+%\begin{acknowledgement}
+
+\textbf{Acknowledgements.} The authors are grateful to Thomas K\"uhne for insightful discussions.
 The research was funded by the Natural Sciences and Engineering Research Council of Canada through the Discovery Grant (RGPIN-2016-05059).
@@
 
-\end{acknowledgement}
+%\end{acknowledgement}
 
-\bibliography{almo-langevin}
-%\input{almo-langevin.ref}
-
-%\else
-
-\begin{suppinfo}
-
-\subsection{Dependence of the forces on the electron localization radius}
-
-The ALMO forces converge to the reference forces as the electron localization radius increases (Figure~\ref{fig:forcecomp}).
-
-\begin{figure}[h!]
-\includegraphics[trim={0cm 0cm 0.1cm 0.1cm},clip,width=13cm]{DeltaForceComparison_ALMO_SCF.eps}
-\caption{\label{fig:forcecomp} Dependence of the Euclidean norm of the instantaneous error $\langle \| \delta \vec{f_{i}}(t) \| \rangle_{i}$ (red line) on the ALMO localization radius $R_c$ expressed in units of atomic van der Waals radii.
-Black line is the magnitude of the time average of the instantaneous error vector.
-%calculated as the average difference between fully converged Born-Oppenheimer forces and fully converged ALMO DFT forces, averaged over all dimensions and all molecules for each time step.
-Forces are computed with the PBE XC functional, TZV2P basis set, and $\epsilon_{\text{SCF}} = 10^{-7}$~a.u. for the configurations from a 30-ps trajectory generated with delocalized AIMD at T=298~K.}
-\end{figure}
-
-%\begin{figure}[!htb]
-%\includegraphics[trim={1.7cm 0.5cm 3.3cm 0.1cm},clip,width=\textwidth]{strongscaling_linear.eps}
-%\caption{\label{fig:strongscaling_linear} Linear plots showing LS behaviour of all ALMO methods for liquid water at 298 K. 
-%Calculations were performed with the PBE $\mathrm{E_{XC}}$ functional and TZV2P basis set, averaged over 10 MD steps. 
-%The localization radius was set to $R_{c} = 1.6$ vdWR and $\epsilon_{\text{SCF}} = 10^{-2}$~a.u.}
-%\end{figure}
-
-%\begin{figure}[!htb]
-%\includegraphics[trim={2.4cm 0cm 3.0cm 1cm},clip,width=8.6cm]{Diffusion_Comparison.eps}
-%\caption{\label{fig:diffusion} Calculated mean-squared displacement function of water using ALMO DFT with $\epsilon_{\text{SCF}} = 10^{-2}$~a.u. and $R_{c} = 1.6$ vdWR.
-%Compared with equivalent MD run using OT DFT as a reference.
-%Calculations were performed at 298 K with the dispersion-corrected PBE $\mathrm{E_{XC}}$ functional and TZV2P basis set, using 125 water molecules for 30 ps.}
-%\end{figure}
-
-\end{suppinfo}
-
-%\fi % SI - main text switch
+\bibliography{manuscript}
 

### 2018-05-26 JCP/manuscript.tex
JCP: update

@@
 
+\usepackage{xr}
 \usepackage{graphicx}% Include figure files
@@
 %\bibliographystyle{abbrv}
+
+\externaldocument[SI-]{SI}
 
@@
 Accuracy and efficiency of ALMO AIMD was tested using liquid water as an example. 
-%RZK: cite our own water paper
 This system is challenging because intermolecular electron delocalization is a critical component of hydrogen bonding and must be described correctly to reproduce static and dynamical properties of liquid water.
-%To calculate properties of water at ambient conditions, 
 A periodic cell containing 125 molecules was simulated for $\SI{30}{\ps}$ at $T=\SI{298}{\K}$ and a constant density of $\SI{1.01}{\g\cdot\cm^{-3}}$. 
@@
 %(Figure~\ref{fig:forcecomp} in the Supplemental Material).
-(Figure~SZZZ in the Supporting Information).
+(Figure~\ref{SI-fig:forcecomp} in the Supporting Information).
 
@@
 The crossover point between ALMO AIMD and cubically scaling methods lies in the region of 256 molecules---length scales routinely accessible with AIMD today.
-For comparison, the crossover point between the cubically-scaling DFT and a representative LS method that computes a sparse DM as the matrix sign function of the effective Hamiltonian~\cite{a:ls-dm-sign} is expected to lie well above 10,000 water molecules (Figures 5a and S5 in Ref.~\onlinecite{:a:almo-ls}).
+For comparison, the crossover point between the cubically-scaling DFT and a representative LS method that computes a sparse DM as the matrix sign function of the effective Hamiltonian~\cite{a:ls-dm-sign} is expected to lie well above 10,000 water molecules (Figures 5a and S5 in Ref.~\onlinecite{a:almo-ls}).
 

### 2018-05-27 JCP/manuscript.tex
Modified Figure 2 + Updated Refs
Made Figure 2 Longer, and added the two requested references to the .bib file.
minor corrections before merge

@@
 Nevertheless, it is capable to reproduce correct dynamical properties of a system as long as $\gamma$ is set to a small value and  partially optimized ALMOs remain close to the ground state resulting in $\Delta \ll \gamma$.
-An alternative approach to performing microcanonical MO-based LS AIMD would be to combine ALMO DFT with the extended Lagrangian method described in Ref.~\onlinecite{Niklasson2008} that has been shown to conserve the total energy in DM-based simulations~\cite{Cawkwell2012}.
+An alternative approach to performing microcanonical MO-based LS AIMD would be to combine ALMO DFT with the extended Lagrangian method described in Ref.~\onlinecite{Niklasson2008}.
+This method has been shown to conserve the total energy in LS DM-based AIMD simulations~\cite{Cawkwell2012}.
 

### 2018-05-28 JCP/manuscript.tex
minor corrections after merge
Minor changes before submitting the review

@@
 To achieve a robust optimization in the problematic second stage, it is important to keep the delocalization component of the trial wavefunction orthogonal to the fixed orbitals obtained in the first stage.
-For mathematical details, see the ALMO SCF method in Ref.~\citenum{a:almo-ls}. ALMO DFT agorithms are presented in the Supporting Information.
+For mathematical details, see the ALMO SCF method in Ref.~\citenum{a:almo-ls}. ALMO DFT agorithms are presented in the Supplementary Material.
 
@@
 A periodic cell containing 125 molecules was simulated for $\SI{30}{\ps}$ at $T=\SI{298}{\K}$ and a constant density of $\SI{1.01}{\g\cdot\cm^{-3}}$. 
-The initial configurations of the water boxes were created by equilibrating the system for 20~ps using the fully converged orbitals. 
-Ricci-Ciccotti algorithm~\cite{Ricci2003} was used to integrate the Langevin equation (see Supporting Information).
+The initial configurations of the water boxes were created by equilibrating the system for 20~ps using fully converged orbitals. 
+Ricci-Ciccotti algorithm~\cite{Ricci2003} was used to integrate the Langevin equation (see Supplementary Material).
 We found that $\gamma = \SI{e-3}{\per\fs}$ is large enough to thermostat the system efficiently and small enough not to significantly affect dynamical properties of liquid water.
@@
 \begin{figure}
-\includegraphics[trim={0.5cm 0cm 0.7cm 0.1cm},clip,width=8.6cm]{2.eps}
+\includegraphics[trim={0.cm 0.cm 0.cm 0.1cm},clip,width=8.6cm]{2.eps}
 \caption{\label{fig:randomforce} 
@@
 We established that the main source of error in forces for this system is the loose convergence criterion and not the finite $R_c$: fully converged ALMO SCF calculations remove the oscillating component of $\delta f$ (Figure~\ref{fig:randomforce}b).
-We also verified that the ALMO forces converge to the reference forces in the limit $R_{c} \rightarrow \infty$
-%(Figure~\ref{fig:forcecomp} in the Supplemental Material).
-(Figure~\ref{SI-fig:forcecomp} in the Supporting Information).
+We also verified that the ALMO forces converge to the reference forces in the limit $R_{c} \rightarrow \infty$ (Figure~\ref{SI-fig:forcecomp} in Supplementary Material).
 
@@
 
-%Convergence of approximate forces to the reference forces with the electron localization radius.
+See supplementary material for the detailed description of ALMO AIMD algorithms and dependence of the atomic forces on the electron localization radius.
 
@@
 
-\textbf{Acknowledgements.} The authors are grateful to Thomas K\"uhne for insightful discussions.
+%\textbf{Acknowledgements.} 
+The authors are grateful to Thomas K\"uhne for insightful discussions.
 The research was funded by the Natural Sciences and Engineering Research Council of Canada through the Discovery Grant (RGPIN-2016-05059).

### 2018-05-28 JCP/manuscript.tex
Re-edit to figure 2
Changed height of panels A and B in figure 2 (again). Should be about 3.3cm total.

--- 2018-05-27 JCP/manuscript.tex
@@
 A periodic cell containing 125 molecules was simulated for $\SI{30}{\ps}$ at $T=\SI{298}{\K}$ and a constant density of $\SI{1.01}{\g\cdot\cm^{-3}}$. 
-The initial configurations of the water boxes were created by equilibrating the system for 20~ps using the fully converged orbitals. 
+The initial configurations of the water boxes were created by equilibrating the system for 20~ps using fully converged orbitals. 
 Ricci-Ciccotti algorithm~\cite{Ricci2003} was used to integrate the Langevin equation (see Supporting Information).
@@
 \begin{figure}
-\includegraphics[trim={0.5cm 0cm 0.7cm 0.1cm},clip,width=8.6cm]{2.eps}
+\includegraphics[trim={0cm 0cm 0cm 0cm},clip,width=8.6cm]{2.eps}
 \caption{\label{fig:randomforce} 

### 2018-05-28 JCP/SI.tex

@@
 \title{
-Supporting Information: \\
+Supplementary Material: \\
 Compact orbitals enable low-cost linear-scaling \emph{ab initio} molecular dynamics for weakly-interacting systems
@@
 
-As described in Refs.~\citenum{almo-ls}, ALMO DFT is designed to be computationally efficient for large systems that can be logically partitioned into small fragments---molecules, atoms, ions---not connected to each other by strong covalent bonds.
+As described in Refs.~\citenum{a:almo-ls}, ALMO DFT is designed to be computationally efficient for large systems that can be logically partitioned into small fragments---molecules, atoms, ions---not connected to each other by strong covalent bonds.
 ALMO DFT differs from the conventional DFT only by constraints imposed on electrons: all electrons and AOs in the system are assigned to fragments and each electron is allowed to delocalize only over its own fragment and all neighboring fragments within a pre-defined cutoff radius $R_c$ (Figure~\ref{fig:blocks}A).
@@
 Functions with names starting with \texttt{DFT} are conventional and are not shown.
-There are two functions that refer to the direct inversion in the iterative subspace (DIIS) method~\cite{ZZZ}.
+There are two functions that refer to the direct inversion in the iterative subspace (DIIS) method.
 \texttt{DIIS.push} updates the DIIS matrix and enables \texttt{DIIS.extrapolate} to generate a new variable that, hopefully, minimizes the error.
@@
 The efficiency of both SCF stages is greatly improved by the availability of a good initial guess for the localized electronic descriptors.
-The initial guess is generated using an algorithm (Figure~\ref{fig:guess}) inspired by the always stable predictor-corrector method of Kolafa~\cite{Kolafa2004}.
+The initial guess is generated using an algorithm (Figure~\ref{fig:guess}) inspired by the always stable predictor-corrector method of Kolafa~\cite{Kolafa2003}.
 
@@
 \end{algorithm}
-\caption{\label{fig:guess} Initial guess for the localized electronic descriptors. The algorithm is inspired by always stable predictor-corrector method of Kolafa~\cite{Kolafa2004}.}
+\caption{\label{fig:guess} Initial guess for the localized electronic descriptors. The algorithm is inspired by always stable predictor-corrector method of Kolafa~\cite{Kolafa2003}.}
 \end{figure}
@@
 
+\bibliography{SI}
+
 \end{document}

### 2018-05-28 JCP/manuscript.tex

@@
 To achieve a robust optimization in the problematic second stage, it is important to keep the delocalization component of the trial wavefunction orthogonal to the fixed orbitals obtained in the first stage.
-For mathematical details, see the ALMO SCF method in Ref.~\citenum{a:almo-ls}. ALMO DFT agorithms are presented in the Supporting Information.
 
@@
 The initial configurations of the water boxes were created by equilibrating the system for 20~ps using fully converged orbitals. 
-Ricci-Ciccotti algorithm~\cite{Ricci2003} was used to integrate the Langevin equation (see Supporting Information).
-We found that $\gamma = \SI{e-3}{\per\fs}$ is large enough to thermostat the system efficiently and small enough not to significantly affect dynamical properties of liquid water.
 %
@@
 %
-Analysis of $\delta \vec{f_{i}}(t)$ shows that the error indeed resembles Gaussian white noise.
-The mean of the error is zero (black line in Figure~\ref{fig:randomforce}a).
-Its ACF decays rapidly (Figure~\ref{fig:randomforce}c) so that the errors can be considered uncorrelated on time scale of $\SI{50}{\fs}$.
-Thus the main assumption behind our approach to ALMO AIMD is justified for liquid water.
-We established that the main source of error in forces for this system is the loose convergence criterion and not the finite $R_c$: fully converged ALMO SCF calculations remove the oscillating component of $\delta f$ (Figure~\ref{fig:randomforce}b).
-We also verified that the ALMO forces converge to the reference forces in the limit $R_{c} \rightarrow \infty$
-%(Figure~\ref{fig:forcecomp} in the Supplemental Material).
-(Figure~\ref{SI-fig:forcecomp} in the Supporting Information).
 
@@
 
-%Convergence of approximate forces to the reference forces with the electron localization radius.
-
 %\end{suppinfo}
@@
 %\begin{acknowledgement}
-
-\textbf{Acknowledgements.} The authors are grateful to Thomas K\"uhne for insightful discussions.
-The research was funded by the Natural Sciences and Engineering Research Council of Canada through the Discovery Grant (RGPIN-2016-05059).
-The authors are grateful to Compute Canada and McGill HPC Centre for computer time.
 
